### 2022-01-27 Manuscript.tex
Copied the manuscript file from the main branch

\documentclass[preprint,12pt,TurnOnLineNumbers]{JASAnew}

\pdfoutput=1
\hyphenation{echo-sounders}
\renewcommand{\figurename}{Figure}

\newcommand{\ek}{Simrad EK80}

\newcommand{\timesym}{t}
\newcommand{\freqsym}{f}
\newcommand{\samplesymt}{n}
\newcommand{\samplesymf}{m}
\newcommand{\genidxsym}{i}

\newcommand{\channelsym}{u}
\newcommand{\nchannels}{N_{\textrm{u}}}

\newcommand{\stagesym}{v}
\newcommand{\nstages}{N_{\textrm{v}}}

\newcommand{\fs}{f_{\textrm{s}}}
\newcommand{\fsdec}{f_{\textrm{s,dec}}}
\newcommand{\fstart}{f_{\textrm{start}}}
\newcommand{\fstop}{f_{\textrm{stop}}}
\newcommand{\fc}{f_{\textrm{c}}}
\newcommand{\fn}{f_{\textrm{n}}}

\newcommand{\zrxe}{z_{\textrm{rx,e}}}
\newcommand{\ztde}{z_{\textrm{td,e}}}

\newcommand{\ptxe}{p_{\textrm{tx,e}}}
\newcommand{\prxe}{p_{\textrm{rx,e}}}

\newcommand{\ntd}{n_{\textrm{td}}}
\newcommand{\tnom}{\tau}
\newcommand{\teff}{\tau_{\textrm{eff}}}

\newcommand{\ytxe}{y_{\textrm{tx,e}}}
\newcommand{\ytxa}{y_{\textrm{tx,a}}}
\newcommand{\yrxa}{y_{\textrm{rx,a}}}
\newcommand{\yrxe}{y_{\textrm{rx,e}}}

\newcommand{\ytx}{y_{\textrm{tx}}}
\newcommand{\ytxnorm}{\tilde{y}_{\textrm{tx}}}
\newcommand{\ytxfd}{y_{\textrm{tx,f,d}}}
\newcommand{\yrx}{y_{\textrm{rx}}}
\newcommand{\yrxorg}{y_{\textrm{rx,org}}}
\newcommand{\ymf}{y_{\textrm{mf}}}

\newcommand{\ytd}{y_{\textrm{td}}}

\newcommand{\ypc}{y_{\textrm{pc}}}
\newcommand{\ypctarget}{y_{\textrm{pc,t}}}
\newcommand{\ypcspread}{y_{\textrm{pc,s}}}
\newcommand{\ymfauto}{y_{\textrm{mf,auto}}}
\newcommand{\ymfautored}{y_{\textrm{mf,auto,red}}}

\newcommand{\ptxauto}{p_{\textrm{tx,auto}}}

\newcommand{\ypctargetf}{Y_{\textrm{pc,t}}}
\newcommand{\ypctargetnormf}{\tilde{Y}_{\textrm{pc,t}}}
\newcommand{\ypcvolumef}{Y_{\textrm{pc,v}}}
\newcommand{\ypcvolumenormf}{\tilde{Y}_{\textrm{pc,v}}}

\newcommand{\ymfautof}{Y_{\textrm{mf,auto}}}
\newcommand{\ymfautoredf}{Y_{\textrm{mf,auto,red}}}

\newcommand{\prxetf}{P_{\textrm{rx,e,t}}}
\newcommand{\prxevf}{P_{\textrm{rx,e,v}}}

\newcommand{\fstage}{k}
\newcommand{\decfac}{D}
\newcommand{\lfl}{L_{\textrm{fl}}}
\newcommand{\hfl}{h_{\textrm{fl}}}
\newcommand{\hannw}{w}
\newcommand{\hannwnorm}{\tilde{\hannw}}
\newcommand{\nw}{N_{\hannw}}
\newcommand{\hannwpart}{\gamma}
\newcommand{\tslide}{t_w}

\newcommand{\bs}{\sigma_{\textrm{bs}}}
\newcommand{\mysp}{S_p}
\newcommand{\ts}{\textrm{TS}}
\newcommand{\sv}{S_{\textrm{v}}}

\newcommand{\range}{r}
\newcommand{\rangeref}{r_0}
\newcommand{\athw}{\phi}
\newcommand{\along}{\theta}
\newcommand{\gain}{g}
\newcommand{\gainzero}{g_0}
\newcommand{\eqang}{\psi}

\newcommand{\rtarget}{r_{\textrm{target}}}
\newcommand{\alongtarget}{\theta_{\textrm{target}}}
\newcommand{\athwtarget}{\phi_{\textrm{target}}}

\newcommand{\wlen}{\lambda}
\newcommand{\cw}{c}
\newcommand{\absorp}{\alpha}

\newcommand{\dft}{\textrm{DFT}}
\newcommand{\ndft}{{N_{\textrm{DFT}}}}
\newcommand{\ndftw}{_{\nw}}
\newcommand{\atan}{\textrm{arctan2}}
\newcommand{\anglefalong}{\gamma_\along}
\newcommand{\anglefathw}{\gamma_\athw}

\newcommand{\sigmabs}{\sigma_{\textrm{bs}}}

\begin{document}

\title[]{Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder}

\author{Lars Nonboe Andersen}
\email{lars.nonboe.andersen@kongsberg.com}
\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}

\author{Dezhang Chu}
\affiliation{Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd. E. Seattle, WA, 98112, USA}

\author{Harald Heimvoll}
\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}

\author{Rolf Korneliussen}
\author{Gavin J. Macaulay}
\author{Egil Ona}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

\begin{abstract}
The use of quantitative broadband echosounders for biological studies and surveys offers considerable advantages over narrowband echosounders. These include improved spectral-based target identification and significantly increased ability to resolve individual targets. Biological studies and surveys typically require accurate measures of backscatter strength and we present here a systematic and comprehensive explanation of how to derive quantitative estimates of target strength and volume backscattering, as a function of frequency from broadband echosounder signals.

\end{abstract}

\maketitle

\section{Introduction}

Management of fisheries resources typically requires knowledge of age and abundance structure over many years. For numerous fish species and stocks, abundance estimates can be obtained by spatially extensive surveys using quantitative echosounder systems, which measure the backscatter from fish \citep{Simmonds2005Fisheries}. An important property of scientific echosounders used for fisheries surveys is accurate measurement of the backscatter amplitude - this is of lesser importance for other echosounder applications, such as bathymetry, presence/absence of objects, and spatial properties of objects. Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most) and when several are operated simultaneously at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help to categorize the backscatter into species or target categories. This is termed the multi-frequency approach \citep{Korneliussen2002operational, holliday1980}.

The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to fisheries applications is an obvious enhancement. Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}. These benefits can lead to improved backscatter categorisation \citep{Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic, korneliussen2018} and hence more accurate abundance estimates.

There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}. None of these systems have achieved widespread use in quantitative fisheries surveys presumably because the commercially-available systems have lacked one or more of the features that are very useful for an operational quantitative fisheries survey echosounder. These comprise (a) a large dynamic range receiver that does not saturate in normal conditions; (b) operating frequencies that are useful for detecting fish at several hundred metres range (between approximately 20 and 200 kHz); (c) split-aperture for detection of angle of arrival of echoes from individual organisms and calibration spheres; (d) capability to perform in-situ calibrations of the amplitude response; (e) a relatively easy transition from existing survey echosounders; (f) useful at typical ship survey speeds (e.g., 5 m s \textsuperscript{-1}). A recent addition to the set of commercially-available broadband echosounders is the Simrad EK80, which meets all of these features.

The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals. Without loss of generality, we use the \ek\ as an example since it is currently the most commonly used broadband echosounder in the fisheries acoustics field. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in fisheries acoustics.

Our presentation uses nomenclature and approaches that are commonly used for narrowband echosounder systems, which were derived from radar processing \citep{cook1967}. In particular, the expressions for target strength ($\ts$) and volume backscattering strength ($\sv$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.

\section{Signal flow and initial processing}

\subsection{System overview}
A basic echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals. During transmission the program defines the signals which are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water. The acoustic signals propagate through the water, are reflected or scattered by objects in the water, and propagate back to the transducer. During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, and processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig. \ref{fi:ek_sys}). Many types of transmit signals are feasible - this paper considers only the transmission of linear frequency modulated signals (also known as linear chirps).

\begin{figure}
\includegraphics[width=16cm]{Figure1}
\caption{\label{fi:ek_sys}Signal and data flow in the \ek{} system. An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage (lower middle) and display and analysis (lower right). Note that all signals are complex-valued time-series.}
\end{figure}
\subsection{Internal signal processing}

The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, that is converted to an analogue electric signal, $\ytxe(\timesym)$, by the transceiver and applied to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ (Fig. \ref{fi:ek_sys}) where $\samplesymt$ is a sample index in the discrete time domain and $\timesym$ is time for the analog signal. Any returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by corresponding receiver channels, $\channelsym$, in the transceiver. For a split-aperture echosounder system, there are typically three or four channels in the system (a minimum of three are necessary for estimating the angle of arriving echoes). Here, we focus on a four-channel system.

The received electric signal, $\yrxe(\timesym,\channelsym)$, from each channel, $\channelsym$, is pre-amplified, filtered by an analog anti-aliasing filter, and digitized in the transceiver at a frequency of $\fs$, creating the digital signal, $\yrxorg(\samplesymt,\channelsym)$.

To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hfl(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$. The individual filter coefficients for each filter and decimation stage are indexed by $i$. The output signal from each channel, $\channelsym$, from each filter and decimation stage, $\stagesym$, is then given by:
%
\begin{equation}
\label{eq:yrx}
\yrx(\samplesymt,\channelsym,\stagesym) = \left( \yrx(\samplesymt,\channelsym,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\yrx(\samplesymt,\channelsym,0)$ is set to $\yrxorg(\samplesymt,\channelsym)$, being the signal before decimation, $*$ indicates convolution and $\nstages$ is the total number of filter stages. The output signal from the final filter and decimation stage, $\yrx(\samplesymt,\channelsym,\nstages)$, is shortened to $\yrx(\samplesymt,\channelsym)$ for convenience. For the output signal, $\yrx(\samplesymt,\channelsym)$, the decimated sampling rate, $\fsdec$, is given by:
%
\begin{equation}
\label{eq:fsdec}
\fsdec = \fs\prod_{\stagesym=1}^{\nstages} \frac{1}{\decfac(\stagesym)}.
\end{equation}
%
The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data. The filtered and decimated complex samples from each transducer channel, $\yrx(\samplesymt,\channelsym)$, are considered raw data and are recorded together with additional data such as from position and motion sensors and system configuration data in raw data files for display and analysis by processing software.

\subsection{Pulse compression}
To increase signal-to-noise ratio and resolution along the acoustic beam a matched filter may be applied to the raw data samples \citep{turin1960}. This technique is also known as pulse compression \citep{klauder1960}. One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in Eq. \ref{eq:yrx}. The normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, is given by:
%
\begin{equation}
\label{eq:ytxnorm}
\ytxnorm(\samplesymt) = \frac{\ytx(\samplesymt)}{\textrm{max}(\ytx(\samplesymt))}\end{equation}
%
where $\textrm{max}$ is the maximum value of $\ytx(\samplesymt)$. The filtered and decimated output signal, $\ytxnorm(\samplesymt,\stagesym)$, from each filter stage, $\stagesym$, using the normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, as the input signal, is given by:
%
\begin{equation}
\label{eq:FilterStagesTX}
\ytxnorm(\samplesymt,\stagesym) = \left( \ytxnorm(\samplesymt,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
\stagesym = 1,\ldots,\nstages,
\end{equation}
%
where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$ and $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$. The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$.

To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal with the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power. The pulse compressed signal, $\ypc(\samplesymt,\channelsym)$, then becomes
\begin{equation}
\label{eq:PulseComp}
\ypc(\samplesymt,\channelsym) = \frac{\yrx(\samplesymt,\channelsym)*\ymf^*(-\samplesymt)}{||\ymf||^2_2},
\end{equation}
%
where $||\ymf||$ indicates the $l^2$-norm of $\ymf$, also known as the Euclidean norm. The received power samples are then used to estimate target strength and volume backscattering strength. For estimating received power samples, the mean signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, will be used:
%
\begin{equation}
\label{eq:SumSig}
\ypc(\samplesymt) = \frac{1}{\nchannels} \sum_{\channelsym = 1}^{\nchannels} \ypc(\samplesymt,\channelsym).
\end{equation}
%
Compensation of echo strength for position in the acoustic beam requires an estimate of the echo arrival angle. This is obtained using the split-aperture method \citep{burdic1991}, which for broadband pules can be implemented with the angle values contained in the complex-valued $\ypc(\samplesymt)$ data, in combination with knowledge of transducer sector geometry. The principle is demonstrated with a transducer that is divided into four quadrants (Fig. \ref{fi:trd_quad}). In this example the summed signals from four halves (1+2, 2+3, 3+4, 4+1) are calculated as:
\begin{eqnarray}
\label{eq:SumHalves}
y_{\textrm{pc,fore}}(\samplesymt) & = & \frac{1}{2} \left( \ypc(\samplesymt,3)+\ypc(\samplesymt,4) \right),\\
y_{\textrm{pc,aft}}(\samplesymt) & = & \frac{1}{2} \left( \ypc(\samplesymt,1)+\ypc(\samplesymt,2) \right),\\
y_{\textrm{pc,star}}(\samplesymt) & = & \frac{1}{2} \left( \ypc(\samplesymt,1)+\ypc(\samplesymt,4) \right),\\
y_{\textrm{pc,port}}(\samplesymt) & = & \frac{1}{2} \left( \ypc(\samplesymt,2)+\ypc(\samplesymt,3) \right),
\end{eqnarray}
%
where fore, aft, star(board), and port indicate the relevant transducer halves.
\begin{figure}
\includegraphics{Figure2}
\caption{\label{fi:trd_quad}Transducer divided into four quadrants. The labels are directions often used when a transducer is mounted on a ship.}
\end{figure}
%
\subsection{Auto correlation function}
The auto correlation function of the matched filter signal, $\ymfauto(\samplesymt)$, that will be used in a later processing step is defined as:
\begin{equation}
\label{eq:TXAuto}
\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}.
\end{equation}

\subsection{Power and angle samples}
The transceiver measures voltage over a load, $\zrxe$, connected in series with the transducer impedance, $\ztde$. When calculating various acoustic properties a system gain parameter will be used which assumes a matched receiver load. The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig. \ref{fi:impedances}) is given by: 
\begin{equation}
\label{eq:prx}
\prxe(\samplesymt) = \nchannels\left( \frac{|\ypc(\samplesymt)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{\zrxe} \right)^2 \frac{1}{|\ztde|}.
\end{equation}

Forward/aft and port/starboard phase angles of target echoes are estimated by combining the transducer half signals thus: 
%
\begin{eqnarray}
\label{eq:phase1}
y_\along(\samplesymt) & = & y_{\textrm{pc,fore}}(\samplesymt) y_{\textrm{pc,aft}}^*(\samplesymt), \\
y_\athw(\samplesymt) & = & y_{\textrm{pc,star}}(\samplesymt) y_{\textrm{pc,port}}^*(\samplesymt),
\end{eqnarray}
%
where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$. The physical echo arrival angles ($\along$ and $\athw$) are then given by:
%
\begin{eqnarray}
\label{eq:phase2}
\along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right).
\end{eqnarray}
%
where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}. The inverse sine is indicated by $\arcsin$, the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$. As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
%
\begin{figure}
\includegraphics[width=\reprintcolumnwidth]{Figure3}
\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
\end{figure}

\section{Target strength}

Echoes from single targets are often characterised by their $\ts$, which is related to the differential backscattering cross section, $\bs$, via
%
\begin{equation}
\label{eq:TS_bs}
\ts = 10\log_{10}\left(\frac{\bs}{\rangeref^2}\right),
\end{equation}
%
where $\log_{10}$ is the logarithm with base 10 and $\rangeref$ is 1 m.

The power-budget equation (i.e., sonar equation) for a single target \citep[Formulation D, ][]{lunde2016} at frequency $\freqsym$ is:
%
\begin{equation}
\label{eq:TS}
\ts(\freqsym) = 10\log_{10}(\prxetf(\freqsym)) + 40\log_{10}(\range) + 2\absorp(\freqsym) \range 
- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along,\athw,\freqsym)}{16\pi^2} \right),
\end{equation}
%
where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain along the main acoustic axis, incorporating beam compensation based on the estimated target bearing, $(\along,\athw)$.

The point scattering strength, $\mysp(\samplesymt)$, is estimated by applying Eq. \ref{eq:TS} to the received digitized power samples using the on-axis gain value with $f$ set to the centre frequency of the broadband pulse, $\fc$: 
\begin{equation}
\label{eq:Sp}
\mysp(\samplesymt) = 10\log_{10}(\prxe(\samplesymt)) + 40\log_{10}(\range(\samplesymt)) 
+ 2\absorp(\fc) \range(\samplesymt) - 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \gainzero^2(\fc)}{16\pi^2} \right),
\end{equation}
%
noting that $\mysp(\samplesymt)$ is an average over frequency of all echoes from single or multiple targets received at sample $\samplesymt$.

Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated.

From the pulse compressed data, $\ypc(\samplesymt)$, the samples before and after the peak echo level corresponding to an echo from a single target are extracted to obtain the signal, $\ypctarget(\samplesymt)$, noting that the number of samples before the peak can differ from those after the peak. From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$. Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$. If the number of samples around the target is more than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used. If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create a reduced autocorrelation signal, $\ymfautored(\samplesymt)$.

The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
\begin{eqnarray}
\label{eq:DFT_Target_Auto}
\ypctargetf(\samplesymf) & = & \dft_\ndft(\ypctarget(\samplesymt)),\\
\ymfautoredf(\samplesymf) & = & \dft_\ndft(\ymfautored(\samplesymt)),
\end{eqnarray}
where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, is then calculated by: 
%
\begin{equation}
\label{eq:DFT_Target_Auto_Norm}
\ypctargetnormf(\samplesymf) = \frac{\ypctargetf(\samplesymf)} {\ymfautoredf(\samplesymf)}.
\end{equation}
%
Assuming, as a first approximation, that the impedances of the transceiver and transducer are independent of frequency, the received power into a matched load, $\prxetf(\samplesymf)$, is then estimated by:
\begin{equation}
\label{eq:prx_FFT_target}
\prxetf(\samplesymf) = \nchannels\left( \frac{|\ypctargetnormf(\samplesymf)|}{2\sqrt{2}} \right)^2 
\left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \frac{1}{|\ztde|}, % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
noting that any variation of impedance with frequency will be reflected in the $\gainzero$ obtained from the calibration process.

Target strength can then be estimated using Eq. \ref{eq:TS}, but with $\freqsym$ replaced by the discrete index of frequency, $\samplesymf$:
\begin{equation}
\label{eq:TS_f}
\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range) + 2\absorp(\samplesymf)\range 
- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along,\athw,\samplesymf)}{16\pi^2} \right).
\end{equation}

\section{Volume backscattering strength}

Echoes from multiple scatterers can be quantified using volume backscattering strength, $\sv$, being the density of backscattering cross sections, and is given by:
%
\begin{equation}
\label{eq:sv}
\sv  =  10\log_{10}\frac{\sum\bs}{V}.
\end{equation}
%
where $V$ is the volume occupied by the scattering targets. The power-budget equation for multiple targets is then:
%
\begin{equation}
\label{eq:sv_f}
\sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range) + 2\absorp(\freqsym)\range 
- 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym)}{32\pi^2} \right), 
\end{equation}
%
where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, and $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum. Note that $\range$ is the range to the centre of the range volume covered by $\tslide$, and the two-way equivalent beam angle, $\eqang$, is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
\begin{equation}
\label{eq:PsiFc}
\eqang(f) = \eqang(\fn)\left(\frac{\fn}{f}\right)^2.
\end{equation}

Volume backscattering samples compressed over the operational frequency band are estimated by applying Eq. \ref{eq:sv_f} to the received digitized power samples using the on-axis gain value with $f$ set to the centre frequency of the broadband pulse, $\fc$:
\begin{equation}
\label{eq:Sv}
\sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range(\samplesymt)) + 2\absorp(\fc)\range(\samplesymt) 
- 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \cw \teff \eqang(\fc) \gainzero^2(\fc)}{32\pi^2} \right),
\end{equation}
noting that $\sv(\samplesymt)$ is an average over frequency of all echoes received at sample $\samplesymt$. In this case, the time window, $\tslide$, is the effective pulse duration, $\teff$, resulting from pulse compression. The effective pulse duration is defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, and is estimated from digitized data via:
%
\begin{eqnarray}
\label{eq:TauEff}
\ptxauto(\samplesymt) & = & |\ymfauto(\samplesymt)|^2,\\
\teff & = & \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
\end{eqnarray}
where $\ptxauto$ is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$. For an ideal system, i.e., no tapering at the rising and trailing edges of the transmitted signal, the effective pulse duration is the same as the transmit pulse duration.

To estimate $\sv$ as a function of frequency a Fourier transform is used, repeatedly applied via a sliding window in range. However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r^2$, implemented as the $20\log_{10}(\range)$ term in Eq. \ref{eq:Sv}) from the beginning of the window to the end can be significant, particularly for short range measurements. Thus, compensation for spreading loss is performed before applying the discrete Fourier transform. Absorption loss compensation is also range dependent (and frequency dependent), but since absorption loss compensation is insignificant for typical operating frequencies at short ranges and the difference in absorption loss compensation between the beginning and the end of the sliding window is insignificant at longer ranges, compensation for absorption loss is performed after applying the discrete Fourier transform.

Compensation of spherical spreading loss requires compensation of received power by a factor of $r^2$, and hence compensation of amplitude by a factor of $\range$:
%
\begin{equation}
\label{eq:spreadcomp}
\ypcspread(\samplesymt) = \ypc(\samplesymt)\range(\samplesymt).
\end{equation}
%
where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading. A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hanning window, $\hannw(\genidxsym)$. The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution. We suggest that it be at least twice the pulse duration and for computational efficiency reasons should result in a number of samples, $\nw$, which is a power of 2.

The normalised Hanning window, $\hannwnorm$, is given by: 
%
\begin{equation}
\label{eq:hannw}
\hannwnorm(\genidxsym) = \frac{\hannw(\genidxsym)}{\left( \frac{||\hannw||_2}{\sqrt{\nw}} \right)}, i = \frac{-\nw}{2}, \ldots, \frac{\nw}{2}
\end{equation}
%
and the discrete Fourier transform of the windowed data, $\ypcvolumef(\samplesymf)$, is then obtained from:
%
\begin{equation}
\label{eq:FFT_volume}
\ypcvolumef(\samplesymf) = \dft_\ndft 
\left( \hannwnorm(\genidxsym) \left(\ypcspread (\genidxsym+\samplesymt) \left[ u(\genidxsym + \frac{\nw}{2}) - u(\genidxsym - \frac{\nw}{2}) \right] \right) \right),
\end{equation}
%
where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window. The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
%
\begin{equation}
\label{eq:FFT_TX_Auto}
\ymfautof(\samplesymf)  =  \dft_\ndft (\ymfauto(\samplesymt)).
\end{equation}
%
The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
%
\begin{equation}
\label{eq:FFT_volume_norm}
\ypcvolumenormf(\samplesymf) = \frac{\ypcvolumef(\samplesymf)}{\ymfautof(\samplesymf)},
\end{equation}
%
and received power into a matched load, $\prxevf(\samplesymf)$, is estimated from:
%
\begin{equation}
\label{eq:prx_FFT_volume}
\prxevf(\samplesymf) = \nchannels \left( \frac{|\ypcvolumenormf(\samplesymf)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \\
\frac{1}{|\ztde|}. % All channels (*4), matched load (/2), and effective values (/sqrt(2))
\end{equation}
%
Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
%
\begin{equation}
\label{eq:Sv_FFT}
\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range \\
- 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
\end{equation}

\section{Illustrative examples}

A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}. This is visible in the $\ts$ (Fig. \ref{fi:ts_example}a, \ref{fi:ts_example}b) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}. The amplitude of the backscatter signal also clearly shows these nulls (Fig. \ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies. The marked increase in range resolution is apparent once pulse compression has been applied (Fig. \ref{fi:ts_example}d), as are the temporal effects of the pulse compression operation.

A metallic sphere is a rather simple and ideal scatterer and we also present $\sv$ from a school of Atlantic mackerel (\textit{Scomber scombrus}) (Fig. \ref{fi:sv_example}a). The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example. In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig. \ref{fi:sv_example}b, \ref{fi:sv_example}c), although in regions where the fish density decreases (e.g., top left of the school), single target echoes become visible.

\begin{figure}
\includegraphics[width=16cm]{Figure4}
\caption{\label{fi:ts_example}(color online) Target strength of 57.2 mm (a) and 22.0 mm (b) diameter tungsten carbide (with 6\% cobalt binder) calibration spheres, theoretical (solid line) and measured and derived using Eq. (\ref{eq:TS_f}) (dotted line). Echogram of non pulse compressed backscatter [dB re 1 W] (c) and pulse compressed backscatter [dB re 1 $\textrm{m}^2$] (d) from 57.2 mm (lower band) and 22.0 mm (upper band) diameter tungsten carbide spheres using a 2.048 ms duration linear frequency modulated pulse (160--260 kHz). The range from the spheres to the transducer varied during the time period shown.}
\end{figure}

\begin{figure}
\includegraphics[width=16cm]{Figure5}
\caption{\label{fi:sv_example}(color online) (a): Mean frequency response (solid line) and standard error (dashed line) from a school of Atlantic mackerel (\textit{Scomber scombrus}) from three simultaneously operating linear frequency modulated echosounder channels (90--170, 160--260, and 280--450 kHz). Echogram of backscattered power [dB re 1 W] (b) and pulse compressed $\sv$ [dB re 1 $m^{-1}$] (c) from the 160--260 kHz pulse. The school is the trapezoid-shaped object between 25 and 90 m below the surface. The echo from the seafloor is visible at 130 m depth.} \end{figure}

\section{Discussion}
Obtaining quantitative broadband data is more complicated than for narrowband echosounders, due in part to the need to account for frequency dependence in most variables and parameters, and the need for an increased understanding of digital signal processing techniques. For example, the \ek{} currently provides complex demodulated \citep{hasan1983} voltages for each transducer sector rather than derived quantities (e.g., the envelope of a pulse compressed signal and the frequency response of single targets and volumes) - these outputs must instead be calculated as proposed in this paper. This was an implementation decision made to allow for more flexible use of the echosounder data, and to ease the development of new processing methodologies. This decision has several disadvantages, such as the significantly increased data quantity and markedly higher amount of computation required to simply display an echogram. These can be ameliorated to some degree by processing the data into a more directly useful form before storage. The advantages of having access to unprocessed data were considered to out-weigh these disadvantages due to the potential benefits of more sophisticated uses of the acoustic data, especially for a tool that is only beginning to be applied to the field of fisheries acosutics.

The methodology to process broadband data has been presented in a general form in the previous sections without any accord to engineering limits. However, any physical implementation introduces operational constraints. As an example, a system designed for shipboard installations with ample access to electrical and computer processing power and large data storage will typically have different operational contraints compared to a system intended for autonomous platforms where electrical power and computing resources can be severely limited. In addition, the transducer is a significant constraint on the operational parameters of an echosounder and is usually the main determinant of the usable transceiver operating parameters (such as transmit bandwidth, maximum transmit power, pulse duration, ping rate, etc).

The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years. For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}. The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals. Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013, lavery2017}. We note that these methods do not provide an operational method to estimate $\teff$ or $\eqang(\freqsym)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.

The processing equations and methodology presented in this paper have been implemented in version 1.12.4 and earlier of the \ek{} software.

\section{Conclusion}

A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented, with reference to the \ek{} echosounder.

\section{Data Availability Statement}

The data associated with this article are available on request from the authors.

%\bibliography{ek80refs}

\begin{thebibliography}{39}
\def\enquote#1{``#1,''}
\def\plainquote#1{``#1''}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\providecommand{\dourl}[1]{\href{http://#1}{\nolinkurl{#1}}}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\noopsort}[1]{}
\providecommand{\switchargs}[2]{#2#1}
  \def\eatspace #1{#1}
\providecommand{\dodoi}[1]{doi: \href{http://dx.doi.org/#1}{\nolinkurl{#1}}}

\bibitem[{Barr \emph{et~al.}(2002)Barr, Coombs, Doonan, and
  McMillian}]{Barr2002Target}
\bibinfo{author}{Barr, R.}, \bibinfo{author}{Coombs, R.},
  \bibinfo{author}{Doonan, I.},  and \bibinfo{author}{McMillian, P.}
  (\textbf{\bibinfo{year}{2002}}). \enquote{\bibinfo{title}{Target
  identification of oreos and associated species}} \bibinfo{type}{Final
  {{Research Report}} for the {{Ministry}} of {{Fisheries Research Project
  OEO2000}}/{{01B}}, {{Objective}} 1} ,
  \dourl{fs.fish.govt.nz/Page.aspx?pk=113&dk=22653}, \bibinfo{note}{(Last
  viewed October 21, 2019)}.

\bibitem[{Bassett \emph{et~al.}(2018)Bassett, De~Robertis, and
  Wilson}]{bassettBroadbandEchosounderMeasurements2018}
\bibinfo{author}{Bassett, C.}, \bibinfo{author}{De~Robertis, A.},  and
  \bibinfo{author}{Wilson, C.~D.} (\textbf{\bibinfo{year}{2018}}).
  \enquote{\bibinfo{title}{Broadband echosounder measurements of the frequency
  response of fishes and euphausiids in the {{Gulf}} of {{Alaska}}}}
  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{75}(3),
  \bibinfo{pages}{1131--1142}, \dodoi{10.1093/icesjms/fsx204}.

\bibitem[{Briseño-Avena \emph{et~al.}(2015)Briseño-Avena, Roberts, Franks,
  and Jaffe}]{Briseno-Avena2015ZOOPS}
\bibinfo{author}{Briseño-Avena, C.}, \bibinfo{author}{Roberts, P. L.~D.},
  \bibinfo{author}{Franks, P. J.~S.},  and \bibinfo{author}{Jaffe, J.~S.}
  (\textbf{\bibinfo{year}{2015}}).
  \enquote{\bibinfo{title}{{ZOOPS-O\textsuperscript{2}: A broadband echosounder
  with coordinated stereo optical imaging for observing plankton in situ}}}
  \bibinfo{journal}{Methods in Oceanography} \textbf{12},
  \bibinfo{pages}{36--54}, \dodoi{10.1016/j.mio.2015.07.001}.

\bibitem[{Burdic(1991)}]{burdic1991}
\bibinfo{author}{Burdic, W.~S.} (\textbf{\bibinfo{year}{1991}}).
  \emph{\bibinfo{title}{Underwater Acoustic System Analysis}}
  (\bibinfo{publisher}{{Prentice Hall}}).

\bibitem[{Chu and Stanton(1998)}]{Chu1998Application}
\bibinfo{author}{Chu, D.},  and \bibinfo{author}{Stanton, T.}
  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{{Application of
  pulse compression techniques to broadband acoustic scattering by live
  individual zooplankton}}} \bibinfo{journal}{The Journal of the Acoustical
  Society of America} \textbf{104}(1), \bibinfo{pages}{39–55},
  \dodoi{10.1121/1.424056}.

\bibitem[{Chu \emph{et~al.}(1992)Chu, Stanton, and Wiebe}]{chu1992}
\bibinfo{author}{Chu, D.}, \bibinfo{author}{Stanton, T.},  and
  \bibinfo{author}{Wiebe, P.} (\textbf{\bibinfo{year}{1992}}).
  \enquote{\bibinfo{title}{Frequency dependence of sound backscattering from
  live individual zooplankton}} \bibinfo{journal}{ICES Journal of Marine
  Science} \textbf{49}(1), \bibinfo{pages}{97--106},
  \dodoi{10.1093/icesjms/49.1.97}.

\bibitem[{Conti and Demer(2003)}]{Conti2003Wide-bandwidth}
\bibinfo{author}{Conti, S.~G.},  and \bibinfo{author}{Demer, D.~A.}
  (\textbf{\bibinfo{year}{2003}}). \enquote{\bibinfo{title}{{Wide-bandwidth
  acoustical characterization of anchovy and sardine from reverberation
  measurements in an echoic tank}}} \bibinfo{journal}{ICES Journal of Marine
  Science} \textbf{60}(3), \bibinfo{pages}{617--624},
  \dodoi{10.1016/S1054-3139(03)00056-0}.

\bibitem[{Cook and Bernfield(1967)}]{cook1967}
\bibinfo{author}{Cook, C.},  and \bibinfo{author}{Bernfield, M.}
  (\textbf{\bibinfo{year}{1967}}). \emph{\bibinfo{title}{Radar {{Signals}}:
  {{An Introduction}} to {{Theory}} and {{Application}}}}
  (\bibinfo{publisher}{{Academic Press, Inc.}}).

\bibitem[{Demer \emph{et~al.}(2015)Demer, Berger, Bernasconi, Boswell, Chu,
  Domokos, Dunford, F\"assler, Gauthier, Hufnagle, Jech, Bouffant,
  {Lebourges-Dhaussy}, Lurton, Macaulay, Perrot, Ryan, {Parker-Stetter},
  Stienessen, Weber, and Williamson}]{demerCalibrationAcousticInstruments2015}
\bibinfo{author}{Demer, D.~A.}, \bibinfo{author}{Berger, L.},
  \bibinfo{author}{Bernasconi, M.}, \bibinfo{author}{Boswell, K.~M.},
  \bibinfo{author}{Chu, D.}, \bibinfo{author}{Domokos, R.},
  \bibinfo{author}{Dunford, A.~J.}, \bibinfo{author}{F\"assler, S. M.~M.},
  \bibinfo{author}{Gauthier, S.}, \bibinfo{author}{Hufnagle, L.~T.},
  \bibinfo{author}{Jech, J.~M.}, \bibinfo{author}{Bouffant, N.},
  \bibinfo{author}{{Lebourges-Dhaussy}, A.}, \bibinfo{author}{Lurton, X.},
  \bibinfo{author}{Macaulay, G.~J.}, \bibinfo{author}{Perrot, Y.},
  \bibinfo{author}{Ryan, T.~E.}, \bibinfo{author}{{Parker-Stetter}, S.},
  \bibinfo{author}{Stienessen, S.}, \bibinfo{author}{Weber, T.~C.},  and
  \bibinfo{author}{Williamson, N.~J.} (\textbf{\bibinfo{year}{2015}}).
  \enquote{\bibinfo{title}{Calibration of acoustic instruments}}
  \bibinfo{type}{{{ICES Cooperative Research Report No}}. 326} ,
  \dodoi{10.17895/ices.pub.5494}.

\bibitem[{Denny and Simpson(1998)}]{dennyBroadbandAcousticFish1998}
\bibinfo{author}{Denny, G.},  and \bibinfo{author}{Simpson, P.}
  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{A broadband
  acoustic fish identification system}} \bibinfo{journal}{The Journal of the
  Acoustical Society of America} \textbf{103}(5), \bibinfo{pages}{3069--3069},
  \dodoi{10.1121/1.422851}.

\bibitem[{Ehrenberg(1979)}]{ehrenberg1979}
\bibinfo{author}{Ehrenberg, J.~E.} (\textbf{\bibinfo{year}{1979}}).
  \enquote{\bibinfo{title}{A comparative analysis of {\emph{in situ}} methods
  for directly measuring the acoustic target strength of individual fish}}
  \bibinfo{journal}{IEEE Journal of Oceanic Engineering} \textbf{4}(4),
  \bibinfo{pages}{141--152}, \dodoi{10.1109/JOE.1979.1145434}.

\bibitem[{Ehrenberg and Torkelson(2000)}]{ehrenbergFMSlideChirp2000}
\bibinfo{author}{Ehrenberg, J.~E.},  and \bibinfo{author}{Torkelson, T.~C.}
  (\textbf{\bibinfo{year}{2000}}). \enquote{\bibinfo{title}{{{FM}} slide
  (chirp) signals: A technique for significantly improving the signal-to-noise
  performance in hydroacoustics assessment systems}}
  \bibinfo{journal}{Fisheries Research} \textbf{47}, \bibinfo{pages}{193--199},
  \dodoi{10.1016/S0165-7836(00)00169-7}.

\bibitem[{Foote \emph{et~al.}(2005)Foote, Atkins, Francis, and
  Knutsen}]{Foote2005Measuring}
\bibinfo{author}{Foote, K.~G.}, \bibinfo{author}{Atkins, P.~R.},
  \bibinfo{author}{Francis, D. T.~I.},  and \bibinfo{author}{Knutsen, T.}
  (\textbf{\bibinfo{year}{2005}}). \enquote{\bibinfo{title}{{Measuring echo
  spectra of marine organisms over a wide bandwidth}}}
  \bibinfo{address}{Proceedings of the International Conference on Underwater Acoustic Measurements: Technologies and Results, Heraklion, Crete}.

\bibitem[{Forland \emph{et~al.}(2014)Forland, Hobæk, and
  Korneliussen}]{Forland2014Scattering}
\bibinfo{author}{Forland, T.~N.}, \bibinfo{author}{Hobæk, H.},  and
  \bibinfo{author}{Korneliussen, R.~J.} (\textbf{\bibinfo{year}{2014}}).
  \enquote{\bibinfo{title}{{Scattering properties of Atlantic mackerel over a
  wide frequency range}}} \bibinfo{journal}{ICES Journal of Marine Science}
  \textbf{71}(7), \bibinfo{pages}{1904--1912}, \dodoi{10.1093/icesjms/fsu045}.

\bibitem[{Gordon and Zedel(1998)}]{Gordon1998FishMASS}
\bibinfo{author}{Gordon, L.},  and \bibinfo{author}{Zedel, L.}
  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{{FishMASS: what can
  you do with a little bandwidth when you are watching fish?}}}
  \bibinfo{journal}{IEEE Seminar Digests} \textbf{1998}(227),
  \bibinfo{pages}{7}, \dodoi{10.1049/ic:19980186}.

\bibitem[{Hasan(1983)}]{hasan1983}
\bibinfo{author}{Hasan, T.} (\textbf{\bibinfo{year}{1983}}).
  \enquote{\bibinfo{title}{Complex demodulation: {{Some}} theory and
  applications}} in \emph{\bibinfo{booktitle}{Handbook of {{Statistics}}}},
  \bibinfo{volume}{\textbf{3}} (\bibinfo{publisher}{{Elsevier}}), pp.
  \bibinfo{pages}{125--156}, \dodoi{10.1016/S0169-7161(83)03009-6}.

\bibitem[{Hob\ae{}k and
  Forland(2013)}]{hobaekCharacterizationTargetSpheres2013}
\bibinfo{author}{Hob\ae{}k, H.},  and \bibinfo{author}{Forland, T.~N.}
  (\textbf{\bibinfo{year}{2013}}). \enquote{\bibinfo{title}{Characterization of
  {{Target Spheres}} for {{Broad}}-{{Band Calibration}} of {{Acoustic
  Systems}}}} \bibinfo{journal}{Acta Acustica united with Acustica}
  \textbf{99}(3), \bibinfo{pages}{465--476}, \dodoi{10.3813/AAA.918627}.

\bibitem[{Holliday and Pieper(1980)}]{holliday1980}
\bibinfo{author}{Holliday, D.~V.},  and \bibinfo{author}{Pieper, R.~E.}
  (\textbf{\bibinfo{year}{1980}}). \enquote{\bibinfo{title}{Volume scattering
  strengths and zooplankton distributions at acoustic frequencies between 0.5
  and 3 {{MHz}}}} \bibinfo{journal}{The Journal of the Acoustical Society of
  America} \textbf{67}(1), \bibinfo{pages}{135--146}, \dodoi{10.1121/1.384472}.

\bibitem[{Imaizumi \emph{et~al.}(2009)Imaizumi, Abe, Sawada, Matsuda, Akamatsu,
  Suga, Wang, Nishimori, Ogawa, Matsuo, and Ito}]{Imaizumi2009Detection}
\bibinfo{author}{Imaizumi, T.}, \bibinfo{author}{Abe, K.},
  \bibinfo{author}{Sawada, K.}, \bibinfo{author}{Matsuda, A.},
  \bibinfo{author}{Akamatsu, T.}, \bibinfo{author}{Suga, T.},
  \bibinfo{author}{Wang, Y.}, \bibinfo{author}{Nishimori, Y.},
  \bibinfo{author}{Ogawa, S.}, \bibinfo{author}{Matsuo, I.},  and
  \bibinfo{author}{Ito, M.} (\textbf{\bibinfo{year}{2009}}).
  \enquote{\bibinfo{title}{{Detection of in situ fish using broadband
  split-beam system}}} \bibinfo{journal}{Institute of Electronics, Information,
  and Communication Engineers, Tech. Rep. IEICE-US2009-36} \textbf{109}(180),
  \bibinfo{pages}{39--42}.

\bibitem[{Klauder \emph{et~al.}(1960)Klauder, Price, Darlington, and
  Albersheim}]{klauder1960}
\bibinfo{author}{Klauder, J.~R.}, \bibinfo{author}{Price, A.~C.},
  \bibinfo{author}{Darlington, S.},  and \bibinfo{author}{Albersheim, W.~J.}
  (\textbf{\bibinfo{year}{1960}}). \enquote{\bibinfo{title}{The theory and
  design of chirp radars}} \bibinfo{journal}{The Bell System Technical Journal}
  \textbf{39}(4), \bibinfo{pages}{745--808},
  \dodoi{10.1002/j.1538-7305.1960.tb03942.x}.

\bibitem[{Korneliussen(2010)}]{korneliussen2010}
\bibinfo{author}{Korneliussen, R.~J.} (\textbf{\bibinfo{year}{2010}}).
  \enquote{\bibinfo{title}{The acoustic idenification of {{Atlantic}}
  mackerel}} \bibinfo{journal}{ICES Journal of Marine Science} \textbf{67}(8),
  \bibinfo{pages}{1749--1758}, \dodoi{10.1093/icesjms/fsq052}.

\bibitem[{Korneliussen \emph{et~al.}(2018)Korneliussen, Berger, Campanlla,
  Dezhang, Demer, De~Robertis, Domokos, Doray, Fielding, F\"assler, Gauthier,
  Gastauer, Horne, Hutton, Iriarte, Jech, Kloser, Lawson, {Lebourges-Dhaussy},
  McQuinn, Pe\~na, Scoulding, Sakinan, Schaber, Taylor, and
  Thompson}]{korneliussen2018}
\bibinfo{author}{Korneliussen, R.~J.}, \bibinfo{author}{Berger, L.},
  \bibinfo{author}{Campanlla, F.}, \bibinfo{author}{Dezhang, C.},
  \bibinfo{author}{Demer, D.}, \bibinfo{author}{De~Robertis, A.},
  \bibinfo{author}{Domokos, R.}, \bibinfo{author}{Doray, M.},
  \bibinfo{author}{Fielding, S.}, \bibinfo{author}{F\"assler, S. M.~M.},
  \bibinfo{author}{Gauthier, S.}, \bibinfo{author}{Gastauer, S.},
  \bibinfo{author}{Horne, J.}, \bibinfo{author}{Hutton, B.},
  \bibinfo{author}{Iriarte, F.}, \bibinfo{author}{Jech, J.~M.},
  \bibinfo{author}{Kloser, R.}, \bibinfo{author}{Lawson, G.},
  \bibinfo{author}{{Lebourges-Dhaussy}, A.}, \bibinfo{author}{McQuinn, I.},
  \bibinfo{author}{Pe\~na, M.}, \bibinfo{author}{Scoulding, B.},
  \bibinfo{author}{Sakinan, S.}, \bibinfo{author}{Schaber, M.},
  \bibinfo{author}{Taylor, J.~C.},  and \bibinfo{author}{Thompson, C.~H.}
  (\textbf{\bibinfo{year}{2018}}). \enquote{\bibinfo{title}{Target
  classification}} \bibinfo{type}{ICES Cooperative Research Report}
  \bibinfo{number}{No. 344} , \dodoi{doi.org/10.17895/ices.pub.4567}.

\bibitem[{Korneliussen \emph{et~al.}(2016)Korneliussen, Heggelund, Macaulay,
  Patel, Johnsen, and Eliassen}]{Korneliussen2016Acoustic}
\bibinfo{author}{Korneliussen, R.~J.}, \bibinfo{author}{Heggelund, Y.},
  \bibinfo{author}{Macaulay, G.~J.}, \bibinfo{author}{Patel, D.},
  \bibinfo{author}{Johnsen, E.},  and \bibinfo{author}{Eliassen, I.~K.}
  (\textbf{\bibinfo{year}{2016}}). \enquote{\bibinfo{title}{{Acoustic
  identification of marine species using a feature library}}}
  \bibinfo{journal}{Methods in Oceanography} \textbf{17},
  \bibinfo{pages}{187--205}, \dodoi{10.1016/j.mio.2016.09.002}.

\bibitem[{Korneliussen and Ona(2002)}]{Korneliussen2002operational}
\bibinfo{author}{Korneliussen, R.~J.},  and \bibinfo{author}{Ona, E.}
  (\textbf{\bibinfo{year}{2002}}). \enquote{\bibinfo{title}{{An operational
  system for processing and visualizing multi-frequency acoustic data}}}
  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{59}(2),
  \bibinfo{pages}{293--313}, \dodoi{10.1006/jmsc.2001.1168}.

\bibitem[{Lavery \emph{et~al.}(2017)Lavery, Bassett, Lawson, and Jech}]{lavery2017}
\bibinfo{author}{Lavery, A.~C.},  \bibinfo{author}{Bassett, C.},
  \bibinfo{author}{Lawson, G.~L.}, and \bibinfo{author}{Jech. J.~M.}
  (\textbf{\bibinfo{year}{2017}}). \enquote{\bibinfo{title}{{Exploiting signal 
  processing approaches for broadband echosounders}}} \bibinfo{journal}{ICES Journal
  of Marine Science} \textbf{74}(8), \bibinfo{pages}{2262--2275},
  \dodoi{10.1093/icesjms/fsx155}.

\bibitem[{Lunde and Korneliussen(2016)}]{lunde2016}
\bibinfo{author}{Lunde, P.},  and \bibinfo{author}{Korneliussen, R.~J.}
  (\textbf{\bibinfo{year}{2016}}). \enquote{\bibinfo{title}{Power-{{Budget
  Equations}} and {{Calibration Factors}} for {{Fish Abundance Estimation Using
  Scientific Echo Sounder}} and {{Sonar Systems}}}} \bibinfo{journal}{Journal
  of Marine Science and Engineering} \textbf{4}(3), \bibinfo{pages}{43},
  \dodoi{10.3390/jmse4030043}.

\bibitem[{MacLennan(1981)}]{maclennan1981}
\bibinfo{author}{MacLennan, D.~N.} (\textbf{\bibinfo{year}{1981}}).
  \enquote{\bibinfo{title}{The {{Theory}} of {{Solid Spheres}} as {{Sonar
  Calibration Targets}}}} \bibinfo{type}{Scottish {{Fisheries Research Report
  Number}} 22} .

\bibitem[{MacLennan \emph{et~al.}(2002)MacLennan, Fernandes, and
  Dalen}]{MacLennan2002consistent}
\bibinfo{author}{MacLennan, D.~N.}, \bibinfo{author}{Fernandes, P.},  and
  \bibinfo{author}{Dalen, J.} (\textbf{\bibinfo{year}{2002}}).
  \enquote{\bibinfo{title}{{A consistent approach to definitions and symbols in
  fisheries acoustics}}} \bibinfo{journal}{ICES Journal of Marine Science}
  \textbf{59}, \bibinfo{pages}{365--369}, \dodoi{10.1006/jmsc.2001.1158}.

\bibitem[{Simmonds \emph{et~al.}(1996)Simmonds, Armstrong, and
  Copland}]{Simmonds1996Species}
\bibinfo{author}{Simmonds, E.~J.}, \bibinfo{author}{Armstrong, F.},  and
  \bibinfo{author}{Copland, P.~J.} (\textbf{\bibinfo{year}{1996}}).
  \enquote{\bibinfo{title}{{Species identification using wideband backscatter
  with neural network and discriminant analysis}}} \bibinfo{journal}{ICES
  Journal of Marine Science} \textbf{53}(2), \bibinfo{pages}{189–195},
  \dodoi{10.1006/jmsc.1996.0021}.

\bibitem[{Simmonds and MacLennan(2005)}]{Simmonds2005Fisheries}
\bibinfo{author}{Simmonds, J.},  and \bibinfo{author}{MacLennan, D.}
  (\textbf{\bibinfo{year}{2005}}). \emph{\bibinfo{title}{{Fisheries Acoustics.
  Theory and Practice}}}, \bibinfo{edition}{2nd ed.}
  (\bibinfo{publisher}{Blackwell Science}, \bibinfo{address}{Oxford}).

\bibitem[{Skaret \emph{et~al.}(2020)Skaret, Johansen, Johnsen, Fall, Fiksen,
  Englund, Fauchald, Gj{\o}s{\ae}ter, Macaulay, and Johannesen}]{skaret2020}
\bibinfo{author}{Skaret, G.}, \bibinfo{author}{Johansen, G.~O.},
  \bibinfo{author}{Johnsen, E.}, \bibinfo{author}{Fall, J.},
  \bibinfo{author}{Fiksen, {\O}.}, \bibinfo{author}{Englund, G.},
  \bibinfo{author}{Fauchald, P.}, \bibinfo{author}{Gj{\o}s{\ae}ter, H.},
  \bibinfo{author}{Macaulay, G.~J.},  and \bibinfo{author}{Johannesen, E.}
  (\textbf{\bibinfo{year}{2020}}). \enquote{\bibinfo{title}{Diel vertical
  movements determine spatial interactions between cod, pelagic fish and krill
  on an {{Arctic}} shelf bank}} \bibinfo{journal}{Marine Ecology Progress
  Series} \textbf{638}, \bibinfo{pages}{13--23}, \dodoi{10.3354/meps13254}.

\bibitem[{Stanton and Chu(2008)}]{stanton2008}
\bibinfo{author}{Stanton, T.~K.},  and \bibinfo{author}{Chu, D.}
  (\textbf{\bibinfo{year}{2008}}). \enquote{\bibinfo{title}{Calibration of
  broadband active acoustic systems using a single standard spherical target}}
  \bibinfo{journal}{The Journal of the Acoustical Society of America}
  \textbf{124}(1), \bibinfo{pages}{128--136}, \dodoi{10.1121/1.2917387}.

\bibitem[{Stanton \emph{et~al.}(2010)Stanton, Chu, Jech, and
  Irish}]{Stanton2010New}
\bibinfo{author}{Stanton, T.~K.}, \bibinfo{author}{Chu, D.},
  \bibinfo{author}{Jech, J.~M.},  and \bibinfo{author}{Irish, J.~D.}
  (\textbf{\bibinfo{year}{2010}}). \enquote{\bibinfo{title}{{New broadband
  methods for resonance classification and high-resolution imagery of fish with
  swimbladders using a modified commercial broadband echosounder}}}
  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{67}(2),
  \bibinfo{pages}{365--378}, \dodoi{10.1093/icesjms/fsp262}.

\bibitem[{Stanton \emph{et~al.}(2012)Stanton, Sellers, and
  Jech}]{Stanton2012Resonance}
\bibinfo{author}{Stanton, T.~K.}, \bibinfo{author}{Sellers, C.~J.},  and
  \bibinfo{author}{Jech, J.~M.} (\textbf{\bibinfo{year}{2012}}).
  \enquote{\bibinfo{title}{{Resonance classification of mixed assemblages of
  fish with swimbladders using a modified commercial broadband acoustic
  echosounder at 1–6 kHz}}} \bibinfo{journal}{Canadian Journal of Fisheries
  and Aquatic Sciences} \textbf{69}(5), \bibinfo{pages}{854--868},
  \dodoi{10.1139/f2012-013}.

\bibitem[{Traykovski \emph{et~al.}(1998)Traykovski, O'Driscoll, and
  McGhee}]{Traykovski1998Effect}
\bibinfo{author}{Traykovski, L.~M.}, \bibinfo{author}{O'Driscoll, R.},  and
  \bibinfo{author}{McGhee, D.} (\textbf{\bibinfo{year}{1998}}).
  \enquote{\bibinfo{title}{{Effect of orientation on broadband acoustic
  scattering of Antarctic krill \textit{Euphausia superba}: Implications for
  inverting zooplankton spectral acoustic signatures for angle of
  orientation}}} \bibinfo{journal}{The Journal of the Acoustical Society of
  America} \textbf{104}(4), \bibinfo{pages}{2121–2135},
  \dodoi{10.1121/1.423726}.

\bibitem[{Turin(1960)}]{turin1960}
\bibinfo{author}{Turin, G.} (\textbf{\bibinfo{year}{1960}}).
  \enquote{\bibinfo{title}{An introduction to matched filters}}
  \bibinfo{journal}{IRE Transactions on Information Theory} \textbf{6}(3),
  \bibinfo{pages}{311--329}, \dodoi{10.1109/TIT.1960.1057571}.

\bibitem[{Zakharia \emph{et~al.}(1989)Zakharia, Corgiatti, Joly, and
  Person}]{Zakharia1989Wide-band}
\bibinfo{author}{Zakharia, M.}, \bibinfo{author}{Corgiatti, J.},
  \bibinfo{author}{Joly, F.},  and \bibinfo{author}{Person, R.}
  (\textbf{\bibinfo{year}{1989}}). \enquote{\bibinfo{title}{{Wide-band sounder
  for fisheries}}} \bibinfo{journal}{Proceedings of the Institute of Acoustics}
  \textbf{11}, \bibinfo{pages}{274--281}.

\bibitem[{Zakharia \emph{et~al.}(1996)Zakharia, Magand, Hetroit, and
  Diner}]{Zakharia1996Wideband}
\bibinfo{author}{Zakharia, M.~E.}, \bibinfo{author}{Magand, F.},
  \bibinfo{author}{Hetroit, F.},  and \bibinfo{author}{Diner, N.}
  (\textbf{\bibinfo{year}{1996}}). \enquote{\bibinfo{title}{{Wideband sounder
  for fish species identification at sea}}} \bibinfo{journal}{ICES Journal of
  Marine Science} \textbf{53}, \bibinfo{pages}{203–208},
  \dodoi{10.1006/jmsc.1996.0023}.

\bibitem[{Zedel \emph{et~al.}(2003)Zedel, Knutsen, and
  Patro}]{Zedel2003Acoustic}
\bibinfo{author}{Zedel, L.}, \bibinfo{author}{Knutsen, T.},  and
  \bibinfo{author}{Patro, R.} (\textbf{\bibinfo{year}{2003}}).
  \enquote{\bibinfo{title}{{Acoustic Doppler current profiler observations of
  herring movement}}} \bibinfo{journal}{ICES Journal of Marine Science}
  \textbf{60}(4), \bibinfo{pages}{846--859},
  \dodoi{10.1016/s1054-3139(03)00067-5}.

\end{thebibliography}

\end{document}

### 2022-01-27 Manuscript.tex
Updated figure references in Manuscript and figure generation code

@@
 \begin{figure}
-\includegraphics[width=16cm]{Figure1}
+\includegraphics[width=16cm]{Fig_ek_sys}
 \caption{\label{fi:ek_sys}Signal and data flow in the \ek{} system.
@@
 \end{figure}
-\subsection{Internal signal processing}
+\subsection{Signal generation}
 
@@
 Here, we focus on a four-channel system.
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_ytx}
+\caption{\label{fi:ytx} Ideal and enveloped send pulse between 34kHz to 45kHz with a slope of 0.057, as implemented in the EK80. The orange curve illustrate the effect of setting the slope to 0.5.}
+\end{figure}
 
@@
 
+\begin{figure}
+\includegraphics[width=16cm]{Fig_y_mf_n}
+\caption{\label{fi:ytx} The absolute value of the filtered and decimated output signal used in the pulse compression.}
+\end{figure}
+
 To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal with the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power.
@@
 \begin{figure}
-\includegraphics{Figure2}
+\includegraphics{Fig_trd_quad}
 \caption{\label{fi:trd_quad}Transducer divided into four quadrants. The labels are directions often used when a transducer is mounted on a ship.}
@@
 \end{equation}
+
+\begin{figure}
+\includegraphics{Fig_ACF}
+\caption{\label{fi:ACF}.}
+\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=\reprintcolumnwidth]{Figure3}
+\includegraphics[width=\reprintcolumnwidth]{Fig_impedances}
 \caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
@@
 \begin{figure}
-\includegraphics[width=16cm]{Figure4}
+\includegraphics[width=16cm]{Fig_ts_example}
 \caption{\label{fi:ts_example}(color online) Target strength of 57.2 mm (a) and 22.0 mm (b) diameter tungsten carbide (with 6\% cobalt binder) calibration spheres, theoretical (solid line) and measured and derived using Eq.
@@
 \begin{figure}
-\includegraphics[width=16cm]{Figure5}
+%\includegraphics[width=16cm]{Figure5}
 \caption{\label{fi:sv_example}(color online) (a): Mean frequency response (solid line) and standard error (dashed line) from a school of Atlantic mackerel (\textit{Scomber scombrus}) from three simultaneously operating linear frequency modulated echosounder channels (90--170, 160--260, and 280--450 kHz).

### 2022-01-27 main.py

@@ -27,6 +27,7 @@ plt.title(
     .format(f0/1000, f1/1000, slope))
 plt.xlabel('time (ms)')
 plt.ylabel('amplitude')
+plt.savefig('Fig_ytx.png')
 
 # Chapter: Signal reception
 
@@ -48,6 +49,7 @@ plt.plot(np.abs(ekcalc.y_mf_n))
 plt.title('The absolute value of the filtered and decmiated output signal')
 plt.xlabel('samples ()')
 plt.ylabel('amplitude')
+plt.savefig('Fig_y_mf_n.png')
 
 # The autocorrelation function of the matched filter signal
 # ekcalc.y_mf_auto
@@ -57,6 +59,7 @@ plt.plot(np.abs(ekcalc.y_mf_auto))
 plt.title('The autocorrelation function of the filtered and decmiated output signal')
 plt.xlabel('samples')
 plt.ylabel('ACF')
+plt.savefig('Fig_ACF.png')
 
 
 # Calculating the pulse compressed quadrant signals separately on each channel
@@ -79,4 +82,3 @@ p_rx_e = ekcalc.calcPower(y_pc_n)
 y_theta_n, y_phi_n = ekcalc.calcElectricalAngles(y_pc_nu)
 
 
-plt.show()

### 2022-01-27 Manuscript.tex
edited the signal reception chapter

@@
 \end{figure}
+
 \subsection{Signal generation}
 
-The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, that is converted to an analogue electric signal, $\ytxe(\timesym)$, by the transceiver and applied to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ (Fig.
-\ref{fi:ek_sys}) where $\samplesymt$ is a sample index in the discrete time domain and $\timesym$ is time for the analog signal.
-Any returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by corresponding receiver channels, $\channelsym$, in the transceiver.
-For a split-aperture echosounder system, there are typically three or four channels in the system (a minimum of three are necessary for estimating the angle of arriving echoes).
-Here, we focus on a four-channel system.
+The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is a sample index in the discrete time domain.
+Typical BB pulses are linear upsweep pulses windowed by an envelope function.
+In the EK80 a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}.
+This is found in the class \verb|Core.EK80Calculation.generateIdealWindowedSendPulse|, where the parameters 'f0', 'f1', tau, fs and slope are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
+A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
+
+To generate the signal
+
+The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain analogue signal $\ytxe(\timesym)$.
+The analogue and amplified signal is passed on to the transducer to generate the transmitted acoustic signal $\ytxA(\timesym)$ in the water column, where $\timesym$ is the time for the (analog) signal.
+For a split-aperture echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
 
@@
 \end{figure}
+
+\subsection{Signal reception}
+
+ (Fig. \ref{fi:ek_sys})
+
+The returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by corresponding receiver channels, $\channelsym$, in the transceiver.
 

### 2022-01-28 Manuscript.tex
Finalized the signal reception subchapter

@@
 
- (Fig. \ref{fi:ek_sys})
-
 The returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by corresponding receiver channels, $\channelsym$, in the transceiver.
-
 The received electric signal, $\yrxe(\timesym,\channelsym)$, from each channel, $\channelsym$, is pre-amplified, filtered by an analog anti-aliasing filter, and digitized in the transceiver at a frequency of $\fs$, creating the digital signal, $\yrxorg(\samplesymt,\channelsym)$.
@@
 The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data.
-The filtered and decimated complex samples from each transducer channel, $\yrx(\samplesymt,\channelsym)$, are considered raw data and are recorded together with additional data such as from position and motion sensors and system configuration data in raw data files for display and analysis by processing software.
+In the example $N_v=2$.
+The frequency responses of the filters are shown in Figure~\ref{fi:fir}, and the corresponding filter coefficients and decimation factors are found in the test data set.
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_fir}
+\caption{\label{fi:fir} The frequency (power) response of the filters in from our test set.}
+\end{figure}
+
+The original sample data $\yrxorg(\samplesymt,\channelsym)$ are not available in the EK80 data files.
+Instead,the resulting filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
+The data are recorded together with additional data such as from position and motion sensors and system configuration data in raw data files for display and analysis by processing software.
+The filter parameters are in this paper used to generate the transmit pulse for the matched filter, see below.
 
@@
 \label{eq:FilterStagesTX}
-\ytxnorm(\samplesymt,\stagesym) = \left( \ytxnorm(\samplesymt,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
+\ytxnorm(\samplesymt,\stagesym) = \left[ \ytxnorm(\samplesymt,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right]_{\downarrow \decfac(\stagesym)}, 
 \stagesym = 1,\ldots,\nstages,
@@
 where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$ and $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$.
-The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$.
+The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n}).
 
@@
 \includegraphics[width=16cm]{Fig_y_mf_n}
-\caption{\label{fi:ytx} The absolute value of the filtered and decimated output signal used in the pulse compression.}
+\caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal $\ymf(\samplesymt)$ used by the pulse compression.}
 \end{figure}

### 2022-01-28 main.py

@@ -6,11 +6,13 @@ from Core.EK80Calculation import EK80Calculation
 # Test data
 
 # The data file ./data/pyEcholabEK80data.json contain data from one
-# ping from the EK80 echosounder, including the 
+# ping from the EK80 echosounder, including the
 
 ekcalc = EK80Calculation('./data/pyEcholabEK80data.json')
 
+#
 # Chapter: Signal generation
+#
 
 f0 = ekcalc.f0  # frequency (Hz)
 f1 = ekcalc.f1  # frequency (Hz)
@@ -29,13 +31,63 @@ plt.xlabel('time (ms)')
 plt.ylabel('amplitude')
 plt.savefig('Fig_ytx.png')
 
+#
 # Chapter: Signal reception
+#
 
-# The raw signal 'y_rx_org' is stored as complex numbers in a numpy nd array
-# where the first dimension corresponds to one of the four the transducer
-# quadrants adn the second is the sample in time. The complex numbers include
-# the phase of the signal.
-# (or is this the decmiated signal? Unclear.)
+# The raw signal 'y_rx_org' is filtered and decimated to the filter
+# bank and the final signal is stored in 'y_rx_nu'. Note that the
+# raw files from the EK80 implementation does not store the initial
+# or the intermediate steps in the raw file, but the filtered and
+# deciated result, store in the variable 'y_rx_nu'
+
+# The filter coefficients 'h_fl_iv' are accessible through
+# In this implementation there are two filters (N_v=2). The first
+# filter operates on the raw data [y_rx(N,u,0)=y_rx_org].
+ekcalc.fil1s[0].Coefficients
+ekcalc.fil1s[1].Coefficients
+# and the corresponding decimation factor is available through
+ekcalc.fil1s[0].DecimationFactor
+ekcalc.fil1s[1].DecimationFactor
+
+# The frequency response function of the filter is given by its
+# discrete time fourier transform:
+H0 = np.fft.fft(ekcalc.fil1s[0].Coefficients)
+H1 = np.fft.fft(ekcalc.fil1s[1].Coefficients)
+
+# Plot of the frequency response of the filters (power) (in dB)
+# [someone needs to review this, I am not an expert here and I am not
+# sure hwo this works for complex filters and how to deal with the
[…]

### 2022-02-02 Manuscript.tex

@@
 The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n}).
-
+%
 \begin{figure}
@@
 \caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal $\ymf(\samplesymt)$ used by the pulse compression.}
+\end{figure}
+
+The auto correlation function of the matched filter signal, $\ymfauto(\samplesymt)$, that will be used in a later processing step is defined as:
+\begin{equation}
+\label{eq:TXAuto}
+\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}.
+\end{equation}
+
+\begin{figure}
+\includegraphics{Fig_ACF}
+\caption{\label{fi:ACF}.}
 \end{figure}
@@
 %
-\subsection{Auto correlation function}
-The auto correlation function of the matched filter signal, $\ymfauto(\samplesymt)$, that will be used in a later processing step is defined as:
-\begin{equation}
-\label{eq:TXAuto}
-\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}.
-\end{equation}
-
-\begin{figure}
-\includegraphics{Fig_ACF}
-\caption{\label{fi:ACF}.}
-\end{figure}
 

### 2022-02-04 Core/EK80CalculationPaper.py
Added (again) all filter stages to the y_tilde_tx_nv variable. Use the data variable instead of the class EK80CalculationPaper.

@@ -68,9 +68,9 @@ class EK80CalculationPaper(EK80DataContainer):
                 #y_tilde_tx_nv = self.stageFilter(y_tilde_tx_nv, fil1)
                 self.f_s_dec *= 1 / fil1.DecimationFactor
         """
-        # Output signal from the final filter and decimation stage is used
+        # Output signal from the final filter and decimation stage [-1] is used
         # as matched filter
-        y_mf_n = y_tilde_tx_nv
+        y_mf_n = y_tilde_tx_nv[-1]
         self.y_mf_n = y_mf_n
 
         # Create complex conjugated time reversed version and 2-norm of
@@ -97,6 +97,21 @@ class EK80CalculationPaper(EK80DataContainer):
             # decimation
 
             self.frequencies = np.linspace(self.f0, self.f1, self.n_f_points)
+            
+    @staticmethod
+    def calcAutoCorrelation(y_mf_n, f_s_dec):
+        y_mf_n_conj_rev = np.conj(y_mf_n)[::-1]
+        y_mf_twoNormSquared = np.linalg.norm(y_mf_n, 2) ** 2
+        y_mf_n_conj_rev = y_mf_n_conj_rev
+        y_mf_twoNormSquared = y_mf_twoNormSquared
+
+        # Calculate auto correlation function for matched filter
+        y_mf_auto_n = np.convolve(y_mf_n, y_mf_n_conj_rev)/y_mf_twoNormSquared
+
+        # Estimate effective pulse duration
+        p_tx_auto = np.abs(y_mf_auto_n) ** 2
+        tau_eff = np.sum(p_tx_auto) / ((np.max(p_tx_auto)) * f_s_dec)
+        return y_mf_auto_n, tau_eff
 
     @staticmethod
     def calc_fs_dec(filter_v, f_s):
@@ -113,16 +128,16 @@ class EK80CalculationPaper(EK80DataContainer):
 
     @staticmethod
     def calc_y_tx_tilde_nv(y_tilde_tx_n, filter_v):
-
-        y_tilde_tx_nv = y_tilde_tx_n
-
-        # import pdb; pdb.set_trace()
-
+        # Initialize with normalized transmit pulse
+        y_tilde_tx_nv = [y_tilde_tx_n]
+        v = 0
         if filter_v is not None:
-            for filter_v in filter_v:
-                y_tilde_tx_nv = np.convolve(y_tilde_tx_nv,
-                                            filter_v["h_fl_i"],
[…]

### 2022-02-04 Core/EK80DataContainer.py

@@ -30,7 +30,7 @@ class EK80DataContainer:
         self.f0 = xml0['Parameter']['FrequencyStart']
         self.fnom = xml0['Transducer']['Frequency']  # nominal design frequency for the transducer
         self.f1 = xml0['Parameter']['FrequencyEnd']
-
+        self.f_c = (self.f0 + self.f1) / 2.0
         self.tau = xml0['Parameter']['PulseDuration']
         self.slope = xml0['Parameter']['Slope']
         self.sampleInterval = xml0['Parameter']['SampleInterval']

### 2022-02-04 main.py

@@ -1,12 +1,14 @@
 import matplotlib.pyplot as plt
 import numpy as np
 from Core.EK80CalculationPaper import EK80CalculationPaper
+from Core.EK80DataContainer import EK80DataContainer
 
 # Test data
 
 # The data file ./data/pyEcholabEK80data.json contain data from one
 # ping from the EK80 echosounder, including the
-ekcalc = EK80CalculationPaper('./data/pyEcholabEK80data.json')
+#ekcalc = EK80CalculationPaper('./data/pyEcholabEK80data.json')
+ekcalc = EK80DataContainer('./data/pyEcholabEK80data.json')
 
 #
 # Chapter: Signal generation
@@ -60,6 +62,8 @@ f_s = ekcalc.f_s  # The initial sampling frequency
 f_s0 = f_s / ekcalc.filter_v[0]["D"]
 f_s1 = f_s / (
     ekcalc.filter_v[0]["D"]*ekcalc.filter_v[1]["D"])
+# Comment: This should be returned by the filter_v functions:
+f_s_dec = f_s1
 
 # The frequency response function of the filter is given by its
 # discrete time fourier transform:
@@ -99,7 +103,8 @@ y_tilde_tx_n = EK80CalculationPaper.calc_y_tx_tilde_n(y_tx_n)
 
 # Comment from Nils Olav: I added all the steps to a list, but the indexing
 # is different from the notation, i.e. y_tilde_tx_nv[v][i], wheras the paper
-# uses y_tilde_tx_nv[i,v].
+# uses y_tilde_tx_nv[i,v]. I think it is ok, though.
+# Add decimation sampling rates to this function?
 y_tilde_tx_nv = EK80CalculationPaper.calc_y_tx_tilde_nv(
     y_tilde_tx_n, ekcalc.filter_v)
 
@@ -108,17 +113,19 @@ y_tilde_tx_nv = EK80CalculationPaper.calc_y_tx_tilde_nv(
 y_mf_n = y_tilde_tx_nv[-1]
 
 plt.figure()
-plt.plot(np.abs(ekcalc.y_mf_n))
+plt.plot(np.abs(y_mf_n))
 plt.title('The absolute value of the filtered and decimated output signal')
 plt.xlabel('samples ()')
 plt.ylabel('amplitude')
 plt.savefig('./Paper/Fig_y_mf_n.png')
 
[…]

### 2022-02-04 Paper/Manuscript.tex
Moved the tau eff description

--- 2022-02-02 Manuscript.tex
@@
 The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain analogue signal $\ytxe(\timesym)$.
-The analogue and amplified signal is passed on to the transducer to generate the transmitted acoustic signal $\ytxA(\timesym)$ in the water column, where $\timesym$ is the time for the (analog) signal.
+The analogue and amplified signal is passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water column, where $\timesym$ is the time for the (analog) signal.
 For a split-aperture echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
@@
 
-The auto correlation function of the matched filter signal, $\ymfauto(\samplesymt)$, that will be used in a later processing step is defined as:
+The auto correlation function of the matched filter signal and the effective pulse duration, defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, will be used in later processing steps and are defined as
 \begin{equation}
 \label{eq:TXAuto}
-\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}.
+\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}
 \end{equation}
+and
+\begin{equation}
+\label{eq:TauEff}
+\teff = \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
+\end{equation}
+where
+\begin{equation*}
+%\label{eq:TauEff}
+\ptxauto(\samplesymt)  =  |\ymfauto(\samplesymt)|^2
+\end{equation*}
+is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.
+For an ideal system, i.e., no tapering at the rising and trailing edges of the transmitted signal, the effective pulse duration is the same as the transmit pulse duration.
 
@@
 \includegraphics{Fig_ACF}
-\caption{\label{fi:ACF}.}
+\caption{The aoutocorrekation function $\ptxauto$ from the example. The corresponding $\teff=0.1ms$ and $\tnom=2ms$\label{fi:ACF}.}
 \end{figure}
@@
 In this case, the time window, $\tslide$, is the effective pulse duration, $\teff$, resulting from pulse compression.
-The effective pulse duration is defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, and is estimated from digitized data via:
-%
-\begin{eqnarray}
-\label{eq:TauEff}
-\ptxauto(\samplesymt) & = & |\ymfauto(\samplesymt)|^2,\\
-\teff & = & \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
-\end{eqnarray}
-where $\ptxauto$ is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.
-For an ideal system, i.e., no tapering at the rising and trailing edges of the transmitted signal, the effective pulse duration is the same as the transmit pulse duration.
 

### 2022-02-04 main.py
Refactor to conform with naming conv. Renamed to a data object througout. Uses only static methods.

@@ -4,32 +4,23 @@ from Core.EK80CalculationPaper import EK80CalculationPaper
 from Core.EK80DataContainer import EK80DataContainer
 
 # Test data
-
-# The data file ./data/pyEcholabEK80data.json contain data from one
-# ping from the EK80 echosounder, including the
-#ekcalc = EK80CalculationPaper('./data/pyEcholabEK80data.json')
-ekcalc = EK80DataContainer('./data/pyEcholabEK80data.json')
+data = EK80DataContainer('./data/pyEcholabEK80data.json')
 
 #
-# Chapter: Signal generation
+# Chapter IIB: Signal generation
 #
 
 # Generate ideal send pulse
-f0 = ekcalc.f0  # frequency (Hz)
-f1 = ekcalc.f1  # frequency (Hz)
-tau = ekcalc.tau  # time (s)
-fs = ekcalc.f_s  # frequency (Hz)
-slope = ekcalc.slope
 y_tx_n, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
-    f0, f1, tau, fs, slope)
+    data.f0, data.f1, data.tau, data.f_s, data.slope)
 y_tx_n05slope, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
-    f0, f1, tau, fs, .5)
+    data.f0, data.f1, data.tau, data.f_s, .5)
 
 plt.figure()
 plt.plot(t*1000, y_tx_n, t*1000, y_tx_n05slope)
 plt.title(
     'Ideal enveloped send pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
-    .format(f0/1000, f1/1000, slope))
+    .format(data.f0/1000, data.f1/1000, data.slope))
 plt.xlabel('time (ms)')
 plt.ylabel('amplitude')
 plt.savefig('./Paper/Fig_ytx.png')
@@ -42,33 +33,32 @@ plt.savefig('./Paper/Fig_ytx.png')
 # through the filter bank and the final signal is stored in 'y_rx_nu'.
 # Note that the raw files from the EK80 implementation does not store
 # the initial or the intermediate steps in the raw file, but the filtered and
-# decimated result only, The data is stored in the variable 'y_rx_nu'.
+# decimated result only.
 
 # In this implementation there are two filters (N_v=2). The first
[…]

### 2022-02-04 Core/EK80CalculationPaper.py
-fixed f_s_dec function to return the freqs for each step

@@ -47,16 +47,16 @@ class EK80CalculationPaper(EK80DataContainer):
         self.y_tx_n = y_tx_n
         self.y_tx_n_t = t
 
-        self.f_s_dec = EK80CalculationPaper.calc_fs_dec(self.filter_v, self.f_s)
+        self.f_s_dec = EK80CalculationPaper.calcDecmiatedSamplingRate(self.filter_v, self.f_s)
 
         # Normalize ideal transmit pulse
         #y_tilde_tx_n = y_tx_n / np.max(y_tx_n)
 
         # Eq (4)
-        y_tilde_tx_n = EK80CalculationPaper.calc_y_tx_tilde_n(y_tx_n)
+        y_tilde_tx_n = EK80CalculationPaper.calcNormalizedTransmitSignal(y_tx_n)
 
         # Eq (5)
-        y_tilde_tx_nv = EK80CalculationPaper.calc_y_tx_tilde_nv(y_tilde_tx_n, self.filter_v)
+        y_tilde_tx_nv = EK80CalculationPaper.calcFilteredAndDecimatedSignal(y_tilde_tx_n, self.filter_v)
 
         # Filter and decimate ideal transmit pulse trough stage filters
         # and calculate decimated sampling rate
@@ -114,20 +114,22 @@ class EK80CalculationPaper(EK80DataContainer):
         return y_mf_auto_n, tau_eff
 
     @staticmethod
-    def calc_fs_dec(filter_v, f_s):
-        f_s_dec = f_s
+    def calcDecmiatedSamplingRate(filter_v, f_s):
+        f_s_dec = [f_s]
+        v = 0
         if filter_v is not None:
             for filter_v in filter_v:
-                f_s_dec *= 1 / filter_v["D"]
-
+                tmp = f_s_dec[v] / filter_v["D"]
+                f_s_dec.append(tmp)
+                v += 1
         return f_s_dec
-
+    
     @staticmethod
-    def calc_y_tx_tilde_n(y_tx_n):
+    def calcNormalizedTransmitSignal(y_tx_n):
         return y_tx_n / np.max(y_tx_n)
 
     @staticmethod
-    def calc_y_tx_tilde_nv(y_tilde_tx_n, filter_v):
+    def calcFilteredAndDecimatedSignal(y_tilde_tx_n, filter_v):
         # Initialize with normalized transmit pulse
[…]

### 2022-02-04 main.py

@@ -46,14 +46,11 @@ data.filter_v[1]["h_fl_i"]
 data.filter_v[0]["D"]
 data.filter_v[1]["D"]
 
-# Comment: This should be returned by the filter_v functions:
-# and the corresponding samplings frequenies
-f_s = data.f_s  # The initial sampling frequency
-f_s0 = f_s / data.filter_v[0]["D"]
-f_s1 = f_s / (
-    data.filter_v[0]["D"]*data.filter_v[1]["D"])
-# The final sampling freq is stored in f_s_dec
-f_s_dec = f_s1
+# The sampling freq for each filter step
+f_s_dec_v = EK80CalculationPaper.calcDecmiatedSamplingRate(
+    data.filter_v, data.f_s)
+# The final sampling frequency
+f_s_dec = f_s_dec_v[-1]
 
 # The frequency response function of the filter is given by its
 # discrete time fourier transform:
@@ -61,13 +58,13 @@ H0 = np.fft.fft(data.filter_v[0]["h_fl_i"])
 H1 = np.fft.fft(data.filter_v[1]["h_fl_i"])
 
 # Plot of the frequency response of the filters (power) (in dB)
-F0 = np.arange(len(H0))*f_s/(len(H0))
-F1 = np.arange(len(H1))*f_s0/(len(H1))
+F0 = np.arange(len(H0))*f_s_dec_v[0]/(len(H0))
+F1 = np.arange(len(H1))*f_s_dec_v[1]/(len(H1))
 G0 = 20 * np.log10(np.abs(H0))
 # Repeat pattern for the second filter (4 times)
-F1l = np.append(F1, F1+f_s0)
-F1l = np.append(F1l, F1+2*f_s0)
-F1l = np.append(F1l, F1+3*f_s0)
+F1l = np.append(F1, F1+f_s_dec_v[1])
+F1l = np.append(F1l, F1+2*f_s_dec_v[1])
+F1l = np.append(F1l, F1+3*f_s_dec_v[1])
 G1 = 20 * np.log10(np.abs(H1))
 G1l = np.append(G1, G1)
 G1l = np.append(G1l, G1)

### 2022-02-04 Core/EK80CalculationPaper.py
MAJOR CHANGE: Removed all non static funcions from EK80CalculationPaper.py

@@ -4,99 +4,6 @@ from Core.EK80DataContainer import EK80DataContainer
 
 
 class EK80CalculationPaper(EK80DataContainer):
-
-    def __init__(self, jsonfname=None):
-        super().__init__(jsonfname)
-
-
-        # Derived constants
-        # Constant used to calculate power, some factors can be simplified, but this is written for clarity
-        K1 = 4 / ((2 * np.sqrt(2)) ** 2)
-        K2 = (np.abs(self.z_rx_e + self.z_td_e) / self.z_rx_e) ** 2
-        K3 = 1.0 / np.abs(self.z_td_e)
-        self.C1Prx = K1 * K2 * K3
-        self.f_s_dec = 1
-        self.calculateDerivedVariables()
-
-    def calculateDerivedVariables(self):
-
-        """
-        Generate match filtered transmit pulse
-
-        Function to generate the decmiated and filtered transmit
-        pulse for matche filtering
-
-        self.f_c Center frequency for transmit pulse
-        self.y_tx_n Ideal transmit pulse at original sampling rate
-        self.y_tx_n_t Time vector for transmit pulse at original sampling rate
-        self.y_tilde_tx_n Normalized ideal transmit pulse
-        self.y_mf_n Filtered and decimated transmit pulse for matched filter
-        self.y_mf_n_conj_rev Complex conjugated time reversed filter
-        self.y_mf_twoNormSquared Squared 2-norm of mathced filter
-        self.y_mf_auto Auto correlation function for matched filter
-        self.tau_eff effective pulse duration
-
-        """
-
-        # Estmate center frequency for transmit pulse
-        self.f_c = (self.f0 + self.f1) / 2.0
-
-        # Generate ideal transmit pulse at original sampling rate
-        y_tx_n, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
-            self.f0, self.f1, self.tau, self.f_s, self.slope)
-        self.y_tx_n = y_tx_n
-        self.y_tx_n_t = t
-
-        self.f_s_dec = EK80CalculationPaper.calcDecmiatedSamplingRate(self.filter_v, self.f_s)
-
-        # Normalize ideal transmit pulse
-        #y_tilde_tx_n = y_tx_n / np.max(y_tx_n)
-
-        # Eq (4)
-        y_tilde_tx_n = EK80CalculationPaper.calcNormalizedTransmitSignal(y_tx_n)
-
-        # Eq (5)
[…]

### 2022-02-04 Core/EK80DataContainer.py

@@ -73,6 +73,15 @@ class EK80DataContainer:
             self.beam_width_athwartship = None
             self.beam_width_alongship = None
             self.isCalibrated = False
+        # Calibration data (copied from EK80CalculationPaper)
+        if self.frequencies is not None:
+            self.frequencies = np.array(self.frequencies)
+        else:
+            # If no calibration make a frequency vector
+            # This is used to calculate correct frequencies after signal
+            # decimation
+
+            self.frequencies = np.linspace(self.f0, self.f1, self.n_f_points)
             
         self.filter_v = None
         if 'FIL1' in jdict and 'NaN' not in jdict['FIL1']:

### 2022-02-04 Paper/Manuscript.tex
Updated chapter IIa-c

@@
 Typical BB pulses are linear upsweep pulses windowed by an envelope function.
-In the EK80 a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}.
-This is found in the class \verb|Core.EK80Calculation.generateIdealWindowedSendPulse|, where the parameters 'f0', 'f1', tau, fs and slope are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
-A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
-
-To generate the signal
-
 The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain analogue signal $\ytxe(\timesym)$.
@@
 
+In the example a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}).
+This is found in the class \verb|Core.EK80Calculation.generateIdealWindowedSendPulse|, where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
+A tapering of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
 \begin{figure}

### 2022-02-06 Core/EK80DataContainer.py
variable groups have own class loosely reflecting data format

@@ -4,16 +4,116 @@ import json
 from Core.FIL1 import FIL1
 
 
+class Constants:
+    def __init__(self, z_td_e,f_s,n_f_points):
+        self.z_td_e = z_td_e
+        self.f_s = f_s
+        self.n_f_points = n_f_points
+
+class Transceiver:
+    def __init__(self, xml):
+        self.z_rx_e = xml['z_rx_e']
+
+class Parameter:
+    def __init__(self, xml):
+        self.f0 = xml['FrequencyStart']
+        self.f1 = xml['FrequencyEnd']
+        self.f_c = (self.f0 + self.f1) / 2.0
+        self.tau = xml['PulseDuration']
+        self.slope = xml['Slope']
+        self.sampleInterval = xml['SampleInterval']
+        self.ptx = xml['TransmitPower']
+
+class Transducer:
+    def __init__(self, xml):
+        self.fnom = xml['Frequency']  # nominal design frequency for the transducer
+        self.G_fnom = xml['GainNom']
+        self.PSI_fnom = xml['EquivalentBeamAngle']
+        self.angle_offset_alongship_fnom = xml['AngleOffsetAlongship']
+        self.angle_offset_athwartship_fnom = xml['AngleOffsetAthwartship']
+        self.angle_sensitivity_alongship_fnom = xml['AngleSensitivityAlongship']
+        self.angle_sensitivity_athwartship_fnom = xml['AngleSensitivityAthwartship']
+        self.beam_width_alongship_fnom = xml['BeamWidthAlongship']
+        self.beam_width_athwartship_fnom = xml['BeamWidthAthwartship']
+        self.corrSa = xml['SaCorrection']
+
+class Environment:
+    def __init__(self, xml):
+        self.c = xml['SoundSpeed']
+        self.alpha = xml['Alpha']
+        self.temperature = xml['Temperature']
+        self.salinity = xml['Salinity']
+        self.acidity = xml['Acidity']
+        self.latitude = xml['Latitude']
+        self.depth = xml['Depth']
+        self.dropKeelOffset = xml['DropKeelOffset']
+
+class FrequencyPar:
+    def __init__(self, xml=None):
+        # Test if broadband calibration values exists, if not use nominal values and extrapolate
+        if xml['frequencies']:
+            print("Broadband calibration values exists")
+            self.frequencies = xml['frequencies']
+            self.gain = xml['gain']
+            self.angle_offset_athwartship = xml['angle_offset_alongship']
[…]

### 2022-02-10 Core/EK80DataContainer.py
Added umpack method for data subclasses. Now used in main

@@ -46,6 +46,10 @@ class Transceiver:
     def __init__(self, xml):
         self.z_rx_e = xml['z_rx_e']
 
+    def getParameters(self):
+        return self.z_rx_e
+
+
 class Parameter:
     def __init__(self, xml):
         self.f0 = xml['FrequencyStart']
@@ -54,7 +58,10 @@ class Parameter:
         self.tau = xml['PulseDuration']
         self.slope = xml['Slope']
         self.sampleInterval = xml['SampleInterval']
-        self.ptx = xml['TransmitPower']
+        self.p_tx_e = xml['TransmitPower']
+
+    def getParameters(self):
+        return self.f0, self.f1, self.f_c, self.tau, self.slope, self.sampleInterval, self.p_tx_e
 
 class Transducer:
     def __init__(self, xml):
@@ -69,6 +76,12 @@ class Transducer:
         self.beam_width_athwartship_fnom = xml['BeamWidthAthwartship']
         self.corrSa = xml['SaCorrection']
 
+    def getParameters(self):
+        return self.fnom, self.G_fnom, self.PSI_fnom, self.angle_offset_alongship_fnom, \
+               self.angle_offset_athwartship_fnom,self.angle_sensitivity_alongship_fnom, \
+               self.angle_sensitivity_athwartship_fnom, self.beam_width_alongship_fnom, \
+               self.beam_width_athwartship_fnom, self.corrSa
+
 class Environment:
     def __init__(self, xml):
         self.c = xml['SoundSpeed']
@@ -79,6 +92,10 @@ class Environment:
         self.latitude = xml['Latitude']
         self.depth = xml['Depth']
         self.dropKeelOffset = xml['DropKeelOffset']
+
+    def getParameters(self):
+        return self.c, self.alpha, self.temperature, self.salinity, \
+               self.acidity, self.latitude,self.depth, self.dropKeelOffset
 
 class FrequencyPar:
     def __init__(self, xml=None):
@@ -111,6 +128,10 @@ class FrequencyPar:
 
             self.frequencies = np.linspace(self.f0, self.f1, self.n_f_points)
 
+    def getParameters(self):
+        return self.frequencies, self.gain, self.angle_offset_athwartship, self.angle_offset_alongship, \
+               self.beam_width_athwartship, self.beam_width_alongship
+
 
 class Filters:
     def __init__(self, xml=None):
[…]

### 2022-02-10 main.py

@@ -7,7 +7,7 @@ from Core.EK80DataContainer import EK80DataContainer
 data = EK80DataContainer('./data/pyEcholabEK80data.json')
 
 
-
+"""
 cont = data.cont
 trcv = data.trcv
 parm = data.parm
@@ -17,33 +17,46 @@ frqp = data.frqp
 filt = data.filt
 raw3 = data.raw3
 deriv = data.deriv
+"""
 # data.trdu.f_c
 # data.f_c -> I paper: $f_c$
 
+# Unpack variabels
+z_td_e, f_s, n_f_points = data.cont.getParameters()
 
+z_rx_e = data.trcv.getParameters()
+
+f0, f1, f_c, tau, slope, sampleInterval, p_tx_e = data.parm.getParameters()
+
+fnom, G_fnom, PSI_fnom, angle_offset_alongship_fnom, \
+angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
+angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
+beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
+
+c, alpha, temperature, salinity, \
+acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
+frequencies, gain, angle_offset_athwartship, angle_offset_alongship, \
+beam_width_athwartship, beam_width_alongship = data.frqp.getParameters()
+
+filter_v, N_v = data.filt.getParameters()
+
+offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
+
+g_0_f_c, lambda_f_c, PSI_f = data.deriv.getParameters()
 #
 # Chapter IIB: Signal generation
 #
 
 # Generate ideal send pulse
-y_tx_n, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
-    parm.f0,
-    parm.f1,
-    parm.tau,
-    cont.f_s,
-    parm.slope)
+y_tx_n, t = EK80CalculationPaper.generateIdealWindowedSendPulse(f0, f1, tau, f_s, slope)
 
-y_tx_n05slope, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
-    parm.f0,
-    parm.f1,
-    parm.tau,
-    cont.f_s, .5)
[…]

### 2022-02-10 Paper/Manuscript.tex
Added phi/theta plot
Added SED calculations and updated paper accordingly.

@@
 \end{equation}
+%
+\begin{figure}
+\includegraphics[width=\reprintcolumnwidth]{Fig_impedances}
+\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
+\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=\reprintcolumnwidth]{Fig_impedances}
-\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
+\includegraphics[width=\reprintcolumnwidth]{'Fig_theta_phi.png'}
+\caption{\label{fi:theta_phi}Physical angles.}
 \end{figure}
@@
 \ts(\freqsym) = 10\log_{10}(\prxetf(\freqsym)) + 40\log_{10}(\range) + 2\absorp(\freqsym) \range 
-- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along,\athw,\freqsym)}{16\pi^2} \right),
+- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,\freqsym)}{16\pi^2} \right),
 \end{equation}
 %
-where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain along the main acoustic axis, incorporating beam compensation based on the estimated target bearing, $(\along,\athw)$.
+
+where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain along the main acoustic axis, incorporating beam compensation based on the estimated target bearing, $(\along_t,\athw_t)$.
 
@@
 Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated.
+This is typically achieved through a single echo detection algorithm (SED).
+Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
+A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
 
-From the pulse compressed data, $\ypc(\samplesymt)$, the samples before and after the peak echo level corresponding to an echo from a single target are extracted to obtain the signal, $\ypctarget(\samplesymt)$, noting that the number of samples before the peak can differ from those after the peak.
 From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$.

### 2022-02-10 Core/EK80CalculationPaper.py

@@ -263,37 +263,36 @@ class EK80CalculationPaper(EK80DataContainer):
         new_auto = auto[idx_start_auto: idx_stop_auto]
 
         return new_auto
+    
+    @staticmethod
+    def singleTarget(y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
+                     r0, r1, before=0.5, after=0.5):
+        # This is a pseudo single target detector (SED) using the max peak
+        # within a range as the single target detection criteria
 
-    def singleTarget(self, y_pc_n, p_rx_e_n, theta_n, phi_n, r0, r1, before=0.5, after=0.5):
-
-        r_n, _ = self.calcRange()
-
-        # Extract samples for given range limits
-
+        # Get the samples within the sub range defined by before/after
         if r0 is not None and r1 is not None:
             Idx = np.where((r_n >= r0) & (r_n <= r1))
-            r_n = r_n[Idx]
-            y_pc_n = y_pc_n[Idx]
-            p_rx_e_n = p_rx_e_n[Idx]
-            theta_n = theta_n[Idx]
-            phi_n = phi_n[Idx]
+            r_n_sub = r_n[Idx]
+            y_pc_n_sub = y_pc_n[Idx]
+            p_rx_e_n_sub = p_rx_e_n[Idx]
+            theta_n_sub = theta_n[Idx]
+            phi_n_sub = phi_n[Idx]
 
         # Use peak power within given range limits as index for single target
-        # This assumes a single target detection algorithm has been used to
-        # detect a single target and define the range limits for the single target
+        # to get the range, theta and phi for the single target
+        idx_peak_p_rx = np.argmax(p_rx_e_n_sub)
+        r_t = r_n_sub[idx_peak_p_rx]
+        theta_t = theta_n_sub[idx_peak_p_rx]
+        phi_t = phi_n_sub[idx_peak_p_rx]
 
-        idx_peak_p_rx = np.argmax(p_rx_e_n)
-        r = r_n[idx_peak_p_rx]
-        theta = theta_n[idx_peak_p_rx]
-        phi = phi_n[idx_peak_p_rx]
+        # Extract pulse compressed samples "before" and "after" the peak power
+        r_t_begin = r_n_sub - before
+        r_t_end = r_n_sub + after
+        Idx2 = np.where((r_n_sub >= r_t_begin) & (r_n_sub <= r_t_end))
[…]

### 2022-02-10 main.py

@@ -236,8 +236,28 @@ Sp_n = EK80CalculationPaper.calcSp(
     r0,
     r1)
 
-# singleTarget:
-# r_n, _ = self.calcRange() -> EK80CalculationPaper
+# Pick samples from a single target (_t denotes data from single target)
+r_t_n, theta_t_n, phi_t_n, y_pc_t_n = EK80CalculationPaper.singleTarget(
+    y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
+    r0, r1, before, after)
+
+# Pick the reduced number of samples from the mathed filtered decimated send pulse
+# y_mf_auto_red =
+
+# Do the FFT of the pulse compressed signal corresponding to the target
+
+# Do the FFT of the pulse compressed send pulse signal rediced mathed filtered
+# decimated send pulse
+# Y_pc_t_m, Y_mf_auto_red =
+# Y_tilde_pc_t_m =
+
+
+# P_rx_e_t_m
+
+# Calculate the target strength
+# TS_m, f_m = CalcTS
+
+# The
 
 # CalcTSf:
 # L = self.n_f_points -> Transceiver or Parameter?

### 2022-02-10 Paper/Manuscript.tex
Compared paper and refactored code for the single target until the DFT.

@@
 \begin{figure}
-\includegraphics[width=\reprintcolumnwidth]{Fig_impedances}
+\includegraphics[width=16cm]{Fig_impedances}
 \caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
@@
 \begin{figure}
-\includegraphics[width=\reprintcolumnwidth]{'Fig_theta_phi.png'}
+\includegraphics[width=16cm]{Fig_theta_phi}
 \caption{\label{fi:theta_phi}Physical angles.}
@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ (or peak $S_p(n)$??
+current code uses $\ypc$) is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
+%
+\begin{figure}
+%\includegraphics[width=\reprintcolumnwidth]{'Fig_SED'}
+\caption{\label{fi:SED}Illustrate the single target data somehow.}
+\end{figure}
 
@@
 where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
-The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, is then calculated by: 
+The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\ref{fi:SED}) is then calculated by: 
 %
@@
 %
+The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...
+%
+\begin{figure}
+\includegraphics[width=16cm]{Fig_normalizedDFT}
+\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
+\end{figure}
+
 Assuming, as a first approximation, that the impedances of the transceiver and transducer are independent of frequency, the received power into a matched load, $\prxetf(\samplesymf)$, is then estimated by:

### 2022-02-10 Core/EK80CalculationPaper.py

@@ -249,20 +249,20 @@ class EK80CalculationPaper(EK80DataContainer):
         return Svf, svf_range, f
 
     @staticmethod
-    def alignAuto(auto, y_pc_t_n):
-
-        idx_peak_auto = np.argmax(np.abs(auto))
+    def alignAuto(y_mf_auto_n, y_pc_t_n):
+        # The equivalent samples around the peak in the decimated tx signal
+        idx_peak_auto = np.argmax(np.abs(y_mf_auto_n))
         idx_peak_y_pc_t_n = np.argmax(np.abs(y_pc_t_n))
 
         left_samples = idx_peak_y_pc_t_n
         right_samples = len(y_pc_t_n) - idx_peak_y_pc_t_n
 
         idx_start_auto = max(0, idx_peak_auto - left_samples)
-        idx_stop_auto = min(len(auto), idx_peak_auto + right_samples)
+        idx_stop_auto = min(len(y_mf_auto_n), idx_peak_auto + right_samples)
 
-        new_auto = auto[idx_start_auto: idx_stop_auto]
+        y_mf_auto_red_n = y_mf_auto_n[idx_start_auto: idx_stop_auto]
 
-        return new_auto
+        return y_mf_auto_red_n
     
     @staticmethod
     def singleTarget(y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
@@ -294,6 +294,33 @@ class EK80CalculationPaper(EK80DataContainer):
 
         return r_t, theta_t, phi_t, y_pc_t
 
+    @staticmethod
+    def calcDFTforTS(y_pc_t_n, y_mf_auto_red_n, n_f_points, f0, f1, f_s_dec):
+
+        # The number of DFT points inpower of 2
+        N_DFT = int(2 ** np.ceil(np.log2(n_f_points)))
+        # Lars: Is this correct? Why not N_DFT instead of n_f_points, which is
+        # before the 2**n conversion. I think this should be spelled out in the
+        # paper and possibly be rewritten for clarity.
+        f_m_t = np.linspace(f0, f1, n_f_points)
+        # Y_pc_t_m = self.freqtransf(_Y_pc_t_m, self.f_s_dec, f_m)
+        # def freqtransf(FFTvecin, fsdec, fvec=None):
[…]

### 2022-02-10 main.py

@@ -21,6 +21,10 @@ deriv = data.deriv
 # data.trdu.f_c
 # data.f_c -> I paper: $f_c$
 
+# To be considered:
+# f0 -> f_0
+# f1 -> f_1
+
 # Unpack variabels
 z_td_e, f_s, n_f_points = data.cont.getParameters()
 
@@ -43,6 +47,7 @@ filter_v, N_v = data.filt.getParameters()
 offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 
 g_0_f_c, lambda_f_c, PSI_f = data.deriv.getParameters()
+
 #
 # Chapter IIB: Signal generation
 #
@@ -244,48 +249,33 @@ Sp_n = EK80CalculationPaper.calcSp(
     r0,
     r1)
 
-# Pick samples from a single target (_t denotes data from single target)
-r_t_n, theta_t_n, phi_t_n, y_pc_t_n = EK80CalculationPaper.singleTarget(
+# Data from a single target (_t denotes data from single target)
+r_t, theta_t, phi_t, y_pc_t_n = EK80CalculationPaper.singleTarget(
     y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
     r0, r1, before, after)
 
-# Pick the reduced number of samples from the mathed filtered decimated send pulse
-# y_mf_auto_red =
+# Pick the reduced samples from the mathed filtered decimated send pulse
+y_mf_auto_red_n = EK80CalculationPaper.alignAuto(y_mf_auto_n, y_pc_t_n)
+# NB: In the example these are the same. Perhaps chose differently for
+# illustrative purposes?
 
-# Do the FFT of the pulse compressed signal corresponding to the target
+# DFT on the pulse compressed received signal and the pulse compressed
+# send pulse signal rediced mathed filtered
 
-# Do the FFT of the pulse compressed send pulse signal rediced mathed filtered
-# decimated send pulse
-# Y_pc_t_m, Y_mf_auto_red =
-# Y_tilde_pc_t_m =
+Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t = EK80CalculationPaper.calcDFTforTS(
[…]

### 2022-02-10 Core/EK80CalculationPaper.py
Compared paper and refactored code for the power by frequency.

@@ -320,6 +320,12 @@ class EK80CalculationPaper(EK80DataContainer):
         Y_tilde_pc_t_m = Y_pc_t_m/Y_mf_auto_red_m
         
         return Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t
+    
+    @staticmethod
+    def calcPowerFreq(N_u, Y_tilde_pc_t_m, z_td_e, z_rx_e):
+        imp = (np.abs(z_rx_e + z_td_e) / np.abs(z_rx_e)) ** 2 / np.abs(z_td_e)
+        P_rx_e_t_m = N_u * (np.abs(Y_tilde_pc_t_m)/(2 * np.sqrt(2))) ** 2 * imp
+        return P_rx_e_t_m
         
     def calcTSf(self, r, theta, phi, y_pc_t_n):
 

### 2022-02-10 main.py

@@ -262,8 +262,9 @@ y_mf_auto_red_n = EK80CalculationPaper.alignAuto(y_mf_auto_n, y_pc_t_n)
 # DFT on the pulse compressed received signal and the pulse compressed
 # send pulse signal rediced mathed filtered
 
-Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t = EK80CalculationPaper.calcDFTforTS(
-    y_pc_t_n, y_mf_auto_red_n, n_f_points, f0, f1, f_s_dec)
+Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t = \
+    EK80CalculationPaper.calcDFTforTS(y_pc_t_n, y_mf_auto_red_n,
+                                      n_f_points, f0, f1, f_s_dec)
 
 fig, axs = plt.subplots(3)
 fig.suptitle('Normalized DFT')
@@ -277,7 +278,14 @@ axs[2].set_ylabel('Y_tilde_pc_t_m')
 plt.savefig('./Paper/Fig_normalizedDFT.png')
 
 
-# Calculate the target strength of the single target
+# Calculate the power by frequency from a single target
+P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
+    N_u,
+    Y_tilde_pc_t_m,
+    z_td_e,
+    z_rx_e)
+
+# Calculate the target strength
 # TS_m, f_m,  = EK80CalculationPaper.calcTSf(r, theta, phi, y_pc_t_n)
 
 #

### 2022-02-18 main.py
Fixed bug in range extraction for the pseudo SED algorithm.

@@ -249,7 +249,7 @@ alpha_f_c = EK80CalculationPaper.calcAbsorption(
     f_c)
 
 # Move to data container (Ruben)
-# g_theta_phi_f
+# G = np.interp1(X, g_theta_phi_f, Y)
 
 #
 # RUBEN>
@@ -268,9 +268,11 @@ Sp_n = EK80CalculationPaper.calcSp(
     r1)
 
 # Data from a single target (_t denotes data from single target)
-r_t, theta_t, phi_t, y_pc_t_n = EK80CalculationPaper.singleTarget(
-    y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
-    r0, r1, before, after)
+r_t, theta_t, phi_t, y_pc_t_n, dum_p, dum_theta, dum_phi = \
+    EK80CalculationPaper.singleTarget(
+        y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
+        r0, r1, before, after)
+
 
 # Pick the reduced samples from the mathed filtered decimated send pulse
 y_mf_auto_red_n = EK80CalculationPaper.alignAuto(y_mf_auto_n, y_pc_t_n)
@@ -303,11 +305,11 @@ P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
     z_td_e,
     z_rx_e)
 
-
 # Calculate the target strength
-TS_m = EK80CalculationPaper.calcTSf_2(
-    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
-    gain, theta_t, phi_t, f_m_t)
+g_theta_phi_f = 1
+#TS_m = EK80CalculationPaper.calcTSf(
+#    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
+#    g_theta_phi_f, theta_t, phi_t, f_m_t)
 
 
 #

### 2022-02-18 Core/EK80DataContainer.py
Moved methods from EK80CalculationPaper.py to EK80DataContainer.py

@@ -14,19 +14,21 @@ from Core.FIL1 import FIL1
 
 class Derived:
     def __init__(self, frqp, parm, trdu, envr):
-        # Calculate Gfc
-        if frqp.isCalibrated:
-            # Calibrated case
-            self.Gfc = np.interp(parm.f_c, frqp.frequencies, frqp.gain)
-        else:
-            # Uncalibrated case
-            self.Gfc = trdu.G_fnom + 20 * np.log10(frqp.frequencies/trdu.fnom)
-        
-        # Calculate PSI_f
-        self.PSI_f = trdu.PSI_fnom + 20 * np.log10(trdu.fnom/frqp.frequencies)
 
+        self.Gfc = self.calc_Gfc(frqp, parm,trdu)
+        self.PSI_f = trdu.PSI_fnom + 20 * np.log10(trdu.fnom/frqp.frequencies)
         self.g_0_f_c = np.power(self.Gfc/10, 10)
         self.lambda_f_c = envr.c/parm.f_c
+
+    @staticmethod
+    def calc_Gfc(frqp, parm, trdu):
+        if frqp.isCalibrated:
+            # Calibrated case
+            return np.interp(parm.f_c, frqp.frequencies, frqp.gain)
+        else:
+            # Uncalibrated case
+            return trdu.G_fnom + 20 * np.log10(frqp.frequencies / trdu.fnom)
+
 
     def getParameters(self):
         return self.g_0_f_c, self.lambda_f_c, self.PSI_f
@@ -191,4 +193,36 @@ class EK80DataContainer:
 
             self.isCalibrated = self.frqp.isCalibrated
 
+    def calc_angle_offsets_m(self, f):
+        if self.isCalibrated:
+            # Calibrated case
+            angle_offset_alongship_m = np.interp(f, self.frqp.frequencies, self.frqp.angle_offset_alongship)
+            angle_offset_athwartship_m = np.interp(f, self.frqp.frequencies, self.frqp.angle_offset_athwartship)
+        else:
+            # Uncalibrated case
+            angle_offset_alongship_m = self.trdu.angle_offset_alongship_fnom * np.ones(len(f))
+            angle_offset_athwartship_m = self.trdu.angle_offset_athwartship_fnom * np.ones(len(f))
+        return angle_offset_alongship_m, angle_offset_athwartship_m
 
+    def calc_beam_widths_m(self, f):
+        if self.isCalibrated:
+            # Calibrated case
[…]

### 2022-02-18 main.py

@@ -295,10 +295,10 @@ P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
     z_rx_e)
 
 # Calculate the target strength
-g_theta_phi_f = 1
-#TS_m = EK80CalculationPaper.calcTSf(
-#    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
-#    g_theta_phi_f, theta_t, phi_t, f_m_t)
+g_theta_t_phi_t_f_t = data.calc_B_theta_phi_m(theta_t, phi_t, f_m_t)
+TS_m = EK80CalculationPaper.calcTSf(
+    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
+    g_theta_t_phi_t_f_t)
 
 
 #

### 2022-02-18 main.py
Added plot of single target

@@ -268,10 +268,22 @@ Sp_n = EK80CalculationPaper.calcSp(
     r1)
 
 # Data from a single target (_t denotes data from single target)
-r_t, theta_t, phi_t, y_pc_t_n, dum_p, dum_theta, dum_phi = \
+r_t, theta_t, phi_t, y_pc_t_n, dum_p, dum_theta, dum_phi, dum_r = \
     EK80CalculationPaper.singleTarget(
         y_pc_n, p_rx_e_n, theta_n, phi_n, r_n,
         r0, r1, before, after)
+
+fig, axs = plt.subplots(2)
+fig.suptitle('Single target')
+axs[0].plot(dum_r, dum_p)
+axs[0].set_ylabel('Power')
+axs[1].plot(dum_r, dum_theta)
+axs[1].plot([r_t, r_t], [-2, 2])
+axs[1].plot(dum_r, dum_phi)
+axs[1].plot(r_t, phi_t)
+axs[1].set_ylabel('Angles')
+axs[1].set_xlabel('range (m)')
+plt.savefig('./Paper/Fig_singleTarget.png')
 
 
 # Pick the reduced samples from the mathed filtered decimated send pulse
@@ -306,7 +318,8 @@ P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
     z_rx_e)
 
 # Calculate the target strength
-g_theta_phi_f = 1
+#g_theta_t_phi_t_f_t = data.calcGain(theta_t, phi_t, f_m_t)
+
 #TS_m = EK80CalculationPaper.calcTSf(
 #    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
 #    g_theta_phi_f, theta_t, phi_t, f_m_t)

### 2022-02-18 Core/EK80DataContainer.py
forgot to return values

@@ -226,3 +226,5 @@ class EK80DataContainer:
                             beam_width_alongship_m / 2)) ** 2 * \
                                                 (np.abs(phi - angle_offset_athwartship_m) / (
                                                             beam_width_athwartship_m / 2)) ** 2))
+
+        return B_theta_phi_m

### 2022-02-18 Paper/Manuscript.tex

@@
 
+Lars: The peak in the send pulse needs to match with the peak in the signal.
+
 Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated.
@@
 \begin{figure}
-%\includegraphics[width=\reprintcolumnwidth]{'Fig_SED'}
-\caption{\label{fi:SED}Illustrate the single target data somehow.}
+\includegraphics[width=16cm]{Fig_singleTarget}
+\caption{\label{fi:SED} The power (upper) and split beam angles (lower) for the single target. The vertical line corresponds to the range and angles for the single target.}
 \end{figure}
@@
 \label{eq:TS_f}
-\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range) + 2\absorp(\samplesymf)\range 
-- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along,\athw,\samplesymf)}{16\pi^2} \right).
+\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\samplesymf)\range_t 
+- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along_t,\athw_t,f_\samplesymf)}{16\pi^2} \right).
 \end{equation}
+
+Usikker på om vi skal bruka indeksar eller variablar for gain.
 

### 2022-02-18 Core/EK80CalculationPaper.py
First version of TS(f) with figure.

@@ -332,13 +332,13 @@ class EK80CalculationPaper(EK80DataContainer):
         return P_rx_e_t_m
 
     @staticmethod
-    def calcTSf(P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c, g_theta_t_phi_t_f_t):
-        # Look up gain from angles and freq (TBA)
-        G = 1
+    def calcTSf(P_rx_e_t_m, r_t, alpha, p_tx_e,
+                lambda_f_c, g_theta_t_phi_t_f_t):
+
         TS_m = 10*np.log10(P_rx_e_t_m) + 40*np.log10(r_t) + 2*alpha*r_t -\
             10*np.log10(
-            (p_tx_e * lambda_f_c**2 * G ** 2) / (16 * np.pi ** 2)
-            )
+            (p_tx_e * lambda_f_c**2 * g_theta_t_phi_t_f_t ** 2) /
+                (16 * np.pi ** 2))
         return TS_m
     
     """

### 2022-02-18 main.py

@@ -287,17 +287,6 @@ Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t = \
     EK80CalculationPaper.calcDFTforTS(y_pc_t_n, y_mf_auto_red_n,
                                       n_f_points, f0, f1, f_s_dec)
 
-fig, axs = plt.subplots(3)
-fig.suptitle('Normalized DFT')
-axs[0].plot(f_m_t, np.abs(Y_pc_t_m))
-axs[0].set_ylabel('Y_tilde_pc_t_m')
-axs[1].plot(f_m_t, np.abs(Y_mf_auto_red_m))
-axs[1].set_ylabel('Y_mf_auto_red_m')
-axs[2].plot(f_m_t, np.abs(Y_tilde_pc_t_m))
-axs[2].set_xlabel('f (Hz)')
-axs[2].set_ylabel('Y_tilde_pc_t_m')
-plt.savefig('./Paper/Fig_normalizedDFT.png')
-
 
 # Calculate the power by frequency from a single target
 P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
@@ -311,6 +300,21 @@ g_theta_t_phi_t_f_t = data.calc_B_theta_phi_m(theta_t, phi_t, f_m_t)
 TS_m = EK80CalculationPaper.calcTSf(
     P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
     g_theta_t_phi_t_f_t)
+
+
+fig, axs = plt.subplots(5)
+axs[0].plot(f_m_t, np.abs(Y_pc_t_m))
+axs[0].set_ylabel('Y_tilde_pc_t_m')
+axs[1].plot(f_m_t, np.abs(Y_mf_auto_red_m))
+axs[1].set_ylabel('Y_mf_auto_red_m')
+axs[2].plot(f_m_t, np.abs(Y_tilde_pc_t_m))
+axs[2].set_ylabel('Y_tilde_pc_t_m')
+axs[3].plot(f_m_t, g_theta_t_phi_t_f_t)
+axs[3].set_ylabel('gain')
+axs[4].plot(f_m_t, TS_m)
+axs[4].set_xlabel('f (Hz)')
+axs[4].set_ylabel('TS(f)')
+plt.savefig('./Paper/Fig_TS.png')
 
 
 #

### 2022-02-18 main.py
Added comment on the weird gain values

@@ -309,12 +309,13 @@ axs[1].plot(f_m_t, np.abs(Y_mf_auto_red_m))
 axs[1].set_ylabel('Y_mf_auto_red_m')
 axs[2].plot(f_m_t, np.abs(Y_tilde_pc_t_m))
 axs[2].set_ylabel('Y_tilde_pc_t_m')
-axs[3].plot(f_m_t, g_theta_t_phi_t_f_t)
+axs[3].plot(f_m_t, g_theta_t_phi_t_f_t) # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
 axs[3].set_ylabel('gain')
 axs[4].plot(f_m_t, TS_m)
 axs[4].set_xlabel('f (Hz)')
 axs[4].set_ylabel('TS(f)')
 plt.savefig('./Paper/Fig_TS.png')
+
 
 
 #

### 2022-02-18 Core/EK80CalculationPaper.py
Fixed problem with weird gain. Notation might be wrong

@@ -98,12 +98,13 @@ class EK80CalculationPaper(EK80DataContainer):
         
         return C1Prx * np.abs(y_pc) ** 2
 
+    """
     @staticmethod
     def calcRange(sampleInterval, sampleCount, c, offset):
         dr = sampleInterval * c * 0.5
         r = np.array([(offset + i + 1) * dr for i in range(0, sampleCount)])
         return r, dr
-
+    """
     @staticmethod
     def calcSp(
             p_rx_e_n,
@@ -335,6 +336,8 @@ class EK80CalculationPaper(EK80DataContainer):
     def calcTSf(P_rx_e_t_m, r_t, alpha, p_tx_e,
                 lambda_f_c, g_theta_t_phi_t_f_t):
 
+
+
         TS_m = 10*np.log10(P_rx_e_t_m) + 40*np.log10(r_t) + 2*alpha*r_t -\
             10*np.log10(
             (p_tx_e * lambda_f_c**2 * g_theta_t_phi_t_f_t ** 2) /
@@ -523,6 +526,7 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         lf = self.lambda_f(f)
         return 10 * np.log10((self.ptx * self.c * lf ** 2) / (32.0 * np.pi * np.pi))
 
+    """
     @staticmethod
     def calcAbsorption(t, s, d, ph, c, f):
         f = f / 1000
@@ -544,4 +548,4 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         a = f ** 2 * (a1 * p1 * f1 / (f1 ** 2 + f ** 2) + a2 * p2 * f2 / (f2 ** 2 + f ** 2) + a3 * p3)
 
         return a / 1000
-
+    """

### 2022-02-18 Core/EK80DataContainer.py

@@ -193,6 +193,34 @@ class EK80DataContainer:
 
             self.isCalibrated = self.frqp.isCalibrated
 
+    @staticmethod
+    def calcAbsorption(t, s, d, ph, c, f):
+        f = f / 1000
+
+        a1 = (8.86 / c) * 10 ** (0.78 * ph - 5)
+        p1 = 1
+        f1 = 2.8 * (s / 35) ** 0.5 * 10 ** (4 - 1245 / (t + 273))
+
+        a2 = 21.44 * (s / c) * (1 + 0.025 * t)
+        p2 = 1 - 1.37e-4 * d + 6.62e-9 * d ** 2
+        f2 = 8.17 * 10 ** (8 - 1990 / (t + 273)) / (1 + 0.0018 * (s - 35))
+
+        p3 = 1 - 3.83e-5 * d + 4.9e-10 * d ** 2
+
+        a3l = 4.937e-4 - 2.59e-5 * t + 9.11e-7 * t ** 2 - 1.5e-8 * t ** 3
+        a3h = 3.964e-4 - 1.146e-5 * t + 1.45e-7 * t ** 2 - 6.5e-10 * t ** 3
+        a3 = a3l * (t <= 20) + a3h * (t > 20)
+
+        a = f ** 2 * (a1 * p1 * f1 / (f1 ** 2 + f ** 2) + a2 * p2 * f2 / (f2 ** 2 + f ** 2) + a3 * p3)
+
+        return a / 1000
+
+    @staticmethod
+    def  calcRange(sampleInterval, sampleCount, c, offset) :
+        dr = sampleInterval * c * 0.5
+        r = np.array([(offset + i + 1) * dr for i in range(0, sampleCount)])
+        return r, dr
+
     def calc_angle_offsets_m(self, f):
         if self.isCalibrated:
             # Calibrated case
@@ -215,6 +243,14 @@ class EK80DataContainer:
             beam_width_athwartship_m = self.trdu.beam_width_athwartship_fnom * self.trdu.fnom / f
         return beam_width_alongship_m, beam_width_athwartship_m
 
+    def calc_G0_m(self, f):
+        if self.isCalibrated:
+            # Calibrated case
+            return np.interp(f, self.frqp.frequencies, self.frqp.gain)
+        else:
+            # Uncalibrated case
+            return self.trdu.G_fnom + 20 * np.log10(f / self.trdu.fnom)
+
     def calc_B_theta_phi_m(self, theta, phi, f):
[…]

### 2022-02-18 main.py

@@ -199,37 +199,14 @@ r1 = 30
 before = 0.5
 after = 1
 
-#p_tx_e = parm.ptx
-#f_c = parm.f_c
-#g_0_f_c = data.deriv.g_0_f_c
-#PSI_f = data.deriv.PSI_f
-#lambda_f_c = data.deriv.lambda_f_c
-
-# logSpCf - > lambda_f_c
-# power -> p_tx_e
-# Gfc -> g_0_f_c
-# r -> r_n
-# Sp -> S_p_n
-
-#
-# <RUBEN!
-#
-
-
-# Will be written explicitly in EK80CalculationPaper and removed
-# logSpCf = EK80CalculationPaper.calculateCSpfdB(f_c, ptx)
-# Move to EK80DataContainer (Ruben)
-r_n, _ = EK80CalculationPaper.calcRange(
+r_n, _ = EK80DataContainer.calcRange(
     sampleInterval,
     sampleCount,
     c,
     offset)
 
-# alpha_fc = self.calcAbsorption(self.temperature, self.salinity,
-# self.depth, self.acidity, self.c, self.f_c)
 
-# Move to EK80DataContainer (Ruben):
-alpha_f_c = EK80CalculationPaper.calcAbsorption(
+alpha_f_c = EK80DataContainer.calcAbsorption(
     temperature,
     salinity,
     depth,
@@ -296,10 +273,16 @@ P_rx_e_t_m = EK80CalculationPaper.calcPowerFreq(
     z_rx_e)
 
 # Calculate the target strength
+"""
 g_theta_t_phi_t_f_t = data.calc_B_theta_phi_m(theta_t, phi_t, f_m_t)
+G0_m = data.calc_G0_m(f_m_t)
+G_theta_phi_m = G0_m - g_theta_t_phi_t_f_t
+"""
+G_theta_phi_m = data.calc_g(theta_t, phi_t, f_m_t)
 TS_m = EK80CalculationPaper.calcTSf(
     P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
-    g_theta_t_phi_t_f_t)
+    G_theta_phi_m)
+
 
 
 fig, axs = plt.subplots(5)
@@ -309,7 +292,7 @@ axs[1].plot(f_m_t, np.abs(Y_mf_auto_red_m))
 axs[1].set_ylabel('Y_mf_auto_red_m')
 axs[2].plot(f_m_t, np.abs(Y_tilde_pc_t_m))
[…]

### 2022-02-18 Core/EK80DataContainer.py
gain returned as linear values

@@ -269,4 +269,4 @@ class EK80DataContainer:
 
         B_theta_phi_m = self.calc_B_theta_phi_m(theta, phi, f)
         G0_m = self.calc_G0_m(f)
-        return G0_m - B_theta_phi_m
+        return np.power(10,(G0_m - B_theta_phi_m)/10)

### 2022-02-18 Core/EK80DataContainer.py
methods should return values in linear domain

@@ -246,10 +246,13 @@ class EK80DataContainer:
     def calc_G0_m(self, f):
         if self.isCalibrated:
             # Calibrated case
-            return np.interp(f, self.frqp.frequencies, self.frqp.gain)
+            dB_G0 = np.interp(f, self.frqp.frequencies, self.frqp.gain)
         else:
             # Uncalibrated case
-            return self.trdu.G_fnom + 20 * np.log10(f / self.trdu.fnom)
+            dB_G0 = self.trdu.G_fnom + 20 * np.log10(f / self.trdu.fnom)
+
+        return np.power(10,dB_G0/10)
+
 
     def calc_B_theta_phi_m(self, theta, phi, f):
         angle_offset_alongship_m, angle_offset_athwartship_m = self.calc_angle_offsets_m(f)
@@ -263,10 +266,11 @@ class EK80DataContainer:
                                                 (np.abs(phi - angle_offset_athwartship_m) / (
                                                             beam_width_athwartship_m / 2)) ** 2))
 
-        return B_theta_phi_m
+
+        return np.power(10, B_theta_phi_m / 10)
 
     def calc_g(self, theta, phi, f):
 
         B_theta_phi_m = self.calc_B_theta_phi_m(theta, phi, f)
         G0_m = self.calc_G0_m(f)
-        return np.power(10,(G0_m - B_theta_phi_m)/10)
+        return G0_m - B_theta_phi_m

### 2022-02-18 Core/EK80Calculation.py
Refactored gain and beam to lowercase, updated gain equation, updated example_TSf.py

@@ -147,7 +147,7 @@ class EK80Calculation(EK80DataContainer):
 
     def calcSp(self, power, r0=None, r1=None):
 
-        Gfc = self.calc_G0_m(self.f_c)
+        Gfc = self.calc_g0_m(self.f_c)
         PSIfc = self.PSI_f(self.f_c)
         logSpCf = self.calculateCSpfdB(self.f_c)
         r, _ = self.calcRange()
@@ -170,7 +170,7 @@ class EK80Calculation(EK80DataContainer):
 
     def calcSv(self, power, r0=None, r1=None):
 
-        Gfc = self.calc_G0_m(self.f_c)
+        Gfc = self.calc_g0_m(self.f_c)
         PSIfc = self.PSI_f(self.f_c)
         logSvCf = self.calculateCSvfdB(self.f_c)
         r, _ = self.calcRange()
@@ -235,7 +235,7 @@ class EK80Calculation(EK80DataContainer):
         _FFTytxauto = np.fft.fft(self.y_mf_auto_n, n=Nw)
         FFTytxauto = self.freqtransf(_FFTytxauto, self.f_s_dec, f)
 
-        Gf = self.calc_G0_m(f)    # Used only if not calibrated
+        Gf = self.calc_g0_m(f)    # Used only if not calibrated
         PSIf = self.PSI_f(f)
         alpha_f = self.calcAbsorption(self.temperature, self.salinity, self.depth, self.acidity, self.c, f)
         logSvCf = self.calculateCSvfdB(f)
@@ -355,8 +355,8 @@ class EK80Calculation(EK80DataContainer):
         _Y_mf_auto_red_m = np.fft.fft(y_mf_auto_red_n, n=N_DFT)
         Y_mf_auto_red_m = self.freqtransf(_Y_mf_auto_red_m, self.f_s_dec, f_m)
 
-        G0_m = self.calc_G0_m(f_m)
-        B_theta_phi_m = self.calc_B_theta_phi_m(theta, phi, f_m)
+        G0_m = self.calc_g0_m(f_m)
+        B_theta_phi_m = self.calc_b_theta_phi_m(theta, phi, f_m)
         G_theta_phi_m = G0_m - B_theta_phi_m
         alpha_m = self.calcAbsorption(self.temperature, self.salinity, self.depth, self.acidity, self.c, f_m)
         logSpCf = self.calculateCSpfdB(f_m)
@@ -376,7 +376,7 @@ class EK80Calculation(EK80DataContainer):
     def PSI_f(self, f):
[…]

### 2022-02-18 Examples/example_TSf.py

@@ -18,57 +18,118 @@ if __name__ == '__main__':
 
     args = parser.parse_args()
 
-    #ekdata = EK80DataContainer(args.file)
-    ekcalc = EK80CalculationPaper(args.file)
+    # Test data
+    data = EK80DataContainer('../data/pyEcholabEK80data.json')
 
-    y_pc_nu = EK80CalculationPaper.calcPulseCompressedQuadrants(ekcalc.y_rx_nu,
-                                                                ekcalc.y_mf_n)
+    # Unpack variabels
+    z_td_e, f_s, n_f_points = data.cont.getParameters()
 
-    #y_pc_nu = ekcalc.calcPulseCompressedQuadrants(ekcalc.y_rx_nu)
+    z_rx_e = data.trcv.getParameters()
+
+    f0, f1, f_c, tau, slope, sampleInterval, p_tx_e = data.parm.getParameters()
+
+    fnom, G_fnom, PSI_fnom, angle_offset_alongship_fnom, \
+    angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
+    angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
+    beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
+
+    c, alpha, temperature, salinity, \
+    acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
+    frequencies, gain, angle_offset_athwartship, angle_offset_alongship, \
+    beam_width_athwartship, beam_width_alongship = data.frqp.getParameters()
+
+    filter_v, N_v = data.filt.getParameters()
+
+    offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
+
+    g_0_f_c, lambda_f_c, PSI_f = data.deriv.getParameters()
+
+     # Generate ideal send pulse
+    y_tx_n, t = EK80CalculationPaper.generateIdealWindowedSendPulse(
+        f0, f1, tau, f_s, slope)
+
+    # The sampling freq for each filter step
+    f_s_dec_v = EK80CalculationPaper.calcDecmiatedSamplingRate(filter_v, f_s)
+
+    # The final sampling frequency
+    f_s_dec = f_s_dec_v[-1]
+
+    # The normalized ideal transmit signal
[…]

### 2022-02-18 Tests/test_EK80Calculation.py

@@ -108,7 +108,7 @@ class TestEK80Calculation(unittest.TestCase):
 
     def test_Gf(self):
 
-        Gf = self.calc.calc_G0_m(39500)
+        Gf = self.calc.calc_g0_m(39500)
 
         np.testing.assert_almost_equal(Gf,26.11500, 5)
 

### 2022-02-18 main.py

@@ -278,10 +278,10 @@ g_theta_t_phi_t_f_t = data.calc_B_theta_phi_m(theta_t, phi_t, f_m_t)
 G0_m = data.calc_G0_m(f_m_t)
 G_theta_phi_m = G0_m - g_theta_t_phi_t_f_t
 """
-G_theta_phi_m = data.calc_g(theta_t, phi_t, f_m_t)
+g_theta_phi_m = data.calc_g(theta_t, phi_t, f_m_t)
 TS_m = EK80CalculationPaper.calcTSf(
     P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
-    G_theta_phi_m)
+    g_theta_phi_m)
 
 
 
@@ -292,7 +292,7 @@ axs[1].plot(f_m_t, np.abs(Y_mf_auto_red_m))
 axs[1].set_ylabel('Y_mf_auto_red_m')
 axs[2].plot(f_m_t, np.abs(Y_tilde_pc_t_m))
 axs[2].set_ylabel('Y_tilde_pc_t_m')
-axs[3].plot(f_m_t, G_theta_phi_m) # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
+axs[3].plot(f_m_t, g_theta_phi_m) # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
 axs[3].set_ylabel('gain')
 axs[4].plot(f_m_t, TS_m)
 axs[4].set_xlabel('f (Hz)')

### 2022-02-18 Paper/Manuscript.tex

@@
 %
-
-where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain along the main acoustic axis, incorporating beam compensation based on the estimated target bearing, $(\along_t,\athw_t)$.
+where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
 
@@
 
-Lars: The peak in the send pulse needs to match with the peak in the signal.
-
 Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated.
@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ (or peak $S_p(n)$??
-current code uses $\ypc$) is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power \prxe(\samplesymt) within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
@@
 \includegraphics[width=16cm]{Fig_singleTarget}
-\caption{\label{fi:SED} The power (upper) and split beam angles (lower) for the single target. The vertical line corresponds to the range and angles for the single target.}
+\caption{\label{fi:SED} The \prxe(\samplesymt) power (upper) and split beam angles ($\along_t$ and $\athw_t$) (lower) for the single target.
+The vertical line corresponds to the range $r_t$ for the single target.}
 \end{figure}
@@
 The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...
-%
-\begin{figure}
-\includegraphics[width=16cm]{Fig_normalizedDFT}
-\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
-\end{figure}
 
@@
 \end{equation}
-
-Usikker på om vi skal bruka indeksar eller variablar for gain.
+%
+\begin{figure}
+\includegraphics[width=16cm]{Fig_TS}
+\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
+\end{figure}
 

### 2022-02-18 Core/EK80CalculationPaper.py
Changed alpha and lambda

@@ -333,14 +333,14 @@ class EK80CalculationPaper(EK80DataContainer):
         return P_rx_e_t_m
 
     @staticmethod
-    def calcTSf(P_rx_e_t_m, r_t, alpha, p_tx_e,
-                lambda_f_c, g_theta_t_phi_t_f_t):
+    def calcTSf(P_rx_e_t_m, r_t, alpha_m, p_tx_e,
+                lambda_m, g_theta_t_phi_t_f_t):
 
 
 
-        TS_m = 10*np.log10(P_rx_e_t_m) + 40*np.log10(r_t) + 2*alpha*r_t -\
+        TS_m = 10*np.log10(P_rx_e_t_m) + 40*np.log10(r_t) + 2*alpha_m*r_t -\
             10*np.log10(
-            (p_tx_e * lambda_f_c**2 * g_theta_t_phi_t_f_t ** 2) /
+            (p_tx_e * lambda_m**2 * g_theta_t_phi_t_f_t ** 2) /
                 (16 * np.pi ** 2))
         return TS_m
     

### 2022-02-18 Examples/example_TSf.py

@@ -126,6 +126,9 @@ if __name__ == '__main__':
         g_theta_phi_m)
 
     plt.figure()
+    plt.plot(g_theta_phi_m)
+
+    plt.figure()
     plt.title('Tsf for target in layer from {:.1f} m to {:.1f} m'.format(args.r0, args.r1))
     plt.plot(TS_m)
     plt.xlabel('kHz')

### 2022-02-18 main.py

@@ -280,7 +280,7 @@ G_theta_phi_m = G0_m - g_theta_t_phi_t_f_t
 """
 g_theta_phi_m = data.calc_g(theta_t, phi_t, f_m_t)
 TS_m = EK80CalculationPaper.calcTSf(
-    P_rx_e_t_m, r_t, alpha, p_tx_e, lambda_f_c,
+    P_rx_e_t_m, r_t, alpha_m, p_tx_e, lambda_m,
     g_theta_phi_m)
 
 

### 2022-02-22 Paper/Manuscript.tex
Change to r_c for the range for sv in the MS

@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power \prxe(\samplesymt) within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
+
+From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$.
+Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
+If the number of samples around the target is more than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used.
+If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create a reduced autocorrelation signal, $\ymfautored(\samplesymt)$.
 %
@@
 \includegraphics[width=16cm]{Fig_singleTarget}
-\caption{\label{fi:SED} The \prxe(\samplesymt) power (upper) and split beam angles ($\along_t$ and $\athw_t$) (lower) for the single target.
-The vertical line corresponds to the range $r_t$ for the single target.}
+\caption{\label{fi:SED} The $\prxe(\samplesymt)$ power (upper) and split beam angles ($\along_t$ and $\athw_t$) (middle) for the single target.
+The vertical line corresponds to the range $r_t$ for the single target.
+The $\ymfautored(\samplesymt)$ (lower) is the autocorrelation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.}
 \end{figure}
-
-From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$.
-Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
-If the number of samples around the target is more than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used.
-If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create a reduced autocorrelation signal, $\ymfautored(\samplesymt)$.
 
@@
 \label{eq:sv_f}
-\sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range) + 2\absorp(\freqsym)\range 
+\sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range_c) + 2\absorp(\freqsym)\range_r 
 - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym)}{32\pi^2} \right), 
@@
 where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, and $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum.
-Note that $\range$ is the range to the centre of the range volume covered by $\tslide$, and the two-way equivalent beam angle, $\eqang$, is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
+Note that $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and the two-way equivalent beam angle, $\eqang$, is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
 \begin{equation}
@@
 \label{eq:Sv}
-\sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range(\samplesymt)) + 2\absorp(\fc)\range(\samplesymt) 
+\sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range_c(\samplesymt)) + 2\absorp(\fc)\range_c(\samplesymt) 
 - 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \cw \teff \eqang(\fc) \gainzero^2(\fc)}{32\pi^2} \right),
@@
 To estimate $\sv$ as a function of frequency a Fourier transform is used, repeatedly applied via a sliding window in range.
-However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r^2$, implemented as the $20\log_{10}(\range)$ term in Eq.
+However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r_c^2$, implemented as the $20\log_{10}(\range_c)$ term in Eq.
 \ref{eq:Sv}) from the beginning of the window to the end can be significant, particularly for short range measurements.
@@
 
-Compensation of spherical spreading loss requires compensation of received power by a factor of $r^2$, and hence compensation of amplitude by a factor of $\range$:
+Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
 %
@@
 \label{eq:spreadcomp}
-\ypcspread(\samplesymt) = \ypc(\samplesymt)\range(\samplesymt).
+\ypcspread(\samplesymt) = \ypc(\samplesymt)\range_c(\samplesymt).
 \end{equation}
@@
 \label{eq:Sv_FFT}
-\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range \\
+\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range_c \\
 - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).

### 2022-02-22 main.py
Added the functions from the paper in main.py (as comments for now)

@@ -22,14 +22,14 @@ z_rx_e = data.trcv.getParameters()
 f0, f1, f_c, tau, slope, sampleInterval, p_tx_e = data.parm.getParameters()
 
 fnom, G_fnom, PSI_fnom, angle_offset_alongship_fnom, \
-angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
-angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
-beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
+    angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
+    angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
+    beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
 
 c, alpha, temperature, salinity, \
-acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
+    acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
 frequencies, gain, angle_offset_athwartship, angle_offset_alongship, \
-beam_width_athwartship, beam_width_alongship = data.frqp.getParameters()
+    beam_width_athwartship, beam_width_alongship = data.frqp.getParameters()
 
 filter_v, N_v = data.filt.getParameters()
 
@@ -241,8 +241,6 @@ r_t, theta_t, phi_t, y_pc_t_n, dum_p, dum_theta, dum_phi, dum_r = \
 
 # Pick the reduced samples from the mathed filtered decimated send pulse
 y_mf_auto_red_n = EK80CalculationPaper.alignAuto(y_mf_auto_n, y_pc_t_n)
-# NB: In the example these are the same. Perhaps chose differently for
-# illustrative purposes?
 
 fig, axs = plt.subplots(3)
 fig.suptitle('Single target')
@@ -259,16 +257,12 @@ axs[2].set_ylabel(' ')
 axs[2].set_xlabel('range (m)')
 plt.savefig('./Paper/Fig_singleTarget.png')
 
-#plt.plot(r_t, y_mf_auto_red_n)
-#plt.show_fig()
 
 # DFT on the pulse compressed received signal and the pulse compressed
[…]

### 2022-02-22 Paper/Manuscript.tex
Minor mis-spelling

@@
 \label{eq:sv_f}
-\sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range_c) + 2\absorp(\freqsym)\range_r 
+\sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range_c) + 2\absorp(\freqsym)\range_c 
 - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym)}{32\pi^2} \right), 
@@
 %
-where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, and $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum.
-Note that $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and the two-way equivalent beam angle, $\eqang$, is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
+where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang$ is the two-way equivalent beam angle.
+The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
 \begin{equation}

### 2022-02-22 Core/EK80CalculationPaper.py
Refactored code for Hanning window definition.

@@ -539,10 +539,22 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         y_pc_s_n = y_pc_n * r_c_n
         return y_pc_s_n
 
-    #@staticmethod
-    #def = calcHanning(t, pulseduration):
-    #    w_tilde_i = t + pulseduration
-    #    return w_tilde_i
+    @staticmethod
+    def defHanningWindow(c, tau, dr, f_s_dec):
+        """
+            Length of Hanning window currently chosen as 2^k samples for
+            lowest k where 2^k >= 2 * No of samples in pulse
+        """
+        L = ((c * 2 * tau) / dr)  # Number of samples in 2 x pulse duration
+        
+        N_w = int(2 ** np.ceil(np.log2(L)))
+        # or : N_w = np.ceil(2 ** np.log2(L)) - Length of Hanning window
+        t_w = np.arange(0, N_w) * f_s_dec
+        
+        w_i = EK80CalculationPaper.hann(N_w)
+        w_tilde_i = w_i / (np.linalg.norm(w_i) / np.sqrt(N_w))
+        
+        return w_tilde_i, N_w, t_w
 
     
     def calculateCSvfdB(self, f):

### 2022-02-22 main.py

@@ -39,7 +39,9 @@ filter_v, N_v = data.filt.getParameters()
 
 offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 
-# Ruben: the Psi should be the Psi_e_f, check paper and rewrite the data class.
+# Ruben: the Psi should be the Psi_e_f_n, check paper and rewrite the data
+# class. The difference is the (f_n/f)**2 term. It should be Psi_e not Psi.
+# Check paper.
 g_0_f_c, lambda_f_c, Psi_e_f_n = data.deriv.getParameters()
 
 #
@@ -326,6 +328,7 @@ Psi_f = EK80CalculationPaper.calc_Psi_f(Psi_e_f_n, f_n, frequencies)
 # Calculate average Sv
 # TODO: I get zero power in the p_rx_e_n. Fails when doing log10. "Quickfix":
 p_rx_e_n = p_rx_e_n + .0000000000000001
+
 Sv_n = EK80CalculationPaper.calc_Sv(p_rx_e_n, r_c_n, lambda_f_c,
                                     p_tx_e, alpha_f_c, c, tau_eff,
                                     Psi_f_c, g_0_f_c)
@@ -334,10 +337,23 @@ Sv_n = EK80CalculationPaper.calc_Sv(p_rx_e_n, r_c_n, lambda_f_c,
 y_pc_s_n = EK80CalculationPaper.calc_PulseCompSphericalSpread(y_pc_n, r_c_n)
 
 # Hanning window
-# w_tilde_i = EK80CalculationPaper.calcHanning(t, pulseduration)
+# TODO: This needs attention:
+dr = np.median(np.diff(r_c_n))
 
-# calculate the DFT on the pulse compressed signal
+w_tilde_i, N_w, t_w = EK80CalculationPaper.defHanningWindow(c, tau, dr,
+                                                            f_s_dec)
+
+
+# Calculate the DFT on the pulse compressed signal
 # Y_pc_v_m, Y_mf_auto_m, Y_tilde_pc_v_m = EK80CalculationPaper.calcDFTforSv(y_pc_s_n, w_tilde_i, y_mf_auto_n)
+
+fig, axs = plt.subplots(2)
+axs[0].plot(t_w, w_tilde_i)
+axs[0].set_ylabel('w_tilde_i')
+axs[1].plot(t_w, w_tilde_i)
[…]

### 2022-02-22 Core/EK80CalculationPaper.py
Added FFT for sv functions

@@ -540,6 +540,75 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         return y_pc_s_n
 
     @staticmethod
+    def calcDFTforSv(y_pc_s_n, w_tilde_i, y_mf_auto_n, N_w,
+                     n_f_points, f0, f1, f_s_dec, r_c_n, step):
+
+        # The frequency vector for the windowed signal
+        f = np.linspace(f0, f1, n_f_points)
+
+        # Initialize the fft as list (and append for each depth bin)
+        Y_pc_v_m_n = []
+        Y_tilde_pc_v_m_n = []
+
+        
+        # The DFT of the ACf of the mathced filter signal
+        _Y_mf_auto_m = np.fft.fft(y_mf_auto_n, n=N_w)
+        Y_mf_auto_m = EK80CalculationPaper.freqtransf(_Y_mf_auto_m, f_s_dec, f)
+
+        svf_range = []
+        min_sample = 0  # int(r0 / dr)
+        max_sample = len(y_pc_s_n)  # int(r1 / dr)
+
+        bin_start_sample = min_sample
+        last_bin = False
+        n_bins = 0
+        while not last_bin:
+            bin_stop_sample = bin_start_sample + N_w
+
+            # Apply window to signal bin
+            if bin_stop_sample < max_sample:
+                # We have a whole bin use precalculated window
+                yspread_bin = w_tilde_i * y_pc_s_n[
+                    bin_start_sample:bin_stop_sample]
+            else:
+                # We might have partial bin, recalculated window
+                
+                # TODO: We need to present this in the paper. Nils Olav is
+                # sceptic to do this.Should we remove the edge cases? They will
+                # be very different to the "center" cases as they may have much
+                # fewer samples?
+                last_bin = True
+                bin_stop_sample = max_sample
+                sub_yspread = y_pc_s_n[bin_start_sample:bin_stop_sample]
+                w = EK80CalculationPaper.hann(len(sub_yspread))
+                w = w / (np.linalg.norm(w) / np.sqrt(len(sub_yspread)))
+                yspread_bin = w * sub_yspread
+
+            # Find the range for this window
+            bin_center_sample = int((bin_stop_sample + bin_start_sample) / 2)
+            bin_center_range = r_c_n[bin_center_sample]
+            svf_range.append(bin_center_range)
+
[…]

### 2022-02-22 main.py

@@ -341,7 +341,12 @@ w_tilde_i, N_w, t_w = EK80CalculationPaper.defHanningWindow(c, tau, dr,
 
 
 # Calculate the DFT on the pulse compressed signal
-# Y_pc_v_m, Y_mf_auto_m, Y_tilde_pc_v_m = EK80CalculationPaper.calcDFTforSv(y_pc_s_n, w_tilde_i, y_mf_auto_n)
+
+step = 1  # Needs some thoughts...
+Y_pc_v_m_n, Y_mf_auto_m, Y_tilde_pc_v_m_n, svf_range \
+    = EK80CalculationPaper.calcDFTforSv(
+        y_pc_s_n, w_tilde_i, y_mf_auto_n, N_w, n_f_points, f0, f1, f_s_dec,
+        r_c_n, step)
 
 fig, axs = plt.subplots(2)
 axs[0].plot(t_w, w_tilde_i)

### 2022-02-22 Paper/Manuscript.tex
Changed from Psi to Psi_e in MS
Added functions to calculate power(f)

@@
 where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang$ is the two-way equivalent beam angle.
-The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
+The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang_e$ at the nominal frequency, $\fn$:
 \begin{equation}
 \label{eq:PsiFc}
-\eqang(f) = \eqang(\fn)\left(\frac{\fn}{f}\right)^2.
+\eqang(f) = \eqang_e(\fn)\left(\frac{\fn}{f}\right)^2.
 \end{equation}
@@
 
+TODO: Explain edge cases.
+
 The normalised Hanning window, $\hannwnorm$, is given by: 
@@
 %
+TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.
+
 The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:

### 2022-02-22 Core/EK80CalculationPaper.py

@@ -609,6 +609,24 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         return Y_pc_v_m_n, Y_mf_auto_m, Y_tilde_pc_v_m_n, svf_range
 
     @staticmethod
+    def calcPowerFreqforSv(Y_tilde_pc_v_m_n, N_u, z_rx_e, z_td_e):
+
+        # Initialize list of power values by range
+        P_rx_e_v_m_n = []
+
+        # Impedances
+        Z = (np.abs(z_rx_e + z_td_e)/np.abs(z_rx_e)) ** 2 / np.abs(z_td_e)
+        
+        # Loop over list of FFTs along range
+        for Y_tilde_pc_v_m in Y_tilde_pc_v_m_n:
+            P_rx_e_v_m = N_u * (
+                np.abs(Y_tilde_pc_v_m) / (2 * np.sqrt(2))) ** 2 * Z
+            # Append power to list
+            P_rx_e_v_m_n.append(P_rx_e_v_m)
+        
+        return P_rx_e_v_m_n
+
+    @staticmethod
     def defHanningWindow(c, tau, dr, f_s_dec):
         """
             Length of Hanning window currently chosen as 2^k samples for

### 2022-02-22 main.py

@@ -348,6 +348,14 @@ Y_pc_v_m_n, Y_mf_auto_m, Y_tilde_pc_v_m_n, svf_range \
         y_pc_s_n, w_tilde_i, y_mf_auto_n, N_w, n_f_points, f0, f1, f_s_dec,
         r_c_n, step)
 
+# Calculate the power
+P_rx_e_t_m_n = EK80CalculationPaper.calcPowerFreqforSv(
+    Y_tilde_pc_v_m_n, N_u, z_rx_e, z_td_e)
+
+# Calculate the Sv(f)
+# Sv_m = EK80CalculationPaper.calcSvf(P_rx_e_t_m, alpha_m, p_tx_e, lambda_m, t_w, Psi_f, g_0_m)
+
+
 fig, axs = plt.subplots(2)
 axs[0].plot(t_w, w_tilde_i)
 axs[0].set_ylabel('w_tilde_i')
@@ -356,8 +364,3 @@ axs[1].set_ylabel('w_tilde_i')
 axs[1].set_xlabel('t (s)')
 plt.savefig('./Paper/Fig_Svf.png')
 
-# Calculate the power
-# P_rx_e_t_m = EK80CalculationPaper.calcPowerFreqforSv(Y_tilde_pc_v_m, N_u, z_rx_e, z_td_e)
-
-# Calculate the Sv(f)
-# Sv_m = EK80CalculationPaper.calcSvf(P_rx_e_t_m, alpha_m, p_tx_e, lambda_m, t_w, Psi_f, g_0_m)

### 2022-02-22 Core/EK80CalculationPaper.py
Added function for Svf, code runs but numbers are not tested.

@@ -169,7 +169,7 @@ class EK80CalculationPaper(EK80DataContainer):
 
         return FFTvecin[idx]
 
-    def calcSvf(self, y_pc, r0, r1, overlap=0.5):
+    def calcSvf_old(self, y_pc, r0, r1, overlap=0.5):
 
         r, dr = self.calcRange()
         yspread = y_pc * r
@@ -626,6 +626,25 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         
         return P_rx_e_v_m_n
 
+    def calcSvf(P_rx_e_t_m_n, alpha_m, p_tx_e, lambda_m, t_w,
+                Psi_f, g_0_m, c, svf_range):
+        
+        # Initialize list of Svf by range
+        Sv_m_n = []
+
+        G = (p_tx_e * lambda_m ** 2 * c * t_w * Psi_f * g_0_m ** 2)/(
+            32 * np.pi ** 2)
+        n = 0
+        # Loop over list of power values along range
+        for P_rx_e_t_m in P_rx_e_t_m_n:
+            Sv_m = 10 * np.log10(
+                P_rx_e_t_m) + 2 * alpha_m * svf_range[n] - 10 * np.log10(G)
+            # Append power to list
+            n += 1
+            Sv_m_n.append(Sv_m)
+        
+        return Sv_m_n
+
     @staticmethod
     def defHanningWindow(c, tau, dr, f_s_dec):
         """
@@ -636,12 +655,13 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         
         N_w = int(2 ** np.ceil(np.log2(L)))
         # or : N_w = np.ceil(2 ** np.log2(L)) - Length of Hanning window
-        t_w = np.arange(0, N_w) * f_s_dec
+        t_w_n = np.arange(0, N_w) * f_s_dec
+        t_w = N_w * f_s_dec
         
         w_i = EK80CalculationPaper.hann(N_w)
         w_tilde_i = w_i / (np.linalg.norm(w_i) / np.sqrt(N_w))
         
-        return w_tilde_i, N_w, t_w
+        return w_tilde_i, N_w, t_w, t_w_n
 
     
     def calculateCSvfdB(self, f):

### 2022-02-22 main.py

@@ -25,6 +25,8 @@ f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
     angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
+# Ruben: is G_fnom the same as g_0_m?
+g_0_m = G_fnom
 
 c, alpha, temperature, salinity, \
     acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
@@ -308,13 +310,9 @@ plt.savefig('./Paper/Fig_TS.png')
 # Chapter IV: VOLUME BACKSCATTERING STRENGTH
 #
 
-# <Ruben>: Variables that we need:
-#
-# r_c_n : The range vector. Use r_n as dummy for now:
+# TODO: r_c_n : The range vector. Use r_n as dummy for now:
 r_c_n = r_n
 #
-#
-# </Ruben>
 
 # Denne er rekna ut i Linje 19 i dataklassen, men er flytta hit sidan
 # den er i artikkelen
@@ -336,8 +334,8 @@ y_pc_s_n = EK80CalculationPaper.calc_PulseCompSphericalSpread(y_pc_n, r_c_n)
 # TODO: This needs attention:
 dr = np.median(np.diff(r_c_n))
 
-w_tilde_i, N_w, t_w = EK80CalculationPaper.defHanningWindow(c, tau, dr,
-                                                            f_s_dec)
+w_tilde_i, N_w, t_w, t_w_n = EK80CalculationPaper.defHanningWindow(c, tau, dr,
+                                                                   f_s_dec)
 
 
 # Calculate the DFT on the pulse compressed signal
@@ -353,13 +351,19 @@ P_rx_e_t_m_n = EK80CalculationPaper.calcPowerFreqforSv(
     Y_tilde_pc_v_m_n, N_u, z_rx_e, z_td_e)
 
 # Calculate the Sv(f)
-# Sv_m = EK80CalculationPaper.calcSvf(P_rx_e_t_m, alpha_m, p_tx_e, lambda_m, t_w, Psi_f, g_0_m)
+# TODO: alpha_m and lambda_m ok length, Psi_f needs interpolation. Make
[…]

### 2022-02-22 main.py
Added comments

@@ -25,7 +25,7 @@ f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
     angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
-# Ruben: is G_fnom the same as g_0_m?
+# Ruben: is G_fnom the same as g_0_m? is g_0_m same as xml['GainNom']?
 g_0_m = G_fnom
 
 c, alpha, temperature, salinity, \
@@ -40,7 +40,10 @@ offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 # Ruben: the Psi should be the Psi_e_f_n, check paper and rewrite the data
 # class. The difference is the (f_n/f)**2 term. It should be Psi_e not Psi.
 # Check paper.
+# Psi_e_f_n - shoud be calculated here ? se eq. 28
+# xml['EquivalentBeamAngle'] . is this Psi_f_n?
 g_0_f_c, lambda_f_c, Psi_e_f_n = data.deriv.getParameters()
+
 
 #
 # Chapter IIB: Signal generation

### 2022-02-23 main.py
added comments

@@ -26,6 +26,7 @@ f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
 # Ruben: is G_fnom the same as g_0_m? is g_0_m same as xml['GainNom']?
+# g_0_m - vektor med m som frekvens indeks
 g_0_m = G_fnom
 
 c, alpha, temperature, salinity, \
@@ -40,8 +41,9 @@ offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 # Ruben: the Psi should be the Psi_e_f_n, check paper and rewrite the data
 # class. The difference is the (f_n/f)**2 term. It should be Psi_e not Psi.
 # Check paper.
-# Psi_e_f_n - shoud be calculated here ? se eq. 28
+# Psi_e_f_n - shoud be calculated here, in main ? se eq. 28
 # xml['EquivalentBeamAngle'] . is this Psi_f_n?
+# We need to discuss the Psi_e  - e is often us for electrical
 g_0_f_c, lambda_f_c, Psi_e_f_n = data.deriv.getParameters()
 
 

### 2022-02-23 Core/EK80CalculationPaper.py
removed dead code

@@ -668,26 +668,3 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         lf = self.lambda_f(f)
         return 10 * np.log10((self.ptx * self.c * lf ** 2) / (32.0 * np.pi * np.pi))
 
-    """
-    @staticmethod
-    def calcAbsorption(t, s, d, ph, c, f):
-        f = f / 1000
-
-        a1 = (8.86 / c) * 10 ** (0.78 * ph - 5)
-        p1 = 1
-        f1 = 2.8 * (s / 35) ** 0.5 * 10 ** (4 - 1245 / (t + 273))
-
-        a2 = 21.44 * (s / c) * (1 + 0.025 * t)
-        p2 = 1 - 1.37e-4 * d + 6.62e-9 * d ** 2
-        f2 = 8.17 * 10 ** (8 - 1990 / (t + 273)) / (1 + 0.0018 * (s - 35))
-
-        p3 = 1 - 3.83e-5 * d + 4.9e-10 * d ** 2
-
-        a3l = 4.937e-4 - 2.59e-5 * t + 9.11e-7 * t ** 2 - 1.5e-8 * t ** 3
-        a3h = 3.964e-4 - 1.146e-5 * t + 1.45e-7 * t ** 2 - 6.5e-10 * t ** 3
-        a3 = a3l * (t <= 20) + a3h * (t > 20)
-
-        a = f ** 2 * (a1 * p1 * f1 / (f1 ** 2 + f ** 2) + a2 * p2 * f2 / (f2 ** 2 + f ** 2) + a3 * p3)
-
-        return a / 1000
-    """

### 2022-02-23 Core/EK80CalculationPaper.py
removed dead code

@@ -98,13 +98,6 @@ class EK80CalculationPaper(EK80DataContainer):
         
         return C1Prx * np.abs(y_pc) ** 2
 
-    """
-    @staticmethod
-    def calcRange(sampleInterval, sampleCount, c, offset):
-        dr = sampleInterval * c * 0.5
-        r = np.array([(offset + i + 1) * dr for i in range(0, sampleCount)])
-        return r, dr
-    """
     @staticmethod
     def calcSp(
             p_rx_e_n,
@@ -405,22 +398,6 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
             # Calibrated case
             return np.interp(f, self.frequencies, self.gain)
 
-    """
-    def calc_B_theta_phi_m(self, theta, phi, f):
-        angle_offset_alongship_m, angle_offset_athwartship_m = self.calc_angle_offsets_m(f)
-        beam_width_alongship_m, beam_width_athwartship_m = self.calc_beam_widths_m(f)
-
-        B_theta_phi_m = 0.5 * 6.0206 * ((np.abs(theta - angle_offset_alongship_m) / (beam_width_alongship_m / 2)) ** 2 + \
-                                        (np.abs(phi - angle_offset_athwartship_m) / (
-                                                    beam_width_athwartship_m / 2)) ** 2 - \
-                                        0.18 * ((np.abs(theta - angle_offset_alongship_m) / (
-                            beam_width_alongship_m / 2)) ** 2 * \
-                                                (np.abs(phi - angle_offset_athwartship_m) / (
-                                                            beam_width_athwartship_m / 2)) ** 2))
-
-        return B_theta_phi_m
-    """
-
     def calc_g0_m(self, f):
         if self.isCalibrated:
             # Calibrated case
@@ -429,29 +406,7 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
             # Uncalibrated case
             return self.G_fnom + 20 * np.log10(f / self.fnom)
 
-    """
-    def calc_angle_offsets_m(self, f):
-        if self.isCalibrated:
-            # Calibrated case
-            angle_offset_alongship_m = np.interp(f, self.frequencies, self.angle_offset_alongship)
-            angle_offset_athwartship_m = np.interp(f, self.frequencies, self.angle_offset_athwartship)
-        else:
-            # Uncalibrated case
-            angle_offset_alongship_m = self.angle_offset_alongship_fnom * np.ones(len(f))
[…]

### 2022-02-23 Core/EK80DataContainer.py
Range calculation. zero is first range

@@ -225,7 +225,7 @@ class EK80DataContainer:
     @staticmethod
     def  calcRange(sampleInterval, sampleCount, c, offset) :
         dr = sampleInterval * c * 0.5
-        r = np.array([(offset + i + 1) * dr for i in range(0, sampleCount)])
+        r = np.array([(offset + i) * dr for i in range(0, sampleCount)])
         return r, dr
 
     def calc_angle_offsets_m(self, f):

### 2022-02-23 Core/EK80CalculationPaper.py
Gjennomgang med Lars. Lagt inn kommentarar.

@@ -307,7 +307,7 @@ class EK80CalculationPaper(EK80DataContainer):
         N_DFT = int(2 ** np.ceil(np.log2(n_f_points)))
         # Corresponding frequency vector
         
-        f_m_t = np.linspace(f0, f1, n_f_points)
+        f_m_t = np.linspace(f0, f1, n_f_points) # Ta ut denne og legg til f_m_t i argument
         # Y_pc_t_m = self.freqtransf(_Y_pc_t_m, self.f_s_dec, f_m)
         # def freqtransf(FFTvecin, fsdec, fvec=None):
         idxtmp = np.floor(f_m_t / f_s_dec * N_DFT).astype('int')
@@ -521,8 +521,8 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         return 20.0 * np.log10(r) + 2.0 * alpha * r
 
     @staticmethod
-    def calc_Psi_f(Psi_e_f_n, f_n, f_m_t):
-        return Psi_e_f_n * (f_n / f_m_t)**2
+    def calc_Psi_f(Psi_f_n, f_n, f_m_t):
+        return Psi_f_n * (f_n / f_m_t)**2
 
     @staticmethod
     def calc_Sv(p_rx_e_n, r_c_n, lambda_f_c,

### 2022-02-23 main.py

@@ -26,7 +26,8 @@ f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
 # Ruben: is G_fnom the same as g_0_m?
-g_0_m = G_fnom
+# g_0_f_c # Senterverdi chirp, det er denne som er brukt
+# G_fnom # Senterverdi av svingar
 
 c, alpha, temperature, salinity, \
     acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
@@ -40,7 +41,7 @@ offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 # Ruben: the Psi should be the Psi_e_f_n, check paper and rewrite the data
 # class. The difference is the (f_n/f)**2 term. It should be Psi_e not Psi.
 # Check paper.
-g_0_f_c, lambda_f_c, Psi_e_f_n = data.deriv.getParameters()
+g_0_f_c, lambda_f_c, Psi_f_n = data.deriv.getParameters()
 
 #
 # Chapter IIB: Signal generation
@@ -310,29 +311,24 @@ plt.savefig('./Paper/Fig_TS.png')
 # Chapter IV: VOLUME BACKSCATTERING STRENGTH
 #
 
-# TODO: r_c_n : The range vector. Use r_n as dummy for now:
-r_c_n = r_n
-#
 
-# Denne er rekna ut i Linje 19 i dataklassen, men er flytta hit sidan
-# den er i artikkelen
-# TBD: Should we change 'frequencies'? The paper uses 'f'.
-Psi_f = EK80CalculationPaper.calc_Psi_f(Psi_e_f_n, f_n, frequencies)
+# Calculate Psi
+Psi_f_c = EK80CalculationPaper.calc_Psi_f(Psi_f_n, f_n, f_c)
 
 # Calculate average Sv
 # TODO: I get zero power in the p_rx_e_n. Fails when doing log10. "Quickfix":
 p_rx_e_n = p_rx_e_n + .0000000000000001
 
-Sv_n = EK80CalculationPaper.calc_Sv(p_rx_e_n, r_c_n, lambda_f_c,
+Sv_n = EK80CalculationPaper.calc_Sv(p_rx_e_n, r_n, lambda_f_c,
                                     p_tx_e, alpha_f_c, c, tau_eff,
[…]

### 2022-02-23 Core/EK80CalculationPaper.py
Moved the frequency vector f_m to the initial part of main.

@@ -294,16 +294,15 @@ class EK80CalculationPaper(EK80DataContainer):
         return r_t, theta_t, phi_t, y_pc_t, p_rx_t, dum_theta, dum_phi, dum_r 
 
     @staticmethod
-    def calcDFTforTS(y_pc_t_n, y_mf_auto_red_n, n_f_points, f0, f1, f_s_dec):
+    def calcDFTforTS(y_pc_t_n, y_mf_auto_red_n, n_f_points, f_m, f_s_dec):
 
         # The number of DFT points inpower of 2
         N_DFT = int(2 ** np.ceil(np.log2(n_f_points)))
         # Corresponding frequency vector
         
-        f_m_t = np.linspace(f0, f1, n_f_points) # Ta ut denne og legg til f_m_t i argument
         # Y_pc_t_m = self.freqtransf(_Y_pc_t_m, self.f_s_dec, f_m)
         # def freqtransf(FFTvecin, fsdec, fvec=None):
-        idxtmp = np.floor(f_m_t / f_s_dec * N_DFT).astype('int')
+        idxtmp = np.floor(f_m / f_s_dec * N_DFT).astype('int')
         idx = np.mod(idxtmp, N_DFT) + 1
 
         # DFT for the target signal
@@ -317,7 +316,7 @@ class EK80CalculationPaper(EK80DataContainer):
         # The Normalized DFT
         Y_tilde_pc_t_m = Y_pc_t_m/Y_mf_auto_red_m
         
-        return Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, f_m_t
+        return Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m
     
     @staticmethod
     def calcPowerFreq(N_u, Y_tilde_pc_t_m, z_td_e, z_rx_e):
@@ -476,8 +475,8 @@ TS_m = 10 * np.log10(P_rx_e_t_m) + \
         return 20.0 * np.log10(r) + 2.0 * alpha * r
 
     @staticmethod
-    def calc_Psi_f(Psi_f_n, f_n, f_m_t):
-        return Psi_f_n * (f_n / f_m_t)**2
+    def calc_Psi_f(Psi_f_n, f_n, f_m):
+        return Psi_f_n * (f_n / f_m)**2
 
     @staticmethod
     def calc_Sv(p_rx_e_n, r_c_n, lambda_f_c,
[…]

### 2022-02-23 main.py

@@ -6,10 +6,6 @@ from Core.EK80DataContainer import EK80DataContainer
 # Test data
 data = EK80DataContainer('./data/pyEcholabEK80data.json')
 
-
-# data.trdu.f_c
-# data.f_c -> I paper: $f_c$
-
 # Unpack variabels
 z_td_e, f_s, n_f_points = data.cont.getParameters()
 
@@ -17,18 +13,13 @@ z_rx_e = data.trcv.getParameters()
 
 f_0, f_1, f_c, tau, slope, sampleInterval, p_tx_e = data.parm.getParameters()
 
-# To be refactored:
-f0 = f_0
-f1 = f_1
+# The frequency vector for both Ts and Sv (grid for index m)
+f_m = np.linspace(f_0, f_1, n_f_points)
 
 f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
     angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
-
-# g_0_f_c # Gain senterverdi chirp, det er denne som er brukt
-# G_fnom # Gain senterverdi av svingar
-g_0_m = G_fnom
 
 c, alpha, temperature, salinity, \
     acidity, latitude, depth, dropKeelOffset = data.envr.getParameters()
@@ -39,18 +30,16 @@ filter_v, N_v = data.filt.getParameters()
 
 offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 
-# Ruben: the Psi should be the Psi_e_f_n, check paper and rewrite the data
-# class. The difference is the (f_n/f)**2 term. It should be Psi_e not Psi.
-# Check paper.
-<<<<<<< HEAD
-g_0_f_c, lambda_f_c, Psi_f_n = data.deriv.getParameters()
-=======
-# Psi_e_f_n - shoud be calculated here, in main ? se eq. 28
-# xml['EquivalentBeamAngle'] . is this Psi_f_n?
-# We need to discuss the Psi_e  - e is often us for electrical
[…]

### 2022-02-23 Core/EK80DataContainer.py
Revised the Psi calculations

@@ -69,7 +69,7 @@ class Transducer:
     def __init__(self, xml):
         self.f_n = xml['Frequency']  # nominal design frequency for the transducer
         self.G_fnom = xml['GainNom']
-        Psi_f_c = xml['EquivalentBeamAngle']
+        Psi_f_c = xml['EquivalentBeamAngle']  # TODO: I think this is the Psi_f_n and not Psi_f_c. Right?
         self.Psi_f_c = 10 ** (Psi_f_c/10) # Make linear
         self.angle_offset_alongship_fnom = xml['AngleOffsetAlongship']
         self.angle_offset_athwartship_fnom = xml['AngleOffsetAthwartship']

### 2022-02-23 main.py

@@ -16,7 +16,8 @@ f_0, f_1, f_c, tau, slope, sampleInterval, p_tx_e = data.parm.getParameters()
 # The frequency vector for both Ts and Sv (grid for index m)
 f_m = np.linspace(f_0, f_1, n_f_points)
 
-f_n, G_fnom, Psi_f_c, angle_offset_alongship_fnom, \
+# TODO: Double check that it is Psi_f_n and no Psi_f_c
+f_n, G_fnom, Psi_f_n, angle_offset_alongship_fnom, \
     angle_offset_athwartship_fnom, angle_sensitivity_alongship_fnom, \
     angle_sensitivity_athwartship_fnom, beam_width_alongship_fnom, \
     beam_width_alongship_fnom, corrSa = data.trdu.getParameters()
@@ -33,13 +34,18 @@ offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 # Ruben: the Psi should be the Psi_f_n, i.e. the Psi at the nominal frequency
 # for the transducer. The function EK80CalculationPaper.calc_Psi_f will convert
 # to Psi_f_c, which is the center freq for the chirp. g and lambda are already
-# converted to f_c. We also need Psi_m on the f grid.
+# converted to f_c.
 
 g_0_f_c, lambda_f_c, Psi_m = data.deriv.getParameters()
-# xml['EquivalentBeamAngle'] . is this Psi_f_n?
-# we need Psi_m, which is on the f_m grid
-# and lamdba_m on the same grid
-Psi_f_n = Psi_m[10]  # TODO: Fix this.
+
+# Cacluate lambda and alpha on the f_m grid
+lambda_m = data.calc_lambda_f(f_m)
+alpha_m = data.calc_alpha_f(f_m)
+
+# Calculate Psi for f_c and on the f_m grid
+Psi_f_c = EK80CalculationPaper.calc_Psi_f(Psi_f_n, f_n, f_c)
+# TODO: double check this. I think it is ok:
+Psi_m = EK80CalculationPaper.calc_Psi_f(Psi_f_n, f_n, f_m)
 
 #
 # Chapter IIB: Signal generation
@@ -282,8 +288,6 @@ G0_m = data.calc_G0_m(f_m)
 G_theta_phi_m = G0_m - g_theta_t_phi_t_f_t
 """
[…]

### 2022-02-24 Paper/Manuscript.tex
Gjennomgang med Lars. Lagt inn kommentarar.
Fixed on axis gain [g(0,0,f_m)]

--- 2022-02-22 Paper/Manuscript.tex
@@
 \ts(\freqsym) = 10\log_{10}(\prxetf(\freqsym)) + 40\log_{10}(\range) + 2\absorp(\freqsym) \range 
-- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,\freqsym)}{16\pi^2} \right),
+- 10\log_{10}\left( \frac{\ptxe \wlen^2(\freqsym) \gain^2(\along_t,\athw_t,\freqsym)}{16\pi^2} \right),
 \end{equation}
@@
 
+Comment from Nils Olav: Using $m$ as an argument to a function is not ideal.
+If we use the function argument perhaps using $f_m$ would be more logical, i.e.
+$\lambda(f_m)$.
+Alternatively using $lambda_m$ is possible.
+The latter is perhaps more in line with the code since the parameter is an array windexed with $m$ and not a function.
+$Psi(f)$, on the the other hand is derived from a function, see below.
+
 \section{Volume backscattering strength}
@@
 \sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range_c \\
-- 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
+- 10\log_{10}\left( \frac{\ptxe \wlen^2(\samplesymf) \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
 \end{equation}

### 2022-02-24 main.py

@@ -36,7 +36,13 @@ offset, sampleCount, y_rx_nu, N_u, y_rx_nu = data.raw3.getParameters()
 # to Psi_f_c, which is the center freq for the chirp. g and lambda are already
 # converted to f_c.
 
-g_0_f_c, lambda_f_c, Psi_m = data.deriv.getParameters()
+g_0_f_c, lambda_f_c, _ = data.deriv.getParameters()
+
+# TODO: Consider to use _f_m instead of _m, depending on the
+# way it will be written in the paper, for the following parameters:
+
+# The on axix gain as a function of f_m
+g_0_m = 1
 
 # Cacluate lambda and alpha on the f_m grid
 lambda_m = data.calc_lambda_f(f_m)
@@ -318,6 +324,8 @@ plt.savefig('./Paper/Fig_TS.png')
 # TODO: I get zero power in the p_rx_e_n. Fails when doing log10. "Quickfix":
 p_rx_e_n = p_rx_e_n + .0000000000000001
 
+# TODO: Range equal to zero will not work. either remove first sample or
+# reconsider the range vector (log10(0) does not exist)
 Sv_n = EK80CalculationPaper.calc_Sv(p_rx_e_n, r_n, lambda_f_c,
                                     p_tx_e, alpha_f_c, c, tau_eff,
                                     Psi_f_c, g_0_f_c)
@@ -347,14 +355,9 @@ P_rx_e_t_m_n = EK80CalculationPaper.calcPowerFreqforSv(
     Y_tilde_pc_v_m_n, N_u, z_rx_e, z_td_e)
 
 # Calculate the Sv(f)
-# TODO: alpha_m and lambda_m ok length, Psi_f needs interpolation. Make
-# it scalar for now (it is also negative???, change sign to make the code run):
-Psi_f_dum = 1  # TODO: Fix this
-g_0_m = 1
-
 Sv_m_n = EK80CalculationPaper.calcSvf(P_rx_e_t_m_n,
                                       alpha_m, p_tx_e, lambda_m, t_w,
-                                      Psi_f_dum, g_0_m, c, svf_range)
+                                      Psi_m, g_0_m, c, svf_range)
         
 
 fig, axs = plt.subplots(2)

### 2022-02-24 Core/EK80CalculationPaper.py
Added code for calculating g_0_m and made a hack for out of bounds for the new data.

@@ -160,6 +160,10 @@ class EK80CalculationPaper(EK80DataContainer):
         idxtmp = np.floor(fvec / fsdec * nfft).astype('int')
         idx = np.mod(idxtmp, nfft) + 1
 
+        # TODO: Hack to avoid out of bounds. Same as for DFT for TS.
+        idx[idx == nfft] = idx[idx == nfft] - 1
+
+
         return FFTvecin[idx]
 
     def calcSvf_old(self, y_pc, r0, r1, overlap=0.5):
@@ -303,7 +307,19 @@ class EK80CalculationPaper(EK80DataContainer):
         # Y_pc_t_m = self.freqtransf(_Y_pc_t_m, self.f_s_dec, f_m)
         # def freqtransf(FFTvecin, fsdec, fvec=None):
         idxtmp = np.floor(f_m / f_s_dec * N_DFT).astype('int')
+        
+        # TODO: Nils Olav comment: If we add +1 we will make an error
+        # in cases where we wrap the frequencies (?). The IDX will then
+        # go from 1024 to 1 instead of 1023 to 1 Since 1024 is out of bounds
+        # The index 0 is missed which I suppose is the DC term and OK?
+        # We are clearly doing something wrong here. There
+        # are also duplicates in idx indicating nearest neighbour imputation.
+        # This needs some thought.
+       
         idx = np.mod(idxtmp, N_DFT) + 1
+        # TODO: Hack to make the code run (do -1 on the index that is out
+        # of bounds)
+        idx[idx == N_DFT] = idx[idx == N_DFT] - 1
 
         # DFT for the target signal
         _Y_pc_t_m = np.fft.fft(y_pc_t_n, n=N_DFT)

### 2022-02-24 main.py

@@ -3,8 +3,13 @@ import numpy as np
 from Core.EK80CalculationPaper import EK80CalculationPaper
 from Core.EK80DataContainer import EK80DataContainer
 
+#
 # Test data
+#
+
 data = EK80DataContainer('./data/pyEcholabEK80data.json')
+# data = EK80DataContainer('./data/CRIMAC_SphereCentre.json')  # Ts sphere
+# data = EK80DataContainer('./data/CRIMAC_Svf.json')  # School
 
 # Unpack variabels
 z_td_e, f_s, n_f_points = data.cont.getParameters()
@@ -42,7 +47,7 @@ g_0_f_c, lambda_f_c, _ = data.deriv.getParameters()
 # way it will be written in the paper, for the following parameters:
 
 # The on axix gain as a function of f_m
-g_0_m = 1
+g_0_m = data.calc_g(0, 0, f_m)
 
 # Cacluate lambda and alpha on the f_m grid
 lambda_m = data.calc_lambda_f(f_m)

### 2022-03-16 Paper/Manuscript.tex
Rewritten introduction in Manuscript.tex v01

@@
 
+\author{Geir Pedersen}
 \author{Rolf Korneliussen}
@@
 \author{Egil Ona}
+\author{Nils Olav Handegard}
 \affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
@@
 
-Management of fisheries resources typically requires knowledge of age and abundance structure over many years.
-For numerous fish species and stocks, abundance estimates can be obtained by spatially extensive surveys using quantitative echosounder systems, which measure the backscatter from fish \citep{Simmonds2005Fisheries}.
-An important property of scientific echosounders used for fisheries surveys is accurate measurement of the backscatter amplitude - this is of lesser importance for other echosounder applications, such as bathymetry, presence/absence of objects, and spatial properties of objects.
-Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most) and when several are operated simultaneously at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help to categorize the backscatter into species or target categories.
-This is termed the multi-frequency approach \citep{Korneliussen2002operational, holliday1980}.
+Active acoustic remote sensing is an essential method for marine ecosystem research and monitoring.
+Echosounders provide unique information on marine organisms on a wide range of spatial and temporal scales, and on ecosystem components from plankton to large predators.
 
-The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to fisheries applications is an obvious enhancement.
-Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}.
-These benefits can lead to improved backscatter categorisation \citep{Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic, korneliussen2018} and hence more accurate abundance estimates.
+Management of fisheries resources have successfully employed quantitative single-beam echosounders for abundance estimates for decades, obtained through spatially extensive surveys mapping the distribution and abundance of species of fishes (Simmonds and MacLennan, 2005).
+Echosounders are increasingly used for the study of not only the distribution and abundance of commercially important fishes, but also the distribution of other trophic levels, and interactions among ecosystem components and drivers at the level of population, foodweb and ecosystem (Godø et al., 2014).
 
-There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
-None of these systems have achieved widespread use in quantitative fisheries surveys presumably because the commercially-available systems have lacked one or more of the features that are very useful for an operational quantitative fisheries survey echosounder.
-These comprise (a) a large dynamic range receiver that does not saturate in normal conditions; (b) operating frequencies that are useful for detecting fish at several hundred metres range (between approximately 20 and 200 kHz); (c) split-aperture for detection of angle of arrival of echoes from individual organisms and calibration spheres; (d) capability to perform in-situ calibrations of the amplitude response; (e) a relatively easy transition from existing survey echosounders; (f) useful at typical ship survey speeds (e.g., 5 m s \textsuperscript{-1}).
-A recent addition to the set of commercially-available broadband echosounders is the Simrad EK80, which meets all of these features.
+Single-beam echosounders have provided important insight in studies of predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
 
-The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals.
-Without loss of generality, we use the \ek\ as an example since it is currently the most commonly used broadband echosounder in the fisheries acoustics field.
-By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in fisheries acoustics.
+Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help in the categorizing the backscatter into species or functional groups.
+This is termed the multi-frequency approach (Holliday and Pieper, 1980; Korneliussen and Ona, 2002).
+
+The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to applications in ecology is an obvious enhancement.
+Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
+These benefits can lead to improved backscatter categorization (R.
+Korneliussen, 2018; R.
+J.
+Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning (see e.g.
+Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).
+
+There have been several scientific broadband echosounder systems developed for laboratory use (Chu et al., 1992; Conti and Demer, 2003; Forland et al., 2014), some prototype or custom-made systems (Barr et al., 2002; Brise~no-Avena et al., 2015; Foote et al., 2005; Imaizumi et al., 2009; Simmonds et al., 1996; Zakharia et al., 1989, 1996) and some commercially available systems (Denny and Simpson, 1998; Ehrenberg and Torkelson, 2000; Gordon and Zedel, 1998; Stanton et al., 2010; Zedel et al., 2003).
+
+The increasing availability of commercial systems specifically for ecosystem acoustics will vastly increase the use of broadband pulses in aquatic ecology, as these systems are more user-friendly  and with capabilities which ease the transition from existing narrowband scientific echosounders (e.g.
+quantitative data output and a familiar data format supported by the major software suites for marine ecosystem acoustics).
+
+The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals with associated code, and to provide a code base for further development and research in broadband acoustic ecology.
+
+\section{Signal flow and initial processing}
+
+\subsection{Accompanying code}
+The code accompanying this paper is written in Python and is available on GitHub.
+All processing steps with respective figures in the paper can be reproduced by running main script (Main.py).
+Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field.
+By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.
 
@@
 In particular, the expressions for target strength ($\ts$) and volume backscattering strength ($\sv$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.
-
-\section{Signal flow and initial processing}
 

### 2022-03-16 Paper/Manuscript.tex
Updated introduction

@@
 
+\author{Geir Pedersen}
+\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
+
 \author{Dezhang Chu}
@@
 
-Active acoustic remote sensing is an essential method for marine ecosystem research and monitoring.
-Echosounders provide unique information on marine organisms on a wide range of spatial and temporal scales, and on ecosystem components from plankton to large predators.
+Active acoustics is an efficient method for remote sensing of marine ecosystems.
+These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km (Makris) to small scale behaviour patterns (refs).
+These methods are also a crucial component of fisheries resource management.
+Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperature sonar, omnidirectional sonars, acoustic imaging sonars as well as echosounders (refs).
 
-Management of fisheries resources have successfully employed quantitative single-beam echosounders for abundance estimates for decades, obtained through spatially extensive surveys mapping the distribution and abundance of species of fishes (Simmonds and MacLennan, 2005).
-Echosounders are increasingly used for the study of not only the distribution and abundance of commercially important fishes, but also the distribution of other trophic levels, and interactions among ecosystem components and drivers at the level of population, foodweb and ecosystem (Godø et al., 2014).
+Single- beam echosounders have extensively been used in fisheries management and ecosystem monitoring.
+Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful.
+Single-beam echosounders have provided important information for fisheries resource management (Simmonds and MacLennan, 2005) as well as insight in ecological processes for the aquatic ecosystem (Godø 2014).
 
-Single-beam echosounders have provided important insight in studies of predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
+Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz).
+The method has been used to  study predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009; Godø...), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
 
-Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help in the categorizing the backscatter into species or functional groups.
-This is termed the multi-frequency approach (Holliday and Pieper, 1980; Korneliussen and Ona, 2002).
-
-The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to applications in ecology is an obvious enhancement.
-Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
+Recent echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems, i.e.
+a continuous frequency coverage over a wide frequency band.
+provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
 These benefits can lead to improved backscatter categorization (R.
@@
 J.
-Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning (see e.g.
+Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning
+(skriv om: e.g.
 Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).
@@
 
-The increasing availability of commercial systems specifically for ecosystem acoustics will vastly increase the use of broadband pulses in aquatic ecology, as these systems are more user-friendly  and with capabilities which ease the transition from existing narrowband scientific echosounders (e.g.
-quantitative data output and a familiar data format supported by the major software suites for marine ecosystem acoustics).
-
-The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals with associated code, and to provide a code base for further development and research in broadband acoustic ecology.
+To tap the full potential of broadband echosounder data, further methods for processing such data must be developed.
+Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
+Different send pulses can be envisioned as well as various methods for acoustic target classification.
+The objective of this paper is to present a systematic and comprehensive explanation of how to process calibrated broadband echosounder data.
+By making associated code available, we provide a starting point for further developing signal processing methods for applying broadband acoustic to aquatic ecology.
 
@@
 \subsection{Accompanying code}
-The code accompanying this paper is written in Python and is available on GitHub.
+The code accompanying this paper is written in Python and is available as supplementary materials.
 All processing steps with respective figures in the paper can be reproduced by running main script (Main.py).
@@
 \section{Discussion}
-Obtaining quantitative broadband data is more complicated than for narrowband echosounders, due in part to the need to account for frequency dependence in most variables and parameters, and the need for an increased understanding of digital signal processing techniques.
-For example, the \ek{} currently provides complex demodulated \citep{hasan1983} voltages for each transducer sector rather than derived quantities (e.g., the envelope of a pulse compressed signal and the frequency response of single targets and volumes) - these outputs must instead be calculated as proposed in this paper.
-This was an implementation decision made to allow for more flexible use of the echosounder data, and to ease the development of new processing methodologies.
-This decision has several disadvantages, such as the significantly increased data quantity and markedly higher amount of computation required to simply display an echogram.
-These can be ameliorated to some degree by processing the data into a more directly useful form before storage.
-The advantages of having access to unprocessed data were considered to out-weigh these disadvantages due to the potential benefits of more sophisticated uses of the acoustic data, especially for a tool that is only beginning to be applied to the field of fisheries acosutics.
-
-The methodology to process broadband data has been presented in a general form in the previous sections without any accord to engineering limits.
-However, any physical implementation introduces operational constraints.
-As an example, a system designed for shipboard installations with ample access to electrical and computer processing power and large data storage will typically have different operational contraints compared to a system intended for autonomous platforms where electrical power and computing resources can be severely limited.
-In addition, the transducer is a significant constraint on the operational parameters of an echosounder and is usually the main determinant of the usable transceiver operating parameters (such as transmit bandwidth, maximum transmit power, pulse duration, ping rate, etc).
-
 The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
@@
 The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
+
+Modern machine learning techniques are well suited for further processing of broadband echosounder data.
+Deep learning algorithms, including deep clustering, provide powerful methods to tease out patterns from high dimensional and complex data.
+By providing similar step sizes across frequencies it is possible to adapt this code to for an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
 

### 2022-03-16 Paper/Manuscript.tex
Updated plots

@@
 
-To estimate $\sv$ as a function of frequency a Fourier transform is used, repeatedly applied via a sliding window in range.
-However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r_c^2$, implemented as the $20\log_{10}(\range_c)$ term in Eq.
-\ref{eq:Sv}) from the beginning of the window to the end can be significant, particularly for short range measurements.
-Thus, compensation for spreading loss is performed before applying the discrete Fourier transform.
-Absorption loss compensation is also range dependent (and frequency dependent), but since absorption loss compensation is insignificant for typical operating frequencies at short ranges and the difference in absorption loss compensation between the beginning and the end of the sliding window is insignificant at longer ranges, compensation for absorption loss is performed after applying the discrete Fourier transform.
-
 Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
@@
 \end{equation}
+
 %
@@
 \end{equation}
+
+By rep pres of a set of centre samples n Sv values can be presented as a functon of range and frequencty for each ping
+Teh range fo sentre samples n could be chosen to halv the window length or any other grid that the user prefer the data presented to be in.
+This can be usefull when combining the Sv(f) across a range of transducers.
+In thode our ex we have simply chosen n to be on the same order of the range resoluteion (fig.~\ref{Fig_Sv_m_n}).
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_Sv_m_n}
+\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($m$) and range ($n$) for a single ping}
+\end{figure}
+
+For echo integration purposes it is common to integrate the Sv over a range, in our example we integrate the Sv from x-x m covering a school of non-swimbladdered fish (fig).
+We note that we normally require avg over several indivvidal pils to uptain an unbiased estimate of Sv for echo integration purposes.
+Here we present te sv(f) over the layer for our example ping for illustrative purposesn (fig.~\ref{Fig_Sv_m}).
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_Sv_m}
+\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($m$) averaged over a range intervall covering a fish school }
+\end{figure}
 
@@
 The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
+
+In this implementation (ref code) several choices have been made.
+Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature, where other approaches could been equally valid.
+These include, as examples,
+Transmit pulse
+Tapering of the transmit pulse
+Decimation factors and filters
+When Convoling the recieved signal with the transmit pulse - edge cases
+Assuemd four sector transducer
+Matched load - hvor stor effekt å gå utover 
+Assumed an impedance (trancuder)
+Effective pulse duration...
+Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...
+Window siz
+Edge cases - removed
+
+To estimate $\sv$ as a function of frequency a Fourier transform is used, repeatedly applied via a sliding window in range.
+However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r_c^2$, implemented as the $20\log_{10}(\range_c)$ term in Eq.
+\ref{eq:Sv}) from the beginning of the window to the end can be significant, particularly for short range measurements.
+Thus, compensation for spreading loss is performed before applying the discrete Fourier transform.
+Absorption loss compensation is also range dependent (and frequency dependent), but since absorption loss compensation is insignificant for typical operating frequencies at short ranges and the difference in absorption loss compensation between the beginning and the end of the sliding window is insignificant at longer ranges, compensation for absorption loss is performed after applying the discrete Fourier transform.
+
+...
 

### 2022-03-16 plots.py

@@ -117,13 +117,14 @@ def plot_TS(f_m,Y_pc_t_m,Y_mf_auto_red_m,Y_tilde_pc_t_m,g_theta_phi_m,TS_m):
 def plotSvf(f_m,Sv_m_n,svf_range):
     plt.figure()
     _f = f_m / 1000
-    plt.imshow(Sv_m_n, extent=[_f[0], _f[-1], svf_range[-1], svf_range[0]], origin='upper', vmin=-180, vmax=-120,
+    plt.imshow(Sv_m_n, extent=[_f[0], _f[-1], svf_range[-1], svf_range[0]], origin='upper',
                interpolation=None)
     plt.colorbar()
     plt.title('Echogram [Sv]')
     plt.xlabel('Frequency [kHz]')
     plt.ylabel('Range [m]')
     plt.axis('auto')
+    plt.savefig('./Paper/Fig_Sv_m_n.png')
 
     plt.figure()
     # Plot Sv(f) in one depth in the middle of layer
@@ -134,6 +135,7 @@ def plotSvf(f_m,Sv_m_n,svf_range):
     plt.xlabel('Frequency [kHz]')
     plt.ylabel('Sv')
     plt.grid()
+
 
     indices = np.where(np.logical_and(svf_range >= 60, svf_range <= 70))
     # returns (array([3, 4, 5]),)
@@ -152,7 +154,7 @@ def plotSvf(f_m,Sv_m_n,svf_range):
     plt.xlabel('Frequency [kHz]')
     plt.ylabel('Range [m]')
     plt.grid()
-
+    plt.savefig('./Paper/Fig_Sv_m.png')
     # Store Sv(f) and f for further analysis
     SvfOut = np.concatenate((f_m[np.newaxis],Sv_m_n), axis=0)
     # np.save('Svf.npy',SvfOut)

### 2022-03-17 Paper/Manuscript.tex
Updates to manuscript, added figures

--- 2022-03-16 Paper/Manuscript.tex
@@
 \author{Geir Pedersen}
-\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
+\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
 
@@
 
-\author{Geir Pedersen}
 \author{Rolf Korneliussen}
@@
 \author{Egil Ona}
+\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
+
+\author{Ruben Patel}
+\affiliation{Codelab, Bergen, Norway}
+
 \author{Nils Olav Handegard}
-\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
+\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
 
@@
 
+To illustrate each step of the process to obtain $TS(f)$ we use a single ping from a data file where a single metallic calibration sphere is suspended approximately 6 m below a 120 kHz transducer (fig.~\ref{Fig_TS_echogram}).
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_TS_echogram}
+\caption{\label{Fig_TS_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file.
+The red vertical line indicates ping to be processed in this section for TS(f).
+Single target calibration sphere is seen at approximately 6 m range and the seafloor at approximately 12 m.}
+\end{figure}
+
 Echoes from single targets are often characterised by their $\ts$, which is related to the differential backscattering cross section, $\bs$, via
@@
 
+A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
+This is visible in the $\ts$ (Fig.
+\ref{fi:SED}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
+The amplitude of the backscatter signal also clearly shows these nulls (Fig.
+\ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
+
 Comment from Nils Olav: Using $m$ as an argument to a function is not ideal.
@@
 
+As a metallic sphere is a rather simple and ideal scatterer, we present $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) in the processes to obtain $Sv((f)$.
+
 Echoes from multiple scatterers can be quantified using volume backscattering strength, $\sv$, being the density of backscattering cross sections, and is given by:
@@
 
+\begin{figure}
+\includegraphics[width=16cm]{Fig_Sv_echogram}
+\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file. The red vertical line indicates ping to be processed in this section.}
+\end{figure}
+
 Volume backscattering samples compressed over the operational frequency band are estimated by applying Eq.
@@
 In this case, the time window, $\tslide$, is the effective pulse duration, $\teff$, resulting from pulse compression.
+
+Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
+%
+\begin{equation}
+\label{eq:spreadcomp}
+\ypcspread(\samplesymt) = \ypc(\samplesymt)\range_c(\samplesymt).
+\end{equation}
+%
+where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading.
+A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$.
+The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution.
+We suggest that it be at least twice the pulse duration and for computational efficiency reasons should result in a number of samples, $\nw$, which is a power of 2.
+
+TODO: Explain edge cases.
+
+The normalised Hann window, $\hannwnorm$, is given by: 
+%
+\begin{equation}
+\label{eq:hannw}
+\hannwnorm(\genidxsym) = \frac{\hannw(\genidxsym)}{\left( \frac{||\hannw||_2}{\sqrt{\nw}} \right)}, i = \frac{-\nw}{2}, \ldots, \frac{\nw}{2}
+\end{equation}
+%
+and the discrete Fourier transform of the windowed data, $\ypcvolumef(\samplesymf)$, is then obtained from:
+%
+\begin{equation}
+\label{eq:FFT_volume}
+\ypcvolumef(\samplesymf) = \dft_\ndft 
+\left( \hannwnorm(\genidxsym) \left(\ypcspread (\genidxsym+\samplesymt) \left[ u(\genidxsym + \frac{\nw}{2}) - u(\genidxsym - \frac{\nw}{2}) \right] \right) \right),
+\end{equation}
+%
+where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window.
+The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
+%
+\begin{equation}
+\label{eq:FFT_TX_Auto}
+\ymfautof(\samplesymf)  =  \dft_\ndft (\ymfauto(\samplesymt)).
+\end{equation}
+
+%
+TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.
+
+The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
+%
+\begin{equation}
+\label{eq:FFT_volume_norm}
+\ypcvolumenormf(\samplesymf) = \frac{\ypcvolumef(\samplesymf)}{\ymfautof(\samplesymf)},
+\end{equation}
+%
+and received power into a matched load, $\prxevf(\samplesymf)$, is estimated from:
+%
+\begin{equation}
+\label{eq:prx_FFT_volume}
+\prxevf(\samplesymf) = \nchannels \left( \frac{|\ypcvolumenormf(\samplesymf)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \\
+\frac{1}{|\ztde|}. % All channels (*4), matched load (/2), and effective values (/sqrt(2))
+\end{equation}
+%
+Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
+%
+\begin{equation}
+\label{eq:Sv_FFT}
+\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range_c \\
+- 10\log_{10}\left( \frac{\ptxe \wlen^2(\samplesymf) \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
+\end{equation}
+
+By selecting a set of centre samples $n$, $Sv$ values can be presented as a function of range (n) and frequency (m) for each ping.
+The range for centre samples n could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
+This can be useful when combining the Sv(f) across a range of transducers.
+In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_Sv_m_n}
+\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($m$) and range ($n$) for a single ping}
+\end{figure}
+
+For echo integration purposes it is common to integrate the Sv over a range, in our example we integrate the Sv from 15 to 34 m covering a school of non-swimbladdered fish (fig).
+We note that we normally require averaging over several individual pings to obtain an unbiased estimate of Sv for echo integration purposes.
+Here we present the sv(f) over the layer for our example ping for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
+Even though this is for a single ping, we still observe a positive slope of the frequency response indicative of non-swimbladdered fish.
+
+\begin{figure}
+\includegraphics[width=16cm]{Fig_Sv_avg}
+\caption{\label{Fig_Sv_avg} Sv as a function of frequency ($m$) averaged over a range interval covering a fish school }
+\end{figure}
+
+The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example.
+In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (fig.~\ref{Fig_Sv_echogram}).
+
+\section{Discussion}
+The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
+For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
+The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
+
+In this implementation (ref code) several choices have been made.
+Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature, where other approaches could been equally valid.
+These include, as examples,\\
+Transmit pulse\\
+Tapering of the transmit pulse\\
+Decimation factors and filters\\
+When Convoling the recieved signal with the transmit pulse - edge cases\\
+Assuemd four sector transducer\\
+Matched load - hvor stor effekt å gå utover \\
+Assumed an impedance (trancuder)\\
+Effective pulse duration...\\
+Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...\\
+Window size\\
+Edge cases - removed\\
 
@@
 
-Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
-%
-\begin{equation}
-\label{eq:spreadcomp}
-\ypcspread(\samplesymt) = \ypc(\samplesymt)\range_c(\samplesymt).
-\end{equation}
-%
-where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading.
-A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hanning window, $\hannw(\genidxsym)$.
-The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution.
-We suggest that it be at least twice the pulse duration and for computational efficiency reasons should result in a number of samples, $\nw$, which is a power of 2.
-
-TODO: Explain edge cases.
-
-The normalised Hanning window, $\hannwnorm$, is given by: 
-%
-\begin{equation}
-\label{eq:hannw}
-\hannwnorm(\genidxsym) = \frac{\hannw(\genidxsym)}{\left( \frac{||\hannw||_2}{\sqrt{\nw}} \right)}, i = \frac{-\nw}{2}, \ldots, \frac{\nw}{2}
-\end{equation}
-%
-and the discrete Fourier transform of the windowed data, $\ypcvolumef(\samplesymf)$, is then obtained from:
-%
-\begin{equation}
-\label{eq:FFT_volume}
-\ypcvolumef(\samplesymf) = \dft_\ndft 
-\left( \hannwnorm(\genidxsym) \left(\ypcspread (\genidxsym+\samplesymt) \left[ u(\genidxsym + \frac{\nw}{2}) - u(\genidxsym - \frac{\nw}{2}) \right] \right) \right),
-\end{equation}
-%
-where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window.
-The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
-%
-\begin{equation}
-\label{eq:FFT_TX_Auto}
-\ymfautof(\samplesymf)  =  \dft_\ndft (\ymfauto(\samplesymt)).
-\end{equation}
-%
-TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.
-
-The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
-%
-\begin{equation}
-\label{eq:FFT_volume_norm}
-\ypcvolumenormf(\samplesymf) = \frac{\ypcvolumef(\samplesymf)}{\ymfautof(\samplesymf)},
-\end{equation}
-%
-and received power into a matched load, $\prxevf(\samplesymf)$, is estimated from:
-%
-\begin{equation}
-\label{eq:prx_FFT_volume}
-\prxevf(\samplesymf) = \nchannels \left( \frac{|\ypcvolumenormf(\samplesymf)|}{2\sqrt{2}} \right)^2 \left( \frac{|\zrxe+\ztde|}{|\zrxe|}\right)^2 \\
-\frac{1}{|\ztde|}. % All channels (*4), matched load (/2), and effective values (/sqrt(2))
-\end{equation}
-%
-Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
-%
-\begin{equation}
-\label{eq:Sv_FFT}
-\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range_c \\
-- 10\log_{10}\left( \frac{\ptxe \wlen^2(\samplesymf) \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
-\end{equation}
-
-\section{Illustrative examples}
-
-A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
-This is visible in the $\ts$ (Fig.
-\ref{fi:ts_example}a, \ref{fi:ts_example}b) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
-The amplitude of the backscatter signal also clearly shows these nulls (Fig.
-\ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
-The marked increase in range resolution is apparent once pulse compression has been applied (Fig.
-\ref{fi:ts_example}d), as are the temporal effects of the pulse compression operation.
-
-A metallic sphere is a rather simple and ideal scatterer and we also present $\sv$ from a school of Atlantic mackerel (\textit{Scomber scombrus}) (Fig.
-\ref{fi:sv_example}a).
-The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example.
-In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.
-\ref{fi:sv_example}b, \ref{fi:sv_example}c), although in regions where the fish density decreases (e.g., top left of the school), single target echoes become visible.
-
-\begin{figure}
-\includegraphics[width=16cm]{Fig_ts_example}
-\caption{\label{fi:ts_example}(color online) Target strength of 57.2 mm (a) and 22.0 mm (b) diameter tungsten carbide (with 6\% cobalt binder) calibration spheres, theoretical (solid line) and measured and derived using Eq.
-(\ref{eq:TS_f}) (dotted line).
-Echogram of non pulse compressed backscatter [dB re 1 W] (c) and pulse compressed backscatter [dB re 1 $\textrm{m}^2$] (d) from 57.2 mm (lower band) and 22.0 mm (upper band) diameter tungsten carbide spheres using a 2.048 ms duration linear frequency modulated pulse (160--260 kHz).
-The range from the spheres to the transducer varied during the time period shown.}
-\end{figure}
-
-\begin{figure}
-%\includegraphics[width=16cm]{Figure5}
-\caption{\label{fi:sv_example}(color online) (a): Mean frequency response (solid line) and standard error (dashed line) from a school of Atlantic mackerel (\textit{Scomber scombrus}) from three simultaneously operating linear frequency modulated echosounder channels (90--170, 160--260, and 280--450 kHz).
-Echogram of backscattered power [dB re 1 W] (b) and pulse compressed $\sv$ [dB re 1 $m^{-1}$] (c) from the 160--260 kHz pulse.
-The school is the trapezoid-shaped object between 25 and 90 m below the surface.
-The echo from the seafloor is visible at 130 m depth.} \end{figure}
-
-\section{Discussion}
-The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
-For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
-The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
+...
 

### 2022-03-17 plots.py

@@ -11,9 +11,9 @@ def plot_ytx(f_0, f_1, tau, f_s,y_tx_n,slope):
 
     plt.figure()
     plt.plot(t * 1000, y_tx_n, t * 1000, y_tx_n05slope)
-    plt.title(
-        'Ideal windowed transmit pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
-            .format(f_0 / 1000, f_1 / 1000, slope))
+    #plt.title(
+    #    'Ideal windowed transmit pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
+    #        .format(f_0 / 1000, f_1 / 1000, slope))
     plt.xlabel('time (ms)')
     plt.ylabel('amplitude')
     plt.savefig('./Paper/Fig_ytx.png')
@@ -50,7 +50,7 @@ def plot_fir(filter_v,f_s_dec_v,f_0, f_1):
 def plot_y_mf_n(y_mf_n):
     plt.figure()
     plt.plot(np.abs(y_mf_n))
-    plt.title('The absolute value of the filtered and decimated output signal')
+    #plt.title('The absolute value of the filtered and decimated output signal')
     plt.xlabel('samples ()')
     plt.ylabel('amplitude')
     plt.savefig('./Paper/Fig_y_mf_n.png')
@@ -59,7 +59,7 @@ def plot_y_mf_n(y_mf_n):
 def plot_ACF(y_mf_auto_n):
     plt.figure()
     plt.plot(np.abs(y_mf_auto_n))
-    plt.title('The autocorrelation function of the matched filter.')
+    #plt.title('The autocorrelation function of the matched filter.')
     plt.xlabel('Samples')
     plt.ylabel('ACF')
     plt.savefig('./Paper/Fig_ACF.png')
@@ -70,7 +70,7 @@ def plot_theta_phi(theta_n,phi_n):
     plt.figure()
     plt.plot(theta_n)
     plt.plot(phi_n)
-    plt.title('The physical angles.')
+    #plt.title('The physical angles.')
     plt.xlabel(' ')
     plt.ylabel('Angles')
     plt.savefig('./Paper/Fig_theta_phi.png')
@@ -78,7 +78,7 @@ def plot_theta_phi(theta_n,phi_n):
 
 def plt_single_target(dum_r, dum_p,dum_theta,r_t,dum_phi,phi_t,y_mf_auto_red_n):
[…]

### 2022-03-17 plots.py
Updates to manuscript, added figures

@@ -16,7 +16,7 @@ def plot_ytx(f_0, f_1, tau, f_s,y_tx_n,slope):
     #        .format(f_0 / 1000, f_1 / 1000, slope))
     plt.xlabel('time (ms)')
     plt.ylabel('amplitude')
-    plt.savefig('./Paper/Fig_ytx.png')
+    plt.savefig('./Paper/Fig_ytx.png',dpi=300)
 
 
 def plot_fir(filter_v,f_s_dec_v,f_0, f_1):
@@ -44,7 +44,7 @@ def plot_fir(filter_v,f_s_dec_v,f_0, f_1):
     plt.xlabel('frequency (Hz)')
     plt.ylabel('Gain (dB)')
     plt.xlim([50000, 210000])
-    plt.savefig('./Paper/Fig_fir.png')
+    plt.savefig('./Paper/Fig_fir.png',dpi=300)
 
 
 def plot_y_mf_n(y_mf_n):
@@ -53,7 +53,7 @@ def plot_y_mf_n(y_mf_n):
     #plt.title('The absolute value of the filtered and decimated output signal')
     plt.xlabel('samples ()')
     plt.ylabel('amplitude')
-    plt.savefig('./Paper/Fig_y_mf_n.png')
+    plt.savefig('./Paper/Fig_y_mf_n.png',dpi=300)
 
 
 def plot_ACF(y_mf_auto_n):
@@ -62,7 +62,7 @@ def plot_ACF(y_mf_auto_n):
     #plt.title('The autocorrelation function of the matched filter.')
     plt.xlabel('Samples')
     plt.ylabel('ACF')
-    plt.savefig('./Paper/Fig_ACF.png')
+    plt.savefig('./Paper/Fig_ACF.png',dpi=300)
 
 
 def plot_theta_phi(theta_n,phi_n):
@@ -73,7 +73,7 @@ def plot_theta_phi(theta_n,phi_n):
     #plt.title('The physical angles.')
     plt.xlabel(' ')
     plt.ylabel('Angles')
-    plt.savefig('./Paper/Fig_theta_phi.png')
+    plt.savefig('./Paper/Fig_theta_phi.png',dpi=300)
 
 
 def plt_single_target(dum_r, dum_p,dum_theta,r_t,dum_phi,phi_t,y_mf_auto_red_n):
[…]

### 2022-03-18 Paper/Manuscript.tex
Small fixes to paper

@@
 \documentclass[preprint,12pt,TurnOnLineNumbers]{JASAnew}
+
+\usepackage{color,soul}
 
@@
 Active acoustics is an efficient method for remote sensing of marine ecosystems.
-These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km (Makris) to small scale behaviour patterns (refs).
+These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km (Duane et al., 2021) to small scale behaviour patterns (Handegard et al., 2003).
 These methods are also a crucial component of fisheries resource management.
-Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperature sonar, omnidirectional sonars, acoustic imaging sonars as well as echosounders (refs).
-
-Single- beam echosounders have extensively been used in fisheries management and ecosystem monitoring.
+Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperture sonar, omnidirectional sonars, acoustic imaging sonars as well as single- and multibeam echosounders (\hl{refs}).
+
+Single-beam echosounders have extensively been used in fisheries management and ecosystem monitoring.
 Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful.
-Single-beam echosounders have provided important information for fisheries resource management (Simmonds and MacLennan, 2005) as well as insight in ecological processes for the aquatic ecosystem (Godø 2014).
+Single-beam echosounders have provided important information for fisheries resource management (Simmonds and MacLennan, 2005) as well as insight in ecological processes for the aquatic ecosystem (Godø, 2014).
 
 Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz).
-The method has been used to  study predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009; Godø...), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
+The method has been used to  study predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009; Godø, 2014), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
 
@@
 a continuous frequency coverage over a wide frequency band.
-provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
+This provides significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
 These benefits can lead to improved backscatter categorization (R.
@@
 J.
-Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning
-(skriv om: e.g.
-Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).
+Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning.
+Early work with recent commercial systems indicate the potential improvement in detection near boundaries (Lavery et al., 2017) and remote sizing of fish (Kubilius et al., 2020;), capabilities of fine-scale observations of interactions between prey and predators (Skaret et al., 2020), the potential for species discrimination (Basset et al., 2018; Blanluet et al., 2019; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021), and the potential for enhanced information on mesopelagic layers (Cotter et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).
 
@@
 
-To tap the full potential of broadband echosounder data, further methods for processing such data must be developed.
+To tap the full potential of broadband echosounder data, further methods for processing such data must be developed (Lavery et al., 2017; Basset et al., 2018).
 Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
@@
 In the example a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}).
-This is found in the class \verb|Core.EK80Calculation.generateIdealWindowedSendPulse|, where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
+This is found in the class \verb|Calculation.generateIdealWindowedTransmitSignal|, where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
 A tapering of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
@@
 \includegraphics[width=16cm]{Fig_ytx}
-\caption{\label{fi:ytx} Ideal and enveloped send pulse between 34kHz to 45kHz with a slope of 0.057, as implemented in the EK80. The orange curve illustrate the effect of setting the slope to 0.5.}
+\caption{\label{fi:ytx} Ideal and enveloped transmit pulse between 34 kHz to 45 kHz with a slope of 0.057, as implemented in the EK80.
+The orange curve illustrate the effect of setting the slope to 0.5.}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_fir}
-\caption{\label{fi:fir} The frequency (power) response of the filters in from our test set.}
+\caption{\label{fi:fir} The frequency (power) response of the filters in our test set.}
 \end{figure}
@@
 The original sample data $\yrxorg(\samplesymt,\channelsym)$ are not available in the EK80 data files.
-Instead,the resulting filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
+Instead, the resulting filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
 The data are recorded together with additional data such as from position and motion sensors and system configuration data in raw data files for display and analysis by processing software.
@@
 \includegraphics{Fig_ACF}
-\caption{The aoutocorrekation function $\ptxauto$ from the example. The corresponding $\teff=0.1ms$ and $\tnom=2ms$\label{fi:ACF}.}
+\caption{The autocorrelation function $\ptxauto$ from the example. The corresponding $\teff=0.1 ms$ and $\tnom=2 ms$\label{fi:ACF}.}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_theta_phi}
-\caption{\label{fi:theta_phi}Physical angles.}
+\caption{\label{fi:theta_phi}Physical angles $\along$ and $\athw$.}
 \end{figure}
@@
 
-To illustrate each step of the process to obtain $TS(f)$ we use a single ping from a data file where a single metallic calibration sphere is suspended approximately 6 m below a 120 kHz transducer (fig.~\ref{Fig_TS_echogram}).
+To illustrate each step of the process to obtain $TS(m)$ we use a single ping from a data file where a single metallic calibration sphere with 35 mm diameter (WC35) is suspended approximately 6 m below a 120 kHz transducer (fig.~\ref{Fig_TS_echogram}).
 
@@
 \caption{\label{Fig_TS_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file.
-The red vertical line indicates ping to be processed in this section for TS(f).
+The red vertical line indicates ping to be processed in this section for TS(m).
 Single target calibration sphere is seen at approximately 6 m range and the seafloor at approximately 12 m.}
@@
 %
-The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...
+\hl{The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...}
 
@@
 The amplitude of the backscatter signal also clearly shows these nulls (Fig.
-\ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
-
-Comment from Nils Olav: Using $m$ as an argument to a function is not ideal.
+\ref{fi:SED}), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
+
+\hl{Comment from Nils Olav: Using $m$ as an argument to a function is not ideal.
 If we use the function argument perhaps using $f_m$ would be more logical, i.e.
@@
 The latter is perhaps more in line with the code since the parameter is an array windexed with $m$ and not a function.
-$Psi(f)$, on the the other hand is derived from a function, see below.
+$Psi(f)$, on the the other hand is derived from a function, see below.}
 
@@
 
-As a metallic sphere is a rather simple and ideal scatterer, we present $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) in the processes to obtain $Sv((f)$.
+As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) to illustrate the processes to obtain $Sv(m)$.
 
@@
 
-TODO: Explain edge cases.
+\hl{TODO: Explain edge cases.}
 
@@
 %
-TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.
+\hl{TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.}
 
@@
 
-By selecting a set of centre samples $n$, $Sv$ values can be presented as a function of range (n) and frequency (m) for each ping.
-The range for centre samples n could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
-This can be useful when combining the Sv(f) across a range of transducers.
+By selecting a set of centre samples $n$, $Sv$ values can be presented as a function of range ($n$) and frequency ($m$) for each ping.
+The range for centre samples $n$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
+This can be useful when combining the $Sv(m)$ across a range of transducers.
 In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).
@@
 
-For echo integration purposes it is common to integrate the Sv over a range, in our example we integrate the Sv from 15 to 34 m covering a school of non-swimbladdered fish (fig).
-We note that we normally require averaging over several individual pings to obtain an unbiased estimate of Sv for echo integration purposes.
-Here we present the sv(f) over the layer for our example ping for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
+For acoustic abundance estimation and classification purposes it is common to integrate the Sv over a range, in our example we integrate $Sv$ from 15 to 34 m covering a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}).
+We note that we normally require averaging over several pings to obtain an unbiased estimate of $Sv$ for echo integration purposes.
+Here we present the $Sv(m)$ over the layer for our example ping for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
 Even though this is for a single ping, we still observe a positive slope of the frequency response indicative of non-swimbladdered fish.
@@
 \includegraphics[width=16cm]{Fig_Sv_avg}
-\caption{\label{Fig_Sv_avg} Sv as a function of frequency ($m$) averaged over a range interval covering a fish school }
+\caption{\label{Fig_Sv_avg} $Sv$ as a function of frequency ($m$) averaged over a range interval covering a fish school }
 \end{figure}

### 2022-03-18 Paper/Manuscript.tex
Small fixes to paper

@@
 
-As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) to illustrate the processes to obtain $Sv(m)$.
+As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $S_v(m)$.
 

### 2022-03-21 Paper/Manuscript.tex
Small changes to text

@@
 These include, as examples,\\
+How to handle calibration, interpolation, missing calibration\\
 Transmit pulse\\

### 2022-03-22 Paper/Manuscript.tex
Move bibliography to the .bib file method

@@
 
-%\bibliography{ek80refs}
-
-\begin{thebibliography}{39}
-\def\enquote#1{``#1,''}
-\def\plainquote#1{``#1''}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-\providecommand{\dourl}[1]{\href{http://#1}{\nolinkurl{#1}}}
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\noopsort}[1]{}
-\providecommand{\switchargs}[2]{#2#1}
-  \def\eatspace #1{#1}
-\providecommand{\dodoi}[1]{doi: \href{http://dx.doi.org/#1}{\nolinkurl{#1}}}
-
-\bibitem[{Barr \emph{et~al.}(2002)Barr, Coombs, Doonan, and
-  McMillian}]{Barr2002Target}
-\bibinfo{author}{Barr, R.}, \bibinfo{author}{Coombs, R.},
-  \bibinfo{author}{Doonan, I.},  and \bibinfo{author}{McMillian, P.}
-  (\textbf{\bibinfo{year}{2002}}). \enquote{\bibinfo{title}{Target
-  identification of oreos and associated species}} \bibinfo{type}{Final
-  {{Research Report}} for the {{Ministry}} of {{Fisheries Research Project
-  OEO2000}}/{{01B}}, {{Objective}} 1} ,
-  \dourl{fs.fish.govt.nz/Page.aspx?pk=113&dk=22653}, \bibinfo{note}{(Last
-  viewed October 21, 2019)}.
-
-\bibitem[{Bassett \emph{et~al.}(2018)Bassett, De~Robertis, and
-  Wilson}]{bassettBroadbandEchosounderMeasurements2018}
-\bibinfo{author}{Bassett, C.}, \bibinfo{author}{De~Robertis, A.},  and
-  \bibinfo{author}{Wilson, C.~D.} (\textbf{\bibinfo{year}{2018}}).
-  \enquote{\bibinfo{title}{Broadband echosounder measurements of the frequency
-  response of fishes and euphausiids in the {{Gulf}} of {{Alaska}}}}
-  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{75}(3),
-  \bibinfo{pages}{1131--1142}, \dodoi{10.1093/icesjms/fsx204}.
-
-\bibitem[{Briseño-Avena \emph{et~al.}(2015)Briseño-Avena, Roberts, Franks,
-  and Jaffe}]{Briseno-Avena2015ZOOPS}
-\bibinfo{author}{Briseño-Avena, C.}, \bibinfo{author}{Roberts, P. L.~D.},
-  \bibinfo{author}{Franks, P. J.~S.},  and \bibinfo{author}{Jaffe, J.~S.}
-  (\textbf{\bibinfo{year}{2015}}).
-  \enquote{\bibinfo{title}{{ZOOPS-O\textsuperscript{2}: A broadband echosounder
-  with coordinated stereo optical imaging for observing plankton in situ}}}
-  \bibinfo{journal}{Methods in Oceanography} \textbf{12},
-  \bibinfo{pages}{36--54}, \dodoi{10.1016/j.mio.2015.07.001}.
-
-\bibitem[{Burdic(1991)}]{burdic1991}
-\bibinfo{author}{Burdic, W.~S.} (\textbf{\bibinfo{year}{1991}}).
-  \emph{\bibinfo{title}{Underwater Acoustic System Analysis}}
-  (\bibinfo{publisher}{{Prentice Hall}}).
-
-\bibitem[{Chu and Stanton(1998)}]{Chu1998Application}
-\bibinfo{author}{Chu, D.},  and \bibinfo{author}{Stanton, T.}
-  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{{Application of
-  pulse compression techniques to broadband acoustic scattering by live
-  individual zooplankton}}} \bibinfo{journal}{The Journal of the Acoustical
-  Society of America} \textbf{104}(1), \bibinfo{pages}{39–55},
-  \dodoi{10.1121/1.424056}.
-
-\bibitem[{Chu \emph{et~al.}(1992)Chu, Stanton, and Wiebe}]{chu1992}
-\bibinfo{author}{Chu, D.}, \bibinfo{author}{Stanton, T.},  and
-  \bibinfo{author}{Wiebe, P.} (\textbf{\bibinfo{year}{1992}}).
-  \enquote{\bibinfo{title}{Frequency dependence of sound backscattering from
-  live individual zooplankton}} \bibinfo{journal}{ICES Journal of Marine
-  Science} \textbf{49}(1), \bibinfo{pages}{97--106},
-  \dodoi{10.1093/icesjms/49.1.97}.
-
-\bibitem[{Conti and Demer(2003)}]{Conti2003Wide-bandwidth}
-\bibinfo{author}{Conti, S.~G.},  and \bibinfo{author}{Demer, D.~A.}
-  (\textbf{\bibinfo{year}{2003}}). \enquote{\bibinfo{title}{{Wide-bandwidth
-  acoustical characterization of anchovy and sardine from reverberation
-  measurements in an echoic tank}}} \bibinfo{journal}{ICES Journal of Marine
-  Science} \textbf{60}(3), \bibinfo{pages}{617--624},
-  \dodoi{10.1016/S1054-3139(03)00056-0}.
-
-\bibitem[{Cook and Bernfield(1967)}]{cook1967}
-\bibinfo{author}{Cook, C.},  and \bibinfo{author}{Bernfield, M.}
-  (\textbf{\bibinfo{year}{1967}}). \emph{\bibinfo{title}{Radar {{Signals}}:
-  {{An Introduction}} to {{Theory}} and {{Application}}}}
-  (\bibinfo{publisher}{{Academic Press, Inc.}}).
-
-\bibitem[{Demer \emph{et~al.}(2015)Demer, Berger, Bernasconi, Boswell, Chu,
-  Domokos, Dunford, F\"assler, Gauthier, Hufnagle, Jech, Bouffant,
-  {Lebourges-Dhaussy}, Lurton, Macaulay, Perrot, Ryan, {Parker-Stetter},
-  Stienessen, Weber, and Williamson}]{demerCalibrationAcousticInstruments2015}
-\bibinfo{author}{Demer, D.~A.}, \bibinfo{author}{Berger, L.},
-  \bibinfo{author}{Bernasconi, M.}, \bibinfo{author}{Boswell, K.~M.},
-  \bibinfo{author}{Chu, D.}, \bibinfo{author}{Domokos, R.},
-  \bibinfo{author}{Dunford, A.~J.}, \bibinfo{author}{F\"assler, S. M.~M.},
-  \bibinfo{author}{Gauthier, S.}, \bibinfo{author}{Hufnagle, L.~T.},
-  \bibinfo{author}{Jech, J.~M.}, \bibinfo{author}{Bouffant, N.},
-  \bibinfo{author}{{Lebourges-Dhaussy}, A.}, \bibinfo{author}{Lurton, X.},
-  \bibinfo{author}{Macaulay, G.~J.}, \bibinfo{author}{Perrot, Y.},
-  \bibinfo{author}{Ryan, T.~E.}, \bibinfo{author}{{Parker-Stetter}, S.},
-  \bibinfo{author}{Stienessen, S.}, \bibinfo{author}{Weber, T.~C.},  and
-  \bibinfo{author}{Williamson, N.~J.} (\textbf{\bibinfo{year}{2015}}).
-  \enquote{\bibinfo{title}{Calibration of acoustic instruments}}
-  \bibinfo{type}{{{ICES Cooperative Research Report No}}. 326} ,
-  \dodoi{10.17895/ices.pub.5494}.
-
-\bibitem[{Denny and Simpson(1998)}]{dennyBroadbandAcousticFish1998}
-\bibinfo{author}{Denny, G.},  and \bibinfo{author}{Simpson, P.}
-  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{A broadband
-  acoustic fish identification system}} \bibinfo{journal}{The Journal of the
-  Acoustical Society of America} \textbf{103}(5), \bibinfo{pages}{3069--3069},
-  \dodoi{10.1121/1.422851}.
-
-\bibitem[{Ehrenberg(1979)}]{ehrenberg1979}
-\bibinfo{author}{Ehrenberg, J.~E.} (\textbf{\bibinfo{year}{1979}}).
-  \enquote{\bibinfo{title}{A comparative analysis of {\emph{in situ}} methods
-  for directly measuring the acoustic target strength of individual fish}}
-  \bibinfo{journal}{IEEE Journal of Oceanic Engineering} \textbf{4}(4),
-  \bibinfo{pages}{141--152}, \dodoi{10.1109/JOE.1979.1145434}.
-
-\bibitem[{Ehrenberg and Torkelson(2000)}]{ehrenbergFMSlideChirp2000}
-\bibinfo{author}{Ehrenberg, J.~E.},  and \bibinfo{author}{Torkelson, T.~C.}
-  (\textbf{\bibinfo{year}{2000}}). \enquote{\bibinfo{title}{{{FM}} slide
-  (chirp) signals: A technique for significantly improving the signal-to-noise
-  performance in hydroacoustics assessment systems}}
-  \bibinfo{journal}{Fisheries Research} \textbf{47}, \bibinfo{pages}{193--199},
-  \dodoi{10.1016/S0165-7836(00)00169-7}.
-
-\bibitem[{Foote \emph{et~al.}(2005)Foote, Atkins, Francis, and
-  Knutsen}]{Foote2005Measuring}
-\bibinfo{author}{Foote, K.~G.}, \bibinfo{author}{Atkins, P.~R.},
-  \bibinfo{author}{Francis, D. T.~I.},  and \bibinfo{author}{Knutsen, T.}
-  (\textbf{\bibinfo{year}{2005}}). \enquote{\bibinfo{title}{{Measuring echo
-  spectra of marine organisms over a wide bandwidth}}}
-  \bibinfo{address}{Proceedings of the International Conference on Underwater Acoustic Measurements: Technologies and Results, Heraklion, Crete}.
-
-\bibitem[{Forland \emph{et~al.}(2014)Forland, Hobæk, and
-  Korneliussen}]{Forland2014Scattering}
-\bibinfo{author}{Forland, T.~N.}, \bibinfo{author}{Hobæk, H.},  and
-  \bibinfo{author}{Korneliussen, R.~J.} (\textbf{\bibinfo{year}{2014}}).
-  \enquote{\bibinfo{title}{{Scattering properties of Atlantic mackerel over a
-  wide frequency range}}} \bibinfo{journal}{ICES Journal of Marine Science}
-  \textbf{71}(7), \bibinfo{pages}{1904--1912}, \dodoi{10.1093/icesjms/fsu045}.
-
-\bibitem[{Gordon and Zedel(1998)}]{Gordon1998FishMASS}
-\bibinfo{author}{Gordon, L.},  and \bibinfo{author}{Zedel, L.}
-  (\textbf{\bibinfo{year}{1998}}). \enquote{\bibinfo{title}{{FishMASS: what can
-  you do with a little bandwidth when you are watching fish?}}}
-  \bibinfo{journal}{IEEE Seminar Digests} \textbf{1998}(227),
-  \bibinfo{pages}{7}, \dodoi{10.1049/ic:19980186}.
-
-\bibitem[{Hasan(1983)}]{hasan1983}
-\bibinfo{author}{Hasan, T.} (\textbf{\bibinfo{year}{1983}}).
-  \enquote{\bibinfo{title}{Complex demodulation: {{Some}} theory and
-  applications}} in \emph{\bibinfo{booktitle}{Handbook of {{Statistics}}}},
-  \bibinfo{volume}{\textbf{3}} (\bibinfo{publisher}{{Elsevier}}), pp.
-  \bibinfo{pages}{125--156}, \dodoi{10.1016/S0169-7161(83)03009-6}.
-
-\bibitem[{Hob\ae{}k and
-  Forland(2013)}]{hobaekCharacterizationTargetSpheres2013}
-\bibinfo{author}{Hob\ae{}k, H.},  and \bibinfo{author}{Forland, T.~N.}
-  (\textbf{\bibinfo{year}{2013}}). \enquote{\bibinfo{title}{Characterization of
-  {{Target Spheres}} for {{Broad}}-{{Band Calibration}} of {{Acoustic
-  Systems}}}} \bibinfo{journal}{Acta Acustica united with Acustica}
-  \textbf{99}(3), \bibinfo{pages}{465--476}, \dodoi{10.3813/AAA.918627}.
-
-\bibitem[{Holliday and Pieper(1980)}]{holliday1980}
-\bibinfo{author}{Holliday, D.~V.},  and \bibinfo{author}{Pieper, R.~E.}
-  (\textbf{\bibinfo{year}{1980}}). \enquote{\bibinfo{title}{Volume scattering
-  strengths and zooplankton distributions at acoustic frequencies between 0.5
-  and 3 {{MHz}}}} \bibinfo{journal}{The Journal of the Acoustical Society of
-  America} \textbf{67}(1), \bibinfo{pages}{135--146}, \dodoi{10.1121/1.384472}.
-
-\bibitem[{Imaizumi \emph{et~al.}(2009)Imaizumi, Abe, Sawada, Matsuda, Akamatsu,
-  Suga, Wang, Nishimori, Ogawa, Matsuo, and Ito}]{Imaizumi2009Detection}
-\bibinfo{author}{Imaizumi, T.}, \bibinfo{author}{Abe, K.},
-  \bibinfo{author}{Sawada, K.}, \bibinfo{author}{Matsuda, A.},
-  \bibinfo{author}{Akamatsu, T.}, \bibinfo{author}{Suga, T.},
-  \bibinfo{author}{Wang, Y.}, \bibinfo{author}{Nishimori, Y.},
-  \bibinfo{author}{Ogawa, S.}, \bibinfo{author}{Matsuo, I.},  and
-  \bibinfo{author}{Ito, M.} (\textbf{\bibinfo{year}{2009}}).
-  \enquote{\bibinfo{title}{{Detection of in situ fish using broadband
-  split-beam system}}} \bibinfo{journal}{Institute of Electronics, Information,
-  and Communication Engineers, Tech. Rep. IEICE-US2009-36} \textbf{109}(180),
-  \bibinfo{pages}{39--42}.
-
-\bibitem[{Klauder \emph{et~al.}(1960)Klauder, Price, Darlington, and
-  Albersheim}]{klauder1960}
-\bibinfo{author}{Klauder, J.~R.}, \bibinfo{author}{Price, A.~C.},
-  \bibinfo{author}{Darlington, S.},  and \bibinfo{author}{Albersheim, W.~J.}
-  (\textbf{\bibinfo{year}{1960}}). \enquote{\bibinfo{title}{The theory and
-  design of chirp radars}} \bibinfo{journal}{The Bell System Technical Journal}
-  \textbf{39}(4), \bibinfo{pages}{745--808},
-  \dodoi{10.1002/j.1538-7305.1960.tb03942.x}.
-
-\bibitem[{Korneliussen(2010)}]{korneliussen2010}
-\bibinfo{author}{Korneliussen, R.~J.} (\textbf{\bibinfo{year}{2010}}).
-  \enquote{\bibinfo{title}{The acoustic idenification of {{Atlantic}}
-  mackerel}} \bibinfo{journal}{ICES Journal of Marine Science} \textbf{67}(8),
-  \bibinfo{pages}{1749--1758}, \dodoi{10.1093/icesjms/fsq052}.
-
-\bibitem[{Korneliussen \emph{et~al.}(2018)Korneliussen, Berger, Campanlla,
-  Dezhang, Demer, De~Robertis, Domokos, Doray, Fielding, F\"assler, Gauthier,
-  Gastauer, Horne, Hutton, Iriarte, Jech, Kloser, Lawson, {Lebourges-Dhaussy},
-  McQuinn, Pe\~na, Scoulding, Sakinan, Schaber, Taylor, and
-  Thompson}]{korneliussen2018}
-\bibinfo{author}{Korneliussen, R.~J.}, \bibinfo{author}{Berger, L.},
-  \bibinfo{author}{Campanlla, F.}, \bibinfo{author}{Dezhang, C.},
-  \bibinfo{author}{Demer, D.}, \bibinfo{author}{De~Robertis, A.},
-  \bibinfo{author}{Domokos, R.}, \bibinfo{author}{Doray, M.},
-  \bibinfo{author}{Fielding, S.}, \bibinfo{author}{F\"assler, S. M.~M.},
-  \bibinfo{author}{Gauthier, S.}, \bibinfo{author}{Gastauer, S.},
-  \bibinfo{author}{Horne, J.}, \bibinfo{author}{Hutton, B.},
-  \bibinfo{author}{Iriarte, F.}, \bibinfo{author}{Jech, J.~M.},
-  \bibinfo{author}{Kloser, R.}, \bibinfo{author}{Lawson, G.},
-  \bibinfo{author}{{Lebourges-Dhaussy}, A.}, \bibinfo{author}{McQuinn, I.},
-  \bibinfo{author}{Pe\~na, M.}, \bibinfo{author}{Scoulding, B.},
-  \bibinfo{author}{Sakinan, S.}, \bibinfo{author}{Schaber, M.},
-  \bibinfo{author}{Taylor, J.~C.},  and \bibinfo{author}{Thompson, C.~H.}
-  (\textbf{\bibinfo{year}{2018}}). \enquote{\bibinfo{title}{Target
-  classification}} \bibinfo{type}{ICES Cooperative Research Report}
-  \bibinfo{number}{No. 344} , \dodoi{doi.org/10.17895/ices.pub.4567}.
-
-\bibitem[{Korneliussen \emph{et~al.}(2016)Korneliussen, Heggelund, Macaulay,
-  Patel, Johnsen, and Eliassen}]{Korneliussen2016Acoustic}
-\bibinfo{author}{Korneliussen, R.~J.}, \bibinfo{author}{Heggelund, Y.},
-  \bibinfo{author}{Macaulay, G.~J.}, \bibinfo{author}{Patel, D.},
-  \bibinfo{author}{Johnsen, E.},  and \bibinfo{author}{Eliassen, I.~K.}
-  (\textbf{\bibinfo{year}{2016}}). \enquote{\bibinfo{title}{{Acoustic
-  identification of marine species using a feature library}}}
-  \bibinfo{journal}{Methods in Oceanography} \textbf{17},
-  \bibinfo{pages}{187--205}, \dodoi{10.1016/j.mio.2016.09.002}.
-
-\bibitem[{Korneliussen and Ona(2002)}]{Korneliussen2002operational}
-\bibinfo{author}{Korneliussen, R.~J.},  and \bibinfo{author}{Ona, E.}
-  (\textbf{\bibinfo{year}{2002}}). \enquote{\bibinfo{title}{{An operational
-  system for processing and visualizing multi-frequency acoustic data}}}
-  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{59}(2),
-  \bibinfo{pages}{293--313}, \dodoi{10.1006/jmsc.2001.1168}.
-
-\bibitem[{Lavery \emph{et~al.}(2017)Lavery, Bassett, Lawson, and Jech}]{lavery2017}
-\bibinfo{author}{Lavery, A.~C.},  \bibinfo{author}{Bassett, C.},
-  \bibinfo{author}{Lawson, G.~L.}, and \bibinfo{author}{Jech. J.~M.}
-  (\textbf{\bibinfo{year}{2017}}). \enquote{\bibinfo{title}{{Exploiting signal 
-  processing approaches for broadband echosounders}}} \bibinfo{journal}{ICES Journal
-  of Marine Science} \textbf{74}(8), \bibinfo{pages}{2262--2275},
-  \dodoi{10.1093/icesjms/fsx155}.
-
-\bibitem[{Lunde and Korneliussen(2016)}]{lunde2016}
-\bibinfo{author}{Lunde, P.},  and \bibinfo{author}{Korneliussen, R.~J.}
-  (\textbf{\bibinfo{year}{2016}}). \enquote{\bibinfo{title}{Power-{{Budget
-  Equations}} and {{Calibration Factors}} for {{Fish Abundance Estimation Using
-  Scientific Echo Sounder}} and {{Sonar Systems}}}} \bibinfo{journal}{Journal
-  of Marine Science and Engineering} \textbf{4}(3), \bibinfo{pages}{43},
-  \dodoi{10.3390/jmse4030043}.
-
-\bibitem[{MacLennan(1981)}]{maclennan1981}
-\bibinfo{author}{MacLennan, D.~N.} (\textbf{\bibinfo{year}{1981}}).
-  \enquote{\bibinfo{title}{The {{Theory}} of {{Solid Spheres}} as {{Sonar
-  Calibration Targets}}}} \bibinfo{type}{Scottish {{Fisheries Research Report
-  Number}} 22} .
-
-\bibitem[{MacLennan \emph{et~al.}(2002)MacLennan, Fernandes, and
-  Dalen}]{MacLennan2002consistent}
-\bibinfo{author}{MacLennan, D.~N.}, \bibinfo{author}{Fernandes, P.},  and
-  \bibinfo{author}{Dalen, J.} (\textbf{\bibinfo{year}{2002}}).
-  \enquote{\bibinfo{title}{{A consistent approach to definitions and symbols in
-  fisheries acoustics}}} \bibinfo{journal}{ICES Journal of Marine Science}
-  \textbf{59}, \bibinfo{pages}{365--369}, \dodoi{10.1006/jmsc.2001.1158}.
-
-\bibitem[{Simmonds \emph{et~al.}(1996)Simmonds, Armstrong, and
-  Copland}]{Simmonds1996Species}
-\bibinfo{author}{Simmonds, E.~J.}, \bibinfo{author}{Armstrong, F.},  and
-  \bibinfo{author}{Copland, P.~J.} (\textbf{\bibinfo{year}{1996}}).
-  \enquote{\bibinfo{title}{{Species identification using wideband backscatter
-  with neural network and discriminant analysis}}} \bibinfo{journal}{ICES
-  Journal of Marine Science} \textbf{53}(2), \bibinfo{pages}{189–195},
-  \dodoi{10.1006/jmsc.1996.0021}.
-
-\bibitem[{Simmonds and MacLennan(2005)}]{Simmonds2005Fisheries}
-\bibinfo{author}{Simmonds, J.},  and \bibinfo{author}{MacLennan, D.}
-  (\textbf{\bibinfo{year}{2005}}). \emph{\bibinfo{title}{{Fisheries Acoustics.
-  Theory and Practice}}}, \bibinfo{edition}{2nd ed.}
-  (\bibinfo{publisher}{Blackwell Science}, \bibinfo{address}{Oxford}).
-
-\bibitem[{Skaret \emph{et~al.}(2020)Skaret, Johansen, Johnsen, Fall, Fiksen,
-  Englund, Fauchald, Gj{\o}s{\ae}ter, Macaulay, and Johannesen}]{skaret2020}
-\bibinfo{author}{Skaret, G.}, \bibinfo{author}{Johansen, G.~O.},
-  \bibinfo{author}{Johnsen, E.}, \bibinfo{author}{Fall, J.},
-  \bibinfo{author}{Fiksen, {\O}.}, \bibinfo{author}{Englund, G.},
-  \bibinfo{author}{Fauchald, P.}, \bibinfo{author}{Gj{\o}s{\ae}ter, H.},
-  \bibinfo{author}{Macaulay, G.~J.},  and \bibinfo{author}{Johannesen, E.}
-  (\textbf{\bibinfo{year}{2020}}). \enquote{\bibinfo{title}{Diel vertical
-  movements determine spatial interactions between cod, pelagic fish and krill
-  on an {{Arctic}} shelf bank}} \bibinfo{journal}{Marine Ecology Progress
-  Series} \textbf{638}, \bibinfo{pages}{13--23}, \dodoi{10.3354/meps13254}.
-
-\bibitem[{Stanton and Chu(2008)}]{stanton2008}
-\bibinfo{author}{Stanton, T.~K.},  and \bibinfo{author}{Chu, D.}
-  (\textbf{\bibinfo{year}{2008}}). \enquote{\bibinfo{title}{Calibration of
-  broadband active acoustic systems using a single standard spherical target}}
-  \bibinfo{journal}{The Journal of the Acoustical Society of America}
-  \textbf{124}(1), \bibinfo{pages}{128--136}, \dodoi{10.1121/1.2917387}.
-
-\bibitem[{Stanton \emph{et~al.}(2010)Stanton, Chu, Jech, and
-  Irish}]{Stanton2010New}
-\bibinfo{author}{Stanton, T.~K.}, \bibinfo{author}{Chu, D.},
-  \bibinfo{author}{Jech, J.~M.},  and \bibinfo{author}{Irish, J.~D.}
-  (\textbf{\bibinfo{year}{2010}}). \enquote{\bibinfo{title}{{New broadband
-  methods for resonance classification and high-resolution imagery of fish with
-  swimbladders using a modified commercial broadband echosounder}}}
-  \bibinfo{journal}{ICES Journal of Marine Science} \textbf{67}(2),
-  \bibinfo{pages}{365--378}, \dodoi{10.1093/icesjms/fsp262}.
-
-\bibitem[{Stanton \emph{et~al.}(2012)Stanton, Sellers, and
-  Jech}]{Stanton2012Resonance}
-\bibinfo{author}{Stanton, T.~K.}, \bibinfo{author}{Sellers, C.~J.},  and
-  \bibinfo{author}{Jech, J.~M.} (\textbf{\bibinfo{year}{2012}}).
-  \enquote{\bibinfo{title}{{Resonance classification of mixed assemblages of
-  fish with swimbladders using a modified commercial broadband acoustic
-  echosounder at 1–6 kHz}}} \bibinfo{journal}{Canadian Journal of Fisheries
-  and Aquatic Sciences} \textbf{69}(5), \bibinfo{pages}{854--868},
-  \dodoi{10.1139/f2012-013}.
-
-\bibitem[{Traykovski \emph{et~al.}(1998)Traykovski, O'Driscoll, and
-  McGhee}]{Traykovski1998Effect}
-\bibinfo{author}{Traykovski, L.~M.}, \bibinfo{author}{O'Driscoll, R.},  and
-  \bibinfo{author}{McGhee, D.} (\textbf{\bibinfo{year}{1998}}).
-  \enquote{\bibinfo{title}{{Effect of orientation on broadband acoustic
-  scattering of Antarctic krill \textit{Euphausia superba}: Implications for
-  inverting zooplankton spectral acoustic signatures for angle of
-  orientation}}} \bibinfo{journal}{The Journal of the Acoustical Society of
-  America} \textbf{104}(4), \bibinfo{pages}{2121–2135},
-  \dodoi{10.1121/1.423726}.
-
-\bibitem[{Turin(1960)}]{turin1960}
-\bibinfo{author}{Turin, G.} (\textbf{\bibinfo{year}{1960}}).
-  \enquote{\bibinfo{title}{An introduction to matched filters}}
-  \bibinfo{journal}{IRE Transactions on Information Theory} \textbf{6}(3),
-  \bibinfo{pages}{311--329}, \dodoi{10.1109/TIT.1960.1057571}.
-
-\bibitem[{Zakharia \emph{et~al.}(1989)Zakharia, Corgiatti, Joly, and
-  Person}]{Zakharia1989Wide-band}
-\bibinfo{author}{Zakharia, M.}, \bibinfo{author}{Corgiatti, J.},
-  \bibinfo{author}{Joly, F.},  and \bibinfo{author}{Person, R.}
-  (\textbf{\bibinfo{year}{1989}}). \enquote{\bibinfo{title}{{Wide-band sounder
-  for fisheries}}} \bibinfo{journal}{Proceedings of the Institute of Acoustics}
-  \textbf{11}, \bibinfo{pages}{274--281}.
-
-\bibitem[{Zakharia \emph{et~al.}(1996)Zakharia, Magand, Hetroit, and
-  Diner}]{Zakharia1996Wideband}
-\bibinfo{author}{Zakharia, M.~E.}, \bibinfo{author}{Magand, F.},
-  \bibinfo{author}{Hetroit, F.},  and \bibinfo{author}{Diner, N.}
-  (\textbf{\bibinfo{year}{1996}}). \enquote{\bibinfo{title}{{Wideband sounder
-  for fish species identification at sea}}} \bibinfo{journal}{ICES Journal of
-  Marine Science} \textbf{53}, \bibinfo{pages}{203–208},
-  \dodoi{10.1006/jmsc.1996.0023}.
-
-\bibitem[{Zedel \emph{et~al.}(2003)Zedel, Knutsen, and
-  Patro}]{Zedel2003Acoustic}
-\bibinfo{author}{Zedel, L.}, \bibinfo{author}{Knutsen, T.},  and
-  \bibinfo{author}{Patro, R.} (\textbf{\bibinfo{year}{2003}}).
-  \enquote{\bibinfo{title}{{Acoustic Doppler current profiler observations of
-  herring movement}}} \bibinfo{journal}{ICES Journal of Marine Science}
-  \textbf{60}(4), \bibinfo{pages}{846--859},
-  \dodoi{10.1016/s1054-3139(03)00067-5}.
-
-\end{thebibliography}
+\bibliography{Reference_database}
 

### 2022-03-22 Paper/Manuscript.tex
Revert to alphabetical order for authors
Convert some hard-coded references to \citep's

@@
 
-\author{Geir Pedersen}
-\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
-
 \author{Dezhang Chu}
@@
 Seattle, WA, 98112, USA}
+
+\author{Nils Olav Handegard}
+\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
 
@@
 
-\author{Nils Olav Handegard}
+\author{Geir Pedersen}
 \affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
@@
 
-To tap the full potential of broadband echosounder data, further methods for processing such data must be developed (Lavery et al., 2017; Basset et al., 2018).
+To tap the full potential of broadband echosounder data, further methods for processing such data must be developed \citep{lavery2017,bassett_broadband_2018}.
 Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
@@
 The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
-For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
+For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
 The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

### 2022-03-22 Paper/Manuscript.tex
Replace hard-code refs with \citep's and fix some bib things

@@
 Active acoustics is an efficient method for remote sensing of marine ecosystems.
-These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km (Duane et al., 2021) to small scale behaviour patterns (Handegard et al., 2003).
+These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km \citep{duane_quantification_2021} to small scale behaviour patterns \citep{handegard_avoidance_2003}.
 These methods are also a crucial component of fisheries resource management.
@@
 Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful.
-Single-beam echosounders have provided important information for fisheries resource management (Simmonds and MacLennan, 2005) as well as insight in ecological processes for the aquatic ecosystem (Godø, 2014).
+Single-beam echosounders have provided important information for fisheries resource management \citep{Simmonds2005Fisheries} as well as insight in ecological processes for the aquatic ecosystem (Godø, 2014).
 
 Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz).
-The method has been used to  study predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009; Godø, 2014), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).
+The method has been used to  study predator-prey interactions \citep{benoit-bird_prey_2003, friedlaender_diel_2009, veit_aggregation_1993}, interactions between marine organisms and the physical environment \citep{crawford_water_2012, freeman_diel_2004, georgiadis_coralligene_2009, ludsin_hypoxia_avoidance_2009} (+Godø, 2014), abundance and distribution of prey and predators \citep{thompson_winter_1991} and non-fish species or group of species \citep{brierley_acoustic_2001}, and the individual or group behavior of marine organisms \citep{axelsen_await_2000, benoit-bird_diel_2001, kaartvedt_use_2009}.
 
@@
 a continuous frequency coverage over a wide frequency band.
-This provides significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000).
-These benefits can lead to improved backscatter categorization (R.
-Korneliussen, 2018; R.
-J.
-Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning.
-Early work with recent commercial systems indicate the potential improvement in detection near boundaries (Lavery et al., 2017) and remote sizing of fish (Kubilius et al., 2020;), capabilities of fine-scale observations of interactions between prey and predators (Skaret et al., 2020), the potential for species discrimination (Basset et al., 2018; Blanluet et al., 2019; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021), and the potential for enhanced information on mesopelagic layers (Cotter et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).
+This provides significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}.
+These benefits can lead to improved backscatter categorization \citep{korneliussen2018, Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic} and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning.
+Early work with recent commercial systems indicate the potential improvement in detection near boundaries \citep{lavery2017} and remote sizing of fish \citep{kubilius_remote_2020}, capabilities of fine-scale observations of interactions between prey and predators \citep{skaret_diel_2020}, the potential for species discrimination \citep{bassett_broadband_2018, blanluet_characterization_2019, benoit-bird_exploring_2020} (+Hasegawa et al., 2021; Gugele et al., 2021), and the potential for enhanced information on mesopelagic layers \citep{cotter_classification_2021, agersted_application_2021, khodabandeloo_estimating_2021, khodabandeloo_mesopelagic_2021}.
 
-There have been several scientific broadband echosounder systems developed for laboratory use (Chu et al., 1992; Conti and Demer, 2003; Forland et al., 2014), some prototype or custom-made systems (Barr et al., 2002; Brise~no-Avena et al., 2015; Foote et al., 2005; Imaizumi et al., 2009; Simmonds et al., 1996; Zakharia et al., 1989, 1996) and some commercially available systems (Denny and Simpson, 1998; Ehrenberg and Torkelson, 2000; Gordon and Zedel, 1998; Stanton et al., 2010; Zedel et al., 2003).
+There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
 
@@
 The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
-For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
+For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret_diel_2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
 The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

### 2022-03-22 Paper/Manuscript.tex
Move two more refs to using citep

@@
 These benefits can lead to improved backscatter categorization \citep{korneliussen2018, Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic} and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning.
-Early work with recent commercial systems indicate the potential improvement in detection near boundaries \citep{lavery2017} and remote sizing of fish \citep{kubilius_remote_2020}, capabilities of fine-scale observations of interactions between prey and predators \citep{skaret_diel_2020}, the potential for species discrimination \citep{bassett_broadband_2018, blanluet_characterization_2019, benoit-bird_exploring_2020} (+Hasegawa et al., 2021; Gugele et al., 2021), and the potential for enhanced information on mesopelagic layers \citep{cotter_classification_2021, agersted_application_2021, khodabandeloo_estimating_2021, khodabandeloo_mesopelagic_2021}.
+Early work with recent commercial systems indicate the potential improvement in detection near boundaries \citep{lavery2017} and remote sizing of fish \citep{kubilius_remote_2020}, capabilities of fine-scale observations of interactions between prey and predators \citep{skaret_diel_2020}, the potential for species discrimination \citep{bassett_broadband_2018, blanluet_characterization_2019, benoit-bird_exploring_2020, gugele_differentiation_2021, hasegawa_situ_2021}, and the potential for enhanced information on mesopelagic layers \citep{cotter_classification_2021, agersted_application_2021, khodabandeloo_estimating_2021, khodabandeloo_mesopelagic_2021}.
 

### 2022-03-22 Paper/Manuscript.tex
fix duplicate figure reference label

@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively.
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
+\ref{fi:TS}.
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
@@
 where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
-The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\ref{fi:SED}) is then calculated by: 
+The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (Fig. \ref{fi:TS}) is then calculated by: 
 %
@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
+\caption{\label{fi:TS}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
 \end{figure}
@@
 This is visible in the $\ts$ (Fig.
-\ref{fi:SED}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
+\ref{fi:TS}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
 The amplitude of the backscatter signal also clearly shows these nulls (Fig.
-\ref{fi:SED}), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
+\ref{fi:TS}), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
 

### 2022-03-22 Paper/Manuscript.tex
Split an overlong equation and highlighted comment
Tidy up a list

@@
 %
-\hl{The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...}
+\hl{The corresponding frequencies f\_m\_t (?) are calculate as follows...
+
+See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...}
 
@@
 \label{eq:Sv}
-\sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range_c(\samplesymt)) + 2\absorp(\fc)\range_c(\samplesymt) 
+\begin{split}
+\sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range_c(\samplesymt)) + 2\absorp(\fc)\range_c(\samplesymt) \\
 - 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \cw \teff \eqang(\fc) \gainzero^2(\fc)}{32\pi^2} \right),
+\end{split}
 \end{equation}
@@
 Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature, where other approaches could been equally valid.
-These include, as examples,\\
-How to handle calibration, interpolation, missing calibration\\
-Transmit pulse\\
-Tapering of the transmit pulse\\
-Decimation factors and filters\\
-When Convoling the recieved signal with the transmit pulse - edge cases\\
-Assuemd four sector transducer\\
-Matched load - hvor stor effekt å gå utover \\
-Assumed an impedance (trancuder)\\
-Effective pulse duration...\\
-Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...\\
-Window size\\
-Edge cases - removed\\
+These include, as examples,
+\begin{itemize}
+\item How to handle calibration, interpolation, missing calibration
+\item Transmit pulse
+\item Tapering of the transmit pulse
+\item Decimation factors and filters
+\item When Convoling the recieved signal with the transmit pulse - edge cases
+\item Assuemd four sector transducer
+\item Matched load - hvor stor effekt å gå utover 
+\item Assumed an impedance (trancuder)
+\item Effective pulse duration...
+\item Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...
+\item Window size
+\item Edge cases - removed
+\end{itemize}
 

### 2022-04-11 Paper/Manuscript.tex
Updated Introduction and Discussion
Direct copy from Word doc on Teams. No edits to text itself yet.

@@
 Active acoustics is an efficient method for remote sensing of marine ecosystems.
-These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km \citep{duane_quantification_2021} to small scale behaviour patterns \citep{handegard_avoidance_2003}.
-These methods are also a crucial component of fisheries resource management.
-Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperture sonar, omnidirectional sonars, acoustic imaging sonars as well as single- and multibeam echosounders (\hl{refs}).
+These methods have been used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}.
+As early as 1935, Oscars Sund observed the distribution of spawning cod in the Lofoten area using a single beam echosounder \citep{sund_echo_1935}.
+The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}.
+More recently, the methods have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
 
-Single-beam echosounders have extensively been used in fisheries management and ecosystem monitoring.
-Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful.
-Single-beam echosounders have provided important information for fisheries resource management \citep{Simmonds2005Fisheries} as well as insight in ecological processes for the aquatic ecosystem (Godø, 2014).
+Several different configurations of acoustic instruments are available, including multibeam sonars \citep{gerlotto_two_1999}, synthetic aperture sonar, omnidirectional sonars \citep{misund_improved_1996}, acoustic imaging sonars \citep{jaffe_ftv_1995} as well as single- and multibeam echosounders \citep{trenkel_new_2008}.
+Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and methodology for using these systems are well developed.
+The ability for accurate calibration of the system has made these especially useful for providing data for fisheries management \citep{Simmonds2005Fisheries}.
 
-Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz).
-The method has been used to  study predator-prey interactions \citep{benoit-bird_prey_2003, friedlaender_diel_2009, veit_aggregation_1993}, interactions between marine organisms and the physical environment \citep{crawford_water_2012, freeman_diel_2004, georgiadis_coralligene_2009, ludsin_hypoxia_avoidance_2009} (+Godø, 2014), abundance and distribution of prey and predators \citep{thompson_winter_1991} and non-fish species or group of species \citep{brierley_acoustic_2001}, and the individual or group behavior of marine organisms \citep{axelsen_await_2000, benoit-bird_diel_2001, kaartvedt_use_2009}.
+Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems.
+This provides significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and  improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
 
-Recent echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems, i.e.
-a continuous frequency coverage over a wide frequency band.
-This provides significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}.
-These benefits can lead to improved backscatter categorization \citep{korneliussen2018, Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic} and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning.
-Early work with recent commercial systems indicate the potential improvement in detection near boundaries \citep{lavery2017} and remote sizing of fish \citep{kubilius_remote_2020}, capabilities of fine-scale observations of interactions between prey and predators \citep{skaret_diel_2020}, the potential for species discrimination \citep{bassett_broadband_2018, blanluet_characterization_2019, benoit-bird_exploring_2020, gugele_differentiation_2021, hasegawa_situ_2021}, and the potential for enhanced information on mesopelagic layers \citep{cotter_classification_2021, agersted_application_2021, khodabandeloo_estimating_2021, khodabandeloo_mesopelagic_2021}.
+Broad banded echosounders have been pursued actively for ecology in particularly based on the promise of improved categorization of species or group of species.
+Using high frequency broadband acoustics (150-600 kHz) Lavery et al \citep{lavery_measurements_2010} showed that broadband helped reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
+\citep{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) with frequency content covering the resonance of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages.
+\citep{blanluet_characterization_2019} used broadband, with ground truthing, to identify that the composition of two Sound Scattering Layers (SSL) detected in the Bay of Biscay in springtime was characterized and identifying fine scale heterogeneity within the SSLs.
+While unsuccessful in discriminating between several species of fishes with large  swimbladders during the Alaska pollock survey, \citep{bassett_broadband_2018} showed that broadband may be helpful in characterizing smaller fishes with swimbladder and euphausiids.
+\citep{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband data (45-170 kHz), while attempts to classify using narrowband multifrequency failed.
+
+The increased spatial resolution is another aspect of broadband acoustics that are actively being used both for improved categorization and increased ability to resolve targets near each other or near boundaries such as the seafloor.
+\citep{lavery2017} explored different broadband pulse shapes to increase successful resolution of single targets in proximity as well as near boundaries in tank experiments.
+By applying high frequency broadband pulses to fish-like artificial targets, \citep{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment.
+Utilizing the increased spatial resolution \citep{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.
+
+An equally interesting outcome from the increased resolution and broad frequency bands is the ability to infer behavioural information on both single individuals, groups of individuals, and interaction between individuals or groups.
+\citep{Traykovski1998Effect} was able to extract orientation information about krill from using a combination of high frequency broadband acoustics and acoustic modelling.
+Fine-scale observations of single targets facilitated the study of interactions between prey and predators, capelin and cod, in the Barents sea (\citep{skaret_diel_2020}).
 
@@
 
-To tap the full potential of broadband echosounder data, further methods for processing such data must be developed \citep{lavery2017,bassett_broadband_2018}.
+The theoretical foundation for signal processing comes from radar applications and is earlier documented for broadband echosounders \citep{stanton2008}.
+To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}.
 Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
 Different send pulses can be envisioned as well as various methods for acoustic target classification.
-The objective of this paper is to present a systematic and comprehensive explanation of how to process calibrated broadband echosounder data.
-By making associated code available, we provide a starting point for further developing signal processing methods for applying broadband acoustic to aquatic ecology.
+When translating equations into code, several choices must be made.
+Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature.
+The latter is typically missing in conventional papers, and this makes it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.
+
+The objective of this paper and associated code is to present a systematic and comprehensive description of the data processing steps for calibrated broadband echosounder data.
+The steps include pulse compression, target strength as a function of frequency [TS(f)], and volume backscattering strength as a function of frequency [Sv(f)].
+The intention is that the code will be used to validate implementations in various echosounders and post processing software, provide a starting point for further developing signal processing, and serve as a learning resource for broadbanded acoustic data processing.
+We also note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.
 
@@
 \section{Discussion}
+
+The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps.
+The inclusion of a corresponding code ensures a complete description.
+One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code may then serve as a starting point for further development.
+We also envision that the paper and code will be used both for educational purposes and for understanding the standard signal processing.
+
+Choices
+
+The contribution includes all the minor details required for the processing steps to work in practice.
+This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows.
+How to handle gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices.
+When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases, and, in our implementation, we simply chose to remove the edge cases.
+We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
+The signal processing also assumes a matched load impedance for the transducer and transceiver.
+Our objective is not to provide an evaluation of these choices, but to serve as a reference implementation when benchmarking the choices.
+
+$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals.
+These features includes size, target classification and is used to extract behaviour through tracking.
+In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified.
+For this to work in proactive a robust single target detector is required.
+Several single echo detection algorithms exist, and various single echo detection algorithms may be required depending on situation.
+Typical SED algorithms are based on traditional single frequency pulses, and utilizing the information in the broadband data improved SEDs may be envisioned, but this is outside the scope of this paper.
+
+$\sv(\freqsym)$ is a key parameter for echo integration.
+To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range.
+The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution.
+Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform.
+Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical operating frequencies at short ranges and between the beginning and the end of the window at longer ranges.
+The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
+The choice of window also allow the data to be cast on a predefined range-frequency grid, which can be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
+
+The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals.
+Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
+We note that these methods do not provide an operational method to estimate $\teff$ or $\Phi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
+
+A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented.
+Providing an open-source resource for anyone interested in learning and developing the techniques further.
+
+Old Discussion below
+
 The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.

### 2022-04-19 Paper/Manuscript.tex
Removed old discussion

@@
 
-Choices
-
 The contribution includes all the minor details required for the processing steps to work in practice.
@@
 
-Old Discussion below
-
-The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years.
-For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret_diel_2020} and to detect objects adjacent to boundaries \citep{lavery2017}.
-The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
-
-In this implementation (ref code) several choices have been made.
-Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature, where other approaches could been equally valid.
-These include, as examples,
-\begin{itemize}
-\item How to handle calibration, interpolation, missing calibration
-\item Transmit pulse
-\item Tapering of the transmit pulse
-\item Decimation factors and filters
-\item When Convoling the recieved signal with the transmit pulse - edge cases
-\item Assuemd four sector transducer
-\item Matched load - hvor stor effekt å gå utover 
-\item Assumed an impedance (trancuder)
-\item Effective pulse duration...
-\item Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...
-\item Window size
-\item Edge cases - removed
-\end{itemize}
-
-To estimate $\sv$ as a function of frequency a Fourier transform is used, repeatedly applied via a sliding window in range.
-However, the duration of this sliding window can be so long that the difference in spherical spreading loss compensation ($r_c^2$, implemented as the $20\log_{10}(\range_c)$ term in Eq.
-\ref{eq:Sv}) from the beginning of the window to the end can be significant, particularly for short range measurements.
-Thus, compensation for spreading loss is performed before applying the discrete Fourier transform.
-Absorption loss compensation is also range dependent (and frequency dependent), but since absorption loss compensation is insignificant for typical operating frequencies at short ranges and the difference in absorption loss compensation between the beginning and the end of the sliding window is insignificant at longer ranges, compensation for absorption loss is performed after applying the discrete Fourier transform.
-
-...
-
-Modern machine learning techniques are well suited for further processing of broadband echosounder data.
-Deep learning algorithms, including deep clustering, provide powerful methods to tease out patterns from high dimensional and complex data.
-By providing similar step sizes across frequencies it is possible to adapt this code to for an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
-
-The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals.
-Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013, lavery2017}.
-We note that these methods do not provide an operational method to estimate $\teff$ or $\eqang(\freqsym)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
-
-The processing equations and methodology presented in this paper have been implemented in version 1.12.4 and earlier of the \ek{} software.
-
 \section{Conclusion}

### 2022-04-19 Paper/Manuscript.tex
Changes to TS(f) and Sv(f) for frequency parameter

@@
 \end{equation}
-%
-\hl{The corresponding frequencies f\_m\_t (?) are calculate as follows...
-
-See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...}
 
@@
 
-Target strength can then be estimated using Eq. \ref{eq:TS}, but with $\freqsym$ replaced by the discrete index of frequency, $\samplesymf$:
+Target strength can then be estimated using Eq. \ref{eq:TS}:
 \begin{equation}
 \label{eq:TS_f}
-\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\samplesymf)\range_t 
-- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along_t,\athw_t,f_\samplesymf)}{16\pi^2} \right).
+\ts(\freqsym) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\freqsym)\range_t - 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,f)}{16\pi^2} \right)
+\end{equation}
+where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using
+\begin{equation}
+\label{eq:m(f)}
+\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor \mod \ndft
 \end{equation}
@@
 \ref{fi:TS}), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
-
-\hl{Comment from Nils Olav: Using $m$ as an argument to a function is not ideal.
-If we use the function argument perhaps using $f_m$ would be more logical, i.e.
-$\lambda(f_m)$.
-Alternatively using $lambda_m$ is possible.
-The latter is perhaps more in line with the code since the parameter is an array windexed with $m$ and not a function.
-$Psi(f)$, on the the other hand is derived from a function, see below.}
 
@@
 
-\hl{TODO: Explain edge cases.}
-
 The normalised Hann window, $\hannwnorm$, is given by: 
@@
 
-%
-\hl{TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.}
-
 The normalized discrete Fourier transform of the windowed data, $\ypcvolumenormf(\samplesymf)$, is then given by:
@@
 Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
+
 %
@@
 \label{eq:Sv_FFT}
-\sv(\samplesymf) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\samplesymf) \range_c \\
-- 10\log_{10}\left( \frac{\ptxe \wlen^2(\samplesymf) \cw \tslide \eqang(\samplesymf) \gainzero^2(\samplesymf) }{32\pi^2} \right).
+\sv(\freqsym) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\freqsym) \range_c - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym) }{32\pi^2} \right).
+\end{equation}
+where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using
+\begin{equation}
+\label{eq:m(f)}
+\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor \mod \ndft
 \end{equation}

### 2022-04-20 Paper/Manuscript.tex
Add files via upload

@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
+\caption{\label{fi:TS}The discrete fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b).
+The normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c) and finally the estimated TS(f) (d).}
 \end{figure}

### 2022-04-20 Paper/Manuscript.tex
Various modifications in paper

@@
 \author{Lars Nonboe Andersen}
-\email{lars.nonboe.andersen@kongsberg.com}
+\email{lars.nonboe.andersen@km.kongsberg.com}
 \affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}
@@
 The steps include pulse compression, target strength as a function of frequency [TS(f)], and volume backscattering strength as a function of frequency [Sv(f)].
-The intention is that the code will be used to validate implementations in various echosounders and post processing software, provide a starting point for further developing signal processing, and serve as a learning resource for broadbanded acoustic data processing.
+The intention is that the code will be used to validate implementations in various relevant data processing software, provide a starting point for further developing signal processing, and serve as a learning resource for broadbanded acoustic data processing.
 We also note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.
@@
 The code accompanying this paper is written in Python and is available as supplementary materials.
-All processing steps with respective figures in the paper can be reproduced by running main script (Main.py).
+All processing steps with respective figures in the paper can be reproduced by running the main scripts (TSf.py and Svf.py).
 Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field.
@@
 
-In the example a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}).
-This is found in the class \verb|Calculation.generateIdealWindowedTransmitSignal|, where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
+In the example a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
 A tapering of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
@@
 \caption{\label{fi:ytx} Ideal and enveloped transmit pulse between 34 kHz to 45 kHz with a slope of 0.057, as implemented in the EK80.
-The orange curve illustrate the effect of setting the slope to 0.5.}
+The orange curve illustrates the effect of setting the slope to 0.5.}
 \end{figure}
@@
 We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
-The signal processing also assumes a matched load impedance for the transducer and transceiver.
-Our objective is not to provide an evaluation of these choices, but to serve as a reference implementation when benchmarking the choices.
+When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length $\ndft$ of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
+Our objective is not to provide an evaluation of all these choices, but to serve as a reference implementation when benchmarking the choices.
 

### 2022-04-20 plots.py
TS figures are now in kHz

@@ -38,12 +38,12 @@ def plotfir(filter_v, f_s_dec_v, f_0, f_1):
     G1l = np.append(G1l, G1)
     G1l = np.append(G1l, G1)
     plt.figure()
-    plt.plot(F0, G0,
-             F1l, G1l,
-             [f_0, f_1], [-140, -140])
-    plt.xlabel('frequency (Hz)')
+    plt.plot(F0/1000, G0,
+             F1l/1000, G1l,
+             [f_0/1000, f_1/1000], [-140, -140])
+    plt.xlabel('frequency (kHz)')
     plt.ylabel('Gain (dB)')
-    plt.xlim([50000, 210000])
+    plt.xlim([50000/1000, 210000/1000])
     plt.savefig('./Paper/Fig_fir.png',dpi=300)
 
 
@@ -95,17 +95,17 @@ def plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red
 
 def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     fig, axs = plt.subplots(5)
-    axs[0].plot(f_m, np.abs(Y_pc_t_m))
+    axs[0].plot(f_m/1000, np.abs(Y_pc_t_m))
     axs[0].set_ylabel('Y_tilde_pc_t_m')
-    axs[1].plot(f_m, np.abs(Y_mf_auto_red_m))
+    axs[1].plot(f_m/1000, np.abs(Y_mf_auto_red_m))
     axs[1].set_ylabel('Y_mf_auto_red_m')
-    axs[2].plot(f_m, np.abs(Y_tilde_pc_t_m))
+    axs[2].plot(f_m/1000, np.abs(Y_tilde_pc_t_m))
     axs[2].set_ylabel('Y_tilde_pc_t_m')
-    axs[3].plot(f_m,
+    axs[3].plot(f_m/1000,
                 g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
     axs[3].set_ylabel('gain')
-    axs[4].plot(f_m, TS_m)
-    axs[4].set_xlabel('f (Hz)')
+    axs[4].plot(f_m/1000, TS_m)
+    axs[4].set_xlabel('f (kHz)')
     axs[4].set_ylabel('TS(f)')
     plt.savefig('./Paper/Fig_TS.png',dpi=300)
 

### 2022-04-20 plots.py
Figure 11

@@ -96,16 +96,15 @@ def plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red
 def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     fig, axs = plt.subplots(5)
     axs[0].plot(f_m/1000, np.abs(Y_pc_t_m))
-    axs[0].set_ylabel('Y_tilde_pc_t_m')
+    axs[0].set_ylabel(r'$Y_{pc,t}(m)$')
     axs[1].plot(f_m/1000, np.abs(Y_mf_auto_red_m))
-    axs[1].set_ylabel('Y_mf_auto_red_m')
+    axs[1].set_ylabel(r'$Y_{mf,auto,red}(m)$')
     axs[2].plot(f_m/1000, np.abs(Y_tilde_pc_t_m))
-    axs[2].set_ylabel('Y_tilde_pc_t_m')
-    axs[3].plot(f_m/1000,
-                g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
+    axs[2].set_ylabel(r'$\tilde{Y}_{pc,t}(m)$')
+    axs[3].plot(f_m/1000,g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
     axs[3].set_ylabel('gain')
     axs[4].plot(f_m/1000, TS_m)
-    axs[4].set_xlabel('f (kHz)')
+    axs[4].set_xlabel('f (Hz)')
     axs[4].set_ylabel('TS(f)')
     plt.savefig('./Paper/Fig_TS.png',dpi=300)
 

### 2022-04-20 plots.py
Attemt to fix plotTS

@@ -96,11 +96,13 @@ def plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red
 def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     fig, axs = plt.subplots(5)
     axs[0].plot(f_m/1000, np.abs(Y_pc_t_m))
-    axs[0].set_ylabel('Y_tilde_pc_t_m')
+    axs[0].set_ylabel('|$Y_{pc,t}(m)$|')
+    #axs[0].set_ylabel('Y_tilde_pc_t_m')
     axs[1].plot(f_m/1000, np.abs(Y_mf_auto_red_m))
-    axs[1].set_ylabel('Y_mf_auto_red_m')
+    axs[1].set_ylabel('|$Y_{mf,auto,red}(m)$|')
+    #axs[1].set_ylabel('Y_mf_auto_red_m')
     axs[2].plot(f_m/1000, np.abs(Y_tilde_pc_t_m))
-    axs[2].set_ylabel('Y_tilde_pc_t_m')
+    axs[2].set_ylabel('|$\\tilde{Y}_{pc,t}(m)$|')
     axs[3].plot(f_m/1000,
                 g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
     axs[3].set_ylabel('gain')

### 2022-04-20 Paper/Manuscript.tex
Updated Figure 11

@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS}The discrete fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b).
-The normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c) and finally the estimated TS(f) (d).}
+\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c) and the estimated TS(f) (d).}
 \end{figure}
@@
 
-As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $S_v(m)$.
+As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $S_v(m)$.
 

### 2022-04-20 plots.py

@@ -104,7 +104,7 @@ def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     axs[3].plot(f_m/1000,g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
     axs[3].set_ylabel('gain')
     axs[4].plot(f_m/1000, TS_m)
-    axs[4].set_xlabel('f (Hz)')
+    axs[4].set_xlabel('f (kHz)')
     axs[4].set_ylabel('TS(f)')
     plt.savefig('./Paper/Fig_TS.png',dpi=300)
 

### 2022-04-21 Paper/Manuscript.tex
Now with references, figure 11 with labeling

@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c) and the estimated TS(f) (d).}
+\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c), gain (d) and the estimated TS(f) (e).}
 \end{figure}

### 2022-04-21 plots.py

@@ -94,6 +94,14 @@ def plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red
 
 
 def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
+    
+    
+    def text_coords(axs=None,scalex=0.9,scaley=0.9):
+        xlims = axs.get_xlim()
+        ylims = axs.get_ylim()
+        return {'x':scalex*np.diff(xlims)+xlims[0],
+                'y':scaley*np.diff(ylims)+ylims[0]}
+    
     fig, axs = plt.subplots(5)
     axs[0].plot(f_m/1000, np.abs(Y_pc_t_m))
 
@@ -107,6 +115,15 @@ def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     axs[4].plot(f_m/1000, TS_m)
     axs[4].set_xlabel('f (kHz)')
     axs[4].set_ylabel('TS(f)')
+
+    scalex = [0.02,0.02,0.02,0.02,0.02]
+    scaley = [0.75,0.75,0.75,0.75,0.75]
+    labels = ['(a)','(b)','(c)','(d)','(e)']
+
+    # f,ax = plt.subplots(2,2)
+    for sx,sy,a,l in zip(scalex,scaley,np.ravel(axs),labels):
+        a.text(s=l,**text_coords(axs=a,scalex=sx,scaley=sy))
+    
     plt.savefig('./Paper/Fig_TS.png',dpi=300)
 
     # Store TS(f) and f for further analysis

### 2022-04-25 Paper/Manuscript.tex
Edits for readability in Introduction

@@
 
-Active acoustics is an efficient method for remote sensing of marine ecosystems.
-These methods have been used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}.
-As early as 1935, Oscars Sund observed the distribution of spawning cod in the Lofoten area using a single beam echosounder \citep{sund_echo_1935}.
+Active acoustics is an efficient method for remote sensing of marine ecosystems and is used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}.
+As early as 1935, Oscar Sund observed the distribution of spawning cod in the Lofoten area using a single beam echosounder \citep{sund_echo_1935}.
 The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}.
-More recently, the methods have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
+More recently, active acoustics have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
 
 Several different configurations of acoustic instruments are available, including multibeam sonars \citep{gerlotto_two_1999}, synthetic aperture sonar, omnidirectional sonars \citep{misund_improved_1996}, acoustic imaging sonars \citep{jaffe_ftv_1995} as well as single- and multibeam echosounders \citep{trenkel_new_2008}.
-Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and methodology for using these systems are well developed.
+Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and the methodologies for using these systems are well developed.
 The ability for accurate calibration of the system has made these especially useful for providing data for fisheries management \citep{Simmonds2005Fisheries}.
 
-Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems.
+Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), compared to conventional single frequency systems.
 This provides significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
 
-Broad banded echosounders have been pursued actively for ecology in particularly based on the promise of improved categorization of species or group of species.
-Using high frequency broadband acoustics (150-600 kHz) Lavery et al \citep{lavery_measurements_2010} showed that broadband helped reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
-\citep{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) with frequency content covering the resonance of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages.
-\citep{blanluet_characterization_2019} used broadband, with ground truthing, to identify that the composition of two Sound Scattering Layers (SSL) detected in the Bay of Biscay in springtime was characterized and identifying fine scale heterogeneity within the SSLs.
-While unsuccessful in discriminating between several species of fishes with large  swimbladders during the Alaska pollock survey, \citep{bassett_broadband_2018} showed that broadband may be helpful in characterizing smaller fishes with swimbladder and euphausiids.
-\citep{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband data (45-170 kHz), while attempts to classify using narrowband multifrequency failed.
+Broadband echosounders have been actively investigated for ecological research, based on the expectation of improved categorization of species or group of species.
+Using high frequency broadband acoustics (150-600 kHz) \citet{lavery_measurements_2010} showed that broadband signals helped to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
+\citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages of broadband signals.
+\citet{blanluet_characterization_2019} used broadband, with ground truthing, to identify that the composition of two sound scattering layers (SSL) detected in the Bay of Biscay in springtime was able to be characterized, and also identified fine scale heterogeneity within the SSLs.
+While unsuccessful at discriminating between several species of fishes with large swimbladders during the Alaska pollock survey, \citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids.
+\citet{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz), while attempts to classify using multiple narrowband signals failed.
 
-The increased spatial resolution is another aspect of broadband acoustics that are actively being used both for improved categorization and increased ability to resolve targets near each other or near boundaries such as the seafloor.
-\citep{lavery2017} explored different broadband pulse shapes to increase successful resolution of single targets in proximity as well as near boundaries in tank experiments.
-By applying high frequency broadband pulses to fish-like artificial targets, \citep{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment.
-Utilizing the increased spatial resolution \citep{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.
+The increased spatial resolution is another aspect of broadband acoustics that is being actively used both for improved categorization and the increased ability to resolve targets near each other or near boundaries such as the seafloor.
+\citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments.
+By applying high frequency broadband pulses to fish-like artificial targets, \citet{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment.
+Utilizing the increased spatial resolution \citet{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.
 
 An equally interesting outcome from the increased resolution and broad frequency bands is the ability to infer behavioural information on both single individuals, groups of individuals, and interaction between individuals or groups.
-\citep{Traykovski1998Effect} was able to extract orientation information about krill from using a combination of high frequency broadband acoustics and acoustic modelling.
-Fine-scale observations of single targets facilitated the study of interactions between prey and predators, capelin and cod, in the Barents sea (\citep{skaret_diel_2020}).
+\citet{Traykovski1998Effect} was able to extract orientation information about krill by using a combination of high frequency broadband acoustics and acoustic modelling.
+Fine-scale observations of single targets facilitated the study of interactions between prey and predators (capelin and cod) in the Barents sea \citep{skaret_diel_2020}.
 
 There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
+Commercially available systems are now produced by several manufacturers.
 
-The theoretical foundation for signal processing comes from radar applications and is earlier documented for broadband echosounders \citep{stanton2008}.
+The theoretical foundation for signal processing comes from radar applications and has been documented for broadband echosounders \citep{stanton2008}.
 To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}.
 Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
-Different send pulses can be envisioned as well as various methods for acoustic target classification.
-When translating equations into code, several choices must be made.
-Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature.
-The latter is typically missing in conventional papers, and this makes it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.
+Different transmission pulse forms can be envisioned as well as various methods for acoustic target classification.
+When translating equations into computer processing code, several choices must be made - some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature.
+The latter is typically missing in the literature, making it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.
 
 The objective of this paper and associated code is to present a systematic and comprehensive description of the data processing steps for calibrated broadband echosounder data.
-The steps include pulse compression, target strength as a function of frequency [TS(f)], and volume backscattering strength as a function of frequency [Sv(f)].
-The intention is that the code will be used to validate implementations in various relevant data processing software, provide a starting point for further developing signal processing, and serve as a learning resource for broadbanded acoustic data processing.
-We also note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.
+The steps include pulse compression, target strength as a function of frequency [$\ts(\freqsym)$], and volume backscattering strength as a function of frequency [$\sv(\freqsym)$].
+The intention is that the code will be used to validate implementations in various relevant data processing software, provide a starting point for further developing active acoustic broadband signal processing, and to serve as a learning resource.
 
@@
 Providing an open-source resource for anyone interested in learning and developing the techniques further.
+We note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.
 

### 2022-04-26 Paper/Manuscript.tex
Edits to sections II, III, and IV

@@
 
+\newcommand{\code}[1]{\texttt{#1}}
+
 \begin{document}
@@
 The latter is typically missing in the literature, making it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.
+By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.
 
@@
 \subsection{Accompanying code}
-The code accompanying this paper is written in Python and is available as supplementary materials.
-All processing steps with respective figures in the paper can be reproduced by running the main scripts (TSf.py and Svf.py).
-Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field.
-By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.
+The code accompanying this paper is written in the Python programming language and is available in the supplementary materials.
+All processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
+Without loss of generality, we use the Simrad EK80 echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
 
@@
 \subsection{System overview}
-A basic echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals.
-During transmission the program defines the signals which are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water.
+A basic quantitative echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals.
+During transmission the program defines the signals that are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water.
 The acoustic signals propagate through the water, are reflected or scattered by objects in the water, and propagate back to the transducer.
-During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, and processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig.
+During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig.
 \ref{fi:ek_sys}).
-Many types of transmit signals are feasible - this paper considers only the transmission of linear frequency modulated signals (also known as linear chirps).
+Many types of transmit signals are feasible - this paper considers only linear frequency modulated signals (also known as linear chirps).
 
@@
 The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is a sample index in the discrete time domain.
-Typical BB pulses are linear upsweep pulses windowed by an envelope function.
-The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain analogue signal $\ytxe(\timesym)$.
-The analogue and amplified signal is passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water column, where $\timesym$ is the time for the (analog) signal.
+Typical broadband pulses are linear upsweep pulses windowed by an envelope function.
+The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain an analogue signal $\ytxe(\timesym)$.
+The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water, where $\timesym$ is the time for the (analog) signal.
 For a split-aperture echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
@@
 \includegraphics[width=16cm]{Fig_ytx}
-\caption{\label{fi:ytx} Ideal and enveloped transmit pulse between 34 kHz to 45 kHz with a slope of 0.057, as implemented in the EK80.
-The orange curve illustrates the effect of setting the slope to 0.5.}
+\caption{\label{fi:ytx} Linear chirp pulses from 34 kHz to 45 kHz with a slope of 0.057 (blue) and 0.5 (orange), as implemented in the EK80.}
 \end{figure}
@@
 To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hfl(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$.
-The individual filter coefficients for each filter and decimation stage are indexed by $i$.
+The individual filter coefficients for each filter and decimation stage are indexed by $\genidxsym$.
 The output signal from each channel, $\channelsym$, from each filter and decimation stage, $\stagesym$, is then given by:
@@
 In the example $N_v=2$.
-The frequency responses of the filters are shown in Figure~\ref{fi:fir}, and the corresponding filter coefficients and decimation factors are found in the test data set.
+The frequency responses of the filters are shown in Figure~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set.
 
@@
 The original sample data $\yrxorg(\samplesymt,\channelsym)$ are not available in the EK80 data files.
-Instead, the resulting filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
-The data are recorded together with additional data such as from position and motion sensors and system configuration data in raw data files for display and analysis by processing software.
-The filter parameters are in this paper used to generate the transmit pulse for the matched filter, see below.
+Instead, the filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
+The data are recorded in computer data files for display and analysis by processing software.
+Additional information, such as that from position and motion sensors and system configuration data is also included in the files.
 
@@
 \includegraphics[width=16cm]{Fig_y_mf_n}
-\caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal $\ymf(\samplesymt)$ used by the pulse compression.}
+\caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, that is used for the pulse compression.}
 \end{figure}
@@
 The received power samples are then used to estimate target strength and volume backscattering strength.
-For estimating received power samples, the mean signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, will be used:
+For estimating received power samples, the mean signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, is used:
 %
@@
 The transceiver measures voltage over a load, $\zrxe$, connected in series with the transducer impedance, $\ztde$.
-When calculating various acoustic properties a system gain parameter will be used which assumes a matched receiver load.
+When calculating various acoustic properties, a system gain parameter will be used which assumes a matched receiver load.
 The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig.
@@
 %
-where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}.
-The inverse sine is indicated by $\arcsin$, the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$.
+where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}.
+The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$.
 As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
@@
 
-To illustrate each step of the process to obtain $TS(m)$ we use a single ping from a data file where a single metallic calibration sphere with 35 mm diameter (WC35) is suspended approximately 6 m below a 120 kHz transducer (fig.~\ref{Fig_TS_echogram}).
+To illustrate each step of the process in obtaining $TS(m)$ we use a single ping from a data file where a 35 mm diameter tungsten carbide calibration sphere (WC35) was suspended approximately 6 m below a 120 kHz transducer (Fig.~\ref{Fig_TS_echogram}).
 
@@
 This is typically achieved through a single echo detection algorithm (SED).
-Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets.
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
-\ref{fi:TS}.
+Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes that occur in actual targets (as opposed to ideal point targets).
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
+\ref{fi:TS}).
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
@@
 \caption{\label{fi:SED} The $\prxe(\samplesymt)$ power (upper) and split beam angles ($\along_t$ and $\athw_t$) (middle) for the single target.
-The vertical line corresponds to the range $r_t$ for the single target.
+The orange vertical line corresponds to the range $r_t$ for the single target.
 The $\ymfautored(\samplesymt)$ (lower) is the autocorrelation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.}
@@
 
-As a metallic sphere is a rather simple and ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $S_v(m)$.
+As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
 
@@
 \includegraphics[width=16cm]{Fig_Sv_echogram}
-\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file. The red vertical line indicates ping to be processed in this section.}
+\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($samplesymf$) and range ($samplesymt$) for a echogram raw data file.
+The red vertical line indicates the ping processed in this section.}
 \end{figure}
@@
 
-By selecting a set of centre samples $n$, $Sv$ values can be presented as a function of range ($n$) and frequency ($m$) for each ping.
-The range for centre samples $n$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
-This can be useful when combining the $Sv(m)$ across a range of transducers.
+By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($freqsym$) for each ping.
+The range for centre samples $\samplesymt$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
+This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
 In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).
@@
 \includegraphics[width=16cm]{Fig_Sv_m_n}
-\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($m$) and range ($n$) for a single ping}
+\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a single ping}
 \end{figure}
 
-For acoustic abundance estimation and classification purposes it is common to integrate the Sv over a range, in our example we integrate $Sv$ from 15 to 34 m covering a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}).
-We note that we normally require averaging over several pings to obtain an unbiased estimate of $Sv$ for echo integration purposes.
-Here we present the $Sv(m)$ over the layer for our example ping for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
-Even though this is for a single ping, we still observe a positive slope of the frequency response indicative of non-swimbladdered fish.
+For acoustic abundance estimation and classification purposes it is common to integrate the $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}).
+It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
+Even though this is for a single ping it is still possible to observe a positive slope of the frequency response that is indicative of non-swimbladdered fish.
 
@@
 \includegraphics[width=16cm]{Fig_Sv_avg}
-\caption{\label{Fig_Sv_avg} $Sv$ as a function of frequency ($m$) averaged over a range interval covering a fish school }
+\caption{\label{Fig_Sv_avg} $\sv$ as a function of frequency ($\samplesymf$) averaged over a range interval covering a fish school.}
 \end{figure}

### 2022-04-26 Paper/Manuscript.tex
Edits to Discussion & tweaks to rest of doc.

@@
 All processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
-Without loss of generality, we use the Simrad EK80 echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
+Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
 
@@
 
-As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
+As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
 
@@
 Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
-
 %
@@
 \end{equation}
-where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using
-\begin{equation}
-\label{eq:m(f)}
-\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor \mod \ndft
-\end{equation}
+where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using Eq. {\ref{eq:m(f)}}.
 
@@
 This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
-In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).
+In our example we have simply chosen n to be on the same order of the range resolution (Fig.~\ref{Fig_Sv_m_n}).
 
@@
 For acoustic abundance estimation and classification purposes it is common to integrate the $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}).
-It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (fig.~\ref{Fig_Sv_avg}).
+It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (Fig.~\ref{Fig_Sv_avg}).
 Even though this is for a single ping it is still possible to observe a positive slope of the frequency response that is indicative of non-swimbladdered fish.
@@
 The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example.
-In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (fig.~\ref{Fig_Sv_echogram}).
+In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_echogram}).
 
@@
 The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps.
-The inclusion of a corresponding code ensures a complete description.
-One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code may then serve as a starting point for further development.
-We also envision that the paper and code will be used both for educational purposes and for understanding the standard signal processing.
+The inclusion of corresponding computer code ensures a more complete description.
+One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code can serve as a starting point for this.
+We also envision that the paper and code be used both for educational purposes and for understanding the standard broadband signal processing.
 
@@
 This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows.
-How to handle gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices.
-When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases, and, in our implementation, we simply chose to remove the edge cases.
+Handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices.
+When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases and in our implementation we chose to remove the edge cases.
 We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
-When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length $\ndft$ of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
+When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
 Our objective is not to provide an evaluation of all these choices, but to serve as a reference implementation when benchmarking the choices.
@@
 $\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals.
-These features includes size, target classification and is used to extract behaviour through tracking.
+These features includes size, target classification and can be used to extract behaviour through tracking.
 In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified.
-For this to work in proactive a robust single target detector is required.
-Several single echo detection algorithms exist, and various single echo detection algorithms may be required depending on situation.
-Typical SED algorithms are based on traditional single frequency pulses, and utilizing the information in the broadband data improved SEDs may be envisioned, but this is outside the scope of this paper.
+For this to work well in practise, a robust single target detector is required.
+Several SED algorithms exist and differing such algorithms may be required depending on the situation.
+Typical SED algorithms are based on traditional single frequency pulses and by utilizing the additional information in broadband echoes improved SEDs may be envisioned, but this is outside the scope of this paper.
+One existing improvement is that broadband pulse compression is itself a type of single target detector.
 
@@
 Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform.
-Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical operating frequencies at short ranges and between the beginning and the end of the window at longer ranges.
+Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies at short ranges and between the beginning and the end of the window at longer ranges.
 The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
-The choice of window also allow the data to be cast on a predefined range-frequency grid, which can be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
+The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
 
-The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals.
+The formulation presented in this paper requires several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$.
 Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
@@
 
-A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented.
-Providing an open-source resource for anyone interested in learning and developing the techniques further.
-We note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.
+A set of equations and associated computer code for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented.
+This provides a resource for those interested in learning and further developing broadband processing techniques.
+The processing equations and methodology presented in this paper are similar to those implements in version 1.12.4 and earlier of the \ek{} software.
 

### 2022-04-26 Paper/Manuscript.tex
fix a few missing symbols uses

@@
 \end{figure}
-
+%
 A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
@@
 \includegraphics[width=16cm]{Fig_Sv_echogram}
-\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($samplesymf$) and range ($samplesymt$) for a echogram raw data file.
+\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a echogram raw data file.
 The red vertical line indicates the ping processed in this section.}
@@
 \includegraphics[width=16cm]{Fig_Sv_m_n}
-\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a single ping}
+\caption{\label{Fig_Sv_m_n} $\sv$ as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a single ping.}
 \end{figure}

### 2022-05-05 Paper/Manuscript.tex
Implemented Haralds comments

@@
 
-In the example a linear sweep enveloped by a Hann function is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
+In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
 A tapering of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
@@
 \includegraphics[width=16cm]{Fig_fir}
-\caption{\label{fi:fir} The frequency (power) response of the filters in our test set.}
+\caption{\label{fi:fir} The frequency (power) response of the filters in our test set.
+The blue and orange curve represents the filter response of the first and second filter.
+The horizontal green line indicates the frequency range of the send pulse.}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_TS_echogram}
-\caption{\label{Fig_TS_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file.
-The red vertical line indicates ping to be processed in this section for TS(m).
-Single target calibration sphere is seen at approximately 6 m range and the seafloor at approximately 12 m.}
+\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere.
+The sphere is located at approximately 6 m range and the seafloor at approximately 12 m.
+The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
 \end{figure}
@@
 The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
-\ref{fi:TS}).
+\ref{fi:SED}).
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
@@
 \includegraphics[width=16cm]{Fig_Sv_echogram}
-\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a echogram raw data file.
-The red vertical line indicates the ping processed in this section.}
+\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file.
+A fish school is seen between 15 and 35 m range, and the sea flor is seen at approximately 50 m.
+The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
 \end{figure}
@@
 where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading.
-A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$.
+A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hanning window, $\hannw(\genidxsym)$.
 The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution.
@@
 
-The normalised Hann window, $\hannwnorm$, is given by: 
+The normalised Hanning window, $\hannwnorm$, is given by: 
 %

### 2022-05-05 Paper/Manuscript.tex
Added rolfs and gavings comments, except figure comments.

@@
 is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.
-For an ideal system, i.e., no tapering at the rising and trailing edges of the transmitted signal, the effective pulse duration is the same as the transmit pulse duration.
+For an ideal system, i.e., no tapering at the rising and trailing edges of the in-water signal, the effective pulse duration is the same as the transmit pulse duration.
 
@@
 \caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere.
-The sphere is located at approximately 6 m range and the seafloor at approximately 12 m.
+The sphere is located at approximately 6 m range.
 The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
@@
 
-As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
+As a metallic sphere is a rather simple and ideal scatterer suitable for investigating TS of single scatterers, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
 
@@
 \caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file.
-A fish school is seen between 15 and 35 m range, and the sea flor is seen at approximately 50 m.
+The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea flor is seen at approximately 50 m.
 The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
@@
 The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
-The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
+The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by deep learning methods \citep[e.g.]{brautaset_acoustic_2020}.
 

### 2022-05-05 Paper/Manuscript.tex
Lars changes added.

@@
 Active acoustics is an efficient method for remote sensing of marine ecosystems and is used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}.
-As early as 1935, Oscar Sund observed the distribution of spawning cod in the Lofoten area using a single beam echosounder \citep{sund_echo_1935}.
+As early as 1935, \citet{sund_echo_1935} observed the distribution of spawning cod in the Lofoten area using a single beam echosounder.
 The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}.
@@
 \citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages of broadband signals.
-\citet{blanluet_characterization_2019} used broadband, with ground truthing, to identify that the composition of two sound scattering layers (SSL) detected in the Bay of Biscay in springtime was able to be characterized, and also identified fine scale heterogeneity within the SSLs.
+\citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to identify that the composition of two sound scattering layers (SSL) detected in the Bay of Biscay in springtime was able to be characterized, and also identified fine scale heterogeneity within the SSLs.
 While unsuccessful at discriminating between several species of fishes with large swimbladders during the Alaska pollock survey, \citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids.
-\citet{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz), while attempts to classify using multiple narrowband signals failed.
+\citet{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz), while attempts to classify species using multiple narrowband signals failed.
 
@@
 
-An equally interesting outcome from the increased resolution and broad frequency bands is the ability to infer behavioural information on both single individuals, groups of individuals, and interaction between individuals or groups.
-\citet{Traykovski1998Effect} was able to extract orientation information about krill by using a combination of high frequency broadband acoustics and acoustic modelling.
+An equally interesting outcome from the increased resolution and broad frequency information is the ability to infer behavioural information for single individuals, groups of individuals, and interaction between individuals or groups.
+\citet{Traykovski1998Effect} was able to extract orientation information for krill by using a combination of high frequency broadband acoustics and acoustic modelling.
 Fine-scale observations of single targets facilitated the study of interactions between prey and predators (capelin and cod) in the Barents sea \citep{skaret_diel_2020}.
@@
 To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}.
-Examples include adapting pulses to improve target separation as well as further parameters from the individual targets.
+Examples include adapting pulses to improve target separation and obtain additional parameters to describe the individual targets.
 Different transmission pulse forms can be envisioned as well as various methods for acoustic target classification.
@@
 The steps include pulse compression, target strength as a function of frequency [$\ts(\freqsym)$], and volume backscattering strength as a function of frequency [$\sv(\freqsym)$].
-The intention is that the code will be used to validate implementations in various relevant data processing software, provide a starting point for further developing active acoustic broadband signal processing, and to serve as a learning resource.
+The intention is that the code will be used as a starting point for implementations in various relevant data processing software and for further developing active acoustic broadband signal processing, and to serve as a learning resource.
 
@@
 \caption{\label{fi:ek_sys}Signal and data flow in the \ek{} system.
-An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage (lower middle) and display and analysis (lower right).
-Note that all signals are complex-valued time-series.}
+An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage (lower middle) and display and analysis (lower right).}
 \end{figure}
@@
 
-The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is a sample index in the discrete time domain.
+The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is the sample index in the discrete time domain.
 Typical broadband pulses are linear upsweep pulses windowed by an envelope function.
-The generated signal $\ytxe(\timesym)$ is converted to an analogue electric signal and amplified by the transceiver to obtain an analogue signal $\ytxe(\timesym)$.
-The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water, where $\timesym$ is the time for the (analog) signal.
+The generated signal is converted to an analogue electric signal $\ytxe(\timesym)$ and amplified by the transceiver to obtain the analogue signal $\ytxe(\timesym)$, where $\timesym$ is the time for the signal.
+The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water.
 For a split-aperture echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
@@
 In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
-A tapering of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
+A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
 \begin{figure}
@@
 %
-where $\yrx(\samplesymt,\channelsym,0)$ is set to $\yrxorg(\samplesymt,\channelsym)$, being the signal before decimation, $*$ indicates convolution and $\nstages$ is the total number of filter stages.
+where $\yrx(\samplesymt,\channelsym,0)$ is set to $\yrxorg(\samplesymt,\channelsym)$, being the signal before decimation, $*$ indicates convolution, $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$, and $\nstages$ is the total number of filter stages.
 The output signal from the final filter and decimation stage, $\yrx(\samplesymt,\channelsym,\nstages)$, is shortened to $\yrx(\samplesymt,\channelsym)$ for convenience.
@@
 %
-where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$ and $\downarrow$ indicates decimation by the factor $\decfac(\stagesym)$.
+where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$.
 The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n}).
@@
 is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.
-For an ideal system, i.e., no tapering at the rising and trailing edges of the in-water signal, the effective pulse duration is the same as the transmit pulse duration.
 
@@
 The received power samples are then used to estimate target strength and volume backscattering strength.
-For estimating received power samples, the mean signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, is used:
+For estimating received power samples, the average signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, is used:
 %
@@
 \along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
-\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right).
+\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right),
 \end{eqnarray}
 %
-where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}.
+where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:theta_phi}) and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}.
 The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$.
@@
 \end{figure}
-
+\clearpage
 \section{Target strength}
@@
 %
-where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
+where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption coefficient, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
 
 The point scattering strength, $\mysp(\samplesymt)$, is estimated by applying Eq.
-\ref{eq:TS} to the received digitized power samples using the on-axis gain value with $f$ set to the centre frequency of the broadband pulse, $\fc$:
+\ref{eq:TS} to the received digitized power samples using the on-axis gain value and $f$ set to the centre frequency of the broadband pulse, $\fc$:
 \begin{equation}
@@
 \ref{fi:TS}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
-The amplitude of the backscatter signal also clearly shows these nulls (Fig.
-\ref{fi:TS}), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.
 
@@
 
-As a metallic sphere is a rather simple and ideal scatterer suitable for investigating TS of single scatterers, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.
+To illustrate calculation of volume backscattering strength as a function of frequency,  $\sv(f)$, we use data collected from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer.
 
@@
 \sv(\samplesymt)  =  10\log_{10}(\prxe(\samplesymt)) + 20\log_{10}(\range_c(\samplesymt)) + 2\absorp(\fc)\range_c(\samplesymt) \\
-- 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \cw \teff \eqang(\fc) \gainzero^2(\fc)}{32\pi^2} \right),
+- 10\log_{10}\left( \frac{\ptxe \wlen^2(\fc) \cw \teff \eqang(\fc) \gainzero^2(\fc)}{32\pi^2} \right).
 \end{split}
 \end{equation}
-noting that $\sv(\samplesymt)$ is an average over frequency of all echoes received at sample $\samplesymt$.
+%noting that $\sv(\samplesymt)$ is an average over frequency of all echoes received at sample $\samplesymt$.
 In this case, the time window, $\tslide$, is the effective pulse duration, $\teff$, resulting from pulse compression.
@@
 This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
-In our example we have simply chosen n to be on the same order of the range resolution (Fig.~\ref{Fig_Sv_m_n}).
+In our example we have simply chosen the set of centre samples as the original range samples (Fig.~\ref{Fig_Sv_m_n}).
 
@@
 
-For acoustic abundance estimation and classification purposes it is common to integrate the $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}).
+For acoustic abundance estimation and classification purposes it is common to integrate $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}).
 It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (Fig.~\ref{Fig_Sv_avg}).
@@
 \includegraphics[width=16cm]{Fig_Sv_avg}
-\caption{\label{Fig_Sv_avg} $\sv$ as a function of frequency ($\samplesymf$) averaged over a range interval covering a fish school.}
+\caption{\label{Fig_Sv_avg} $\sv$ as a function of frequency averaged over a depth interval covering a fish school.}
 \end{figure}
 
-The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example.
+The trend for increasing $\sv$ with frequency is well-known for non-swimbladderred fish \citep{korneliussen2010} and is consistent with the trend observed in this example.
 In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_echogram}).
@@
 The contribution includes all the minor details required for the processing steps to work in practice.
-This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows.
-Handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices.
-When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases and in our implementation we chose to remove the edge cases.
+This includes steps well founded in the literature as well as practical and more ad-hoc choices.
+Handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering.
+When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases and in our implementation we chose to exclude the edge cases.
 We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
 When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
-Our objective is not to provide an evaluation of all these choices, but to serve as a reference implementation when benchmarking the choices.
+Our objective is not to provide an evaluation of all these choices, but to serve as a startin point.
 
 $\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals.
-These features includes size, target classification and can be used to extract behaviour through tracking.
-In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified.
-For this to work well in practise, a robust single target detector is required.
+Features includes size, target classification and, behaviour through tracking.
+In our implementation, calculation of $\ts(\freqsym)$ assumes that a single target has successfully been identified.
+This requires a robust single target detector.
 Several SED algorithms exist and differing such algorithms may be required depending on the situation.
 Typical SED algorithms are based on traditional single frequency pulses and by utilizing the additional information in broadband echoes improved SEDs may be envisioned, but this is outside the scope of this paper.
-One existing improvement is that broadband pulse compression is itself a type of single target detector.
 
@@
 To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range.
-The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution.
+The chosen size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution.
 Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform.
-Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies at short ranges and between the beginning and the end of the window at longer ranges.
+Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies between the beginning and the end of the window for relatively short range windows.
 The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
@@
 Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
-We note that these methods do not provide an operational method to estimate $\teff$ or $\Phi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
+We note that these methods do not provide an operational method to estimate $\teff$ or $\Psi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
 
-A set of equations and associated computer code for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented.
-This provides a resource for those interested in learning and further developing broadband processing techniques.
+A set of equations and associated computer code for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented along with example code, providing a resource for those interested in learning and further developing broadband processing techniques.
 The processing equations and methodology presented in this paper are similar to those implements in version 1.12.4 and earlier of the \ek{} software.
@@
 
-A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented, with reference to the \ek{} echosounder.
+A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented along with example code, with reference to the \ek{} echosounder.
 
@@
 
-The data associated with this article are available on request from the authors.
+The code and data associated with this article are available through GitHub through \url{https://github.com/CRIMAC-WP4-Machine-learning/CRIMAC-Raw-To-Svf-TSf}.
+The code and data for the pre-print is tagged version 0.9.
+The version for the printed paper is 1.0.
+Further developments will have higher version numbers.
 

### 2022-05-09 Paper/Manuscript.tex
Updated figures

@@
 \includegraphics[width=16cm]{Fig_ytx}
-\caption{\label{fi:ytx} Linear chirp pulses from 34 kHz to 45 kHz with a slope of 0.057 (blue) and 0.5 (orange), as implemented in the EK80.}
+\caption{\label{fi:ytx} Envelope of linear chirp pulses from 92 kHz to 158 kHz with a slope of 0.057 (blue) and 0.5 (orange).}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_theta_phi}
-\caption{\label{fi:theta_phi}Physical angles $\along$ and $\athw$.}
+\caption{\label{fi:theta_phi}The physical angles $\along$ and $\athw$ for the target strength example data (see below).
+The single target can be seen around sample number 1000 where the angles are less variable.}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_TS_echogram}
-\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere.
+\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range for a calibration sphere.
 The sphere is located at approximately 6 m range.
-The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
+The red vertical line indicates the ping that is used for illustrating the $\ts(\freqsym)$ processing.}
 \end{figure}
@@
 \includegraphics[width=16cm]{Fig_Sv_echogram}
-\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file.
-The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea flor is seen at approximately 50 m.
-The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
+\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for the raw data file used in the Sv(f) example.
+The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
+The red vertical line indicates the ping that is used for illustrating the $\sv(\freqsym)$ processing.}
 \end{figure}

### 2022-05-09 Svf.py

@@ -158,8 +158,8 @@ theta_n, phi_n = Calculation.calcAngles(
     gamma_phi_f_c)
 
 # Plots for paper
-if do_plot:
-    plotThetaPhi(theta_n, phi_n)
+#if do_plot:
+#    plotThetaPhi(theta_n, phi_n)
 
 
  #

### 2022-05-09 plots.py

@@ -1,7 +1,7 @@
 import matplotlib.pyplot as plt
 import numpy as np
 from Core.Calculation import Calculation
-
+from scipy.signal import hilbert
 
 def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
 
@@ -10,12 +10,12 @@ def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
         f_0, f_1, tau, f_s, .5)
 
     plt.figure()
-    plt.plot(t * 1000, y_tx_n, t * 1000, y_tx_n05slope)
+    plt.plot(t * 1000, np.abs(hilbert(y_tx_n)), t * 1000, np.abs(hilbert(y_tx_n05slope)))
     #plt.title(
     #    'Ideal windowed transmit pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
     #        .format(f_0 / 1000, f_1 / 1000, slope))
-    plt.xlabel('Time [ms]')
-    plt.ylabel('Normalised amplitude')
+    plt.xlabel('Time (ms)')
+    plt.ylabel('Envelope ()')
     plt.savefig('./Paper/Fig_ytx.png',dpi=300)
 
 
@@ -41,9 +41,9 @@ def plotfir(filter_v, f_s_dec_v, f_0, f_1):
     plt.plot(F0/1000, G0,
              F1l/1000, G1l,
              [f_0/1000, f_1/1000], [-140, -140])
-    plt.xlabel('Frequency [kHz]')
-    plt.ylabel('Gain [dB]')
-    plt.xlim([50000/1000, 210000/1000])
+    plt.xlabel('Frequency (kHz)')
+    plt.ylabel('Gain (dB)')
+    plt.xlim([0000/1000, 310000/1000])
     plt.savefig('./Paper/Fig_fir.png',dpi=300)
 
 
@@ -51,8 +51,8 @@ def plotymfn(y_mf_n):
     plt.figure()
     plt.plot(np.abs(y_mf_n))
     #plt.title('The absolute value of the filtered and decimated output signal')
-    plt.xlabel('Samples')
-    plt.ylabel('Normalised amplitude')
+    plt.xlabel('n ()')
+    plt.ylabel('$y_{mf}$ ()')
     plt.savefig('./Paper/Fig_y_mf_n.png',dpi=300)
 
 
[…]

### 2022-07-01 Paper/Manuscript.tex
Final update paper

@@
 \usepackage{color,soul}
+
+\usepackage{lineno}
+\linenumbers
 
@@
 \author{Lars Nonboe Andersen}
-\email{lars.nonboe.andersen@km.kongsberg.com}
 \affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}
@@
 \author{Geir Pedersen}
+\email{geir.pedersen@hi.no}
 \affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
@@
 \begin{abstract}
+1.
 The use of quantitative broadband echosounders for biological studies and surveys offers considerable advantages over narrowband echosounders.
 These include improved spectral-based target identification and significantly increased ability to resolve individual targets.
-Biological studies and surveys typically require accurate measures of backscatter strength and we present here a systematic and comprehensive explanation of how to derive quantitative estimates of target strength and volume backscattering, as a function of frequency from broadband echosounder signals.
+An understanding of current processing steps is required to fully utilize and further develop broadband acoustic methods in marine ecology.\linebreak
+2.
+We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength (TS(f)) and volume backscattering strength (Sv(f)) using data from the EK80 broadband scientific echosounder as examples.
+Although the overall processing steps are described and build on established methods from literature, multiple choices need to be made during implementation.
+\linebreak
+3.
+To highlight and discuss some of these choices and facilitating a common understanding within the community, we have also developed a code which will be made publicly available and open source.
+The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to TS(f) and Sv(f).
+\linebreak
+4.
+This code can serve as a reference for developing own code or implementation in existing processing pipelines, as an educational tool and as a starting point for further development of broadband acoustic methods in fisheries acoustics.\linebreak
+
+Keywords:  \linebreak
 
@@
 The code accompanying this paper is written in the Python programming language and is available in the supplementary materials.
-All processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
+All single ping processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
+Reproduction of figures 9 and 12 requires the original echosounder raw data and are not included.
 Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
@@
 \includegraphics[width=16cm]{Fig_fir}
-\caption{\label{fi:fir} The frequency (power) response of the filters in our test set.
+\caption{\label{fi:fir} The frequency response (filter gain) of the filters in our test set.
 The blue and orange curve represents the filter response of the first and second filter.
-The horizontal green line indicates the frequency range of the send pulse.}
+The horizontal green line indicates the frequency range of the transmit signal.}
 \end{figure}
@@
 \begin{figure}
+
 \includegraphics[width=16cm]{Fig_theta_phi}
@@
 
-To illustrate each step of the process in obtaining $TS(m)$ we use a single ping from a data file where a 35 mm diameter tungsten carbide calibration sphere (WC35) was suspended approximately 6 m below a 120 kHz transducer (Fig.~\ref{Fig_TS_echogram}).
+To illustrate calculation of target strength ($TS$) as a function of frequency, $TS(f)$, we use data collected on a 35 mm diameter tungsten carbide calibration sphere (WC35) suspended approximately 6 m below a 120 kHz transducer (Fig.~\ref{Fig_TS_echogram}).
 
@@
 \includegraphics[width=16cm]{Fig_TS_echogram}
-\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range for a calibration sphere.
+\caption{\label{Fig_TS_echogram} $\mysp$ as a function of ping number and range for a calibration sphere.
 The sphere is located at approximately 6 m range.
@@
 
-The power-budget equation (i.e., sonar equation) for a single target \citep[Formulation D, ][]{lunde2016} at frequency $\freqsym$ is:
+Generalizing the power-budget equation (i.e., sonar equation), given in e.g. \citep{lunde2016}, for broadband signals yields in logarithmic form $TS$ at frequency $\freqsym$:
 %
@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes that occur in actual targets (as opposed to ideal point targets).
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
-\ref{fi:SED}).
+The alongship angle $\along_n$, athwartship angle $\athw_n$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
+\ref{fi:theta_phi}).
 A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
 
-From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$.
+From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$ (Fig.~\ref{fi:SED}).
 Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c), gain (d) and the estimated TS(f) (e).}
+\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c), transducer gain $(10\log_{10}\gain^2(\along_t,\athw_t,\freqsym))$ (d), and the estimated TS(f) (e).}
 \end{figure}

### 2022-07-01 plots.py

@@ -42,7 +42,7 @@ def plotfir(filter_v, f_s_dec_v, f_0, f_1):
              F1l/1000, G1l,
              [f_0/1000, f_1/1000], [-140, -140])
     plt.xlabel('Frequency (kHz)')
-    plt.ylabel('Gain (dB)')
+    plt.ylabel('Filter gain (dB)')
     plt.xlim([0000/1000, 310000/1000])
     plt.savefig('./Paper/Fig_fir.png',dpi=300)
 
@@ -103,7 +103,7 @@ def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
         return {'x':scalex*np.diff(xlims)+xlims[0],
                 'y':scaley*np.diff(ylims)+ylims[0]}
     
-    fig, axs = plt.subplots(5, sharex=True, figsize=(6.4,7))
+    fig, axs = plt.subplots(5, sharex=True, figsize=(6.4,9))
     axs[0].plot(f_m/1000, np.abs(Y_pc_t_m))
 
     axs[0].set_ylabel(r'|$Y_{pc,t}$| ()', fontsize=8)
@@ -112,10 +112,10 @@ def plotTS(f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m):
     axs[2].plot(f_m/1000, np.abs(Y_tilde_pc_t_m))
     axs[2].set_ylabel(r'|$\tilde{Y}_{pc,t}$| ()', fontsize=8)
     axs[3].plot(f_m/1000,g_theta_phi_m)  # weird gain might be tracked down to  xml['angle_offset_alongship'] and xml['angle_offset_alongship']
-    axs[3].set_ylabel('Gain (dB)')
+    axs[3].set_ylabel('Transducer gain (dB)')
     axs[4].plot(f_m/1000, TS_m)
     axs[4].set_xlabel('f (kHz)')
-    axs[4].set_ylabel('TS(f) (dB re $1m^2$)')
+    axs[4].set_ylabel('TS (dB re $1m^2$)')
 
     scalex = [0.02,0.02,0.02,0.02,0.02]
     scaley = [0.75,0.75,0.75,0.75,0.75]

### 2023-09-08 Paper/Supplementary.tex
Add list of symbols in supplementary document
Move the symbol aliases to a separate file so that it can be included into multiple .tex files.

Add supplemenary.tex document that contains the list of symbols from a much earlier version of the manuscript so will have some differences to update (to come).

\documentclass[preprint,12pt,TurnOnLineNumbers]{JASAnew}
\usepackage{longtable}

\usepackage{lineno}
\linenumbers

\pdfoutput=1

\input{symbol_aliases.tex}

\begin{document}

\section{List of symbols}

Units are given for relevant symbols ([1] indicates a dimensionless quantity).

\begin{longtable}{p{0.25\linewidth} p{0.75\linewidth}}
Symbol & Description\\

$\atan(y,x)$ & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
$\cw$ & Sound speed in water [$m s^{-1}$].\\
$\decfac(\stagesym)$ & Integer filter decimation factor [1].\\
$\dft_\ndft(x)$ & The discrete Fourier transform of length $\ndft$ of $x$.\\
FT & The continuous Fourier transform. \\
$\freqsym$ & Frequency [Hz].\\
$\fc$ & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$ [Hz].\\
$\fn$ & Nominal operating frequency of a transducer [Hz].\\
$\fs$ & Analogue to digital sampling frequency [Hz].\\
$\fsdec$ & Decimated sampling frequency [Hz].\\
$\fstart$ & Start frequency of linear upsweep chirp pulse [Hz].\\
$\fstop$ & Stop frequency of linear upsweep chirp pulse [Hz].\\

$\gain(\along,\athw,\freqsym)$ & Transducer gain as a function of echo arrival angle in the beam and frequency [1].\\

$\gain(\along,\athw,\samplesymf)$ & Transducer gain as a function of echo arrival angle in the beam and frequency index [1].\\

$\gainzero(\freqsym)$ & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$ [1].\\

$\hfl(\genidxsym,\stagesym)$ & Complex valued receiving filter coefficients [1].\\

$\Im(x)$ & The imaginary part of $x$.\\

$\samplesymf$ & Sample index in frequency domain [1].\\

$\nchannels$ & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal [1].\\
$\nstages$ & Total number of filter stages [1].\\
$\nw$ & Number of samples used in the sliding Hanning window [1].\\
$\samplesymt$ & Sample index in time domain [1]\\

$\prxetf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$ [W].\\
$\prxetf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a single target [W].\\
$\prxevf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$ [W].\\
$\prxevf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a volume [W$\textrm{m}^2$].\\
$\prxe(\samplesymt)$  & Received electric power in a matched load [W].\\

$\ptxe$ & Transmitted electric power [W].\\
$\ptxauto(\samplesymt)$ & Square of the absolute value of the matched filter autocorrelation function [1].\\

$\Re(x)$ & The real part of $x$.\\
$\range$ & Distance from transducer [m].\\
$\rangeref$ & Reference distance [m]. \\
$\range(\samplesymt)$ & Distance from transducer [m].\\

$\mysp(\samplesymt)$  & Point scattering strength [dB re 1 $\textrm{m}^2$].\\
$\sv$  & Volume backscattering strength [dB re 1 $\textrm{m}^{-1}$].\\
$\sv(\freqsym)$ & Volume backscattering strength at frequency $\freqsym$ [dB re 1 $\textrm{m}^{-1}$].\\
$\sv(\samplesymt)$ & Volume backscattering strength at sample index $\samplesymt$ [dB re $\textrm{m}^{-1}$]\\

$\ts$ & Target strength [dB re 1 $\textrm{m}^2$].\\
$\ts(\freqsym)$ & Target strength at frequency $f$ [dB re 1 $\textrm{m}^2$].\\
$\ts(\samplesymf)$ & Target strength at frequency index $\samplesymf$ [dB re 1 $\textrm{m}^{2}$]\\

$\timesym$ &  Time [s].\\
$\tslide$ & Duration of sliding window for calculating volume backscattering strength [s].\\

$\channelsym$ & Channel number/transducer sector [1].\\
$u(\genidxsym)$ & The Heaviside step function [1]. \\
$V$ & Volume occupied by scattering targets [$\textrm{m}^3$].\\
$\stagesym$ & Filter stage [1].\\

$\hannw$ & The Hanning window function [1].\\
$\hannw(\genidxsym)$ & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$ [1].\\
$\hannwnorm(\genidxsym)$ & Normalised Hanning window [1].\\

$x^*$ & The complex conjugate of $x$.\\
$||x||$ & The $l^2$-norm of $x$, also known as the Euclidean norm.\\

$\ymfautof(\samplesymf)$ & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter [1].\\
$\ymfautoredf(\samplesymf)$ & DFT of the reduced autocorrelation function for the matched filter [1].\\
$\ypctargetf(\samplesymf)$ & DFT of the pulse compressed signal from a single target [V].\\
$\ypctargetnormf(\samplesymf)$ & DFT of the pulse compressed signal from a single target 
normalized by the DFT of the reduced autocorrelation function for the matched filter [Vm].\\
$\ypcvolumef(\samplesymf)$ & DFT of the pulse compressed signal from a volume. Compensated for spreading loss [Vm].\\
$\ypcvolumenormf(\samplesymf)$ & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.  Compensated for spreading loss [V].\\

$\ymf(\samplesymt)$ & Matched filter. Signal used for pulse compression [1].\\
$\ymfauto(\samplesymt)$ & Autocorrelation function for the matched filter [1].\\
$\ymfautored(\samplesymt)$ & Reduced autocorrelation function for the matched filter [1].\\

$\ypc(\samplesymt)$ & Pulse compressed signal averaged over all transducer sectors [V].\\
$\ypc(\samplesymt, \channelsym) $ & Pulse compressed signal from channel $\channelsym$ [V].\\
$y_{\textrm{pc,aft}}(\samplesymt)$ & Pulse compressed signal from the aft transducer half [V].\\
$y_{\textrm{pc,fore}}(\samplesymt)$ & Pulse compressed signal from the forward transducer half [V].\\
$y_{\textrm{pc,port}}(\samplesymt)$ & Pulse compressed signal from the port transducer half [V].\\
$\ypcspread(\samplesymt)$ & Pulse compressed signal compensated for spherical spreading [Vm].\\
$y_{\textrm{pc,star}}(\samplesymt)$ & Pulse compressed signal from the starboard transducer half [V].\\
$\ypctarget(\samplesymt)$ & Pulse compressed signal from a single target [V].\\

$\yrx(\samplesymt,\channelsym)$ & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$ [V].\\
$\yrx(\samplesymt,\channelsym,\stagesym)$ & Received digitised, bandpass filtered, decimated complex signal [V].\\
$\yrxa(\timesym)$ & Analogue acoustic signal received by the transducer [Pa].\\
$\yrxa(\timesym,\channelsym)$ & Analogue acoustic signal received by each transducer sector $\channelsym$ [Pa].\\
$\yrxe(\timesym,\channelsym)$ & Analogue electric signal received by each transceiver channel $\channelsym$ [V].\\

$\yrxorg(\samplesymt,\channelsym)$ & Received digitised signal before the bandpass filtering and decimation stages, $\yrxorg(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,0)$ [V].\\

$\ytx(\samplesymt)$ & Ideal transmitted signal generated from transmit signal properties [V].\\
$\ytxnorm(\samplesymt)$ & Ideal normalized transmitted signal generated from transmit signal properties [1].\\
$\ytxnorm(\samplesymt,\stagesym)$ & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$ [1].\\

$\ytxa(\timesym)$ & Analogue acoustic transmit signal [Pa].\\

$\ytxe(\timesym)$ & Analogue electric transmit signal [V].\\

$y_\along(\samplesymt)$ & Electrical angle along the minor axis of the transducer (alongship when ship-mounted) [$^{\circ}$].\\
$y_\athw(\samplesymt)$ & Electrical angle along the major axis of the transducer (athwartship when ship-mounted) [$^{\circ}$].\\

$\zrxe$ & Receiver electric impedance [$\Omega$].\\
$\ztde$ & Transducer sector electric impedance [$\Omega$].\\

$\absorp(\freqsym)$  & Absorption coefficient at frequency $f$ [dB $\textrm{m}^{-1}$].\\
$\absorp(\samplesymf)$ & Absorption coefficient at frequency index $m$ [dB $\textrm{m}^{-1}$]\\
$\anglefalong$ & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo [1].\\
$\anglefathw$ & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo [1].\\
$\along$ & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
$\along(\samplesymt)$ & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
$\wlen$ &  Acoustic wavelength [m].\\
$\wlen_\samplesymf$ & Acoustic wavelength at frequency index $m$ [m]\\
$\sigmabs$ & Backscattering cross-section [$\textrm{m}^2$].\\
$\tnom$ & Nominal transmit pulse duration [s].\\
$\teff$ & Effective transmit pulse duration [s].\\
$\athw$ & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
$\athw(\samplesymt)$ & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
$\eqang(\freqsym)$ & Two-way equivalent beam angle at frequency $\freqsym$ [sr].\\
\end{longtable}

\end{document}

### 2023-09-08 Paper/Supplementary.tex
Add column for Python variable name to list of symbols table

@@
-\documentclass[preprint,12pt,TurnOnLineNumbers]{JASAnew}
+\documentclass[12pt,a4paper]{article}
 \usepackage{longtable}
-
-\usepackage{lineno}
-\linenumbers
+\usepackage[landscape]{geometry}
 
@@
 
-\begin{longtable}{p{0.25\linewidth} p{0.75\linewidth}}
-Symbol & Description\\
+\begin{longtable}{p{0.15\linewidth} p{0.2\linewidth} p{0.65\linewidth} }
+Symbol & Variable name & Description \\
 
-$\atan(y,x)$ & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
-$\cw$ & Sound speed in water [$m s^{-1}$].\\
-$\decfac(\stagesym)$ & Integer filter decimation factor [1].\\
-$\dft_\ndft(x)$ & The discrete Fourier transform of length $\ndft$ of $x$.\\
-FT & The continuous Fourier transform. \\
-$\freqsym$ & Frequency [Hz].\\
-$\fc$ & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$ [Hz].\\
-$\fn$ & Nominal operating frequency of a transducer [Hz].\\
-$\fs$ & Analogue to digital sampling frequency [Hz].\\
-$\fsdec$ & Decimated sampling frequency [Hz].\\
-$\fstart$ & Start frequency of linear upsweep chirp pulse [Hz].\\
-$\fstop$ & Stop frequency of linear upsweep chirp pulse [Hz].\\
+$\atan(y,x)$ & & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
+$\cw$ & & Sound speed in water [$m s^{-1}$].\\
+$\decfac(\stagesym)$ & & Integer filter decimation factor [1].\\
+$\dft_\ndft(x)$ & & The discrete Fourier transform of length $\ndft$ of $x$.\\
+FT & n/a & The continuous Fourier transform. \\
+$\freqsym$ & & Frequency [Hz].\\
+$\fc$ & & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$ [Hz].\\
+$\fn$ & & Nominal operating frequency of a transducer [Hz].\\
+$\fs$ & & Analogue to digital sampling frequency [Hz].\\
+$\fsdec$ & & Decimated sampling frequency [Hz].\\
+$\fstart$ & & Start frequency of linear upsweep chirp pulse [Hz].\\
+$\fstop$ & & Stop frequency of linear upsweep chirp pulse [Hz].\\
 
-$\gain(\along,\athw,\freqsym)$ & Transducer gain as a function of echo arrival angle in the beam and frequency [1].\\
+$\gain(\along,\athw,\freqsym)$ & & Transducer gain as a function of echo arrival angle in the beam and frequency [1].\\
 
-$\gain(\along,\athw,\samplesymf)$ & Transducer gain as a function of echo arrival angle in the beam and frequency index [1].\\
+$\gain(\along,\athw,\samplesymf)$ & & Transducer gain as a function of echo arrival angle in the beam and frequency index [1].\\
 
-$\gainzero(\freqsym)$ & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$ [1].\\
+$\gainzero(\freqsym)$ & & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$ [1].\\
 
-$\hfl(\genidxsym,\stagesym)$ & Complex valued receiving filter coefficients [1].\\
+$\hfl(\genidxsym,\stagesym)$ & & Complex valued receiving filter coefficients [1].\\
 
-$\Im(x)$ & The imaginary part of $x$.\\
+$\Im(x)$ & & The imaginary part of $x$.\\
 
-$\samplesymf$ & Sample index in frequency domain [1].\\
+$\samplesymf$ & & Sample index in frequency domain [1].\\
 
-$\nchannels$ & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal [1].\\
-$\nstages$ & Total number of filter stages [1].\\
-$\nw$ & Number of samples used in the sliding Hanning window [1].\\
-$\samplesymt$ & Sample index in time domain [1]\\
+$\nchannels$ & & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal [1].\\
+$\nstages$ & & Total number of filter stages [1].\\
+$\nw$ & & Number of samples used in the sliding Hanning window [1].\\
+$\samplesymt$ & & Sample index in time domain [1]\\
 
-$\prxetf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$ [W].\\
-$\prxetf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a single target [W].\\
-$\prxevf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$ [W].\\
-$\prxevf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a volume [W$\textrm{m}^2$].\\
-$\prxe(\samplesymt)$  & Received electric power in a matched load [W].\\
+$\prxetf(\freqsym)$ & & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$ [W].\\
+$\prxetf(\samplesymf)$ & & DFT of the received electric power in a matched load for the signal from a single target [W].\\
+$\prxevf(\freqsym)$ & & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$ [W].\\
+$\prxevf(\samplesymf)$ & & DFT of the received electric power in a matched load for the signal from a volume [W$\textrm{m}^2$].\\
+$\prxe(\samplesymt)$  & & Received electric power in a matched load [W].\\
 
-$\ptxe$ & Transmitted electric power [W].\\
-$\ptxauto(\samplesymt)$ & Square of the absolute value of the matched filter autocorrelation function [1].\\
+$\ptxe$ & & Transmitted electric power [W].\\
+$\ptxauto(\samplesymt)$ & & Square of the absolute value of the matched filter autocorrelation function [1].\\
 
-$\Re(x)$ & The real part of $x$.\\
-$\range$ & Distance from transducer [m].\\
-$\rangeref$ & Reference distance [m]. \\
-$\range(\samplesymt)$ & Distance from transducer [m].\\
+$\Re(x)$ & & The real part of $x$.\\
+$\range$ & & Distance from transducer [m].\\
+$\rangeref$ & & Reference distance [m]. \\
+$\range(\samplesymt)$ & & Distance from transducer [m].\\
 
-$\mysp(\samplesymt)$  & Point scattering strength [dB re 1 $\textrm{m}^2$].\\
-$\sv$  & Volume backscattering strength [dB re 1 $\textrm{m}^{-1}$].\\
-$\sv(\freqsym)$ & Volume backscattering strength at frequency $\freqsym$ [dB re 1 $\textrm{m}^{-1}$].\\
-$\sv(\samplesymt)$ & Volume backscattering strength at sample index $\samplesymt$ [dB re $\textrm{m}^{-1}$]\\
+$\mysp(\samplesymt)$  & & Point scattering strength [dB re 1 $\textrm{m}^2$].\\
+$\sv$  & & Volume backscattering strength [dB re 1 $\textrm{m}^{-1}$].\\
+$\sv(\freqsym)$ & & Volume backscattering strength at frequency $\freqsym$ [dB re 1 $\textrm{m}^{-1}$].\\
+$\sv(\samplesymt)$ & & Volume backscattering strength at sample index $\samplesymt$ [dB re $\textrm{m}^{-1}$]\\
 
-$\ts$ & Target strength [dB re 1 $\textrm{m}^2$].\\
-$\ts(\freqsym)$ & Target strength at frequency $f$ [dB re 1 $\textrm{m}^2$].\\
-$\ts(\samplesymf)$ & Target strength at frequency index $\samplesymf$ [dB re 1 $\textrm{m}^{2}$]\\
+$\ts$ & & Target strength [dB re 1 $\textrm{m}^2$].\\
+$\ts(\freqsym)$ & & Target strength at frequency $f$ [dB re 1 $\textrm{m}^2$].\\
+$\ts(\samplesymf)$ & & Target strength at frequency index $\samplesymf$ [dB re 1 $\textrm{m}^{2}$]\\
 
-$\timesym$ &  Time [s].\\
-$\tslide$ & Duration of sliding window for calculating volume backscattering strength [s].\\
+$\timesym$ & &  Time [s].\\
+$\tslide$ & & Duration of sliding window for calculating volume backscattering strength [s].\\
 
-$\channelsym$ & Channel number/transducer sector [1].\\
-$u(\genidxsym)$ & The Heaviside step function [1]. \\
-$V$ & Volume occupied by scattering targets [$\textrm{m}^3$].\\
-$\stagesym$ & Filter stage [1].\\
+$\channelsym$ & & Channel number/transducer sector [1].\\
+$u(\genidxsym)$ & & The Heaviside step function [1]. \\
+$V$ & & Volume occupied by scattering targets [$\textrm{m}^3$].\\
+$\stagesym$ & & Filter stage [1].\\
 
-$\hannw$ & The Hanning window function [1].\\
-$\hannw(\genidxsym)$ & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$ [1].\\
-$\hannwnorm(\genidxsym)$ & Normalised Hanning window [1].\\
+$\hannw$ & & The Hanning window function [1].\\
+$\hannw(\genidxsym)$ & & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$ [1].\\
+$\hannwnorm(\genidxsym)$ & & Normalised Hanning window [1].\\
 
-$x^*$ & The complex conjugate of $x$.\\
-$||x||$ & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
+$x^*$ & & The complex conjugate of $x$.\\
+$||x||$ & & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
 
-$\ymfautof(\samplesymf)$ & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter [1].\\
-$\ymfautoredf(\samplesymf)$ & DFT of the reduced autocorrelation function for the matched filter [1].\\
-$\ypctargetf(\samplesymf)$ & DFT of the pulse compressed signal from a single target [V].\\
-$\ypctargetnormf(\samplesymf)$ & DFT of the pulse compressed signal from a single target 
+$\ymfautof(\samplesymf)$ & & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter [1].\\
+$\ymfautoredf(\samplesymf)$ & & DFT of the reduced autocorrelation function for the matched filter [1].\\
+$\ypctargetf(\samplesymf)$ & & DFT of the pulse compressed signal from a single target [V].\\
+$\ypctargetnormf(\samplesymf)$ & & DFT of the pulse compressed signal from a single target 
 normalized by the DFT of the reduced autocorrelation function for the matched filter [Vm].\\
-$\ypcvolumef(\samplesymf)$ & DFT of the pulse compressed signal from a volume. Compensated for spreading loss [Vm].\\
-$\ypcvolumenormf(\samplesymf)$ & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
+$\ypcvolumef(\samplesymf)$ & & DFT of the pulse compressed signal from a volume. Compensated for spreading loss [Vm].\\
+$\ypcvolumenormf(\samplesymf)$ & & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
 Compensated for spreading loss [V].\\
 
-$\ymf(\samplesymt)$ & Matched filter. Signal used for pulse compression [1].\\
-$\ymfauto(\samplesymt)$ & Autocorrelation function for the matched filter [1].\\
-$\ymfautored(\samplesymt)$ & Reduced autocorrelation function for the matched filter [1].\\
+$\ymf(\samplesymt)$ & & Matched filter. Signal used for pulse compression [1].\\
+$\ymfauto(\samplesymt)$ & & Autocorrelation function for the matched filter [1].\\
+$\ymfautored(\samplesymt)$ & & Reduced autocorrelation function for the matched filter [1].\\
 
-$\ypc(\samplesymt)$ & Pulse compressed signal averaged over all transducer sectors [V].\\
-$\ypc(\samplesymt, \channelsym) $ & Pulse compressed signal from channel $\channelsym$ [V].\\
-$y_{\textrm{pc,aft}}(\samplesymt)$ & Pulse compressed signal from the aft transducer half [V].\\
-$y_{\textrm{pc,fore}}(\samplesymt)$ & Pulse compressed signal from the forward transducer half [V].\\
-$y_{\textrm{pc,port}}(\samplesymt)$ & Pulse compressed signal from the port transducer half [V].\\
-$\ypcspread(\samplesymt)$ & Pulse compressed signal compensated for spherical spreading [Vm].\\
-$y_{\textrm{pc,star}}(\samplesymt)$ & Pulse compressed signal from the starboard transducer half [V].\\
-$\ypctarget(\samplesymt)$ & Pulse compressed signal from a single target [V].\\
+$\ypc(\samplesymt)$ & & Pulse compressed signal averaged over all transducer sectors [V].\\
+$\ypc(\samplesymt, \channelsym) $ & & Pulse compressed signal from channel $\channelsym$ [V].\\
+$y_{\textrm{pc,aft}}(\samplesymt)$ & & Pulse compressed signal from the aft transducer half [V].\\
+$y_{\textrm{pc,fore}}(\samplesymt)$ & & Pulse compressed signal from the forward transducer half [V].\\
+$y_{\textrm{pc,port}}(\samplesymt)$ & & Pulse compressed signal from the port transducer half [V].\\
+$\ypcspread(\samplesymt)$ & & Pulse compressed signal compensated for spherical spreading [Vm].\\
+$y_{\textrm{pc,star}}(\samplesymt)$ & & Pulse compressed signal from the starboard transducer half [V].\\
+$\ypctarget(\samplesymt)$ & & Pulse compressed signal from a single target [V].\\
 
-$\yrx(\samplesymt,\channelsym)$ & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$ [V].\\
-$\yrx(\samplesymt,\channelsym,\stagesym)$ & Received digitised, bandpass filtered, decimated complex signal [V].\\
-$\yrxa(\timesym)$ & Analogue acoustic signal received by the transducer [Pa].\\
-$\yrxa(\timesym,\channelsym)$ & Analogue acoustic signal received by each transducer sector $\channelsym$ [Pa].\\
-$\yrxe(\timesym,\channelsym)$ & Analogue electric signal received by each transceiver channel $\channelsym$ [V].\\
+$\yrx(\samplesymt,\channelsym)$ & & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$ [V].\\
+$\yrx(\samplesymt,\channelsym,\stagesym)$ & & Received digitised, bandpass filtered, decimated complex signal [V].\\
+$\yrxa(\timesym)$ & & Analogue acoustic signal received by the transducer [Pa].\\
+$\yrxa(\timesym,\channelsym)$ & & Analogue acoustic signal received by each transducer sector $\channelsym$ [Pa].\\
+$\yrxe(\timesym,\channelsym)$ & & Analogue electric signal received by each transceiver channel $\channelsym$ [V].\\
 
-$\yrxorg(\samplesymt,\channelsym)$ & Received digitised signal before the bandpass filtering and decimation stages, $\yrxorg(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,0)$ [V].\\
+$\yrxorg(\samplesymt,\channelsym)$ & & Received digitised signal before the bandpass filtering and decimation stages, $\yrxorg(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,0)$ [V].\\
 
-$\ytx(\samplesymt)$ & Ideal transmitted signal generated from transmit signal properties [V].\\
-$\ytxnorm(\samplesymt)$ & Ideal normalized transmitted signal generated from transmit signal properties [1].\\
-$\ytxnorm(\samplesymt,\stagesym)$ & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$ [1].\\
+$\ytx(\samplesymt)$ & & Ideal transmitted signal generated from transmit signal properties [V].\\
+$\ytxnorm(\samplesymt)$ & & Ideal normalized transmitted signal generated from transmit signal properties [1].\\
+$\ytxnorm(\samplesymt,\stagesym)$ & & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$ [1].\\
 
-$\ytxa(\timesym)$ & Analogue acoustic transmit signal [Pa].\\
+$\ytxa(\timesym)$ & & Analogue acoustic transmit signal [Pa].\\
 
-$\ytxe(\timesym)$ & Analogue electric transmit signal [V].\\
+$\ytxe(\timesym)$ & & Analogue electric transmit signal [V].\\
 
-$y_\along(\samplesymt)$ & Electrical angle along the minor axis of the transducer (alongship when ship-mounted) [$^{\circ}$].\\
-$y_\athw(\samplesymt)$ & Electrical angle along the major axis of the transducer (athwartship when ship-mounted) [$^{\circ}$].\\
+$y_\along(\samplesymt)$ & & Electrical angle along the minor axis of the transducer (alongship when ship-mounted) [$^{\circ}$].\\
+$y_\athw(\samplesymt)$ & & Electrical angle along the major axis of the transducer (athwartship when ship-mounted) [$^{\circ}$].\\
 
-$\zrxe$ & Receiver electric impedance [$\Omega$].\\
-$\ztde$ & Transducer sector electric impedance [$\Omega$].\\
+$\zrxe$ & & Receiver electric impedance [$\Omega$].\\
+$\ztde$ & & Transducer sector electric impedance [$\Omega$].\\
 
-$\absorp(\freqsym)$  & Absorption coefficient at frequency $f$ [dB $\textrm{m}^{-1}$].\\
-$\absorp(\samplesymf)$ & Absorption coefficient at frequency index $m$ [dB $\textrm{m}^{-1}$]\\
-$\anglefalong$ & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo [1].\\
-$\anglefathw$ & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo [1].\\
-$\along$ & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
-$\along(\samplesymt)$ & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
-$\wlen$ &  Acoustic wavelength [m].\\
-$\wlen_\samplesymf$ & Acoustic wavelength at frequency index $m$ [m]\\
-$\sigmabs$ & Backscattering cross-section [$\textrm{m}^2$].\\
-$\tnom$ & Nominal transmit pulse duration [s].\\
-$\teff$ & Effective transmit pulse duration [s].\\
-$\athw$ & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
-$\athw(\samplesymt)$ & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
-$\eqang(\freqsym)$ & Two-way equivalent beam angle at frequency $\freqsym$ [sr].\\
+$\absorp(\freqsym)$  & & Absorption coefficient at frequency $f$ [dB $\textrm{m}^{-1}$].\\
+$\absorp(\samplesymf)$ & & Absorption coefficient at frequency index $m$ [dB $\textrm{m}^{-1}$]\\
+$\anglefalong$ & & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo [1].\\
+$\anglefathw$ & & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo [1].\\
+$\along$ & & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
+$\along(\samplesymt)$ & & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
+$\wlen$ & &  Acoustic wavelength [m].\\
+$\wlen_\samplesymf$ & & Acoustic wavelength at frequency index $m$ [m]\\
+$\sigmabs$ & & Backscattering cross-section [$\textrm{m}^2$].\\
+$\tnom$ & & Nominal transmit pulse duration [s].\\
+$\teff$ & & Effective transmit pulse duration [s].\\
+$\athw$ & & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
+$\athw(\samplesymt)$ & & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
+$\eqang(\freqsym)$ & & Two-way equivalent beam angle at frequency $\freqsym$ [sr].\\
 \end{longtable}

### 2023-09-08 Paper/Supplementary.tex
move units into own column in table of symbols

--- 2023-09-08 Paper/Supplementary.tex
@@
 
-\begin{longtable}{p{0.25\linewidth} p{0.75\linewidth}}
-Symbol & Description\\
+\begin{longtable}{p{0.15\linewidth} p{0.13\linewidth} p{0.12\linewidth} p{0.6\linewidth} }
+Symbol & Variable name & Units & Description \\
 
-$\atan(y,x)$ & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
-$\cw$ & Sound speed in water [$m s^{-1}$].\\
-$\decfac(\stagesym)$ & Integer filter decimation factor [1].\\
-$\dft_\ndft(x)$ & The discrete Fourier transform of length $\ndft$ of $x$.\\
-FT & The continuous Fourier transform. \\
-$\freqsym$ & Frequency [Hz].\\
-$\fc$ & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$ [Hz].\\
-$\fn$ & Nominal operating frequency of a transducer [Hz].\\
-$\fs$ & Analogue to digital sampling frequency [Hz].\\
-$\fsdec$ & Decimated sampling frequency [Hz].\\
-$\fstart$ & Start frequency of linear upsweep chirp pulse [Hz].\\
-$\fstop$ & Stop frequency of linear upsweep chirp pulse [Hz].\\
+$\atan(y,x)$ & -- & -- & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
+$\cw$ & & $m s^{-1}$ & Sound speed in water.\\
+$\decfac(\stagesym)$ & & 1 & Integer filter decimation factor.\\
+$\dft_\ndft(x)$ & & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
+FT & -- & -- & The continuous Fourier transform. \\
+$\freqsym$ & & Hz & Frequency.\\
+$\fc$ & & Hz & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$.\\
+$\fn$ & & Hz & Nominal operating frequency of a transducer.\\
+$\fs$ & & Hz & Analogue to digital sampling frequency.\\
+$\fsdec$ & & Hz & Decimated sampling frequency.\\
+$\fstart$ & & Hz & Start frequency of linear upsweep chirp pulse.\\
+$\fstop$ & & Hz & Stop frequency of linear upsweep chirp pulse.\\
 
-$\gain(\along,\athw,\freqsym)$ & Transducer gain as a function of echo arrival angle in the beam and frequency [1].\\
+$\gain(\along,\athw,\freqsym)$ & & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency.\\
 
-$\gain(\along,\athw,\samplesymf)$ & Transducer gain as a function of echo arrival angle in the beam and frequency index [1].\\
+$\gain(\along,\athw,\samplesymf)$ & & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency index.\\
 
-$\gainzero(\freqsym)$ & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$ [1].\\
+$\gainzero(\freqsym)$ & & 1 & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$.\\
 
-$\hfl(\genidxsym,\stagesym)$ & Complex valued receiving filter coefficients [1].\\
+$\hfl(\genidxsym,\stagesym)$ & & 1 & Complex valued receiving filter coefficients.\\
 
-$\Im(x)$ & The imaginary part of $x$.\\
+$\Im(x)$ & & -- & The imaginary part of $x$.\\
 
-$\samplesymf$ & Sample index in frequency domain [1].\\
+$\samplesymf$ & & 1 & Sample index in frequency domain.\\
 
-$\nchannels$ & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal [1].\\
-$\nstages$ & Total number of filter stages [1].\\
-$\nw$ & Number of samples used in the sliding Hanning window [1].\\
-$\samplesymt$ & Sample index in time domain [1]\\
+$\nchannels$ & & 1 & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal.\\
+$\nstages$ & & 1 & Total number of filter stages.\\
+$\nw$ & & 1 & Number of samples used in the sliding Hanning window.\\
+$\samplesymt$ & & 1 & Sample index in time domain.\\
 
-$\prxetf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$ [W].\\
-$\prxetf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a single target [W].\\
-$\prxevf(\freqsym)$ & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$ [W].\\
-$\prxevf(\samplesymf)$ & DFT of the received electric power in a matched load for the signal from a volume [W$\textrm{m}^2$].\\
-$\prxe(\samplesymt)$  & Received electric power in a matched load [W].\\
+$\prxetf(\freqsym)$ & & W & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$.\\
+$\prxetf(\samplesymf)$ & & W & DFT of the received electric power in a matched load for the signal from a single target.\\
+$\prxevf(\freqsym)$ & & W & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$.\\
+$\prxevf(\samplesymf)$ & & W$\textrm{m}^2$ & DFT of the received electric power in a matched load for the signal from a volume.\\
+$\prxe(\samplesymt)$  & & W & Received electric power in a matched load.\\
 
-$\ptxe$ & Transmitted electric power [W].\\
-$\ptxauto(\samplesymt)$ & Square of the absolute value of the matched filter autocorrelation function [1].\\
+$\ptxe$ & & W & Transmitted electric power.\\
+$\ptxauto(\samplesymt)$ & & 1 & Square of the absolute value of the matched filter autocorrelation function.\\
 
-$\Re(x)$ & The real part of $x$.\\
-$\range$ & Distance from transducer [m].\\
-$\rangeref$ & Reference distance [m]. \\
-$\range(\samplesymt)$ & Distance from transducer [m].\\
+$\Re(x)$ & & -- & The real part of $x$.\\
+$\range$ & & m & Distance from transducer.\\
+$\rangeref$ & & m & Reference distance. \\
+$\range(\samplesymt)$ & & m & Distance from transducer.\\
 
-$\mysp(\samplesymt)$  & Point scattering strength [dB re 1 $\textrm{m}^2$].\\
-$\sv$  & Volume backscattering strength [dB re 1 $\textrm{m}^{-1}$].\\
-$\sv(\freqsym)$ & Volume backscattering strength at frequency $\freqsym$ [dB re 1 $\textrm{m}^{-1}$].\\
-$\sv(\samplesymt)$ & Volume backscattering strength at sample index $\samplesymt$ [dB re $\textrm{m}^{-1}$]\\
+$\mysp(\samplesymt)$  & & dB re 1 $\textrm{m}^2$ & Point scattering strength.\\
+$\sv$  & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength.\\
+$\sv(\freqsym)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at frequency $\freqsym$.\\
+$\sv(\samplesymt)$ & & dB re $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
 
-$\ts$ & Target strength [dB re 1 $\textrm{m}^2$].\\
-$\ts(\freqsym)$ & Target strength at frequency $f$ [dB re 1 $\textrm{m}^2$].\\
-$\ts(\samplesymf)$ & Target strength at frequency index $\samplesymf$ [dB re 1 $\textrm{m}^{2}$]\\
+$\ts$ & & dB re 1 $\textrm{m}^2$ & Target strength.\\
+$\ts(\freqsym)$ & & dB re 1 $\textrm{m}^2$ & Target strength at frequency $f$.\\
+$\ts(\samplesymf)$ & & dB re 1 $\textrm{m}^{2}$ & Target strength at frequency index $\samplesymf$.\\
 
-$\timesym$ &  Time [s].\\
-$\tslide$ & Duration of sliding window for calculating volume backscattering strength [s].\\
+$\timesym$ & & s &  Time.\\
+$\tslide$ & & s & Duration of sliding window for calculating volume backscattering strength.\\
 
-$\channelsym$ & Channel number/transducer sector [1].\\
-$u(\genidxsym)$ & The Heaviside step function [1]. \\
-$V$ & Volume occupied by scattering targets [$\textrm{m}^3$].\\
-$\stagesym$ & Filter stage [1].\\
+$\channelsym$ & & 1 & Channel number/transducer sector.\\
+$u(\genidxsym)$ & & 1 & The Heaviside step function. \\
+$V$ & & $\textrm{m}^3$ & Volume occupied by scattering targets.\\
+$\stagesym$ & & 1 & Filter stage.\\
 
-$\hannw$ & The Hanning window function [1].\\
-$\hannw(\genidxsym)$ & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$ [1].\\
-$\hannwnorm(\genidxsym)$ & Normalised Hanning window [1].\\
+$\hannw$ & & 1 & The Hanning window function.\\
+$\hannw(\genidxsym)$ & & 1 & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
+$\hannwnorm(\genidxsym)$ & & 1 & Normalised Hanning window.\\
 
-$x^*$ & The complex conjugate of $x$.\\
-$||x||$ & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
+$x^*$ & & -- & The complex conjugate of $x$.\\
+$||x||$ & & -- & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
 
-$\ymfautof(\samplesymf)$ & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter [1].\\
-$\ymfautoredf(\samplesymf)$ & DFT of the reduced autocorrelation function for the matched filter [1].\\
-$\ypctargetf(\samplesymf)$ & DFT of the pulse compressed signal from a single target [V].\\
-$\ypctargetnormf(\samplesymf)$ & DFT of the pulse compressed signal from a single target 
-normalized by the DFT of the reduced autocorrelation function for the matched filter [Vm].\\
-$\ypcvolumef(\samplesymf)$ & DFT of the pulse compressed signal from a volume. Compensated for spreading loss [Vm].\\
-$\ypcvolumenormf(\samplesymf)$ & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
-Compensated for spreading loss [V].\\
+$\ymfautof(\samplesymf)$ & & 1 & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter.\\
+$\ymfautoredf(\samplesymf)$ & & 1 & DFT of the reduced autocorrelation function for the matched filter.\\
+$\ypctargetf(\samplesymf)$ & & V & DFT of the pulse compressed signal from a single target.\\
+$\ypctargetnormf(\samplesymf)$ & & Vm & DFT of the pulse compressed signal from a single target 
+normalized by the DFT of the reduced autocorrelation function for the matched filter.\\
+$\ypcvolumef(\samplesymf)$ & & Vm & DFT of the pulse compressed signal from a volume. Compensated for spreading loss.\\
+$\ypcvolumenormf(\samplesymf)$ & & V & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
+Compensated for spreading loss.\\
 
-$\ymf(\samplesymt)$ & Matched filter. Signal used for pulse compression [1].\\
-$\ymfauto(\samplesymt)$ & Autocorrelation function for the matched filter [1].\\
-$\ymfautored(\samplesymt)$ & Reduced autocorrelation function for the matched filter [1].\\
+$\ymf(\samplesymt)$ & & 1 & Matched filter. Signal used for pulse compression.\\
+$\ymfauto(\samplesymt)$ & & 1 & Autocorrelation function for the matched filter.\\
+$\ymfautored(\samplesymt)$ & & 1 & Reduced autocorrelation function for the matched filter.\\
 
-$\ypc(\samplesymt)$ & Pulse compressed signal averaged over all transducer sectors [V].\\
-$\ypc(\samplesymt, \channelsym) $ & Pulse compressed signal from channel $\channelsym$ [V].\\
-$y_{\textrm{pc,aft}}(\samplesymt)$ & Pulse compressed signal from the aft transducer half [V].\\
-$y_{\textrm{pc,fore}}(\samplesymt)$ & Pulse compressed signal from the forward transducer half [V].\\
-$y_{\textrm{pc,port}}(\samplesymt)$ & Pulse compressed signal from the port transducer half [V].\\
-$\ypcspread(\samplesymt)$ & Pulse compressed signal compensated for spherical spreading [Vm].\\
-$y_{\textrm{pc,star}}(\samplesymt)$ & Pulse compressed signal from the starboard transducer half [V].\\
-$\ypctarget(\samplesymt)$ & Pulse compressed signal from a single target [V].\\
+$\ypc(\samplesymt)$ & & V & Pulse compressed signal averaged over all transducer sectors.\\
+$\ypc(\samplesymt, \channelsym) $ & & V & Pulse compressed signal from channel $\channelsym$.\\
+$y_{\textrm{pc,aft}}(\samplesymt)$ & & V & Pulse compressed signal from the aft transducer half.\\
+$y_{\textrm{pc,fore}}(\samplesymt)$ & & V & Pulse compressed signal from the forward transducer half.\\
+$y_{\textrm{pc,port}}(\samplesymt)$ & & V & Pulse compressed signal from the port transducer half.\\
+$\ypcspread(\samplesymt)$ & & Vm & Pulse compressed signal compensated for spherical spreading.\\
+$y_{\textrm{pc,star}}(\samplesymt)$ & & V & Pulse compressed signal from the starboard transducer half.\\
+$\ypctarget(\samplesymt)$ & & V & Pulse compressed signal from a single target.\\
 
-$\yrx(\samplesymt,\channelsym)$ & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$ [V].\\
-$\yrx(\samplesymt,\channelsym,\stagesym)$ & Received digitised, bandpass filtered, decimated complex signal [V].\\
-$\yrxa(\timesym)$ & Analogue acoustic signal received by the transducer [Pa].\\
-$\yrxa(\timesym,\channelsym)$ & Analogue acoustic signal received by each transducer sector $\channelsym$ [Pa].\\
-$\yrxe(\timesym,\channelsym)$ & Analogue electric signal received by each transceiver channel $\channelsym$ [V].\\
+$\yrx(\samplesymt,\channelsym)$ & & V & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$.\\
+$\yrx(\samplesymt,\channelsym,\stagesym)$ & & V & Received digitised, bandpass filtered, decimated complex signal.\\
+$\yrxa(\timesym)$ & & Pa & Analogue acoustic signal received by the transducer.\\
+$\yrxa(\timesym,\channelsym)$ & & Pa & Analogue acoustic signal received by each transducer sector $\channelsym$.\\
+$\yrxe(\timesym,\channelsym)$ & & V & Analogue electric signal received by each transceiver channel $\channelsym$.\\
 
-$\yrxorg(\samplesymt,\channelsym)$ & Received digitised signal before the bandpass filtering and decimation stages, $\yrxorg(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,0)$ [V].\\
+$\yrxorg(\samplesymt,\channelsym)$ & & V & Received digitised signal before the bandpass filtering and decimation stages, $\yrxorg(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,0)$.\\
 
-$\ytx(\samplesymt)$ & Ideal transmitted signal generated from transmit signal properties [V].\\
-$\ytxnorm(\samplesymt)$ & Ideal normalized transmitted signal generated from transmit signal properties [1].\\
-$\ytxnorm(\samplesymt,\stagesym)$ & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$ [1].\\
+$\ytx(\samplesymt)$ & & V & Ideal transmitted signal generated from transmit signal properties.\\
+$\ytxnorm(\samplesymt)$ & & 1 & Ideal normalized transmitted signal generated from transmit signal properties.\\
+$\ytxnorm(\samplesymt,\stagesym)$ & & 1 & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$.\\
 
-$\ytxa(\timesym)$ & Analogue acoustic transmit signal [Pa].\\
+$\ytxa(\timesym)$ & & Pa & Analogue acoustic transmit signal.\\
 
-$\ytxe(\timesym)$ & Analogue electric transmit signal [V].\\
+$\ytxe(\timesym)$ & & V & Analogue electric transmit signal.\\
 
-$y_\along(\samplesymt)$ & Electrical angle along the minor axis of the transducer (alongship when ship-mounted) [$^{\circ}$].\\
-$y_\athw(\samplesymt)$ & Electrical angle along the major axis of the transducer (athwartship when ship-mounted) [$^{\circ}$].\\
+$y_\along(\samplesymt)$ & & $^{\circ}$ & Electrical angle along the minor axis of the transducer (alongship when ship-mounted).\\
+$y_\athw(\samplesymt)$ & & $^{\circ}$ & Electrical angle along the major axis of the transducer (athwartship when ship-mounted).\\
 
-$\zrxe$ & Receiver electric impedance [$\Omega$].\\
-$\ztde$ & Transducer sector electric impedance [$\Omega$].\\
+$\zrxe$ & & $\Omega$ & Receiver electric impedance.\\
+$\ztde$ & & $\Omega$ & Transducer sector electric impedance.\\
 
-$\absorp(\freqsym)$  & Absorption coefficient at frequency $f$ [dB $\textrm{m}^{-1}$].\\
-$\absorp(\samplesymf)$ & Absorption coefficient at frequency index $m$ [dB $\textrm{m}^{-1}$]\\
-$\anglefalong$ & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo [1].\\
-$\anglefathw$ & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo [1].\\
-$\along$ & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
-$\along(\samplesymt)$ & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted) [$^\circ$].\\
-$\wlen$ &  Acoustic wavelength [m].\\
-$\wlen_\samplesymf$ & Acoustic wavelength at frequency index $m$ [m]\\
-$\sigmabs$ & Backscattering cross-section [$\textrm{m}^2$].\\
-$\tnom$ & Nominal transmit pulse duration [s].\\
-$\teff$ & Effective transmit pulse duration [s].\\
-$\athw$ & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
-$\athw(\samplesymt)$ & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted) [$^\circ$].\\
-$\eqang(\freqsym)$ & Two-way equivalent beam angle at frequency $\freqsym$ [sr].\\
+$\absorp(\freqsym)$  & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency $f$.\\
+$\absorp(\samplesymf)$ & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency index $m$\\
+$\anglefalong$ & & 1 & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo.\\
+$\anglefathw$ & & 1 & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo.\\
+$\along$ & & $^\circ$ & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
+$\along(\samplesymt)$ & & $^\circ$ & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
+$\wlen$ & & m &  Acoustic wavelength.\\
+$\wlen_\samplesymf$ & & m & Acoustic wavelength at frequency index $m$\\
+$\sigmabs$ & & $\textrm{m}^2$ & Backscattering cross-section.\\
+$\tnom$ & & s & Nominal transmit pulse duration.\\
+$\teff$ & & s & Effective transmit pulse duration.\\
+$\athw$ & & $^\circ$ & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
+$\athw(\samplesymt)$ & & $^\circ$ & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
+$\eqang(\freqsym)$ & & sr & Two-way equivalent beam angle at frequency $\freqsym$.\\
 \end{longtable}

### 2023-09-11 Paper/Supplementary.tex
Check list of symbols against manuscript
Update list of symbols to match manuscript.
Correct some incorrect symbols in manuscript.

@@
 
+\renewcommand{\code}[1]{\texttt{\detokenize{#1}}}
+
 \begin{document}
@@
 
-Units are given for relevant symbols ([1] indicates a dimensionless quantity).
+(1 indicates a dimensionless quantity and -- indicates not applicable).
 
-\begin{longtable}{p{0.15\linewidth} p{0.13\linewidth} p{0.12\linewidth} p{0.6\linewidth} }
+\begin{longtable}{p{0.15\linewidth} p{0.20\linewidth} p{0.12\linewidth} p{0.5\linewidth} }
 Symbol & Variable name & Units & Description \\
@@
 $\dft_\ndft(x)$ & & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
-FT & -- & -- & The continuous Fourier transform. \\
 $\freqsym$ & & Hz & Frequency.\\
@@
 $\hfl(\genidxsym,\stagesym)$ & & 1 & Complex valued receiving filter coefficients.\\
-
+$\genidxsym$ & & 1 & Generic integer index.\\
 $\Im(x)$ & & -- & The imaginary part of $x$.\\
-
+$j$ & & & The square root of $-1$.\\
 $\samplesymf$ & & 1 & Sample index in frequency domain.\\
@@
 $\rangeref$ & & m & Reference distance. \\
+$\range_c$ & & m & Distance from the transducer to the centre of the range volume covered by $\tslide$.\\
+$\range_c(\samplesymt)$ & & m & $range_c$ at sample number $\samplesymt$.\\
 $\range(\samplesymt)$ & & m & Distance from transducer.\\
@@
 $\ts(\freqsym)$ & & dB re 1 $\textrm{m}^2$ & Target strength at frequency $f$.\\
-$\ts(\samplesymf)$ & & dB re 1 $\textrm{m}^{2}$ & Target strength at frequency index $\samplesymf$.\\
+$\ts(\samplesymf)$ & \code{TS_m} & dB re 1 $\textrm{m}^{2}$ & Target strength at frequency index $\samplesymf$.\\
 
@@
 $\ymfauto(\samplesymt)$ & & 1 & Autocorrelation function for the matched filter.\\
-$\ymfautored(\samplesymt)$ & & 1 & Reduced autocorrelation function for the matched filter.\\
+$\ymfautored(\samplesymt)$ & \code{y_mf_auto_red_n} & 1 & Reduced autocorrelation function for the matched filter.\\
 
@@
 
-$y_\along(\samplesymt)$ & & $^{\circ}$ & Electrical angle along the minor axis of the transducer (alongship when ship-mounted).\\
-$y_\athw(\samplesymt)$ & & $^{\circ}$ & Electrical angle along the major axis of the transducer (athwartship when ship-mounted).\\
+$y_\along(\samplesymt)$ & & rad & Electrical angle along the minor axis of the transducer (alongship when ship-mounted).\\
+$y_\athw(\samplesymt)$ & & rad & Electrical angle along the major axis of the transducer (athwartship when ship-mounted).\\
 
@@
 $\anglefathw$ & & 1 & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo.\\
-$\along$ & & $^\circ$ & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
-$\along(\samplesymt)$ & & $^\circ$ & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
+$\along$ & & rad & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
+$\along(\samplesymt)$ & & rad & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
 $\wlen$ & & m &  Acoustic wavelength.\\
@@
 $\teff$ & & s & Effective transmit pulse duration.\\
-$\athw$ & & $^\circ$ & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
-$\athw(\samplesymt)$ & & $^\circ$ & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
+$\athw$ & & rad & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
+$\athw(\samplesymt)$ & & rad & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
 $\eqang(\freqsym)$ & & sr & Two-way equivalent beam angle at frequency $\freqsym$.\\

### 2023-09-12 Paper/Manuscript.tex
Add list of symbols in supplementary document
Move the symbol aliases to a separate file so that it can be included into multiple .tex files.

Add supplemenary.tex document that contains the list of symbols from a much earlier version of the manuscript so will have some differences to update (to come).

@@
 
-\newcommand{\ek}{Simrad EK80}
-
-\newcommand{\timesym}{t}
-\newcommand{\freqsym}{f}
-\newcommand{\samplesymt}{n}
-\newcommand{\samplesymf}{m}
-\newcommand{\genidxsym}{i}
-
-\newcommand{\channelsym}{u}
-\newcommand{\nchannels}{N_{\textrm{u}}}
-
-\newcommand{\stagesym}{v}
-\newcommand{\nstages}{N_{\textrm{v}}}
-
-\newcommand{\fs}{f_{\textrm{s}}}
-\newcommand{\fsdec}{f_{\textrm{s,dec}}}
-\newcommand{\fstart}{f_{\textrm{start}}}
-\newcommand{\fstop}{f_{\textrm{stop}}}
-\newcommand{\fc}{f_{\textrm{c}}}
-\newcommand{\fn}{f_{\textrm{n}}}
-
-\newcommand{\zrxe}{z_{\textrm{rx,e}}}
-\newcommand{\ztde}{z_{\textrm{td,e}}}
-
-\newcommand{\ptxe}{p_{\textrm{tx,e}}}
-\newcommand{\prxe}{p_{\textrm{rx,e}}}
-
-\newcommand{\ntd}{n_{\textrm{td}}}
-\newcommand{\tnom}{\tau}
-\newcommand{\teff}{\tau_{\textrm{eff}}}
-
-\newcommand{\ytxe}{y_{\textrm{tx,e}}}
-\newcommand{\ytxa}{y_{\textrm{tx,a}}}
-\newcommand{\yrxa}{y_{\textrm{rx,a}}}
-\newcommand{\yrxe}{y_{\textrm{rx,e}}}
-
-\newcommand{\ytx}{y_{\textrm{tx}}}
-\newcommand{\ytxnorm}{\tilde{y}_{\textrm{tx}}}
-\newcommand{\ytxfd}{y_{\textrm{tx,f,d}}}
-\newcommand{\yrx}{y_{\textrm{rx}}}
-\newcommand{\yrxorg}{y_{\textrm{rx,org}}}
-\newcommand{\ymf}{y_{\textrm{mf}}}
-
-\newcommand{\ytd}{y_{\textrm{td}}}
-
-\newcommand{\ypc}{y_{\textrm{pc}}}
-\newcommand{\ypctarget}{y_{\textrm{pc,t}}}
-\newcommand{\ypcspread}{y_{\textrm{pc,s}}}
-\newcommand{\ymfauto}{y_{\textrm{mf,auto}}}
-\newcommand{\ymfautored}{y_{\textrm{mf,auto,red}}}
-
-\newcommand{\ptxauto}{p_{\textrm{tx,auto}}}
-
-\newcommand{\ypctargetf}{Y_{\textrm{pc,t}}}
-\newcommand{\ypctargetnormf}{\tilde{Y}_{\textrm{pc,t}}}
-\newcommand{\ypcvolumef}{Y_{\textrm{pc,v}}}
-\newcommand{\ypcvolumenormf}{\tilde{Y}_{\textrm{pc,v}}}
-
-\newcommand{\ymfautof}{Y_{\textrm{mf,auto}}}
-\newcommand{\ymfautoredf}{Y_{\textrm{mf,auto,red}}}
-
-\newcommand{\prxetf}{P_{\textrm{rx,e,t}}}
-\newcommand{\prxevf}{P_{\textrm{rx,e,v}}}
-
-\newcommand{\fstage}{k}
-\newcommand{\decfac}{D}
-\newcommand{\lfl}{L_{\textrm{fl}}}
-\newcommand{\hfl}{h_{\textrm{fl}}}
-\newcommand{\hannw}{w}
-\newcommand{\hannwnorm}{\tilde{\hannw}}
-\newcommand{\nw}{N_{\hannw}}
-\newcommand{\hannwpart}{\gamma}
-\newcommand{\tslide}{t_w}
-
-\newcommand{\bs}{\sigma_{\textrm{bs}}}
-\newcommand{\mysp}{S_p}
-\newcommand{\ts}{\textrm{TS}}
-\newcommand{\sv}{S_{\textrm{v}}}
-
-\newcommand{\range}{r}
-\newcommand{\rangeref}{r_0}
-\newcommand{\athw}{\phi}
-\newcommand{\along}{\theta}
-\newcommand{\gain}{g}
-\newcommand{\gainzero}{g_0}
-\newcommand{\eqang}{\psi}
-
-\newcommand{\rtarget}{r_{\textrm{target}}}
-\newcommand{\alongtarget}{\theta_{\textrm{target}}}
-\newcommand{\athwtarget}{\phi_{\textrm{target}}}
-
-\newcommand{\wlen}{\lambda}
-\newcommand{\cw}{c}
-\newcommand{\absorp}{\alpha}
-
-\newcommand{\dft}{\textrm{DFT}}
-\newcommand{\ndft}{{N_{\textrm{DFT}}}}
-\newcommand{\ndftw}{_{\nw}}
-\newcommand{\atan}{\textrm{arctan2}}
-\newcommand{\anglefalong}{\gamma_\along}
-\newcommand{\anglefathw}{\gamma_\athw}
-
-\newcommand{\sigmabs}{\sigma_{\textrm{bs}}}
-
-\newcommand{\code}[1]{\texttt{#1}}
+% All the \renewcommand stuff is now in a separate file so that it can be 
+% reused for the list of symbols in the Supplementary document.
+\input{symbol_aliases.tex}
 
@@
 \includegraphics{Fig_ACF}
-\caption{The autocorrelation function $\ptxauto$ from the example. The corresponding $\teff=0.1 ms$ and $\tnom=2 ms$\label{fi:ACF}.}
+\caption{The autocorrelation function $\ptxauto$ from the example. The corresponding $\teff=0.1$ ms and $\tnom=2$ ms\label{fi:ACF}.}
 \end{figure}
@@
 
-Generalizing the power-budget equation (i.e., sonar equation), given in e.g. \citep{lunde2016}, for broadband signals yields in logarithmic form $TS$ at frequency $\freqsym$:
+Generalizing the power-budget equation (i.e., sonar equation) for broadband signals \citep{lunde2016} yields, in logarithmic form, $TS$ at frequency $\freqsym$:
 %
@@
 %
-where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption coefficient, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
+where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp(\freqsym)$ the acoustic absorption coefficient, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain(\along,\athw,\freqsym)$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
 
@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes that occur in actual targets (as opposed to ideal point targets).
-The alongship angle $\along_n$, athwartship angle $\athw_n$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
+The alongship angle $\along(\samplesymt)$, athwartship angle $\athw(\samplesymt)$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
 \ref{fi:theta_phi}).
@@
 If the number of samples around the target is more than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used.
-If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create a reduced autocorrelation signal, $\ymfautored(\samplesymt)$.
+If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create $\ymfautored(\samplesymt)$.
 %
@@
 %
-where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang$ is the two-way equivalent beam angle.
+where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang(\freqsym)$ is the two-way equivalent beam angle.
 The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
@@
 
-By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($freqsym$) for each ping.
-The range for centre samples $\samplesymt$ could be chosen to half the window length or any other grid that the user prefer the data presented to be in.
+By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($\freqsym$) for each ping.
+The range for centre samples $\samplesymt$ could be chosen as half the window length or any other grid that the user prefer the data presented to be in.
 This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
@@
 Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
-We note that these methods do not provide an operational method to estimate $\teff$ or $\Psi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
+We note that these methods do not provide an operational method to estimate $\teff$ or $\eqang(\freqsym)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
 

### 2023-09-12 Paper/Supplementary.tex
Improve description for symbol u

@@
 
-$\gain(\along,\athw,\samplesymf)$ & & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency index.\\
+% $\gain(\along,\athw,\samplesymf)$ & & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency index.\\
 
@@
 
-$\channelsym$ & & 1 & Channel number/transducer sector.\\
+$\channelsym$ & & 1 & Receiver channel number and transducer sector number.\\
 $u(\genidxsym)$ & & 1 & The Heaviside step function. \\
@@
 $\yrxa(\timesym)$ & & Pa & Analogue acoustic signal received by the transducer.\\
-$\yrxa(\timesym,\channelsym)$ & & Pa & Analogue acoustic signal received by each transducer sector $\channelsym$.\\
+% $\yrxa(\timesym,\channelsym)$ & & Pa & Analogue acoustic signal received by each transducer sector $\channelsym$.\\
 $\yrxe(\timesym,\channelsym)$ & & V & Analogue electric signal received by each transceiver channel $\channelsym$.\\
@@
 $\absorp(\freqsym)$  & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency $f$.\\
-$\absorp(\samplesymf)$ & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency index $m$\\
+% $\absorp(\samplesymf)$ & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency index $m$\\
 $\anglefalong$ & & 1 & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo.\\
@@
 $\along$ & & rad & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
-$\along(\samplesymt)$ & & rad & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
+% $\along(\samplesymt)$ & & rad & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
 $\wlen$ & & m &  Acoustic wavelength.\\
@@
 $\athw$ & & rad & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
-$\athw(\samplesymt)$ & & rad & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
+% $\athw(\samplesymt)$ & & rad & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
 $\eqang(\freqsym)$ & & sr & Two-way equivalent beam angle at frequency $\freqsym$.\\

### 2023-09-13 Core/Calculation.py
Change docstring format to Sphinx Numpy style and add more docs
Unknown parts of the docstring are flagged with XXX

@@ -13,17 +13,25 @@ class Calculation(EK80DataContainer):
         """
         Generate the ideal windowed transmit signal.
         
-        Parameters:
-            f0 -- the start frequency [Hz]
-            f1 -- the end frequency [Hz]
-            tau -- the transmit signal duration [s]
-            fs -- the signal sample rate [Hz]
-            slope -- the proportion of the pulse that is windowed [1].
-                     This includes the start and end parts of the windowing.
+        Parameters
+        ----------
+            f0 : float
+                The start frequency [Hz]
+            f1 : float
+                The end frequency [Hz]
+            tau : float
+                The transmit signal duration [s]
+            fs : float
+                The signal sample rate [Hz]
+            slope : float
+                The proportion of the pulse that is windowed [1]. This includes the start and end parts of the windowing.
                      
-        Returns:
-            y -- the ideal windowed transmit signal amplitude [1]
-            t -- the time of each value in y [s]
+        Returns
+        -------
+            y : np.array
+                The ideal windowed transmit signal amplitude [1]
+            t : np.arrayfloat
+                The time of each value in y [s]
         
         """
         
@@ -49,14 +57,21 @@ class Calculation(EK80DataContainer):
         """
         Generate a chirp pulse.
         
-        Parameters:
-            t -- times at which to calculate the chirp signal [s]
-            f0 -- start frequency of the pulse [Hz]
-            t1 -- duration of the pulse [s]
-            f1 -- end frequency of the pulse [Hz]
+        Parameters
+        ----------
+            t : float
+                Times at which to calculate the chirp signal [s]
+            f0 : float
+                Start frequency of the pulse [Hz]
+            t1 : float
+                Duration of the pulse [s]
+            f1 : float
+                End frequency of the pulse [Hz]
             
-        Returns:
-            The chirp pulse amplitude at times t [1]
+        Returns
+        -------
+            chirp : np.array
+                The chirp pulse amplitude at times t [1]
         """
         
         a = np.pi * (f1 - f0) / t1
@@ -69,11 +84,15 @@ class Calculation(EK80DataContainer):
         """
         Generate Hann window weights.
         
-        Parameters:
-            L -- the number of samples to use [1]
+        Parameters
+        ----------
+            L : int
+                The number of samples to use [1]
[…]

### 2023-09-13 Core/FIL1.py
Add docstring to FIL1 class

@@ -1,5 +1,23 @@
 class FIL1:
+    """
+    Contains information about a signal processing filter.
+    
+    Attributes
+    ----------
+        noOfCoefficients : int
+            Number of filter coefficients [1]
+        decimationFactor : int
+            Filter decimation factor [1]
+        coefficients : np.array
+            Filter coefficients [1]
+                    
+    """
     def __init__(self, noOfCoefficients, decimationFactor, coefficients):
+        """
+        Initialise the FIL1 class.
+        
+        """
+
         self.NoOfCoefficients = noOfCoefficients
         self.DecimationFactor = decimationFactor
         self.Coefficients = coefficients

### 2023-09-13 Core/Calculation.py
Add template docstrings to all functions

@@ -251,10 +251,31 @@ class Calculation(EK80DataContainer):
 
     @staticmethod
     def calcAverageSignal(y_pc_nu):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        Returns
+        -------
+        
+        """
+        
         return np.sum(y_pc_nu, axis=0) / y_pc_nu.shape[0]
 
     @staticmethod
     def calcTransducerHalves(y_pc_nu):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        Returns
+        -------
+        
+        """
         y_pc_fore_n = 0.5 * (y_pc_nu[2, :] + y_pc_nu[3, :])
         y_pc_aft_n = 0.5 * (y_pc_nu[0, :] + y_pc_nu[1, :])
         y_pc_star_n = 0.5 * (y_pc_nu[0, :] + y_pc_nu[3, :])
@@ -268,6 +289,16 @@ class Calculation(EK80DataContainer):
 
     @staticmethod
     def calcPower(y_pc, z_td_e, z_rx_e):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        Returns
+        -------
+        
+        """
         K1 = 4 / ((2 * np.sqrt(2)) ** 2)
         K2 = (np.abs(z_rx_e + z_td_e) / z_rx_e) ** 2
         K3 = 1.0 / np.abs(z_td_e)
@@ -279,6 +310,16 @@ class Calculation(EK80DataContainer):
 
     @staticmethod
     def calcAngles(y_pc_halves, gamma_theta, gamma_phi):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        Returns
+        -------
+        
+        """
         # Transducers might have different segment configuration
         # Here we assume 4 quadrants
         y_pc_fore_n, y_pc_aft_n, y_pc_star_n, y_pc_port_n = y_pc_halves
@@ -305,6 +346,16 @@ class Calculation(EK80DataContainer):
 
     @staticmethod
     def calcSp(p_rx_e_n, r_n, alpha_f_c, p_tx_e, lambda_f_c, g_0_f_c, r0=None, r1=None):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        Returns
+        -------
+        
+        """
         # Pick a range of the data
         if r0 is not None and r1 is not None:
[…]

### 2023-09-13 Core/EK80DataContainer.py
Add template docstrings to all functions and classes

@@ -1,27 +1,82 @@
 import numpy as np
 import json
-
+"""
+A set of classes that store data from an EK80 .raw file. Classes are:
+    Constants
+    Transceiver
+    Parameter
+    Transducer
+    Environment
+    FrequencyPar
+    Filters
+    Raw3
+    EK80DataContainer
+"""
 
 class Constants:
+    """
+    XXX.
+    
+    Attributes
+    ----------
+    
+    """
     def __init__(self, z_td_e, f_s, n_f_points):
         self.z_td_e = z_td_e
         self.f_s = f_s
         self.n_f_points = n_f_points
 
     def getParameters(self):
+        """
+        XXX.
+        
+        Returns
+        -------
+                       
+        """
+
         return self.z_td_e, self.f_s, self.n_f_points
 
 
 class Transceiver:
+    """
+    XXX.
+    
+    Attributes
+    ----------
+    
+    """
     def __init__(self, xml):
         self.z_rx_e = xml["z_rx_e"]
 
     def getParameters(self):
+        """
+        XXX.
+        
+        Returns
+        -------
+        
+        """
+
         return self.z_rx_e
 
 
 class Parameter:
+    """
+    XXX.
+    
+    Attributes
+    ----------
+    
+    """
     def __init__(self, xml):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        """
         self.f0 = xml["FrequencyStart"]
         self.f1 = xml["FrequencyEnd"]
         self.f_c = (self.f0 + self.f1) / 2.0
@@ -31,6 +86,14 @@ class Parameter:
         self.p_tx_e = xml["TransmitPower"]
 
     def getParameters(self):
+        """
+        XXX.
+        
+        Returns
+        -------
+        
+        """
+
         return (
             self.f0,
             self.f1,
@@ -43,7 +106,23 @@ class Parameter:
 
 
 class Transducer:
+    """
+    XXX.
+    
+    Attributes
+    ----------
+    
+    """
     def __init__(self, xml):
+        """
+        XXX.
+        
+        Parameters
+        ----------
+        
+        
+        """
+
         self.f_n = xml["Frequency"]  # nominal design frequency for the transducer
         self.G_f_n = xml["GainNom"]
         Psi_f_n = xml["EquivalentBeamAngle"]
@@ -57,6 +136,15 @@ class Transducer:
         self.corrSa = xml["SaCorrection"]
 
     def getParameters(self):
+        """
[…]

### 2023-09-13 main.py
Add template docstrings to more files.

@@ -6,6 +6,14 @@ from plots import *
 
 
 def preCalculations(data):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     global z_td_e, z_rx_e, N_u
     global f_0, f_1, tau, f_s, slope, p_tx_e
     global filter_v, f_s_dec_v
@@ -81,6 +89,14 @@ def preCalculations(data):
 
 
 def calcBasics(do_plot):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     global f_s_dec_v, f_s_dec
     global y_tx_n, y_mf_n, y_mf_auto_n, tau_eff
     global y_pc_n, p_rx_e_n, theta_n, phi_n
@@ -183,6 +199,14 @@ def calcBasics(do_plot):
 
 
 def calcTS():
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     global f_m, Y_pc_t_m, Y_mf_auto_red_m, Y_tilde_pc_t_m, g_theta_phi_m, TS_m
     global r_t, theta_t, phi_t, y_pc_t_n, dum_p, dum_theta, dum_phi, dum_r
     global y_mf_auto_red_n
@@ -245,6 +269,14 @@ def calcTS():
 
 
 def calcSv():
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     global p_rx_e_n, f_m, svf_range, Sv_m_n
 
     #

### 2023-09-13 plots.py

@@ -5,6 +5,13 @@ from scipy.signal import hilbert
 
 
 def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
     # Example of ideal windowed transmit signal with slope 0.5
     y_tx_n05slope, t = Calculation.generateIdealWindowedTransmitSignal(
         f_0, f_1, tau, f_s, 0.5
@@ -23,6 +30,14 @@ def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
 
 
 def plotfir(filter_v, f_s_dec_v, f_0, f_1):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     # The frequency response function of the filter is given by its
     # discrete time fourier transform:
     H0 = np.fft.fft(filter_v[0]["h_fl_i"])
@@ -51,6 +66,14 @@ def plotfir(filter_v, f_s_dec_v, f_0, f_1):
 
 
 def plotymfn(y_mf_n):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     plt.figure()
     plt.plot(np.abs(y_mf_n))
     # plt.title('The absolute value of the filtered and decimated output signal')
@@ -60,6 +83,14 @@ def plotymfn(y_mf_n):
 
 
 def plotACF(y_mf_auto_n):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     plt.figure()
     plt.plot(np.abs(y_mf_auto_n))
     # plt.title('The autocorrelation function of the matched filter.')
@@ -69,6 +100,14 @@ def plotACF(y_mf_auto_n):
 
 
 def plotThetaPhi(theta_n, phi_n):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
     # Plot angles
     fig, axs = plt.subplots(2, sharex=True)
     # fig.suptitle('Single target')
@@ -81,6 +120,14 @@ def plotThetaPhi(theta_n, phi_n):
 
 
 def plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red_n):
+    """
+    XXX.
+    
+    Parameters
+    ----------
+    
+    """
+
[…]

### 2023-09-13 Core/Calculation.py
Fix an indentation error

@@ -608,7 +608,7 @@ class Calculation(EK80DataContainer):
         -------
         
         """
-       y_pc_s_n = y_pc_n * r_c_n
+        y_pc_s_n = y_pc_n * r_c_n
 
         return y_pc_s_n
 

### 2023-09-13 Core/EK80DataContainer.py
Fix some grammar in output

@@ -226,7 +226,7 @@ class FrequencyPar:
 
         # Test if broadband calibration values exists, if not use nominal values and extrapolate
         if xml["frequencies"]:
-            print("Broadband calibration values exists")
+            print("Broadband calibration values exist")
             self.frequencies = xml["frequencies"]
             self.gain = xml["gain"]
             self.angle_offset_athwartship = xml["angle_offset_athwartship"]
@@ -237,7 +237,7 @@ class FrequencyPar:
             self.isCalibrated = True
         elif not xml["frequencies"]:
             print(
-                "Broadband calibration values does not exist - use nominal and fit function"
+                "Broadband calibration values do not exist - using nominal and fit function"
             )
             self.frequencies = None
             self.gain = None

### 2023-09-13 Core/Calculation.py
Fix indentation in docstrings

@@ -15,24 +15,24 @@ class Calculation(EK80DataContainer):
         
         Parameters
         ----------
-            f0 : float
-                The start frequency [Hz]
-            f1 : float
-                The end frequency [Hz]
-            tau : float
-                The transmit signal duration [s]
-            fs : float
-                The signal sample rate [Hz]
-            slope : float
-                The proportion of the pulse that is windowed [1]. This includes the start and end parts of the windowing.
-                     
+        f0 : float
+            The start frequency [Hz]
+        f1 : float
+            The end frequency [Hz]
+        tau : float
+            The transmit signal duration [s]
+        fs : float
+            The signal sample rate [Hz]
+        slope : float
+            The proportion of the pulse that is windowed [1]. This includes the start and end parts of the windowing.
+                 
         Returns
         -------
-            y : np.array
-                The ideal windowed transmit signal amplitude [1]
-            t : np.arrayfloat
-                The time of each value in y [s]
-        
+        y : np.array
+            The ideal windowed transmit signal amplitude [1]
+        t : np.arrayfloat
+            The time of each value in y [s]
+    
         """
         
         nsamples = int(np.floor(tau * fs))
@@ -59,19 +59,19 @@ class Calculation(EK80DataContainer):
         
         Parameters
         ----------
-            t : float
-                Times at which to calculate the chirp signal [s]
-            f0 : float
-                Start frequency of the pulse [Hz]
-            t1 : float
-                Duration of the pulse [s]
-            f1 : float
-                End frequency of the pulse [Hz]
+        t : float
+            Times at which to calculate the chirp signal [s]
+        f0 : float
+            Start frequency of the pulse [Hz]
+        t1 : float
+            Duration of the pulse [s]
+        f1 : float
+            End frequency of the pulse [Hz]
             
         Returns
         -------
-            chirp : np.array
-                The chirp pulse amplitude at times t [1]
+        np.array
+            The chirp pulse amplitude at times t [1]
         """
         
         a = np.pi * (f1 - f0) / t1
@@ -86,13 +86,13 @@ class Calculation(EK80DataContainer):
         
         Parameters
         ----------
-            L : int
-                The number of samples to use [1]
-            
+        L : int
+            The number of samples to use [1]
+        
         Returns
         -------
[…]

### 2023-09-13 Core/FIL1.py

@@ -4,13 +4,13 @@ class FIL1:
     
     Attributes
     ----------
-        noOfCoefficients : int
-            Number of filter coefficients [1]
-        decimationFactor : int
-            Filter decimation factor [1]
-        coefficients : np.array
-            Filter coefficients [1]
-                    
+    noOfCoefficients : int
+        Number of filter coefficients [1]
+    decimationFactor : int
+        Filter decimation factor [1]
+    coefficients : np.array
+        Filter coefficients [1]
+                
     """
     def __init__(self, noOfCoefficients, decimationFactor, coefficients):
         """

### 2023-09-13 Core/Calculation.py
Add more docstrings

@@ -291,14 +291,22 @@ class Calculation(EK80DataContainer):
     @staticmethod
     def calcPower(y_pc, z_td_e, z_rx_e):
         """
-        XXX.
+        Calculate the received power into a matched load.
         
         Parameters
         ----------
+        y_pc : np.array
+            Pulse compressed signal [V]
+        z_td_e : float
+            Transducer electrical impedance [Ω]
+        z_rx_e :
+            Receiver electrical impedance [Ω]
         
         Returns
         -------
-        
+        np.array
+            Received electrical power [W]
+            
         """
         K1 = 4 / ((2 * np.sqrt(2)) ** 2)
         K2 = (np.abs(z_rx_e + z_td_e) / z_rx_e) ** 2

### 2023-09-13 Core/Calculation.py
Remove hardcoded assumption that acoustic data is from a 4 receiver/sector system

@@ -289,9 +289,11 @@ class Calculation(EK80DataContainer):
     #
 
     @staticmethod
-    def calcPower(y_pc, z_td_e, z_rx_e):
+    def calcPower(y_pc, z_td_e, z_rx_e, nu):
         """
         Calculate the received power into a matched load.
+        
+        Output received power values of 0.0 are set to 1e-20.
         
         Parameters
         ----------
@@ -301,6 +303,8 @@ class Calculation(EK80DataContainer):
             Transducer electrical impedance [Ω]
         z_rx_e :
             Receiver electrical impedance [Ω]
+        nu : int
+            Number of receiver channels [1]
         
         Returns
         -------
@@ -308,7 +312,7 @@ class Calculation(EK80DataContainer):
             Received electrical power [W]
             
         """
-        K1 = 4 / ((2 * np.sqrt(2)) ** 2)
+        K1 = nu / ((2 * np.sqrt(2)) ** 2)
         K2 = (np.abs(z_rx_e + z_td_e) / z_rx_e) ** 2
         K3 = 1.0 / np.abs(z_td_e)
         C1Prx = K1 * K2 * K3

### 2023-09-13 main.py

@@ -186,7 +186,7 @@ def calcBasics(do_plot):
     #
 
     # Total received power for all channels (all transducer sectors)
-    p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e)
+    p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e, y_pc_nu.shape[0])
 
     # Physical angles
     theta_n, phi_n = Calculation.calcAngles(

### 2023-09-13 Paper/Manuscript.tex
Consistently use splitbeam instead of split-aperture in manuscript text

@@
 The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water.
-For a split-aperture echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
+For a splitbeam echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
 
@@
 Compensation of echo strength for position in the acoustic beam requires an estimate of the echo arrival angle.
-This is obtained using the split-aperture method \citep{burdic1991}, which for broadband pules can be implemented with the angle values contained in the complex-valued $\ypc(\samplesymt)$ data, in combination with knowledge of transducer sector geometry.
+This is obtained using the splitbeam method \citep{burdic1991}, which for broadband pules can be implemented with the angle values contained in the complex-valued $\ypc(\samplesymt)$ data, in combination with knowledge of transducer sector geometry.
 The principle is demonstrated with a transducer that is divided into four quadrants (Fig.

### 2023-09-13 Core/Calculation.py
Fill in some of the docstrings templates

@@ -324,17 +324,26 @@ class Calculation(EK80DataContainer):
     @staticmethod
     def calcAngles(y_pc_halves, gamma_theta, gamma_phi):
         """
-        XXX.
+        Calculate splitbeam angles from 4 quadrant receiver/transducer data.
         
         Parameters
         ----------
+        y_pc_halves : np.array
+            Pulse compressed signal from transducer halves [V]
+        gamma_theta : float
+             Major axis conversion factor from electrical splitbeam angles to physical angles []
+        gamma_phi : float
+            Minor axis conversion factor from electrical splitbeam angles to physical angles []
         
         Returns
         -------
+        theta_n : np.array
+            Major axis physical splitbeam angles [°]
+        phi_n : np.array
+            Minor axis physical splitbeam angles [°]
         
         """
-        # Transducers might have different segment configuration
-        # Here we assume 4 quadrants
+
         y_pc_fore_n, y_pc_aft_n, y_pc_star_n, y_pc_port_n = y_pc_halves
 
         y_theta_n = y_pc_fore_n * np.conj(y_pc_aft_n)
@@ -507,13 +516,21 @@ class Calculation(EK80DataContainer):
 
     def calcg0(self, f):
         """
-        XXX.
+        Calculate gain at given frequencies.
+        
+        Uses calibration data if available, otherwise uses a theoretical 
+        variation of gain with frequency relationship and nominal gain at
+        a nominal frequency.
         
         Parameters
         ----------
+        f : np.array
+            Frequencies to calculate the gain at [Hz]
         
         Returns
         -------
+        np.array
+            Gain values at the given frequencues, `f` [dB]
         
         """
         if self.isCalibrated:
@@ -526,13 +543,23 @@ class Calculation(EK80DataContainer):
     @staticmethod
     def calcg0_calibrated(f, freq, gain):
         """
-        XXX.
+        Interpate frequency dependent gain data onto specified frequencies.
+        
+        Uses one-dimensional piecewise linear interpolation.
         
         Parameters
         ----------
+        f : np.array
+            Frequencies to interpolate on to [Hz]
+        freq : np.array
+            Frequencies at which gain values are provided [Hz]
+        gain : np.array
+            Gain values to use in the interpolation [dB]
         
         Returns
         -------
+        np.array
+            The gain values at frequencies given in `f` [1]
         
         """
         dB_G0 = np.interp(f, freq, gain)
[…]

### 2023-09-14 Paper/Supplementary.tex
Revert "Improve description for symbol u"
This reverts commit b93f5043c33068647469ddbdcd10471b89662f11.
Modification to Supplementary to match paper

@@
 
-$\hfl(\genidxsym,\stagesym)$ & & 1 & Complex valued receiving filter coefficients.\\
+$\hbp(\genidxsym,\stagesym)$ & & 1 & Complex valued receiving filter coefficients.\\
 $\genidxsym$ & & 1 & Generic integer index.\\
@@
 $\sv(\freqsym)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at frequency $\freqsym$.\\
-$\sv(\samplesymt)$ & & dB re $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
+$\sv(\samplesymt)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
 
@@
 
-$\hannw$ & & 1 & The Hanning window function.\\
-$\hannw(\genidxsym)$ & & 1 & The Hanning window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
-$\hannwnorm(\genidxsym)$ & & 1 & Normalised Hanning window.\\
+$\hannw$ & & 1 & The Hann window function.\\
+$\hannw(\genidxsym)$ & & 1 & The Hann window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
+$\hannwnorm(\genidxsym)$ & & 1 & Normalised Hann window.\\
 
 $x^*$ & & -- & The complex conjugate of $x$.\\
-$||x||$ & & -- & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
+$||x||_2$ & & -- & The $l^2$-norm of $x$, also known as the Euclidean norm.\\
 

### 2023-09-14 Paper/Manuscript.tex
Revised version of paper

@@
 \usepackage{lineno}
+\usepackage{xurl}
+
 \linenumbers
@@
 
-% All the \renewcommand stuff is now in a separate file so that it can be 
-% reused for the list of symbols in the Supplementary document.
-\input{symbol_aliases.tex}
+\newcommand{\ek}{Simrad EK80}
+
+\newcommand{\timesym}{t}
+\newcommand{\freqsym}{f}
+\newcommand{\samplesymt}{n}
+\newcommand{\samplesymf}{m}
+\newcommand{\genidxsym}{i}
+
+\newcommand{\channelsym}{u}
+\newcommand{\nchannels}{N_{\textrm{u}}}
+
+\newcommand{\stagesym}{v}
+\newcommand{\nstages}{N_{\textrm{v}}}
+
+\newcommand{\fs}{f_{\textrm{s}}}
+\newcommand{\fsdec}{f_{\textrm{s,dec}}}
+\newcommand{\fstart}{f_{\textrm{start}}}
+\newcommand{\fstop}{f_{\textrm{stop}}}
+\newcommand{\fc}{f_{\textrm{c}}}
+\newcommand{\fn}{f_{\textrm{n}}}
+
+\newcommand{\zrxe}{z_{\textrm{rx,e}}}
+\newcommand{\ztde}{z_{\textrm{td,e}}}
+
+\newcommand{\ptxe}{p_{\textrm{tx,e}}}
+\newcommand{\prxe}{p_{\textrm{rx,e}}}
+
+\newcommand{\ntd}{n_{\textrm{td}}}
+\newcommand{\tnom}{\tau}
+\newcommand{\teff}{\tau_{\textrm{eff}}}
+
+\newcommand{\ytxe}{y_{\textrm{tx,e}}}
+\newcommand{\ytxa}{y_{\textrm{tx,a}}}
+\newcommand{\yrxa}{y_{\textrm{rx,a}}}
+\newcommand{\yrxe}{y_{\textrm{rx,e}}}
+
+\newcommand{\ytx}{y_{\textrm{tx}}}
+\newcommand{\ytxnorm}{\tilde{y}_{\textrm{tx}}}
+\newcommand{\ytxfd}{y_{\textrm{tx,f,d}}}
+\newcommand{\yrx}{y_{\textrm{rx}}}
+\newcommand{\yrxorg}{y_{\textrm{rx,org}}}
+\newcommand{\ymf}{y_{\textrm{mf}}}
+
+\newcommand{\ytd}{y_{\textrm{td}}}
+
+\newcommand{\ypc}{y_{\textrm{pc}}}
+\newcommand{\ypctarget}{y_{\textrm{pc,t}}}
+\newcommand{\ypcspread}{y_{\textrm{pc,s}}}
+\newcommand{\ymfauto}{y_{\textrm{mf,auto}}}
+\newcommand{\ymfautored}{y_{\textrm{mf,auto,red}}}
+
+\newcommand{\ptxauto}{p_{\textrm{tx,auto}}}
+
+\newcommand{\ypctargetf}{Y_{\textrm{pc,t}}}
+\newcommand{\ypctargetnormf}{\tilde{Y}_{\textrm{pc,t}}}
+\newcommand{\ypcvolumef}{Y_{\textrm{pc,v}}}
+\newcommand{\ypcvolumenormf}{\tilde{Y}_{\textrm{pc,v}}}
+
+\newcommand{\ymfautof}{Y_{\textrm{mf,auto}}}
+\newcommand{\ymfautoredf}{Y_{\textrm{mf,auto,red}}}
+
+\newcommand{\prxetf}{P_{\textrm{rx,e,t}}}
+\newcommand{\prxevf}{P_{\textrm{rx,e,v}}}
+
+\newcommand{\fstage}{k}
+\newcommand{\decfac}{D}
+\newcommand{\lfl}{L_{\textrm{fl}}}
+\newcommand{\hfl}{h_{\textrm{fl}}}
+\newcommand{\hbp}{h_{\textrm{bp}}}
+\newcommand{\hannw}{w}
+\newcommand{\hannwnorm}{\tilde{\hannw}}
+\newcommand{\nw}{N_{\hannw}}
+\newcommand{\hannwpart}{\gamma}
+\newcommand{\tslide}{t_w}
+
+\newcommand{\bs}{\sigma_{\textrm{bs}}}
+\newcommand{\mysp}{S_p}
+\newcommand{\ts}{\textrm{TS}}
+\newcommand{\sv}{S_{\textrm{v}}}
+
+\newcommand{\range}{r}
+\newcommand{\rangeref}{r_0}
+\newcommand{\athw}{\phi}
+\newcommand{\along}{\theta}
+\newcommand{\gain}{g}
+\newcommand{\gainzero}{g_0}
+\newcommand{\eqang}{\psi}
+
+\newcommand{\rtarget}{r_{\textrm{target}}}
+\newcommand{\alongtarget}{\theta_{\textrm{target}}}
+\newcommand{\athwtarget}{\phi_{\textrm{target}}}
+
+\newcommand{\wlen}{\lambda}
+\newcommand{\cw}{c}
+\newcommand{\absorp}{\alpha}
+
+\newcommand{\dft}{\textrm{DFT}}
+\newcommand{\ndft}{{N_{\textrm{DFT}}}}
+\newcommand{\ndftw}{_{\nw}}
+\newcommand{\atan}{\textrm{arctan2}}
+\newcommand{\anglefalong}{\gamma_\along}
+\newcommand{\anglefathw}{\gamma_\athw}
+
+\newcommand{\sigmabs}{\sigma_{\textrm{bs}}}
+
+\newcommand{\code}[1]{\texttt{#1}}
 
@@
 \author{Lars Nonboe Andersen}
-\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}
+\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}
 
@@
 \author{Harald Heimvoll}
-\affiliation{Kongsberg Maritime AS, Strandpromenaden 50, 3191, Horten, Norway}
+\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}
 
@@
 1.
-The use of quantitative broadband echosounders for biological studies and surveys offers considerable advantages over narrowband echosounders.
+The use of quantitative broadband echosounders for biological studies and surveys can offer considerable advantages over narrowband echosounders.
 These include improved spectral-based target identification and significantly increased ability to resolve individual targets.
@@
 We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength (TS(f)) and volume backscattering strength (Sv(f)) using data from the EK80 broadband scientific echosounder as examples.
-Although the overall processing steps are described and build on established methods from literature, multiple choices need to be made during implementation.
+Although the overall processing steps are described and build on established methods from the literature, multiple choices need to be made during implementation.
 \linebreak
 3.
-To highlight and discuss some of these choices and facilitating a common understanding within the community, we have also developed a code which will be made publicly available and open source.
+To highlight and discuss some of these choices and facilitate a common understanding within the community, we have also developed a Python code which will be made publicly available and open source.
 The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to TS(f) and Sv(f).
@@
 4.
-This code can serve as a reference for developing own code or implementation in existing processing pipelines, as an educational tool and as a starting point for further development of broadband acoustic methods in fisheries acoustics.\linebreak
+This code can serve as a reference for developing custom code or implementation in existing processing pipelines, as an educational tool and as a starting point for further development of broadband acoustic methods in fisheries acoustics.\linebreak
 
@@
 
-Active acoustics is an efficient method for remote sensing of marine ecosystems and is used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}.
+Echosounders are used for remote sensing marine ecosystems.
 As early as 1935, \citet{sund_echo_1935} observed the distribution of spawning cod in the Lofoten area using a single beam echosounder.
 The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}.
-More recently, active acoustics have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
+More recently, fisheries acoustics sensors have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
 
-Several different configurations of acoustic instruments are available, including multibeam sonars \citep{gerlotto_two_1999}, synthetic aperture sonar, omnidirectional sonars \citep{misund_improved_1996}, acoustic imaging sonars \citep{jaffe_ftv_1995} as well as single- and multibeam echosounders \citep{trenkel_new_2008}.
-Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and the methodologies for using these systems are well developed.
-The ability for accurate calibration of the system has made these especially useful for providing data for fisheries management \citep{Simmonds2005Fisheries}.
-
-Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), compared to conventional single frequency systems.
-This provides significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
-
-Broadband echosounders have been actively investigated for ecological research, based on the expectation of improved categorization of species or group of species.
-Using high frequency broadband acoustics (150-600 kHz) \citet{lavery_measurements_2010} showed that broadband signals helped to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
-\citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages of broadband signals.
-\citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to identify that the composition of two sound scattering layers (SSL) detected in the Bay of Biscay in springtime was able to be characterized, and also identified fine scale heterogeneity within the SSLs.
-While unsuccessful at discriminating between several species of fishes with large swimbladders during the Alaska pollock survey, \citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids.
-\citet{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz), while attempts to classify species using multiple narrowband signals failed.
-
-The increased spatial resolution is another aspect of broadband acoustics that is being actively used both for improved categorization and the increased ability to resolve targets near each other or near boundaries such as the seafloor.
+Today echosounders can produce pulses with a wide and continuous frequency range (broadband pulses), compared to the conventional narrowband systems.
+This provides in most cases significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
+\citet{lavery_measurements_2010} used broadband signals to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
+\citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish to classify size classes.
+\citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to characterize the composition of two sound scattering layers (SSL).
+\citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids.
+\citet{benoit-bird_exploring_2020} were able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz).
 \citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments.
@@
 
-An equally interesting outcome from the increased resolution and broad frequency information is the ability to infer behavioural information for single individuals, groups of individuals, and interaction between individuals or groups.
-\citet{Traykovski1998Effect} was able to extract orientation information for krill by using a combination of high frequency broadband acoustics and acoustic modelling.
-Fine-scale observations of single targets facilitated the study of interactions between prey and predators (capelin and cod) in the Barents sea \citep{skaret_diel_2020}.
+Several scientific broadband echosounder systems have been developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
+More widespread use of broadband acoustics in fisheries and ecosystem research have followed from the upgrade of widely used narrowband systems to gain broadband capabilities.
 
-There have been several scientific broadband echosounder systems developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
-Commercially available systems are now produced by several manufacturers.
+Signal processing methods for broadband echosounders \citep{stanton2008} are based on radar signal processing theory, with further adaptation to echosounders \citep{lavery2017,bassett_broadband_2018}.
+When implementing the equations to computer processing code, some choices are well founded in the signal processing literature whereas others are of a more practical and ad-hoc nature.
+The latter is typically missing in the literature, making it difficult to benchmark new methods and to test the implementation in new echosounders and post processing software.
 
-The theoretical foundation for signal processing comes from radar applications and has been documented for broadband echosounders \citep{stanton2008}.
-To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}.
-Examples include adapting pulses to improve target separation and obtain additional parameters to describe the individual targets.
-Different transmission pulse forms can be envisioned as well as various methods for acoustic target classification.
-When translating equations into computer processing code, several choices must be made - some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature.
-The latter is typically missing in the literature, making it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.
-By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.
-
-The objective of this paper and associated code is to present a systematic and comprehensive description of the data processing steps for calibrated broadband echosounder data.
-The steps include pulse compression, target strength as a function of frequency [$\ts(\freqsym)$], and volume backscattering strength as a function of frequency [$\sv(\freqsym)$].
-The intention is that the code will be used as a starting point for implementations in various relevant data processing software and for further developing active acoustic broadband signal processing, and to serve as a learning resource.
+The objective of this paper and associated code is to provide a benchmark for developing and implementing signal processing algorithms for broadband echosounder.
+We present the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data.
+The steps include pulse compression, target strength as a function of frequency ($\ts(\freqsym)$, dB re 1m$^2$), and volume backscattering strength as a function of frequency ($\sv(\freqsym)$, dB re 1m$^-1$).
+The intention is that the code will be used as a starting point for implementations in various relevant data processing software, for further developing active acoustic broadband signal processing, and to serve as a learning resource.
 
@@
 \subsection{Accompanying code}
-The code accompanying this paper is written in the Python programming language and is available in the supplementary materials.
+The code  accompanying this paper is written in the Python programming language (v3.10) and is available through GitHub (see \ref{dataavstatement} for code and data availability) together with the data used in the examples.
 All single ping processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
-Reproduction of figures 9 and 12 requires the original echosounder raw data and are not included.
+Reproduction of Fig.
+\ref{fi:ts_theta_phi}a and Fig.
+\ref{fi:Fig_Sv_echogram}a requires downloading the original echosounder raw data separately and using the scripts (\code{TSfEchogram.py} and \code{SvfEchogram.py}).
 Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
+The test data sets accompanying the code are representative for the data contained in Simrad EK80 raw files.
+In order to make the code echosounder independent simple json strings are used as input.
+Detailed information on settings used during data collection are available in the code and the data files on GitHub.
 
 Our presentation uses nomenclature and approaches that are commonly used for narrowband echosounder systems, which were derived from radar processing \citep{cook1967}.
-In particular, the expressions for target strength ($\ts$) and volume backscattering strength ($\sv$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.
+In particular, the expressions for target strength ($\ts$, dB re 1m${^2}$) and volume backscattering strength ($\sv$, dB re 1m$^{-1}$) \citep{MacLennan2002consistent} are presented in a similar manner for broadband signals as for narrowband signals.
 
@@
 \ref{fi:ek_sys}).
-Many types of transmit signals are feasible - this paper considers only linear frequency modulated signals (also known as linear chirps).
+Many types of transmit signals are feasible - this paper considers only upsweep linear frequency modulated signals (also known as linear chirps).
 
@@
 \includegraphics[width=16cm]{Fig_ek_sys}
-\caption{\label{fi:ek_sys}Signal and data flow in the \ek{} system.
-An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage (lower middle) and display and analysis (lower right).}
+\caption{\label{fi:ek_sys} Signal and data flow in the \ek{} system.
+An echosounder ping starts with the definition of a transmit signal (upper left) and ends with file storage and display and analysis after matched filtering.}
 \end{figure}
@@
 The controlling computer program generates a short-duration digital transmit signal (a ping), $\ytx(\samplesymt)$, where $\samplesymt$ is the sample index in the discrete time domain.
+The nominal pulse duration, $\tnom$, is defined as the duration of the digital transmit signal $\ytx(\samplesymt)$.
 Typical broadband pulses are linear upsweep pulses windowed by an envelope function.
@@
 
-In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively.
-A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
-\begin{figure}
-\includegraphics[width=16cm]{Fig_ytx}
-\caption{\label{fi:ytx} Envelope of linear chirp pulses from 92 kHz to 158 kHz with a slope of 0.057 (blue) and 0.5 (orange).}
-\end{figure}
+The most commonly used transmit signals (e.g.
+in Simrad EK80) are linear frequency modulated signals with an applied Tukey window function (see e.g.
+\citet{lavery2017} for a more in-depth treatment of this subject).
 
@@
 
-To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hfl(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$.
+To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hbp(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$.
+This processing step is approximately equivalent to a traditional demodulation process.
+See e.g.
+Table 1.2 in \citep{Demer2017} for examples of values.
 The individual filter coefficients for each filter and decimation stage are indexed by $\genidxsym$.
@@
 \label{eq:yrx}
-\yrx(\samplesymt,\channelsym,\stagesym) = \left( \yrx(\samplesymt,\channelsym,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
+\yrx(\samplesymt,\channelsym,\stagesym) = \left( \yrx(\samplesymt,\channelsym,\stagesym-1) * \hbp(\genidxsym,\stagesym) \right)_{\downarrow \decfac(\stagesym)}, 
 \stagesym = 1,\ldots,\nstages,
@@
 %
-The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data.
-In the example $N_v=2$.
-The frequency responses of the filters are shown in Figure~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set.
+The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data \citep{ProakisDSP,CrochireMDSP}.
+The frequency responses of the filters are shown in Fig.
+~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set, where $N_v=2$.
 
@@
 \includegraphics[width=16cm]{Fig_fir}
-\caption{\label{fi:fir} The frequency response (filter gain) of the filters in our test set.
+\caption{\label{fi:fir} Example of frequency response (filter gain) of the filters in our test set.
 The blue and orange curve represents the filter response of the first and second filter.
-The horizontal green line indicates the frequency range of the transmit signal.}
+Note that the blue line is above 0 dB, this is caused by the transition from complex to real values.
+The vertical dashed green lines indicates the frequency range of the transmit signal.}
 \end{figure}
@@
 
-\subsection{Pulse compression}
+\subsection{Matched filtering (pulse compression)}
 To increase signal-to-noise ratio and resolution along the acoustic beam a matched filter may be applied to the raw data samples \citep{turin1960}.
-This technique is also known as pulse compression \citep{klauder1960}.
+This technique is also known as pulse compression \citep{klauder1960}, achieving the average transmitted power of a relatively long pulse while obtaining the range resolution of a shorter pulse.
 One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in Eq.
@@
 \label{eq:FilterStagesTX}
-\ytxnorm(\samplesymt,\stagesym) = \left[ \ytxnorm(\samplesymt,\stagesym-1) * \hfl(\genidxsym,\stagesym) \right]_{\downarrow \decfac(\stagesym)}, 
+\ytxnorm(\samplesymt,\stagesym) = \left[ \ytxnorm(\samplesymt,\stagesym-1) * \hbp(\genidxsym,\stagesym) \right]_{\downarrow \decfac(\stagesym)}, 
 \stagesym = 1,\ldots,\nstages,
@@
 where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$.
-The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n}).
+The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n_ACF}a).
 %
 \begin{figure}
-\includegraphics[width=16cm]{Fig_y_mf_n}
-\caption{\label{fi:y_mf_n} The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, that is used for the pulse compression.}
+\includegraphics[width=16cm]{Fig_y_mf_n_ACF}
+\caption{\label{fi:y_mf_n_ACF} An example where the decimated sampling rate ($\fsdec$, Eq.
+\ref{eq:fsdec}) is 125 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, Eq.
+\ref{eq:TauEff}) is 0.01 ms.
+(a) The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, from the final filter and decimation stage, that is used for the pulse compression.
+(b) The autocorrelation function $\ptxauto(n)$.}
 \end{figure}
 
-The auto correlation function of the matched filter signal and the effective pulse duration, defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, will be used in later processing steps and are defined as
+The auto correlation function of the matched filter signal ($\ymfauto(\samplesymt)$) and the effective pulse duration ($\teff$) will be used in later processing steps.
+
+$\ymfauto(\samplesymt)$ is defined as
+
 \begin{equation}
@@
 \end{equation}
-and
+
+where $"*"$ denotes convolution, $"^*"$ complex conjugate, and $||\ymf||_2$ the $l^2$-norm of $\ymf$, also known as the Euclidean norm (see e.g.
+\citet{PadgettSigProc} and \citet{GhatakR} for background on $l^2$-norm).
+
+$\teff$ is defined as
+
 \begin{equation}
@@
 \end{equation}
+
 where
+
 \begin{equation*}
@@
 \end{equation*}
-is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$.
 
-\begin{figure}
-\includegraphics{Fig_ACF}
-\caption{The autocorrelation function $\ptxauto$ from the example. The corresponding $\teff=0.1$ ms and $\tnom=2$ ms\label{fi:ACF}.}
-\end{figure}
+is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over the duration of the autocorrelation function (Fig. \ref{fi:y_mf_n_ACF}b).
 
-To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal with the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power.
+To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power.
 The pulse compressed signal, $\ypc(\samplesymt,\channelsym)$, then becomes
@@
 %
-where $||\ymf||$ indicates the $l^2$-norm of $\ymf$, also known as the Euclidean norm.
 The received power samples are then used to estimate target strength and volume backscattering strength.
 For estimating received power samples, the average signal, $\ypc(\samplesymt)$, over all transducer sectors, $\nchannels$, is used:
-%
+
 \begin{equation}
@@
 \end{equation}
-%
+
 Compensation of echo strength for position in the acoustic beam requires an estimate of the echo arrival angle.
@@
 The principle is demonstrated with a transducer that is divided into four quadrants (Fig.
-\ref{fi:trd_quad}).
+\ref{fi:Fig_trd_quad_Impedances}a).
 In this example the summed signals from four halves (1+2, 2+3, 3+4, 4+1) are calculated as:
@@
 where fore, aft, star(board), and port indicate the relevant transducer halves.
+
 \begin{figure}
-\includegraphics{Fig_trd_quad}
-\caption{\label{fi:trd_quad}Transducer divided into four quadrants. The labels are directions often used when a transducer is mounted on a ship.}
+\includegraphics[width=16cm]{Fig_trd_quad_Impedances}
+\caption{\label{fi:Fig_trd_quad_Impedances}(a) Transducer divided into four quadrants.
+The labels are directions often used when a transducer is mounted on a ship.
+(b) Equivalent circuit diagram of transducer/transceiver with system impedances.}
 \end{figure}
@@
 The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig.
-\ref{fi:impedances}) is given by:
+\ref{fi:Fig_trd_quad_Impedances}b) is given by:
 \begin{equation}
@@
 %
-\begin{figure}
-\includegraphics[width=16cm]{Fig_impedances}
-\caption{\label{fi:impedances}Equivalent circuit diagram of transducer/transceiver with system impedances.}
-\end{figure}
 
@@
 %
-where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$.
+where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted).
 The physical echo arrival angles ($\along$ and $\athw$) are then given by:
@@
 %
-where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:theta_phi}) and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}.
+where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:ts_theta_phi}b) and are derived from the transducer geometry \citep{Urick3rdPrinciples} and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}.
 The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$.
@@
 %
-\begin{figure}
 
-\includegraphics[width=16cm]{Fig_theta_phi}
-\caption{\label{fi:theta_phi}The physical angles $\along$ and $\athw$ for the target strength example data (see below).
-The single target can be seen around sample number 1000 where the angles are less variable.}
-\end{figure}
 \clearpage
@@
 
-To illustrate calculation of target strength ($TS$) as a function of frequency, $TS(f)$, we use data collected on a 35 mm diameter tungsten carbide calibration sphere (WC35) suspended approximately 6 m below a 120 kHz transducer (Fig.~\ref{Fig_TS_echogram}).
+To illustrate calculation of target strength ($TS$, dB re 1m$^2$) as a function of frequency, $TS(f)$, we use data collected on a 35 mm diameter tungsten carbide calibration sphere (WC35) suspended approximately 5.8 m below a 120 kHz transducer (Fig.~\ref{fi:ts_theta_phi}a).
 
 \begin{figure}
-\includegraphics[width=16cm]{Fig_TS_echogram}
-\caption{\label{Fig_TS_echogram} $\mysp$ as a function of ping number and range for a calibration sphere.
-The sphere is located at approximately 6 m range.
-The red vertical line indicates the ping that is used for illustrating the $\ts(\freqsym)$ processing.}
+\includegraphics[width=16cm]{Fig_TS_echogram_theta_phi}
+\caption{\label{fi:ts_theta_phi} (a) $\mysp$ as a function of ping number and range for a calibration sphere.
+The sphere is located at approximately 5.8 m range.
+The red vertical line indicates the ping that is used for illustrating the $\ts(\freqsym)$ processing.
+(b) The physical angles $\along$ and $\athw$ for the target strength example data (read vertical line in a).
+The single target can be seen around range 5.8 m where the angles are less variable.}
 \end{figure}
@@
 
-Generalizing the power-budget equation (i.e., sonar equation) for broadband signals \citep{lunde2016} yields, in logarithmic form, $TS$ at frequency $\freqsym$:
+Generalizing the power-budget equation (i.e. sonar equation) for broadband signals \citep{lunde2016} yields, in logarithmic form, $TS$ at frequency $\freqsym$:
 %
@@
 
-The point scattering strength, $\mysp(\samplesymt)$, is estimated by applying Eq.
+The point scattering strength (TS prior to target detection and beam pattern compensation), $\mysp(\samplesymt)$, is estimated by applying Eq.
 \ref{eq:TS} to the received digitized power samples using the on-axis gain value and $f$ set to the centre frequency of the broadband pulse, $\fc$:
@@
 %
-noting that $\mysp(\samplesymt)$ is an average over frequency of all echoes from single or multiple targets received at sample $\samplesymt$.
+noting that $\mysp(\samplesymt)$ represents an average over the entire frequency band for all echoes received at sample $\samplesymt$.
 
 Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated.
-This is typically achieved through a single echo detection algorithm (SED).
+This is typically achieved through a single echo detection algorithm (SED, see e.g.
+\citet{OnaCRRTS}).
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes that occur in actual targets (as opposed to ideal point targets).
 The alongship angle $\along(\samplesymt)$, athwartship angle $\athw(\samplesymt)$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
-\ref{fi:theta_phi}).
-A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
+\ref{fi:ts_theta_phi}b).
+A simple pseudo SED algorithm, simply using a threshold, is implemented in the code for illustrative purposes.
 
-From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$ (Fig.~\ref{fi:SED}).
+From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak (to that used for the target signal) are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$ (Fig.~\ref{fi:SED}).
 Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
@@
 \includegraphics[width=16cm]{Fig_singleTarget}
-\caption{\label{fi:SED} The $\prxe(\samplesymt)$ power (upper) and split beam angles ($\along_t$ and $\athw_t$) (middle) for the single target.
-The orange vertical line corresponds to the range $r_t$ for the single target.
-The $\ymfautored(\samplesymt)$ (lower) is the autocorrelation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.}
+\caption{\label{fi:SED} $\ypctarget(\samplesymt)$ (upper figure) and $\ymfautored(\samplesymt)$ (lower figure).
+$\ymfautored(\samplesymt)$ is the autocorrelation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.
+The corresponding split beam angles ($\along_t$ and $\athw_t$) for the single target are shown in (Fig.~\ref{fi:ts_theta_phi}b).}
 \end{figure}
@@
 where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
-The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (Fig. \ref{fi:TS}) is then calculated by: 
+The normalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (Fig. \ref{fi:TS}) is then calculated by: 
 %
@@
 %
-noting that any variation of impedance with frequency will be reflected in the $\gainzero$ obtained from the calibration process.
+noting that potential significant variation of impedance with frequency will be accounted for in the $\gainzero$ obtained from the calibration process.
 
@@
 \label{eq:TS_f}
-\ts(\freqsym) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\freqsym)\range_t - 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,f)}{16\pi^2} \right)
+\ts(\freqsym) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\freqsym)\range_t - 10\log_{10}\left( \frac{\ptxe \wlen^2(\freqsym) \gain^2(\along_t,\athw_t,f)}{16\pi^2} \right)
 \end{equation}
@@
 \label{eq:m(f)}
-\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor \mod \ndft
+\samplesymf = \lfloor \frac{\freqsym}{\fsdec} \ndft \rfloor 
 \end{equation}
+where $\lfloor$ $\rfloor$ represents modulus.
 %
@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS}The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$ (a) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$ (b), the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$ (c), transducer gain $(10\log_{10}\gain^2(\along_t,\athw_t,\freqsym))$ (d), and the estimated TS(f) (e).}
+\caption{\label{fi:TS} (a) The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$, (b) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$, (c) the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$, (d) transducer gain $(10\log_{10}\gain^2(\along_t,\athw_t,\freqsym))$, and (e) the estimated TS(f).}
 \end{figure}
@@
 A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
-This is visible in the $\ts$ (Fig.
+This is visible in the estimated $\ts$ (Fig.
 \ref{fi:TS}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
@@
 
-To illustrate calculation of volume backscattering strength as a function of frequency,  $\sv(f)$, we use data collected from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer.
+To illustrate calculation of volume backscattering strength as a function of frequency,  $\sv(f)$ (dB re 1m$^{-1}$), we use data collected from a school of non-swimbladdered fish (Fig.~\ref{fi:Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer.
 
@@
 %
-where $V$ is the volume occupied by the scattering targets. The power-budget equation for multiple targets is then:
+where $V$ is the ensonified volume occupied by the scattering targets. The power-budget equation for multiple targets is then:
 %
@@
 \sv(\freqsym) = 10\log_{10}(\prxevf(\freqsym)) + 20\log_{10}(\range_c) + 2\absorp(\freqsym)\range_c 
-- 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym)}{32\pi^2} \right), 
+- 10\log_{10}\left( \frac{\ptxe \wlen^2(\freqsym) \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym)}{32\pi^2} \right), 
 \end{equation}
@@
 \begin{figure}
-\includegraphics[width=16cm]{Fig_Sv_echogram}
-\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for the raw data file used in the Sv(f) example.
-The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
-The red vertical line indicates the ping that is used for illustrating the $\sv(\freqsym)$ processing.}
+\includegraphics[width=16cm]{Fig_Sv_echogram_m_n}
+\caption{\label{fi:Fig_Sv_echogram} Illustration of volume backscattering strength as a function of frequency, Sv(f).
+A non-swimbladdered fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
+(a) Sv as a function of ping number and range range for the raw data file used in the Sv(f) example.
+The red vertical line indicates the ping that is used for illustrating the Sv(f) processing used in (b).
+(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.}
 \end{figure}
@@
 where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading.
-A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hanning window, $\hannw(\genidxsym)$.
+A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$.
 The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution.
@@
 
-The normalised Hanning window, $\hannwnorm$, is given by: 
+The normalised Hann window, $\hannwnorm$, is given by: 
 %
@@
 \label{eq:Sv_FFT}
-\sv(\freqsym) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\freqsym) \range_c - 10\log_{10}\left( \frac{\ptxe \wlen^2 \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym) }{32\pi^2} \right).
+\sv(\freqsym) = 10\log_{10}(\prxevf(\samplesymf)) + 2\absorp(\freqsym) \range_c - 10\log_{10}\left( \frac{\ptxe \wlen^2(\freqsym) \cw \tslide \eqang(\freqsym) \gainzero^2(\freqsym) }{32\pi^2} \right).
 \end{equation}
@@
 This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
-In our example we have simply chosen the set of centre samples as the original range samples (Fig.~\ref{Fig_Sv_m_n}).
+In our example we have simply chosen the set of centre samples as the original range samples (Fig.~\ref{fi:Fig_Sv_echogram}).
 
-\begin{figure}
-\includegraphics[width=16cm]{Fig_Sv_m_n}
-\caption{\label{Fig_Sv_m_n} $\sv$ as a function of frequency ($\samplesymf$) and range ($\samplesymt$) for a single ping.}
-\end{figure}
-
-For acoustic abundance estimation and classification purposes it is common to integrate $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{Fig_Sv_echogram}).
+For acoustic abundance estimation and classification purposes it is common to integrate $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{fi:Fig_Sv_echogram}).
 It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (Fig.~\ref{Fig_Sv_avg}).
@@
 \includegraphics[width=16cm]{Fig_Sv_avg}
-\caption{\label{Fig_Sv_avg} $\sv$ as a function of frequency averaged over a depth interval covering a fish school.}
+\caption{\label{Fig_Sv_avg} $\sv$ of the single ping marked in Fig.~\ref{fi:Fig_Sv_echogram}a as a function of frequency averaged over a depth interval covering a fish school.
+A weak positive slope is observed of the frequency response.}
 \end{figure}
 
-The trend for increasing $\sv$ with frequency is well-known for non-swimbladderred fish \citep{korneliussen2010} and is consistent with the trend observed in this example.
-In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_echogram}).
+The trend for increasing $\sv$ with frequency is well-known for fish without swimbladder \citep{korneliussen2010} and is consistent with the trend observed in this example.
+In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_avg}).
 
@@
 
-The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps.
-The inclusion of corresponding computer code ensures a more complete description.
-One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code can serve as a starting point for this.
-We also envision that the paper and code be used both for educational purposes and for understanding the standard broadband signal processing.
-
-The contribution includes all the minor details required for the processing steps to work in practice.
-This includes steps well founded in the literature as well as practical and more ad-hoc choices.
-Handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering.
-When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases and in our implementation we chose to exclude the edge cases.
+The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of the data processing steps.
+The contribution includes all steps well founded in the literature as well as any practical and more ad-hoc choices.
+The choices include handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering.
+When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases (e.g.
+distorted output at the beginning and end of the section of filtered data) and in our implementation we chose to exclude these.
 We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
 When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
-Our objective is not to provide an evaluation of all these choices, but to serve as a startin point.
+Our objective is not to provide an evaluation of all these choices, but to document the baseline for the processing that can be used for benchmarking purposes.
 
@@
 This requires a robust single target detector.
-Several SED algorithms exist and differing such algorithms may be required depending on the situation.
+Several SED algorithms exist and different algorithms may be required depending on the situation.
 Typical SED algorithms are based on traditional single frequency pulses and by utilizing the additional information in broadband echoes improved SEDs may be envisioned, but this is outside the scope of this paper.
@@
 To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range.
-The chosen size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution.
+The chosen size of the window is two times the pulse length, and is a compromise between spatial and frequency resolution.
 Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform.
-Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies between the beginning and the end of the window for relatively short range windows.
+Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for the operating frequencies for typical marine ecosystem echosounder for short range windows.
 The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
@@
 The formulation presented in this paper requires several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$.
-Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
+Methods to estimate these are not within the scope of this paper, but common practice is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
 We note that these methods do not provide an operational method to estimate $\teff$ or $\eqang(\freqsym)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
@@
 
-\section{Data Availability Statement}
+\section{Data Availability Statement} \label{dataavstatement}
 
 The code and data associated with this article are available through GitHub through \url{https://github.com/CRIMAC-WP4-Machine-learning/CRIMAC-Raw-To-Svf-TSf}.
-The code and data for the pre-print is tagged version 0.9.
+The raw data files are available at \url{https://zenodo.org/record/8318274}.
+The code and data for the pre-print is tagged version 0.95.
 The version for the printed paper is 1.0.

### 2023-09-14 Paper/SvfEchogram.py
Added scripts to plot Sv and Sp echograms

@@ -1,0 +1,91 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Nov 12 16:11:12 2020
+Script which reads raw EK80 data and makes the Sv echogram for the JEE paper
+The script requires pyEcholab (RHT-EK80 branch)
+The raw files used in the paper are avilable from
+https://zenodo.org/record/8318274
+@author: Geir Pedersen, IMR
+"""
+
+import matplotlib.pyplot as plt
+from matplotlib.pyplot import figure, show, subplots_adjust, get_cmap
+import sys
+sys.path.append(__file__ + '/../../../../../include')
+import numpy as np
+import sys
+from echolab2.instruments import EK80
+from echolab2.plotting.matplotlib import echogram
+import glob
+
+
+from datetime import datetime
+import pandas as pd
+from plotly.offline import plot
+import plotly.graph_objs as go
+#import plotly.express as px
+import plotly.io as pio
+#from time import time
+import numpy as np
+#import matplotlib.pyplot as plt
+#import matplotlib.dates as mdates
+from datetime import datetime, timedelta, timezone
+from plotly.subplots import make_subplots
+pio.renderers.default = 'png'
+
+# Raw file to read, change this to reflect raw file and where it is stored
+file = 'D:/CRIMAC/CRIMAC_WP1/Svf_new/IMR-D20210507-T074652-Svf.raw'
+
+ek80 = EK80.EK80()
+
+ek80.read_raw(file)
+ek80.__dict__
+
+# # read raw data from second channel (vertical echosounder)
+raw_list = ek80.raw_data[ek80.channel_ids[4]]
+raw_data = raw_list[0]
+
+# del ek80
+
+# get calibration values
+calibration = raw_data.get_calibration()
+
+#  convert to Sv
+cal_obj = raw_data.get_calibration()
+Sv = raw_data.get_Sv(calibation=cal_obj)
+
+# plot a ping
+fig1 = figure()
+sv=plt.plot(Sv[0,:])
+show()
+
+# go from pyEcholab to array
+dfSv=Sv.data
+dfRange=Sv.get_v_axis()
+dfPingTime=Sv.ping_time
+
+# disregard what's after the surface
+#dfSv=dfSv[:,0:1400]
[…]

### 2023-09-14 Paper/TSfEchogram.py

@@ -1,0 +1,92 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Nov 12 16:11:12 2020
+Script which reads raw EK80 data and makes the Sp echogram for the JEE paper
+The script requires pyEcholab (RHT-EK80 branch)
+The raw files used in the paper are avilable from
+https://zenodo.org/record/8318274
+@author: Geir Pedersen, IMR
+"""
+
+import matplotlib.pyplot as plt
+from matplotlib.pyplot import figure, show, subplots_adjust, get_cmap
+import sys
+sys.path.append(__file__ + '/../../../../../include')
+import numpy as np
+import sys
+from echolab2.instruments import EK80
+from echolab2.plotting.matplotlib import echogram
+import glob
+
+
+from datetime import datetime
+import pandas as pd
+from plotly.offline import plot
+import plotly.graph_objs as go
+#import plotly.express as px
+import plotly.io as pio
+#from time import time
+import numpy as np
+#import matplotlib.pyplot as plt
+#import matplotlib.dates as mdates
+from datetime import datetime, timedelta, timezone
+from plotly.subplots import make_subplots
+pio.renderers.default = 'png'
+
+# Raw file to read
+file = 'D:/CRIMAC/CRIMAC_WP1/SphereBeam_New/IMR-D20211215-T143432-TSf.raw'
+
+ek80 = EK80.EK80()
+
+ek80.read_raw(file)
+ek80.__dict__
+
+# # read raw data from second channel (vertical echosounder)
+raw_list = ek80.raw_data[ek80.channel_ids[1]]
+raw_data = raw_list[0]
+
+# get calibration values
+calibration = raw_data.get_calibration()
+
+#  convert to Sv
+cal_obj = raw_data.get_calibration()
+#Sv = raw_data.get_Sv(calibation=cal_obj)
+Sv = raw_data.get_Sp(calibation=cal_obj)
+
+# plot a ping
+fig1 = figure()
+sv=plt.plot(Sv[0,:])
+show()
+
+# go from pyEcholab to array
+dfSv=Sv.data
+dfRange=Sv.get_v_axis()
+dfPingTime=Sv.ping_time
+
+# disregard what's after the surface
[…]

### 2023-09-14 Svf.py
Edited TSf, Svf, plots.py to reflect paper

@@ -79,8 +79,8 @@ psi_m = Calculation.calcpsi(psi_f_n, f_n, f_m)
 y_tx_n, t = Calculation.generateIdealWindowedTransmitSignal(f_0, f_1, tau, f_s, slope)
 
 # Plots for paper
-if do_plot:
-    plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
+# if do_plot:
+#     plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
 
 #
 # Chapter IIC: Signal reception
@@ -154,7 +154,7 @@ y_pc_halves_n = Calculation.calcTransducerHalves(y_pc_nu)
 #
 
 # Total received power for all channels (all transducer sectors)
-p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e)
+p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e,N_u)
 
 # Physical angles
 theta_n, phi_n = Calculation.calcAngles(y_pc_halves_n, gamma_theta_f_c, gamma_phi_f_c)

### 2023-09-14 TSf.py

@@ -79,8 +79,8 @@ psi_m = Calculation.calcpsi(psi_f_n, f_n, f_m)
 y_tx_n, t = Calculation.generateIdealWindowedTransmitSignal(f_0, f_1, tau, f_s, slope)
 
 # Plots for paper
-if do_plot:
-    plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
+# if do_plot:
+#     plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
 
 #
 # Chapter IIC: Signal reception
@@ -154,14 +154,14 @@ y_pc_halves_n = Calculation.calcTransducerHalves(y_pc_nu)
 #
 
 # Total received power for all channels (all transducer sectors)
-p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e)
+p_rx_e_n = Calculation.calcPower(y_pc_n, z_td_e, z_rx_e, N_u)
 
 # Physical angles
 theta_n, phi_n = Calculation.calcAngles(y_pc_halves_n, gamma_theta_f_c, gamma_phi_f_c)
 
 # Plots for paper
 if do_plot:
-    plotThetaPhi(theta_n, phi_n)
+    plotThetaPhi(theta_n, phi_n, dr)
 
 #
 # Chapter III: TARGET STRENGTH
@@ -195,7 +195,7 @@ y_mf_auto_red_n = Calculation.alignAuto(y_mf_auto_n, y_pc_t_n)
 
 # Plots for paper
 if do_plot:
-    plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red_n)
+    plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red_n,y_pc_t_n)
 
 # DFT of target signal, DFT of reduced auto correlation signal, and
 # normalized DFT of target signal

### 2023-09-14 plots.py

@@ -4,29 +4,29 @@ from Core.Calculation import Calculation
 from scipy.signal import hilbert
 
 
-def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
-    """
-    XXX.
+# def plotytx(f_0, f_1, tau, f_s, y_tx_n, slope):
+#     """
+#     XXX.
     
-    Parameters
-    ----------
+#     Parameters
+#     ----------
     
-    """
-    # Example of ideal windowed transmit signal with slope 0.5
-    y_tx_n05slope, t = Calculation.generateIdealWindowedTransmitSignal(
-        f_0, f_1, tau, f_s, 0.5
-    )
+#     """
+#     # Example of ideal windowed transmit signal with slope 0.5
+#     y_tx_n05slope, t = Calculation.generateIdealWindowedTransmitSignal(
+#         f_0, f_1, tau, f_s, 0.5
+#     )
 
-    plt.figure()
-    plt.plot(
-        t * 1000, np.abs(hilbert(y_tx_n)), t * 1000, np.abs(hilbert(y_tx_n05slope))
-    )
-    # plt.title(
-    #    'Ideal windowed transmit pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
-    #        .format(f_0 / 1000, f_1 / 1000, slope))
-    plt.xlabel("Time (ms)")
-    plt.ylabel("Envelope ()")
-    plt.savefig("./Paper/Fig_ytx.png", dpi=300)
+#     plt.figure()
+#     plt.plot(
+#         t * 1000, np.abs(hilbert(y_tx_n)), t * 1000, np.abs(hilbert(y_tx_n05slope))
+#     )
+#     # plt.title(
+#     #    'Ideal windowed transmit pulse.{:.0f}kHz - {:.0f}kHz, slope {:.3f}'
+#     #        .format(f_0 / 1000, f_1 / 1000, slope))
+#     plt.xlabel("Time (ms)")
+#     plt.ylabel("Envelope ()")
+#     plt.savefig("./Paper/Fig_ytx.png", dpi=300)
 
 
 def plotfir(filter_v, f_s_dec_v, f_0, f_1):
@@ -58,11 +58,17 @@ def plotfir(filter_v, f_s_dec_v, f_0, f_1):
     G1l = np.append(G1l, G1)
[…]

### 2023-09-14 main.py
Small fixes

@@ -111,8 +111,8 @@ def calcBasics(do_plot):
     )
 
     # Plots for paper
-    if do_plot:
-        plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
+    # if do_plot:
+    #     plotytx(f_0, f_1, tau, f_s, y_tx_n, slope)
 
     #
     # Chapter IIC: Signal reception
@@ -195,7 +195,7 @@ def calcBasics(do_plot):
 
     # Plots for paper
     if do_plot:
-        plotThetaPhi(theta_n, phi_n)
+        plotThetaPhi(theta_n, phi_n, dr)
 
 
 def calcTS():
@@ -244,7 +244,7 @@ def calcTS():
     y_mf_auto_red_n = Calculation.alignAuto(y_mf_auto_n, y_pc_t_n)
 
     # Plots for paper
-    plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red_n)
+    plotSingleTarget(dum_r, dum_p, dum_theta, r_t, dum_phi, phi_t, y_mf_auto_red_n, y_pc_t_n)
 
     # DFT of target signal, DFT of reduced auto correlation signal, and
     # normalized DFT of target signal

### 2023-09-14 Paper/Manuscript.tex
Small fixes
Misc fixes

@@
 The code  accompanying this paper is written in the Python programming language (v3.10) and is available through GitHub (see \ref{dataavstatement} for code and data availability) together with the data used in the examples.
-All single ping processing steps with respective figures in the paper can be reproduced by running the main scripts (\code{TSf.py} and \code{Svf.py}).
+All single ping processing steps with respective figures in the paper can be reproduced by running the main script, \code{main.py}.
 Reproduction of Fig.
@@
 
-where $"*"$ denotes convolution, $"^*"$ complex conjugate, and $||\ymf||_2$ the $l^2$-norm of $\ymf$, also known as the Euclidean norm (see e.g.
-\citet{PadgettSigProc} and \citet{GhatakR} for background on $l^2$-norm).
+where $"*"$ denotes convolution, $"^*"$ complex conjugate, and $||\ymf||_2$ the $l^2$-norm of $\ymf$, also known as the Euclidean norm.
+See e.g.
+\citet{PadgettSigProc} and \citet{GhatakR} for background on $l^2$-norm.
 

### 2023-09-18 Paper/Supplementary.tex
Adding variable names

@@
 $\atan(y,x)$ & -- & -- & The four quadrant inverse tangent which returns values in the interval $[-\pi,\pi]$ , inclusive.\\
-$\cw$ & & $m s^{-1}$ & Sound speed in water.\\
-$\decfac(\stagesym)$ & & 1 & Integer filter decimation factor.\\
-$\dft_\ndft(x)$ & & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
+$\cw$ & c & $m s^{-1}$ & Sound speed in water.\\
+$\decfac(\stagesym)$ & filter\_v[n]['D'] & 1 & Integer filter decimation factor.\\
+$\dft_\ndft(x)$ & N\_DFT & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
 $\freqsym$ & & Hz & Frequency.\\
 $\fc$ & & Hz & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$.\\
-$\fn$ & & Hz & Nominal operating frequency of a transducer.\\
-$\fs$ & & Hz & Analogue to digital sampling frequency.\\
-$\fsdec$ & & Hz & Decimated sampling frequency.\\
-$\fstart$ & & Hz & Start frequency of linear upsweep chirp pulse.\\
-$\fstop$ & & Hz & Stop frequency of linear upsweep chirp pulse.\\
+$\fn$ & f\_n & Hz & Nominal operating frequency of a transducer.\\
+$\fs$ & f\_s & Hz & Analogue to digital sampling frequency.\\
+$\fsdec$ & f\_s\_dec & Hz & Decimated sampling frequency.\\
+$\fstart$ & f\_1 & Hz & Start frequency of linear upsweep chirp pulse.\\
+$\fstop$ & f\_2 & Hz & Stop frequency of linear upsweep chirp pulse.\\
 
-$\gain(\along,\athw,\freqsym)$ & & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency.\\
+$\gain(\along,\athw,\freqsym)$ & g\_theta\_phi\_m & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency.\\
 
@@
 
-$\gainzero(\freqsym)$ & & 1 & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$.\\
+$\gainzero(\freqsym)$ & g\_0\_m & 1 & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$.\\
 
-$\hbp(\genidxsym,\stagesym)$ & & 1 & Complex valued receiving filter coefficients.\\
+$\hbp(\genidxsym,\stagesym)$ & filter\_v[n]['h\_fl\_i'] & 1 & Complex valued receiving filter coefficients.\\
 $\genidxsym$ & & 1 & Generic integer index.\\
@@
 $j$ & & & The square root of $-1$.\\
-$\samplesymf$ & & 1 & Sample index in frequency domain.\\
+$\samplesymf$ & m & 1 & Sample index in frequency domain.\\
 
-$\nchannels$ & & 1 & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal.\\
+$\nchannels$ & N\_u & 1 & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal.\\
 $\nstages$ & & 1 & Total number of filter stages.\\
 $\nw$ & & 1 & Number of samples used in the sliding Hanning window.\\
-$\samplesymt$ & & 1 & Sample index in time domain.\\
+$\samplesymt$ & n & 1 & Sample index in time domain.\\
 
 $\prxetf(\freqsym)$ &  & W & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$.\\
-$\prxetf(\samplesymf)$ & & W & DFT of the received electric power in a matched load for the signal from a single target.\\
+$\prxetf(\samplesymf)$ & P\_rx\_e\_t\_m & W & DFT of the received electric power in a matched load for the signal from a single target.\\
 $\prxevf(\freqsym)$ & & W & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$.\\
-$\prxevf(\samplesymf)$ & & W$\textrm{m}^2$ & DFT of the received electric power in a matched load for the signal from a volume.\\
-$\prxe(\samplesymt)$  & & W & Received electric power in a matched load.\\
+$\prxevf(\samplesymf)$ & P\_rx\_e\_v\_m & W$\textrm{m}^2$ & DFT of the received electric power in a matched load for the signal from a volume.\\
+$\prxe(\samplesymt)$  & p\_rx\_e\_n & W & Received electric power in a matched load.\\
 
-$\ptxe$ & & W & Transmitted electric power.\\
-$\ptxauto(\samplesymt)$ & & 1 & Square of the absolute value of the matched filter autocorrelation function.\\
+$\ptxe$ & p\_tx\_e & W & Transmitted electric power.\\
+$\ptxauto(\samplesymt)$ & p\_tx\_auto & 1 & Square of the absolute value of the matched filter autocorrelation function.\\
 
@@
 $\range_c(\samplesymt)$ & & m & $range_c$ at sample number $\samplesymt$.\\
-$\range(\samplesymt)$ & & m & Distance from transducer.\\
+$\range(\samplesymt)$ & r\_n & m & Distance from transducer.\\
 
-$\mysp(\samplesymt)$  & & dB re 1 $\textrm{m}^2$ & Point scattering strength.\\
+$\mysp(\samplesymt)$  & Sp\_n & dB re 1 $\textrm{m}^2$ & Point scattering strength.\\
 $\sv$  & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength.\\
 $\sv(\freqsym)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at frequency $\freqsym$.\\
-$\sv(\samplesymt)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
+$\sv(\samplesymt)$ & Sv\_n & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
 
@@
 $\timesym$ & & s &  Time.\\
-$\tslide$ & & s & Duration of sliding window for calculating volume backscattering strength.\\
+$\tslide$ & t\_w & s & Duration of sliding window for calculating volume backscattering strength.\\
 

### 2023-09-18 Paper/Supplementary.tex
Adding variable names

@@
 $\cw$ & c & $m s^{-1}$ & Sound speed in water.\\
-$\decfac(\stagesym)$ & filter\_v[n]['D'] & 1 & Integer filter decimation factor.\\
-$\dft_\ndft(x)$ & N\_DFT & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
+$\decfac(\stagesym)$ & \code{filter_v[n]['D']} & 1 & Integer filter decimation factor.\\
+$\dft_\ndft(x)$ & & -- & The discrete Fourier transform of length $\ndft$ of $x$.\\
 $\freqsym$ & & Hz & Frequency.\\
 $\fc$ & & Hz & Frequency at centre of linear upsweep chirp pulse. Equivalent to $\fstart + \frac{|\fstop - \fstart|}{2}$.\\
-$\fn$ & f\_n & Hz & Nominal operating frequency of a transducer.\\
-$\fs$ & f\_s & Hz & Analogue to digital sampling frequency.\\
-$\fsdec$ & f\_s\_dec & Hz & Decimated sampling frequency.\\
-$\fstart$ & f\_1 & Hz & Start frequency of linear upsweep chirp pulse.\\
-$\fstop$ & f\_2 & Hz & Stop frequency of linear upsweep chirp pulse.\\
+$\fn$ & \code{f_n} & Hz & Nominal operating frequency of a transducer.\\
+$\fs$ & \code{f_s} & Hz & Analogue to digital sampling frequency.\\
+$\fsdec$ & \code{f_s_dec} & Hz & Decimated sampling frequency.\\
+$\fstart$ & \code{f_1} & Hz & Start frequency of linear upsweep chirp pulse.\\
+$\fstop$ & \code{f_2} & Hz & Stop frequency of linear upsweep chirp pulse.\\
 
-$\gain(\along,\athw,\freqsym)$ & g\_theta\_phi\_m & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency.\\
+$\gain(\along,\athw,\freqsym)$ & \code{g_theta_phi_m} & 1 & Transducer gain as a function of echo arrival angle in the beam and frequency.\\
 
@@
 
-$\gainzero(\freqsym)$ & g\_0\_m & 1 & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$.\\
+$\gainzero(\freqsym)$ & \code{g_0_m} & 1 & Transducer gain along the main acoustic axis, i.e. $\gain(\along, \athw,\freqsym)$ where $\along = \athw = 0$.\\
 
-$\hbp(\genidxsym,\stagesym)$ & filter\_v[n]['h\_fl\_i'] & 1 & Complex valued receiving filter coefficients.\\
+$\hbp(\genidxsym,\stagesym)$ & \code{filter_v[n]['h_fl_i']} & 1 & Complex valued receiving filter coefficients.\\
 $\genidxsym$ & & 1 & Generic integer index.\\
@@
 $j$ & & & The square root of $-1$.\\
-$\samplesymf$ & m & 1 & Sample index in frequency domain.\\
+$\samplesymf$ & \code{m} & 1 & Sample index in frequency domain.\\
 
-$\nchannels$ & N\_u & 1 & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal.\\
+$\nchannels$ & \code{N_u} & 1 & Total number of transducer sectors/transceiver channels that are used to receive and process the acoustic signal.\\
 $\nstages$ & & 1 & Total number of filter stages.\\
@@
 $\prxetf(\freqsym)$ &  & W & FT of the received electric power in a matched load for the signal from a single target at frequency $\freqsym$.\\
-$\prxetf(\samplesymf)$ & P\_rx\_e\_t\_m & W & DFT of the received electric power in a matched load for the signal from a single target.\\
+$\prxetf(\samplesymf)$ & \code{P_rx_e_t_m} & W & DFT of the received electric power in a matched load for the signal from a single target.\\
 $\prxevf(\freqsym)$ & & W & FT of the received electric power in a matched load for the signal from a volume at frequency $\freqsym$.\\
-$\prxevf(\samplesymf)$ & P\_rx\_e\_v\_m & W$\textrm{m}^2$ & DFT of the received electric power in a matched load for the signal from a volume.\\
-$\prxe(\samplesymt)$  & p\_rx\_e\_n & W & Received electric power in a matched load.\\
+$\prxevf(\samplesymf)$ & \code{P_rx_e_v_m} & W$\textrm{m}^2$ & DFT of the received electric power in a matched load for the signal from a volume.\\
+$\prxe(\samplesymt)$  & \code{p_rx_e_n} & W & Received electric power in a matched load.\\
 
-$\ptxe$ & p\_tx\_e & W & Transmitted electric power.\\
-$\ptxauto(\samplesymt)$ & p\_tx\_auto & 1 & Square of the absolute value of the matched filter autocorrelation function.\\
+$\ptxe$ & \code{p_tx_e} & W & Transmitted electric power.\\
+$\ptxauto(\samplesymt)$ & \code{p_tx_auto} & 1 & Square of the absolute value of the matched filter autocorrelation function.\\
 
@@
 $\range_c(\samplesymt)$ & & m & $range_c$ at sample number $\samplesymt$.\\
-$\range(\samplesymt)$ & r\_n & m & Distance from transducer.\\
+$\range(\samplesymt)$ & \code{r_n} & m & Distance from transducer.\\
 
@@
 $\hannw$ & & 1 & The Hann window function.\\
-$\hannw(\genidxsym)$ & & 1 & The Hann window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
-$\hannwnorm(\genidxsym)$ & & 1 & Normalised Hann window.\\
+$\hannw(\genidxsym)$ & w\_i & 1 & The Hann window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
+$\hannwnorm(\genidxsym)$ & w\_tilde\_i & 1 & Normalised Hann window.\\
 
@@
 
-$\ymfautof(\samplesymf)$ & & 1 & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter.\\
-$\ymfautoredf(\samplesymf)$ &  & 1 & DFT of the reduced autocorrelation function for the matched filter.\\
-$\ypctargetf(\samplesymf)$ & & V & DFT of the pulse compressed signal from a single target.\\
-$\ypctargetnormf(\samplesymf)$ & & Vm & DFT of the pulse compressed signal from a single target 
+$\ymfautof(\samplesymf)$ & Y\_mf\_auto\_m & 1 & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter.\\
+$\ymfautoredf(\samplesymf)$ & Y\_mf\_auto\_m\_red & 1 & DFT of the reduced autocorrelation function for the matched filter.\\
+$\ypctargetf(\samplesymf)$ & Y\_pc\_t\_m & V & DFT of the pulse compressed signal from a single target.\\
+$\ypctargetnormf(\samplesymf)$ & Y\_tilde\_pc\_t\_m & Vm & DFT of the pulse compressed signal from a single target 
 normalized by the DFT of the reduced autocorrelation function for the matched filter.\\
-$\ypcvolumef(\samplesymf)$ & & Vm & DFT of the pulse compressed signal from a volume. Compensated for spreading loss.\\
-$\ypcvolumenormf(\samplesymf)$ & & V & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
+$\ypcvolumef(\samplesymf)$ & Y\_pc\_v\_m & Vm & DFT of the pulse compressed signal from a volume. Compensated for spreading loss.\\
+$\ypcvolumenormf(\samplesymf)$ & \code{Y\_tilde\_pc\_v\_m} & V & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
 Compensated for spreading loss.\\

### 2023-09-18 Paper/Supplementary.tex
Adding variable names

@@
 
-$\mysp(\samplesymt)$  & Sp\_n & dB re 1 $\textrm{m}^2$ & Point scattering strength.\\
+$\mysp(\samplesymt)$  & \code{Sp_n} & dB re 1 $\textrm{m}^2$ & Point scattering strength.\\
 $\sv$  & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength.\\
 $\sv(\freqsym)$ & & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at frequency $\freqsym$.\\
-$\sv(\samplesymt)$ & Sv\_n & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
+$\sv(\samplesymt)$ & \code{Sv_n} & dB re 1 $\textrm{m}^{-1}$ & Volume backscattering strength at sample index $\samplesymt$.\\
 
@@
 $\hannw$ & & 1 & The Hann window function.\\
-$\hannw(\genidxsym)$ & w\_i & 1 & The Hann window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
-$\hannwnorm(\genidxsym)$ & w\_tilde\_i & 1 & Normalised Hann window.\\
+$\hannw(\genidxsym)$ & \code{w_i} & 1 & The Hann window function for index i, defined by $w(i) = 0.5(1+\cos (2\pi i /N_w)), -N_w/2 \leq i \leq N_w/2$.\\
+$\hannwnorm(\genidxsym)$ & \code{w_tilde_i} & 1 & Normalised Hann window.\\
 
@@
 
-$\ymfautof(\samplesymf)$ & Y\_mf\_auto\_m & 1 & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter.\\
-$\ymfautoredf(\samplesymf)$ & Y\_mf\_auto\_m\_red & 1 & DFT of the reduced autocorrelation function for the matched filter.\\
-$\ypctargetf(\samplesymf)$ & Y\_pc\_t\_m & V & DFT of the pulse compressed signal from a single target.\\
-$\ypctargetnormf(\samplesymf)$ & Y\_tilde\_pc\_t\_m & Vm & DFT of the pulse compressed signal from a single target 
+$\ymfautof(\samplesymf)$ & \code{Y_mf_auto_m} & 1 & Discrete Fourier transform (DFT) of the autocorrelation function for the matched filter.\\
+$\ymfautoredf(\samplesymf)$ & \code{Y_mf_auto_m_red} & 1 & DFT of the reduced autocorrelation function for the matched filter.\\
+$\ypctargetf(\samplesymf)$ & \code{Y_pc_t_m} & V & DFT of the pulse compressed signal from a single target.\\
+$\ypctargetnormf(\samplesymf)$ & \code{Y_tilde_pc_t_m} & Vm & DFT of the pulse compressed signal from a single target 
 normalized by the DFT of the reduced autocorrelation function for the matched filter.\\
-$\ypcvolumef(\samplesymf)$ & Y\_pc\_v\_m & Vm & DFT of the pulse compressed signal from a volume. Compensated for spreading loss.\\
-$\ypcvolumenormf(\samplesymf)$ & \code{Y\_tilde\_pc\_v\_m} & V & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
+$\ypcvolumef(\samplesymf)$ & \code{Y_pc_v_m} & Vm & DFT of the pulse compressed signal from a volume. Compensated for spreading loss.\\
+$\ypcvolumenormf(\samplesymf)$ & \code{Y_tilde_pc_v_m} & V & DFT of the pulse compressed signal from a single volume normalized by the DFT of the reduced autocorrelation function for the matched filter.
 Compensated for spreading loss.\\
 
-$\ymf(\samplesymt)$ & & 1 & Matched filter. Signal used for pulse compression.\\
-$\ymfauto(\samplesymt)$ & & 1 & Autocorrelation function for the matched filter.\\
+$\ymf(\samplesymt)$ & \code{y_mf_n} & 1 & Matched filter. Signal used for pulse compression.\\
+$\ymfauto(\samplesymt)$ & \code{y_mf_auto_n} & 1 & Autocorrelation function for the matched filter.\\
 $\ymfautored(\samplesymt)$ & \code{y_mf_auto_red_n} & 1 & Reduced autocorrelation function for the matched filter.\\
 
-$\ypc(\samplesymt)$ & & V & Pulse compressed signal averaged over all transducer sectors.\\
-$\ypc(\samplesymt, \channelsym) $ & & V & Pulse compressed signal from channel $\channelsym$.\\
-$y_{\textrm{pc,aft}}(\samplesymt)$ & & V & Pulse compressed signal from the aft transducer half.\\
-$y_{\textrm{pc,fore}}(\samplesymt)$ & & V & Pulse compressed signal from the forward transducer half.\\
-$y_{\textrm{pc,port}}(\samplesymt)$ & & V & Pulse compressed signal from the port transducer half.\\
-$\ypcspread(\samplesymt)$ & & Vm & Pulse compressed signal compensated for spherical spreading.\\
-$y_{\textrm{pc,star}}(\samplesymt)$ & & V & Pulse compressed signal from the starboard transducer half.\\
-$\ypctarget(\samplesymt)$ & & V & Pulse compressed signal from a single target.\\
+$\ypc(\samplesymt)$ & \code{y_pc_n} & V & Pulse compressed signal averaged over all transducer sectors.\\
+$\ypc(\samplesymt, \channelsym) $ & \code{y_pc_nu} & V & Pulse compressed signal from channel $\channelsym$.\\
+$y_{\textrm{pc,aft}}(\samplesymt)$ & \code{y_pc_aft_n} & V & Pulse compressed signal from the aft transducer half.\\
+$y_{\textrm{pc,fore}}(\samplesymt)$ & \code{y_pc_fore_n} & V & Pulse compressed signal from the forward transducer half.\\
+$y_{\textrm{pc,port}}(\samplesymt)$ & \code{y_pc_port_n} & V & Pulse compressed signal from the port transducer half.\\
+$\ypcspread(\samplesymt)$ & \code{y_pc_s_n} & Vm & Pulse compressed signal compensated for spherical spreading.\\
+$y_{\textrm{pc,star}}(\samplesymt)$ & \code{y_pc_star_n} & V & Pulse compressed signal from the starboard transducer half.\\
+$\ypctarget(\samplesymt)$ & \code{y_pc_t_n} & V & Pulse compressed signal from a single target.\\
 
-$\yrx(\samplesymt,\channelsym)$ & & V & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$.\\
+$\yrx(\samplesymt,\channelsym)$ & \code{y_rx_nu} & V & Received digitised, bandpass filtered, decimated complex signal after the final filter stage, $\yrx(\samplesymt,\channelsym) = \yrx(\samplesymt,\channelsym,\nstages)$.\\
 $\yrx(\samplesymt,\channelsym,\stagesym)$ & & V & Received digitised, bandpass filtered, decimated complex signal.\\
@@
 
-$\ytx(\samplesymt)$ & & V & Ideal transmitted signal generated from transmit signal properties.\\
-$\ytxnorm(\samplesymt)$ & & 1 & Ideal normalized transmitted signal generated from transmit signal properties.\\
-$\ytxnorm(\samplesymt,\stagesym)$ & & 1 & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$.\\
+$\ytx(\samplesymt)$ & \code{y_tx_n} & V & Ideal transmitted signal generated from transmit signal properties.\\
+$\ytxnorm(\samplesymt)$ & \code{y_tilde_tx_n} & 1 & Ideal normalized transmitted signal generated from transmit signal properties.\\
+$\ytxnorm(\samplesymt,\stagesym)$ & \code{y_tilde_tx_nv} & 1 & Ideal normalized transmitted signal generated from transmit signal properties after application of filter stage $\stagesym$.\\
 
@@
 
-$y_\along(\samplesymt)$ & & rad & Electrical angle along the minor axis of the transducer (alongship when ship-mounted).\\
-$y_\athw(\samplesymt)$ & & rad & Electrical angle along the major axis of the transducer (athwartship when ship-mounted).\\
+$y_\along(\samplesymt)$ & \code{y_theta_n} & rad & Electrical angle along the minor axis of the transducer (alongship when ship-mounted).\\
+$y_\athw(\samplesymt)$ & \code{y_phi_n} & rad & Electrical angle along the major axis of the transducer (athwartship when ship-mounted).\\
 
-$\zrxe$ & & $\Omega$ & Receiver electric impedance.\\
-$\ztde$ & & $\Omega$ & Transducer sector electric impedance.\\
+$\zrxe$ & \code{z_rx_e} & $\Omega$ & Receiver electric impedance.\\
+$\ztde$ & \code{z_td_e} & $\Omega$ & Transducer sector electric impedance.\\
 
-$\absorp(\freqsym)$  & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency $f$.\\
+$\absorp(\freqsym)$  & \code{alpha_m} & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency $f$.\\
 % $\absorp(\samplesymf)$ & & dB $\textrm{m}^{-1}$ & Absorption coefficient at frequency index $m$\\
-$\anglefalong$ & & 1 & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo.\\
-$\anglefathw$ & & 1 & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo.\\
-$\along$ & & rad & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
+$\anglefalong$ & \code{gamma_theta} & 1 & Conversion factor between phase difference in signals from the fore and aft transducer halves and the physical arrival angle of an echo.\\
+$\anglefathw$ & \code{gamma_phi} & 1 & Conversion factor between phase difference in signals from the port and starboard transducer halves and the physical arrival angle of an echo.\\
+$\along$ & \code{theta_n} & rad & Angle coordinates along the minor axis of the transducer (alongship when ship-mounted).\\
 % $\along(\samplesymt)$ & & rad & Angle of arrival of target echo along the minor axis of the transducer (alongship when ship-mounted).\\
 $\wlen$ & & m &  Acoustic wavelength.\\
-$\wlen_\samplesymf$ & & m & Acoustic wavelength at frequency index $m$\\
+$\wlen_\samplesymf$ & \code{lambda_m} & m & Acoustic wavelength at frequency index $m$\\
 $\sigmabs$ & & $\textrm{m}^2$ & Backscattering cross-section.\\
-$\tnom$ & & s & Nominal transmit pulse duration.\\
-$\teff$ & & s & Effective transmit pulse duration.\\
-$\athw$ & & rad & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
+$\tnom$ & \code{tau} & s & Nominal transmit pulse duration.\\
+$\teff$ & \code{tau_eff} & s & Effective transmit pulse duration.\\
+$\athw$ & \code{phi} & rad & Angle coordinates along the major axis of the transducer (athwartship when ship-mounted).\\
 % $\athw(\samplesymt)$ & & rad & Angle of arrival of target echo along the major axis of the transducer (athwartship when ship-mounted).\\
-$\eqang(\freqsym)$ & & sr & Two-way equivalent beam angle at frequency $\freqsym$.\\
+$\eqang(\freqsym)$ & \code{psi_m} & sr & Two-way equivalent beam angle at frequency $\freqsym$.\\
 \end{longtable}

### 2023-09-18 main.py
Removed old json files. Renamed SphereBeam to TSf

@@ -7,10 +7,16 @@ from plots import *
 
 def preCalculations(data):
     """
-    XXX.
+    Reads echosounder data, single ping, from datafile (json), performs
+    pre-calculations.
     
     Parameters
     ----------
+    
+    Returns
+    -------
+    All input data for further calculations.
+    
     
     """
 

### 2023-09-19 Paper/Manuscript.tex
Added information to json files and file reader, changed json file names

@@
 In order to make the code echosounder independent simple json strings are used as input.
-Detailed information on settings used during data collection are available in the code and the data files on GitHub.
+Detailed information on settings used during data collection is available in the code and the data files on GitHub.
 
@@
 where fore, aft, star(board), and port indicate the relevant transducer halves.
+In scientific echoshounder data fore-aft angles are often labelled  alongship angles and  port-starboard angles athwartship angles.
 

### 2023-09-19 TSf.py

@@ -2,7 +2,7 @@ from Core.EK80DataContainer import EK80DataContainer
 from plots import *
 
 do_plot = True
-file = r"./Data/CRIMAC_SphereBeam.json"
+file = r"./Data/CRIMAC_TSf.json"
 
 data = EK80DataContainer(file)  # TS sphere
 

### 2023-09-19 Tools/readpyecholabPing.py

@@ -8,12 +8,21 @@ Requires that pyecholab is installed
 https://github.com/CI-CMG/pyEcholab
 
 Usage:
-python readpyecholabPing --channel 2 --pingNo 11 --inputFile E:/IMR-CRIMAC-EK80-FM-38kHz-WC22-WC381.raw --outputFile ..\Data\pyEcholabEK80data.json
+python readpyecholabPing --channel 2 --pingNo 11 --inputFile E:/IMR-D20210507-T074652-Svf.raw --outputFile ..\Data\pyEcholabEK80data.json
+
 or if you are within Spyder or similar:
+
 run readpyecholabPing --channel 2 --pingNo 11 --inputFile E:/IMR-CRIMAC-EK80-FM-38kHz-WC22-WC381.raw --outputFile ..\Data\pyEcholabEK80data.json
-Reads channel 2, in this case 38 kHz, and ping number 11 in the raw file
-run readpyecholabPing --channel 2 --pingNo 11 --inputFile C:/Users/a32685/Documents/Projects/2020_CRIMAC/CRIMACHackEx/cal-babak-D20201120-T080856.raw
-GP 01.06.2021
+
+Reads channel 2, in this case 38 kHz, and ping number 11 in the raw file.
+
+Numbering starts at 1 (channel and pingNo).
+
+The json files used in the code and paper were created with:
+run readpyecholabPing --channel 4 --pingNo 514 --inputFile ..\RawData\IMR-D20210507-T074652-Svf.raw --outputFile ..\Data\CRIMAC_Svf.json
+run readpyecholabPing --channel 2 --pingNo 510 --inputFile ..\RawData\IMR-D20211215-T143432-TSf.raw --outputFile ..\Data\CRIMAC_TSf.json
+
+GP 08.09.2023
 """
 
 import json
@@ -77,6 +86,10 @@ raw_data = raw_list[0]
 Type = raw_data.configuration[0]["transceiver_type"]
 Frequency = raw_data.configuration[0]["transducer_frequency"]
 filenameRaw = raw_data.configuration[0]["file_name"]
+transTrans = raw_data.configuration[0]["channel_id"]
+application = raw_data.configuration[0]["application_name"]
+applicationVersion = raw_data.configuration[0]["application_version"]
+fileFormatVersion = raw_data.configuration[0]["file_format_version"]
 
 # get calibration values
 calibration = raw_data.get_calibration()
[…]

### 2023-09-19 main.py

@@ -337,7 +337,7 @@ if __name__ == "__main__":
     )
     ap.add_argument(
         "--tfile",
-        default="./data/CRIMAC_SphereBeam.json",
+        default="./data/CRIMAC_TSf.json",
         help="File containing raw data for TS calculation",
     )
     ap.add_argument(

### 2023-09-19 Paper/Manuscript.tex
Updated manuscript

@@
 The red vertical line indicates the ping that is used for illustrating the Sv(f) processing used in (b).
-(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.}
+(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.
+In this example the decimated sampling rate ($\fsdec$, Eq.
+\ref{eq:fsdec}) is 93.75 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, Eq.
+\ref{eq:TauEff}) is 0.02 ms.}
 \end{figure}

### 2023-09-19 Tools/readpyecholabPing.py

@@ -238,6 +238,7 @@ data = {
         "Centre Frequency": Frequency,
         "Channel Number": Channel+1,        
         "Pulse Duration": round(PulseDurationPing*10**3,4),
+        "Ramping / Slope": round(Slope,4),
         "Ping Number": PingNo+1,
     },
     "XML0": {

### 2023-09-19 main.py

@@ -321,11 +321,6 @@ def calcSv():
     # Plots for paper
     plotSvf(f_m, Sv_m_n, svf_range)
 
-
-# Example Usage :
-#   main.py --tfile ./data/CRIMAC_SphereBeam.json --sfile ./data/CRIMAC_Svf.json
-#   main.py --tfile ./data/CRIMAC_SphereBeam.json --sfile ./data/CRIMAC_Svf.json --plots false
-#   main.py --tfile ./data/CRIMAC_SphereBeam.json --sfile none
 if __name__ == "__main__":
     # Handle command line parameters
     #

### 2023-09-27 Paper/Manuscript.tex
Updated manuscript

--- 2023-09-14 Paper/Manuscript.tex
@@
 \usepackage{color,soul}
-
 \usepackage{lineno}
 \usepackage{xurl}
+\usepackage{hyperref}
 
@@
 
-\author{Lars Nonboe Andersen}
-\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}
-
-\author{Dezhang Chu}
-\affiliation{Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd.
-E.
-Seattle, WA, 98112, USA}
-
-\author{Nils Olav Handegard}
-\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
-
-\author{Harald Heimvoll}
-\affiliation{Kongsberg Discovery AS, Strandpromenaden 50, 3191, Horten, Norway}
-
-\author{Rolf Korneliussen}
-\author{Gavin J. Macaulay}
-\author{Egil Ona}
-\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
-
-\author{Ruben Patel}
-\affiliation{Codelab, Bergen, Norway}
-
-\author{Geir Pedersen}
-\email{geir.pedersen@hi.no}
-\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}
-
 \begin{abstract}
@@
 
-Keywords:  \linebreak
-
 \end{abstract}
@@
 The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}.
-More recently, fisheries acoustics sensors have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
+More recently, fisheries acoustics sensors have been deployed on a wide range of platforms including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.
 
-Today echosounders can produce pulses with a wide and continuous frequency range (broadband pulses), compared to the conventional narrowband systems.
-This provides in most cases significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
-\citet{lavery_measurements_2010} used broadband signals to reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure.
+Today, echosounders can produce pulses with a wide and continuous frequency range (broadband pulses), compared to the conventional narrowband systems.
+This provides in most cases significantly better along-beam (range) resolution, a higher signal-to-noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and improved frequency resolution for backscatter categorization \citep{korneliussen2018}.
+\citet{lavery_measurements_2010} used broadband signals to reduce ambiguities in the interpretation of acoustic scattering from zooplankton and oceanic microstructure.
 \citet{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) that included the resonance frequency of swimbladdered fish to classify size classes.
-\citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing, to characterize the composition of two sound scattering layers (SSL).
-\citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids.
+\citet{blanluet_characterization_2019} used broadband acoustics, with ground truthing (nets and video), to characterize the composition of two sound scattering layers (SSL).
+\citet{bassett_broadband_2018} showed that broadband signals may be helpful in characterizing smaller fishes with swimbladders and euphausiids (15-150 kHz).
 \citet{benoit-bird_exploring_2020} were able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband signals (45-170 kHz).
-\citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments.
-By applying high frequency broadband pulses to fish-like artificial targets, \citet{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment.
-Utilizing the increased spatial resolution \citet{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.
+\citet{lavery2017} explored different broadband pulse shapes to increase the ability to resolve adjacent single targets as well as near boundaries in tank experiments (15-400 kHz).
+By applying high-frequency broadband pulses to fish-like artificial targets, \citet{kubilius_remote_2020, kubilius_remote_2023} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled \textit{ex situ} environment (45-90 and 160-260 kHz).
+Using the increased range resolution, \citet{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder \textit{in situ} (45-260 kHz).
+Using narrow (18, 38 kHz) and broadband acoustic data (70-280 kHz) with a mixed species scattering model,  \citet{loranger_broadband_2022} demonstrated estimation of mean length and total biomass of longfin squid and mackerel.
 
 Several scientific broadband echosounder systems have been developed for laboratory use \citep{Conti2003Wide-bandwidth, Forland2014Scattering, chu1992}, some prototype or custom-made systems \citep{Zakharia1989Wide-band, Zakharia1996Wideband, Simmonds1996Species, Foote2005Measuring, Imaizumi2009Detection, Briseno-Avena2015ZOOPS, Barr2002Target} and some commercially available systems \citep{Gordon1998FishMASS, Zedel2003Acoustic, Stanton2010New, ehrenbergFMSlideChirp2000, dennyBroadbandAcousticFish1998}.
-More widespread use of broadband acoustics in fisheries and ecosystem research have followed from the upgrade of widely used narrowband systems to gain broadband capabilities.
+More widespread use of broadband acoustics in fisheries and ecosystem research has followed from the upgrade of widely used narrowband systems to gain broadband capabilities.
 
 Signal processing methods for broadband echosounders \citep{stanton2008} are based on radar signal processing theory, with further adaptation to echosounders \citep{lavery2017,bassett_broadband_2018}.
-When implementing the equations to computer processing code, some choices are well founded in the signal processing literature whereas others are of a more practical and ad-hoc nature.
-The latter is typically missing in the literature, making it difficult to benchmark new methods and to test the implementation in new echosounders and post processing software.
+When implementing the equations to computer processing code, some choices are well founded in the signal processing literature, whereas others are of a more practical and ad-hoc nature.
+The latter is typically missing in the literature, making it difficult to benchmark new methods and to test the implementation in new echosounders and post-processing software.
 
-The objective of this paper and associated code is to provide a benchmark for developing and implementing signal processing algorithms for broadband echosounder.
+The objective of this paper and the associated code is to provide a benchmark for developing and implementing signal processing algorithms for broadband echosounders.
 We present the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data.
-The steps include pulse compression, target strength as a function of frequency ($\ts(\freqsym)$, dB re 1m$^2$), and volume backscattering strength as a function of frequency ($\sv(\freqsym)$, dB re 1m$^-1$).
+The steps include pulse compression, target strength as a function of frequency ($\ts(\freqsym)$, dB re 1m$^2$), and volume backscattering strength as a function of frequency ($\sv(\freqsym)$, dB re 1m$^{-1}$).
 The intention is that the code will be used as a starting point for implementations in various relevant data processing software, for further developing active acoustic broadband signal processing, and to serve as a learning resource.
@@
 \subsection{Accompanying code}
-The code  accompanying this paper is written in the Python programming language (v3.10) and is available through GitHub (see \ref{dataavstatement} for code and data availability) together with the data used in the examples.
-All single ping processing steps with respective figures in the paper can be reproduced by running the main script, \code{main.py}.
-Reproduction of Fig.
-\ref{fi:ts_theta_phi}a and Fig.
-\ref{fi:Fig_Sv_echogram}a requires downloading the original echosounder raw data separately and using the scripts (\code{TSfEchogram.py} and \code{SvfEchogram.py}).
-Without loss of generality, we use the \ek{} echosounder as an example since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
-The test data sets accompanying the code are representative for the data contained in Simrad EK80 raw files.
-In order to make the code echosounder independent simple json strings are used as input.
-Detailed information on settings used during data collection are available in the code and the data files on GitHub.
+The code  accompanying this paper is written in the Python programming language (v3.10) and is available through GitHub (see \autoref{dataavstatement} for code and data availability) together with the data used in the examples.
+All single-ping processing steps with respective figures in the paper can be reproduced by running the main script, \code{main.py}.
+Reproduction of \autoref{fi:ts_theta_phi}a and \autoref{fi:Fig_Sv_echogram}a requires downloading the original echosounder raw data separately and using the scripts (\code{TSfEchogram.py} and \code{SvfEchogram.py}).
+
+Without loss of generality, we use the \ek{} echosounder as an example, since it is currently the most commonly used broadband system in the marine ecosystem acoustics field.
+The test data sets accompanying the code are representative of the data contained in Simrad EK80 raw files.
+In order to make the code echosounder independent, simple json strings are used as input.
+Detailed information on the settings used during data collection is available in the code and data files on GitHub.
 
@@
 \subsection{System overview}
-A basic quantitative echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records received signals.
-During transmission the program defines the signals that are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water.
+A basic quantitative echosounder system consists of a transducer, a transceiver, and a computer program that controls the operation of the transceiver and records the received signals.
+During transmission, the program defines the signals that are created as electric signals in the transceiver, converted to acoustic signals by the transducer and transmitted into the water.
 The acoustic signals propagate through the water, are reflected or scattered by objects in the water, and propagate back to the transducer.
-During reception the transducer converts the received acoustic signals to electric signals, which are received, pre-amplified, filtered, digitized, processed in the transceiver, and then transferred to the controlling program for further data processing and storage (Fig.
-\ref{fi:ek_sys}).
+During reception, the transducer converts the received acoustic signals to electric signals, which are received, preamplified, filtered, digitized, processed in the transceiver, and then transferred to the controlling program for further data processing and storage ( \autoref{fi:ek_sys}).
 Many types of transmit signals are feasible - this paper considers only upsweep linear frequency modulated signals (also known as linear chirps).
@@
 The analogue and amplified signals are passed on to the transducer to generate the transmitted acoustic signal $\ytxa(\timesym)$ in the water.
-For a splitbeam echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
+For a split-beam echosounder system, there are typically three or four channels to allow estimation of the angle of arriving echoes, and the signal is typically transmitted with equal power across the channels.
 
@@
 
-The returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by corresponding receiver channels, $\channelsym$, in the transceiver.
+The returning acoustic signal, $\yrxa(\timesym)$, is received by each transducer sector, $\channelsym$, and converted to an analog electric signal, $\yrxe(\timesym,\channelsym)$, in the transducer and received by the corresponding receiver channels, $\channelsym$, in the transceiver.
 The received electric signal, $\yrxe(\timesym,\channelsym)$, from each channel, $\channelsym$, is pre-amplified, filtered by an analog anti-aliasing filter, and digitized in the transceiver at a frequency of $\fs$, creating the digital signal, $\yrxorg(\samplesymt,\channelsym)$.
 
-To remove noise and reduce the quantity of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hbp(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$.
+To remove noise and reduce the amount of data, the sampled signal from each channel is filtered and decimated in multiple stages, $\stagesym$, using complex bandpass filters, $\hbp(\genidxsym,\stagesym)$, and decimation factors, $\decfac(\stagesym)$.
 This processing step is approximately equivalent to a traditional demodulation process.
@@
 The characteristics of the bandpass filter and decimation factors are chosen with regard to the desired operating bandwidth, noise suppression levels, impulse response duration, and other common filter characteristics, with the aim of maintaining sufficient information in the data \citep{ProakisDSP,CrochireMDSP}.
-The frequency responses of the filters are shown in Fig.
-~\ref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set, where $N_v=2$.
+The frequency responses of the filters are shown in \autoref{fi:fir} and the corresponding filter coefficients and decimation factors are given in the test data set, where $N_v=2$.
 
@@
 \caption{\label{fi:fir} Example of frequency response (filter gain) of the filters in our test set.
-The blue and orange curve represents the filter response of the first and second filter.
-Note that the blue line is above 0 dB, this is caused by the transition from complex to real values.
-The vertical dashed green lines indicates the frequency range of the transmit signal.}
+The blue and orange curves represent the filter responses of the first and second filter.
+Note that the blue line is above 0 dB; this is caused by the transition from complex to real values.
+The vertical dashed green lines indicate the frequency range of the transmit signal.}
 \end{figure}
@@
 Instead, the filtered and decimated complex samples from each transducer channel $\yrx(\samplesymt,\channelsym)$ are stored in the data files.
-The data are recorded in computer data files for display and analysis by processing software.
-Additional information, such as that from position and motion sensors and system configuration data is also included in the files.
+Data are recorded in computer data files for display and analysis by processing software.
+Additional information, such as from position and motion sensors and system configuration data, is also included in the files.
 
 \subsection{Matched filtering (pulse compression)}
-To increase signal-to-noise ratio and resolution along the acoustic beam a matched filter may be applied to the raw data samples \citep{turin1960}.
-This technique is also known as pulse compression \citep{klauder1960}, achieving the average transmitted power of a relatively long pulse while obtaining the range resolution of a shorter pulse.
-One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in Eq.
-\ref{eq:yrx}.
+To increase signal-to-noise ratio and resolution along the acoustic beam, a matched filter may be applied to the raw data samples \citep{turin1960}.
+This technique is also known as pulse compression \citep{klauder1960}, which achieves the average transmitted power of a relatively long pulse while obtaining the range resolution of a shorter pulse.
+One approach for a matched filter is to use a normalized version of the ideal transmit signal as the replicate signal, filtered and decimated using the same filters and decimation factors as applied in \autoref{eq:yrx}.
 The normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, is given by:
@@
 where $\textrm{max}$ is the maximum value of $\ytx(\samplesymt)$.
-The filtered and decimated output signal, $\ytxnorm(\samplesymt,\stagesym)$, from each filter stage, $\stagesym$, using the normalized ideal transmit signal, $\ytxnorm(\samplesymt)$, as the input signal, is given by:
+The filtered and decimated output signal, $\ytxnorm(\samplesymt,\stagesym)$, of each filter stage, $\stagesym$, using the ideal normalized transmit signal, $\ytxnorm(\samplesymt)$, as the input signal, is given by:
 %
@@
 where $\ytxnorm(\samplesymt,0)$ is set to $\ytxnorm(\samplesymt)$.
-The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is denoted as $\ymf(\samplesymt)$  (Fig.~\ref{fi:y_mf_n_ACF}a).
+The output signal from the final filter and decimation stage, $\ytxnorm(\samplesymt,\nstages)$, is used as the matched filter and is indicated as $\ymf(\samplesymt)$  (\autoref{fi:y_mf_n_ACF}a).
 %
@@
 \includegraphics[width=16cm]{Fig_y_mf_n_ACF}
-\caption{\label{fi:y_mf_n_ACF} An example where the decimated sampling rate ($\fsdec$, Eq.
-\ref{eq:fsdec}) is 125 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, Eq.
-\ref{eq:TauEff}) is 0.01 ms.
-(a) The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, from the final filter and decimation stage, that is used for the pulse compression.
+\caption{\label{fi:y_mf_n_ACF} An example where the decimated sampling rate ($\fsdec$, \autoref{eq:fsdec}) is 125 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, \autoref{eq:TauEff}) is 0.01 ms.
+(a) The absolute value of the filtered and decimated output signal, $\ymf(\samplesymt)$, from the final filter and decimation stage, which is used for the pulse compression.
 (b) The autocorrelation function $\ptxauto(n)$.}
@@
 
-The auto correlation function of the matched filter signal ($\ymfauto(\samplesymt)$) and the effective pulse duration ($\teff$) will be used in later processing steps.
+The autocorrelation function of the matched filter signal ($\ymfauto(\samplesymt)$) and the effective pulse duration ($\teff$) will be used in later processing steps.
 
@@
 
-is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over the duration of the autocorrelation function (Fig. \ref{fi:y_mf_n_ACF}b).
+is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over the duration of the autocorrelation function (\autoref{fi:y_mf_n_ACF}b).
 
-To perform pulse compression the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power.
+To perform pulse compression, the received signal, $\yrx(\samplesymt,\channelsym)$, is convolved with a complex conjugated and time-reversed version of the matched filter signal, and here also normalized with the $l^2$-norm of the matched filter to maintain received signal power.
 The pulse compressed signal, $\ypc(\samplesymt,\channelsym)$, then becomes
@@
 Compensation of echo strength for position in the acoustic beam requires an estimate of the echo arrival angle.
-This is obtained using the splitbeam method \citep{burdic1991}, which for broadband pules can be implemented with the angle values contained in the complex-valued $\ypc(\samplesymt)$ data, in combination with knowledge of transducer sector geometry.
-The principle is demonstrated with a transducer that is divided into four quadrants (Fig.
-\ref{fi:Fig_trd_quad_Impedances}a).
-In this example the summed signals from four halves (1+2, 2+3, 3+4, 4+1) are calculated as:
+This is obtained using the split-beam method \citep{burdic1991}, which for broadband pules can be implemented with the angle values contained in the complex-valued $\ypc(\samplesymt)$ data, in combination with knowledge of transducer sector geometry.
+The principle is demonstrated with a transducer divided into four quadrants (\autoref{fi:Fig_trd_quad_Impedances}a).
+In this example, the summed signals from four halves (1+2, 2+3, 3+4, 4+1) are calculated as:
 \begin{eqnarray}
@@
 where fore, aft, star(board), and port indicate the relevant transducer halves.
+In scientific echoshounder data, fore-aft angles are often labelled  alongship angles and  port-starboard angles athwartship angles.
 
@@
 The labels are directions often used when a transducer is mounted on a ship.
-(b) Equivalent circuit diagram of transducer/transceiver with system impedances.}
+(b) Equivalent circuit diagram of the transducer/transceiver with the impedances of the system.}
 \end{figure}
@@
 The transceiver measures voltage over a load, $\zrxe$, connected in series with the transducer impedance, $\ztde$.
-When calculating various acoustic properties, a system gain parameter will be used which assumes a matched receiver load.
-The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (Fig.
-\ref{fi:Fig_trd_quad_Impedances}b) is given by:
+When calculating various acoustic properties, a system gain parameter will be used that assumes a matched receiver load.
+The total received power, $\prxe(\samplesymt)$, from all transducer sectors for a matched receiver load (\autoref{fi:Fig_trd_quad_Impedances}b) is given by:
 \begin{equation}
@@
 %
-where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:ts_theta_phi}b) and are derived from the transducer geometry \citep{Urick3rdPrinciples} and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}.
+where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (\autoref{fi:ts_theta_phi}b) and are derived from the geometry of the transducer \citep{Urick3rdPrinciples} and $\fc$ the centre frequency of the chirp pulse \citep{ehrenberg1979}.
 The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$.
-As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
+As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for the port/starboard angles.
 %
@@
 
-To illustrate calculation of target strength ($TS$, dB re 1m$^2$) as a function of frequency, $TS(f)$, we use data collected on a 35 mm diameter tungsten carbide calibration sphere (WC35) suspended approximately 5.8 m below a 120 kHz transducer (Fig.~\ref{fi:ts_theta_phi}a).
+To illustrate the calculation of target strength ($TS$, dB re 1m$^2$) as a function of frequency, $TS(f)$, we use data collected on a 35 mm diameter tungsten carbide calibration sphere (WC35) suspended approximately 5.8 m below a 120 kHz transducer (\autoref{fi:ts_theta_phi}a).
 
@@
 \includegraphics[width=16cm]{Fig_TS_echogram_theta_phi}
-\caption{\label{fi:ts_theta_phi} (a) $\mysp$ as a function of ping number and range for a calibration sphere.
-The sphere is located at approximately 5.8 m range.
-The red vertical line indicates the ping that is used for illustrating the $\ts(\freqsym)$ processing.
+\caption{\label{fi:ts_theta_phi} (a) $\mysp$ as a function of the number of pings and range.
+A calibration sphere (WC35) is located at approximately 5.8 m range.
+The red vertical line indicates the ping that is used to illustrate $\ts(\freqsym)$ processing.
 (b) The physical angles $\along$ and $\athw$ for the target strength example data (read vertical line in a).
-The single target can be seen around range 5.8 m where the angles are less variable.}
+The single target can be seen around the range 5.8 m where the angles are less variable.}
 \end{figure}
@@
 %
-where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp(\freqsym)$ the acoustic absorption coefficient, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain(\along,\athw,\freqsym)$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
+where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range of the target, $\absorp(\freqsym)$ the acoustic absorption coefficient, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain(\along,\athw,\freqsym)$ the transducer gain that incorporates both the on-axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.
 
-The point scattering strength (TS prior to target detection and beam pattern compensation), $\mysp(\samplesymt)$, is estimated by applying Eq.
-\ref{eq:TS} to the received digitized power samples using the on-axis gain value and $f$ set to the centre frequency of the broadband pulse, $\fc$:
+The point scattering strength (TS prior to target detection and beam pattern compensation), $\mysp(\samplesymt)$, is estimated by applying \autoref{eq:TS} to the received digitized power samples using the on-axis gain value and $f$ set to the centre frequency of the broadband pulse, $\fc$:
 \begin{equation}
@@
 Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes that occur in actual targets (as opposed to ideal point targets).
-The alongship angle $\along(\samplesymt)$, athwartship angle $\athw(\samplesymt)$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (Fig.
-\ref{fi:ts_theta_phi}b).
+The alongship angle $\along(\samplesymt)$, athwartship angle $\athw(\samplesymt)$ and sample number $n$ at the \emph{peak} power $\prxe(\samplesymt)$ within the detected target are used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively (\autoref{fi:ts_theta_phi}b).
 A simple pseudo SED algorithm, simply using a threshold, is implemented in the code for illustrative purposes.
 
-From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak (to that used for the target signal) are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$ (Fig.~\ref{fi:SED}).
-Depending on the target scattering characteristics and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
-If the number of samples around the target is more than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used.
+From the autocorrelation function of the matched filter signal, $\ymfauto(\samplesymt)$, the equivalent number of samples around the peak (to that used for the target signal) are extracted to create the reduced autocorrelation signal of the matched filter signal, $\ymfautored(\samplesymt)$ (\autoref{fi:SED}).
+Depending on the scattering characteristics of the target and the distance to any adjacent single targets, the number of samples around the peak echo level in $\ypctarget(\samplesymt)$ that contain the majority of the echo energy can be more or less than the total number of samples around the peak of $\ymfauto(\samplesymt)$.
+If the number of samples around the target is greater than the total number of samples around the peak of $\ymfauto(\samplesymt)$ all samples around the peak of $\ymfauto(\samplesymt)$ are used.
 If the number of samples around the target is less than the total number of samples around the peak of $\ymfauto(\samplesymt)$, this lower number is used to create $\ymfautored(\samplesymt)$.
@@
 \caption{\label{fi:SED} $\ypctarget(\samplesymt)$ (upper figure) and $\ymfautored(\samplesymt)$ (lower figure).
-$\ymfautored(\samplesymt)$ is the autocorrelation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.
-The corresponding split beam angles ($\along_t$ and $\athw_t$) for the single target are shown in (Fig.~\ref{fi:ts_theta_phi}b).}
+$\ymfautored(\samplesymt)$ is the auto-correlation function of the transmit signal reduced to the length of the target signal and aligned with the peak power of the target.
+The corresponding split beam angles ($\along_t$ and $\athw_t$) for the single target are shown in (\autoref{fi:ts_theta_phi}b).}
 \end{figure}
 
-The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
+The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto-correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
 \begin{eqnarray}
@@
 where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
-The normalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (Fig. \ref{fi:TS}) is then calculated by: 
+The normalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\autoref{fi:TS}) is then calculated by: 
 %
@@
 
-Target strength can then be estimated using Eq. \ref{eq:TS}:
+The target strength can then be estimated using \autoref{eq:TS}:
 \begin{equation}
@@
 \end{equation}
-where $\lfloor$ $\rfloor$ represents modulus.
+where $\lfloor$ $\rfloor$ represents the modulus.
 %
@@
 \includegraphics[width=16cm]{Fig_TS}
-\caption{\label{fi:TS} (a) The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$, (b) and the reduced auto correlation signal $\ymfautoredf(\samplesymf)$, (c) the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$, (d) transducer gain $(10\log_{10}\gain^2(\along_t,\athw_t,\freqsym))$, and (e) the estimated TS(f).}
+\caption{\label{fi:TS} (a) The discrete Fourier transform of the target signal $\ypctargetf(\samplesymf)$, (b) and the reduced auto-correlation signal $\ymfautoredf(\samplesymf)$, (c) the normalized discrete Fourier transform of the target signal $\ypctargetnormf(\samplesymf)$, (d) transducer gain $(10\log_{10}\gain^2(\along_t,\athw_t,\freqsym))$, and (e) the estimated TS(f).}
 \end{figure}
 %
-A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
-This is visible in the estimated $\ts$ (Fig.
-\ref{fi:TS}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
+A frequency-modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}.
+This is visible in the estimated $\ts$ (\autoref{fi:TS}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}.
 
@@
 
-To illustrate calculation of volume backscattering strength as a function of frequency,  $\sv(f)$ (dB re 1m$^{-1}$), we use data collected from a school of non-swimbladdered fish (Fig.~\ref{fi:Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer.
+To illustrate calculation of volume backscattering strength as a function of frequency,  $\sv(f)$ (dB re 1m$^{-1}$), we use data collected on a school of fish lacking swimbladder (\autoref{fi:Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer.
 
@@
 %
-where $\prxevf(\freqsym)$ is the received electric power in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used for evaluating the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang(\freqsym)$ is the two-way equivalent beam angle.
+where $\prxevf(\freqsym)$ is the electric power received in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used to evaluate the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang(\freqsym)$ is the two-way equivalent beam angle.
 The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
@@
 \includegraphics[width=16cm]{Fig_Sv_echogram_m_n}
-\caption{\label{fi:Fig_Sv_echogram} Illustration of volume backscattering strength as a function of frequency, Sv(f).
-A non-swimbladdered fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
+\caption{\label{fi:Fig_Sv_echogram} Illustration of the volume backscattering strength as a function of frequency, Sv(f).
+The fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
 (a) Sv as a function of ping number and range range for the raw data file used in the Sv(f) example.
-The red vertical line indicates the ping that is used for illustrating the Sv(f) processing used in (b).
-(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.}
+The red vertical line indicates the ping that is used to illustrate the Sv(f) processing used in (b).
+(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.
+In this example, the decimated sampling rate ($\fsdec$, \autoref{eq:fsdec}) is 93.75 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, \autoref{eq:TauEff}) is 0.02 ms.}
 \end{figure}
 
-Volume backscattering samples compressed over the operational frequency band are estimated by applying Eq.
-\ref{eq:sv_f} to the received digitized power samples using the on-axis gain value with $f$ set to the centre frequency of the broadband pulse, $\fc$:
+Volume backscattering samples compressed over the operational frequency band are estimated by applying \autoref{eq:sv_f} to the received digitized power samples using the on-axis gain value with $f$ set to the centre frequency of the broadband pulse, $\fc$:
 \begin{equation}
@@
 where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading.
-A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$.
+A discrete Fourier transform is performed on the range-compensated pulse-compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$.
 The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution.
@@
 
-The normalised Hann window, $\hannwnorm$, is given by: 
+The normalized Hann window, $\hannwnorm$, is given by: 
 %
@@
 where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window.
-The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
+The discrete Fourier transform of the auto-correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
 %
@@
 %
-Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by:
+Finally, the discretized estimate of $\sv(\freqsym)$, $\sv(\samplesymf)$, is given by the following:
 %
@@
 \end{equation}
-where the sample index $\samplesymf$ corresponding to frequency $\freqsym$ can be estimated using Eq. {\ref{eq:m(f)}}.
+where the sample index $\samplesymf$ corresponding to the frequency $\freqsym$ can be estimated using {\autoref{eq:m(f)}}.
 
 By selecting a set of centre samples $\timesym$, $\sv$ values can be presented as a function of range ($\samplesymt$) and frequency ($\freqsym$) for each ping.
-The range for centre samples $\samplesymt$ could be chosen as half the window length or any other grid that the user prefer the data presented to be in.
+The range for the centre samples $\samplesymt$ could be chosen as half the window length or any other grid that the user prefers the data presented to be in.
 This can be useful when combining the $\sv(\freqsym)$ across a range of transducers.
-In our example we have simply chosen the set of centre samples as the original range samples (Fig.~\ref{fi:Fig_Sv_echogram}).
+In our example, we have simply chosen the set of centre samples as the original range samples (\autoref{fi:Fig_Sv_echogram}).
 
-For acoustic abundance estimation and classification purposes it is common to integrate $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, Fig.~\ref{fi:Fig_Sv_echogram}).
-It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here just one ping is used for illustrative purposes (Fig.~\ref{Fig_Sv_avg}).
+For acoustic abundance estimation and classification purposes, it is common to integrate $\sv$ over a range (15 to 34 m in the example, covering a school of non-swimbladdered fish, \autoref{fi:Fig_Sv_echogram}).
+It is normal to average $\sv$ over several pings to obtain an unbiased estimate, but here only one ping is used for illustrative purposes (\autoref{Fig_Sv_avg}).
 Even though this is for a single ping it is still possible to observe a positive slope of the frequency response that is indicative of non-swimbladdered fish.
@@
 \includegraphics[width=16cm]{Fig_Sv_avg}
-\caption{\label{Fig_Sv_avg} $\sv$ of the single ping marked in Fig.~\ref{fi:Fig_Sv_echogram}a as a function of frequency averaged over a depth interval covering a fish school.
-A weak positive slope is observed of the frequency response.}
+\caption{\label{Fig_Sv_avg} $\sv$ of the single ping marked in \autoref{fi:Fig_Sv_echogram}a as a function of frequency averaged over a depth interval covering a fish school.
+A weak positive slope is observed in the frequency response.}
 \end{figure}
 
-The trend for increasing $\sv$ with frequency is well-known for fish without swimbladder \citep{korneliussen2010} and is consistent with the trend observed in this example.
-In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig.~\ref{Fig_Sv_avg}).
+The trend for increasing $\sv$ with frequency is well known for fish without swimbladder \citep{korneliussen2010} and is consistent with the trend observed in this example.
+In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (\autoref{Fig_Sv_avg}).
 
@@
 The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of the data processing steps.
-The contribution includes all steps well founded in the literature as well as any practical and more ad-hoc choices.
-The choices include handling of gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering.
+The contribution includes all steps well-founded in the literature as well as any practical and more ad-hoc choices.
+Choices include handling gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering.
 When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases (e.g.
 distorted output at the beginning and end of the section of filtered data) and in our implementation we chose to exclude these.
-We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
-When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform and the actual choice will be a compromise between accuracy and computational speed.
+We also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed.
+When estimating $\ts(\freqsym)$ and $\sv(\freqsym)$ the resolution and accuracy will depend on the length, $\ndft$, of the Fourier transform, and the actual choice will be a compromise between accuracy and computational speed.
 Our objective is not to provide an evaluation of all these choices, but to document the baseline for the processing that can be used for benchmarking purposes.
 
-$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals.
-Features includes size, target classification and, behaviour through tracking.
-In our implementation, calculation of $\ts(\freqsym)$ assumes that a single target has successfully been identified.
-This requires a robust single target detector.
-Several SED algorithms exist and different algorithms may be required depending on the situation.
+$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features from single individuals.
+Features include size, target classification, and behaviour through tracking.
+In our implementation, the calculation of $\ts(\freqsym)$ assumes that a single target has been successfully identified.
+This requires a robust single-target detector.
+There are several SED algorithms, and different algorithms may be required depending on the situation.
 Typical SED algorithms are based on traditional single frequency pulses and by utilizing the additional information in broadband echoes improved SEDs may be envisioned, but this is outside the scope of this paper.
@@
 $\sv(\freqsym)$ is a key parameter for echo integration.
-To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range.
-The chosen size of the window is two times the pulse length, and is a compromise between spatial and frequency resolution.
+To estimate $\sv(\freqsym)$ a Fourier transform is used, applied repeatedly via a sliding window in range.
+The chosen size of the window is twice the pulse length, and is a compromise between spatial and frequency resolution.
 Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform.
 Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for the operating frequencies for typical marine ecosystem echosounder for short range windows.
-The compensation for absorption loss is therefore performed after applying the discrete Fourier transform.
+Therefore, the compensation for absorption loss is performed after applying the discrete Fourier transform.
 The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by deep learning methods \citep[e.g.]{brautaset_acoustic_2020}.
 
-The formulation presented in this paper requires several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$.
-Methods to estimate these are not within the scope of this paper, but common practice is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
+The formulation presented in this paper requires several frequency-dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$.
+Methods for estimating these are not within the scope of this paper, but common practice is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}.
 We note that these methods do not provide an operational method to estimate $\teff$ or $\eqang(\freqsym)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.
@@
 A set of equations and associated computer code for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented along with example code, providing a resource for those interested in learning and further developing broadband processing techniques.
-The processing equations and methodology presented in this paper are similar to those implements in version 1.12.4 and earlier of the \ek{} software.
+The processing equations and methodology presented in this paper are similar to those implemented in version 1.12.4 and earlier of the \ek{} software.
 
@@
 
-A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented along with example code, with reference to the \ek{} echosounder.
+A set of equations for calculating calibrated, frequency-dependent, target strength, and volume backscatter from broadband echosounder signals have been presented along with example code, with reference to the \ek{} echosounder.
 

### 2023-09-29 Paper/Manuscript.tex
Error fixes

@@
 
-\title[]{Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder}
+\title[]{Quantitative processing of broadband data as implemented in a scientific split-beam echosounder}
 
@@
 2.
-We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength (TS(f)) and volume backscattering strength (Sv(f)) using data from the EK80 broadband scientific echosounder as examples.
+We describe the steps involved in processing broadband acoustic data from raw data to frequency dependent target strength ($\ts(\freqsym)$) and volume backscattering strength ($\sv(\freqsym)$) using data from the EK80 broadband scientific echosounder as examples.
 Although the overall processing steps are described and build on established methods from the literature, multiple choices need to be made during implementation.
@@
 To highlight and discuss some of these choices and facilitate a common understanding within the community, we have also developed a Python code which will be made publicly available and open source.
-The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to TS(f) and Sv(f).
+The code follows the steps using raw data from two single pings, showing the step-by-step processing from raw data to $\ts(\freqsym)$ and $\sv(\freqsym)$.
 \linebreak
@@
 %
-where $\prxevf(\freqsym)$ is the electric power received in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\tslide$ the duration of the time window, excluding the zero-padded portion if applied, used to evaluate the frequency spectrum, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang(\freqsym)$ is the two-way equivalent beam angle.
+where $\prxevf(\freqsym)$ is the electric power received in a matched load for the signal from a volume at frequency $\freqsym$, $\cw$ the sound speed, $\range_c$ is the range to the centre of the range volume covered by $\tslide$, and $\eqang(\freqsym)$ is the two-way equivalent beam angle.
+$\tslide$ is the duration of the time window, excluding the zero-padded portion if applied, used to evaluate the frequency spectrum.
+Note that for the single frequency, narrowband, case the pulse duration is used, for the broadband case $\tslide$ is used.
 The two-way equivalent beam angle is a function of frequency that is derived from an empirical estimate of $\eqang$ at the nominal frequency, $\fn$:
@@
 The fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m.
-(a) Sv as a function of ping number and range range for the raw data file used in the Sv(f) example.
-The red vertical line indicates the ping that is used to illustrate the Sv(f) processing used in (b).
-(b) Sv(f) as a function of frequency and range for a single ping indicated in (a) with a vertical red line.
+(a) Pulse compressed Sv as a function of ping number and range for the raw data file used in the Sv(f) example.
+The red vertical line indicates the ping that is used to illustrate the Sv(f) processing.
+(b) Sv as a function of frequency and range for the single ping indicated in (a) with a vertical red line.
 In this example, the decimated sampling rate ($\fsdec$, \autoref{eq:fsdec}) is 93.75 kHz, the nominal pulse duration ($\tnom$) is 2 ms, and the effective pulse duration ($\teff$, \autoref{eq:TauEff}) is 0.02 ms.}

### 2023-09-29 Svf.py

@@ -170,9 +170,11 @@ theta_n, phi_n = Calculation.calcAngles(y_pc_halves_n, gamma_theta_f_c, gamma_ph
 
 # Sv estimation parameters
 step = 1  # Step in samples for sliding window
-
+          # In our example (main.py and figures in the paper) we have used step = 1.
+          # Note that step > 1 is possible and that processing time for larget datasets
+          # will be high when step = 1, and that we have not implemented this option here yet.
+          
 # Volume backscattering strength compressed frequency band
-
 Sv_n = Calculation.calcSv(
     p_rx_e_n, r_n, lambda_f_c, p_tx_e, alpha_f_c, c, tau_eff, psi_f_c, g_0_f_c
 )
